### 2023-04-13 src/tex/ms.tex
[showyourwork] first commit

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@ekaterinailin}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
Donec vehicula augue eu neque, morbi tristique senectus et netus et.
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
Phasellus eu tellus sit amet tortor gravida placerat.
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
Praesent eget sem vel leo ultrices bibendum.
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
Curabitur auctor semper nulla donec varius orci eget risus.
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2023-04-17 src/scripts/FIGURE_periodograms_xray_and_optical.py
periodograms added

@@ -1,0 +1,97 @@
+"""
+Python 3.8 - UTF-8
+
+X-ray Loops
+Ekaterina Ilin, 2023
+MIT License
+
+---
+
+This script reads in stacked PN and MOS, and optical monitoring light curves 
+from XMM-Newton, calculates periodograms for both, and plots the power spectra.
+"""
+
+
+import lightkurve as lk
+import pandas as pd
+import matplotlib.pyplot as plt
+
+import paths
+
+if __name__ == "__main__":
+
+    # TIC 277 rotation period in days
+    prot = 0.1900125
+
+    # ----------------------------------------------------------------------------
+    # X-ray periodogram
+
+    # read in stacked XMM light curve
+    df = pd.read_csv(paths.data / "stacked_xray_lightcurve.csv")
+
+    # get the maximum period 
+    time = df.time / 3600. / 24. # convert to days
+    max_period = (time.iloc[-1] - time.iloc[0]) / 2.
+
+    # convert to lightkurve object
+    lc = lk.LightCurve(time=df.time / 3600. / 24., flux=df.normalized_flux)
+
+    # make periodogram in period space
+    pg = lc.to_periodogram(method='lombscargle', minimum_period=0.01,
+                        maximum_period=max_period, oversample_factor=15) 
+
+
+
+    # plot periodogram in period
+    pg.plot(scale="log")
+    plt.axvline(pg.period_at_max_power.value, c="r",
+                linestyle="dotted",
+                label=f"period at max. power: {pg.period_at_max_power.value*24:.3f} h")
+
+    # make vertical lines for rotation period and harmonics
+    for fac in [.5, 1, 2, 4, 8, 16]:
+        plt.axvline(prot / fac)
+        plt.annotate(f"{prot/fac*24:.3f} h", (prot / fac*1.05, 1e-2),
+                    rotation=90, fontsize=11)
+
+    # layout
+    plt.legend(fontsize=11, frameon=True, loc=3)
+    plt.xlim(0.01, max_period)
+
+
+    plt.savefig(paths.figures  / "periodogram_xmm_stacked.png", dpi=300)
+
+    # ----------------------------------------------------------------------------
+    # optical periodogram
+
+    # read in OM data
+    df = pd.read_csv(paths.data / "timeseries.csv")
+
+    # convert to lightkurve object
+    lc = lk.LightCurve(time=df.time / 3600. / 24., flux=df.rate)
+
[…]

### 2023-04-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Title
-\title{An open source scientific article}
+\title{Corona of an M7 fast rotator with a high latitude flare}
 
@@
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-    Phasellus eu tellus sit amet tortor gravida placerat.
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-    Praesent eget sem vel leo ultrices bibendum.
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-    Curabitur auctor semper nulla donec varius orci eget risus.
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-    Duis eget orci sit amet orci dignissim rutrum.
+    Small scale fields
+
+    late M dwarfs
+
+    XMM-Newton observations
+
+    Flare
+
+    Anything unusual about the dwarf
+
 \end{abstract}
@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-Phasellus eu tellus sit amet tortor gravida placerat.
-Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-Praesent eget sem vel leo ultrices bibendum.
-Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-Curabitur auctor semper nulla donec varius orci eget risus.
-Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
+\section{Observations}
 
-Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
-Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
-Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
-Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
-Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
-Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
-Pellentesque cursus luctus mauris \citep{Luger2021}.
+XMM-Newton
 
-\bibliography{bib}
+uninterrupted two rotation periods
+
+MOS1/2 and PN, as well as OM data
+
+\section{Results}
+
+show light curve
+
+show spectra with fits and residuals
+
+\subsection{Rotational variability}
+
+\begin{figure*}[ht!]
+    \script{FIGURE_periodograms_xray_and_optical.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/periodogram_xmm_stacked.png}
+        \caption{
+         MOS and PN light curves stacked Lomb-Scargle periodogram.
+        }
+        \label{fig:coherence_hist}
+    \end{centering}
+\end{figure*}
+
+\begin{figure*}[ht!]
+    \script{FIGURE_periodograms_xray_and_optical.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/periodogram_om.png}
+        \caption{
+         MOS and PN light curves stacked Lomb-Scargle periodogram.
+        }
+        \label{fig:coherence_hist}
+    \end{centering}
+\end{figure*}
+
+periodograms
+
+present simple model and upper limits from the data
+
+\subsection{Coronal temperature and luminosity}
+
+spectral analysis
+
+APEC+APEC model, grsa abundances
+
+\subsection{Flare}
+
+optical and X-ray combined
+
+\section{Discussion}
+
+Ro-Lx relation
+
+\section{Summary and Conclusions}
+
+\section{Acknowledgements}
+\citep{lightkurvecollaboration2018lightkurve}
+
+\section{Data Availability}
+
+\bibliography{references}
 

### 2023-04-17 src/scripts/FIGURE_data_resid.py
added script to plot data resid

@@ -1,0 +1,90 @@
+"""
+Python 3.8 - UTF-8
+
+X-ray Loops
+Ekaterina Ilin, 2023
+MIT License
+
+---
+
+This script reads in stacked the writefits output from XSPEC, and plots
+data, model, and residuals in a two-panel figure.
+"""
+
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+import paths
+
+
+def plot_data_resid(file):
+    """Plot the data, model and residuals in a two panel figure.
+    
+    Parameters:
+    ----------
+    file : str
+        filename in results folder
+    
+    """
+
+    # column names
+    names = ["E [keV]", "dE", "flux [counts/s/keV]",
+             "e_flux", "model [counts/s/keV]"]
+
+    # read in the file
+    _ = pd.read_csv(paths.data / file, delimiter="\s+", skiprows=3, names=names)
+    
+    # split at the NO NO NO line
+    split_at = _[_["E [keV]"]=="NO"].index.values[0]
+
+    # define data and residuals
+    data = _.iloc[:split_at].astype(float)
+    resid = _.iloc[split_at+1:].drop("model [counts/s/keV]", axis=1).astype(float)
+
+    # make a figure with two subplots, one for data and model, and one for residuals
+    fig, ax = plt.subplots(nrows=2, ncols=1, sharex=True, figsize=(7,7),
+                           gridspec_kw={'height_ratios': [2.5, 1],
+                                        'wspace':0, 'hspace':0})
+    # model
+    ax[0].stairs(data["model [counts/s/keV]"], 
+                 np.append(data["E [keV]"].values-data["dE"].values, 
+                           data["E [keV]"].values[-1] + data["dE"].values[-1]), 
+                 color="white", edgecolor="k", linewidth=2)
+
+    # flux with errors
+    ax[0].errorbar(data["E [keV]"], data["flux [counts/s/keV]"],
+                yerr=data["e_flux"], fmt=".", markersize=13, alpha=1, c="olive")
+
+    # residuals
+    ax[1].errorbar(resid["E [keV]"], resid["flux [counts/s/keV]"],
+                yerr=resid["e_flux"], fmt=".", markersize=13, c="olive")
+
+    # zero line for residuals
+    ax[1].axhline(0,c="k")
+
+    # labels
+    ax[1].set_xlabel("E [keV]")
[…]

### 2023-04-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \label{sec:intro}
+
+\begin{table}
+
+    \caption{TIC 277539431 (WISEA J105515.71-735611.3)}
+    \begin{tabular}{l|r}
+         Parameter & Value  \\\hline\hline 
+         Distance & $13.70\pm0.11\,$pc [2] \\
+         $K_s$ mag & $9.666 \pm 0.024$ [3]\\
+         $J$ mag & $10.630 \pm 0.023$ [3] \\
+         SpT & M7 [1]\\
+         Rotation period & $273.618 \pm 0.007\,$min [1]\\
+         $v\sin i$ & $38.6\pm1.0\,$km/s [1] \\
+         Inclination & $87.0^{+2.0}_{-2.4}\,$ deg [1]\\
+         Radius & $0.145\pm0.004\,R_\odot$ [1]\\
+         TESS flaring rate $\beta$ &  \\
+
+    \end{tabular}
+    \newline\footnotesize
+    [1] \citet{ilin2021giant}, [2] \citet{bailer-jones2018estimating}, [3] 2MASS, \citet{skrutskie2006two}
+    \label{tab:modelparams}
+\end{table}
 
@@
         \caption{
-         MOS and PN light curves stacked Lomb-Scargle periodogram.
+         OM light curve Lomb-Scargle periodogram.
         }
@@
 \subsection{Coronal temperature and luminosity}
+
+\begin{figure}[ht!]
+    \script{FIGURE_data_resid.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/pn_onlyflare1_data_resid.png}
+        \caption{
+         Flare only PN spectrum fit
+        }
+        \label{fig:spec_pn_onlyflare}
+    \end{centering}
+\end{figure}
 

### 2023-04-18 src/scripts/FIGURE_lightcurves.py
added the light curves

@@ -1,0 +1,52 @@
+"""
+Python 3.8 - UTF-8
+
+X-ray Loops
+Ekaterina Ilin, 2023
+MIT License
+
+---
+
+This script reads in stacked PN and MOS, and optical monitoring light curves 
+from XMM-Newton, and plots them together.
+"""
+
+
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+import paths
+
+
+if __name__ == "__main__":
+
+    # read in optical data
+    om = pd.read_csv("../data/timeseries.csv")
+    om.time = om.time / 3600. / 24.
+    om.rate = om.rate / np.nanmedian(om.rate)
+
+    # read in X-ray data
+    xray = pd.read_csv("../data/stacked_xray_lightcurve.csv")
+    xray.time = xray.time / 3600. / 24.
+
+    # make the figure
+
+    fix, ax = plt.subplots(2, 1, sharex=True, figsize=(8, 6),
+                        gridspec_kw={'height_ratios': [2, 3],
+                                            'wspace':0, 'hspace':0})
+    ax[0].scatter(om.time, om.rate, label="OM", s=2)
+    ax[1].plot(xray.time, xray.normalized_flux + 1., label="PN + MOS1 + MOS2")
+    ax[0].set_ylim(0,3)
+
+    for a in ax:
+        a.axhline(1, c="grey")
+        a.set_xlim(om.time.min(), xray.time.max())
+        a.legend(loc=1, frameon=True)
+
+
+    ax[0].set_ylabel("normalized flux")
+    ax[1].set_ylabel("normalized flux")
+    ax[1].set_xlabel("time [days]")
+
+    plt.tight_layout()
+    plt.savefig(paths.figures / "lightcurves.png", dpi=300)

### 2023-04-18 src/scripts/FIGURE_lightcurves.py
added an inset showing the small precursor flare in OM

@@ -35,7 +35,20 @@ if __name__ == "__main__":
                         gridspec_kw={'height_ratios': [2, 3],
                                             'wspace':0, 'hspace':0})
     ax[0].scatter(om.time, om.rate, label="OM", s=2)
-    ax[1].plot(xray.time, xray.normalized_flux + 1., label="PN + MOS1 + MOS2")
+    ax[1].plot(xray.time, xray.normalized_flux + 1., label="PN + MOS1 + MOS2", c="olive")
+    
+    
+    # add inset
+    axins = ax[1].inset_axes([0.3, 0.55, 0.4, 0.5])
+    l, r = 600, 750
+    axins.scatter(om.time[l:r], om.rate[l:r], s=1)
+    axins.set_xlim(om.time[l], om.time[r])
+    axins.set_ylim(0, 3)
+    patch, lines = ax[0].indicate_inset_zoom(axins)
+    # [line.set(visible=False) for line in lines]
+
+
+    # layout
     ax[0].set_ylim(0,3)
 
     for a in ax:

### 2023-04-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Author list
-\author{@ekaterinailin}
+\author{Ekaterina Ilin, Katja Poppenh\"ager}
 
@@
 
+Small scale fields of late M dwarfs
+
+We know little about the corona of late M dwarfs
+
+Planets around late M dwarfs exists, see TRAPPIST-1 or LHS 3154~\citep{stefansson2023extreme}
+
+TIC 277 -- edge-on rotation, high-latitude flare
+
 \begin{table}
@@
     \caption{TIC 277539431 (WISEA J105515.71-735611.3)}
-    \begin{tabular}{l|r}
-         Parameter & Value  \\\hline\hline 
+    \begin{tabular}{ll}\hline 
+         Parameter & Value  \\\hline
          Distance & $13.70\pm0.11\,$pc [2] \\
@@
          Radius & $0.145\pm0.004\,R_\odot$ [1]\\
-         TESS flaring rate $\beta$ &  \\
+         TESS flaring rate $\beta$ &  \\\hline
 
@@
 
-show light curve
+Figure~\ref{fig:lightcurves} shows the time series in X-ray and optical.
+
+\begin{figure*}[ht!]
+    \script{FIGURE_lightcurves.py}
+    \begin{centering}
+        \includegraphics[width=0.75\linewidth]{figures/lightcurves.png}
+        \caption{
+         Top panel: Optical Monitoring (OM) light curve. Bottom panel: X-ray (PN, MOS1 and MOS2 combined) light curve. 
+        }
+        \label{fig:lightcurves}
+    \end{centering}
+\end{figure*}
 
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/periodogram_xmm_stacked.png}
+        \includegraphics[width=0.75\linewidth]{figures/periodogram_xmm_stacked.png}
         \caption{
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/periodogram_om.png}
+        \includegraphics[width=0.75\linewidth]{figures/periodogram_om.png}
         \caption{
@@
         \caption{
-         Flare only PN spectrum fit
+         Flare only PN Double apec spectrum fit.
         }

### 2023-04-18 src/scripts/FIGURE_lx_lbol.py
add figure to discussion

@@ -1,0 +1,80 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+import paths
+
+from astropy.table import Table
+
+def tau_wright2018(V, Ks, err=False, eV=None, eKs=None):
+    """Convective turnover time from Wright et al. 2018 using
+    Eq. 5 from that paper.
+
+    Parameters
+    ----------
+    V : float
+        The V magnitude of the star.
+    Ks : float
+        The Ks magnitude of the star.
+    err : bool, optional
+        If True, return the error on the Rossby number.
+        The default is False.
+    eV : float, optional
+        The error on the V magnitude of the star.
+        The default is None.
+    eKs : float, optional
+        The error on the Ks magnitude of the star.
+        The default is None.
+
+    Returns 
+    -------
+    tau : float
+        The convective turnover time of the star.
+    tau_err_high : float
+        The upper error on the convective turnover time.
+    tau_err_low : float
+        The lower error on the convective turnover time.
+
+    """
+
+    tau = 0.64 + 0.25 * (V - Ks)
+
+    if err:
+        tau_err_high = 0.74 + 0.33 * (V + eV - Ks + eKs)
+        tau_err_low = 0.54 + 0.17 * (V - eV - Ks - eKs)
+        return 10**tau, 10**tau_err_high, 10**tau_err_low
+    else:
+        return 10**tau
+
+
+
+if __name__ == "__main__":
+
+    # read Wright et al. 2011 data
+    wright2011 = Table.read(paths.data / "wright2011.fit")
+    wright2011["tau"] = tau_wright2018(wright2011["Vmag"], wright2011["Vmag"] - wright2011["V-K"])
+    wright2011["Ro"] = wright2011["Prot"] / wright2011["tau"]   
+
+    # read in TIC 277 data
+    df = pd.read_csv(paths.data / "apec2_results.csv")
+
+    plt.figure(figsize=(6.5,5))
+    
+    # TIC 277 data
+    plt.errorbar(df.Rossby, df.Lx_Lbol, 
+                xerr=[df.Rossby-df.Rossby_low,df.Rossby_high-df.Rossby], 
+                yerr=df.e_Lx_Lbol, fmt='o', label="TIC 277", alpha=1, c="k")
+    
+    # Wright et al. 2011 data
+    plt.scatter(wright2011["Ro"].value, 10**wright2011["Lx_bol"].value,
[…]

### 2023-04-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
          Distance & $13.70\pm0.11\,$pc [2] \\
+         $G$ mag & \\
+         $G_{BP}$ mag & \\
+         $G_{BP}$ mag & \\
          $K_s$ mag & $9.666 \pm 0.024$ [3]\\
@@
 
-Ro-Lx relation
+\begin{figure}[ht!]
+    \script{FIGURE_lx_lbol.py}
+    \begin{centering}
+        \includegraphics[width=0.75\linewidth]{figures/lx_lbol.png}
+        \caption{
+         Rossby number versus $L_x/L_{bol}$.
+        }
+        \label{fig:lxlbol}
+    \end{centering}
+\end{figure}
 

### 2023-04-19 src/scripts/FIGURE_data_resid.py
added plotting for joint fits

@@ -17,7 +17,76 @@ import matplotlib.pyplot as plt
 import paths
 
 
-def plot_data_resid(file):
+def plot_data_resid_3(file):
+
+
+     # column names
+    names = ["E [keV]", "dE", "flux [counts/s/keV]",
+             "e_flux", "model [counts/s/keV]"]
+
+    # read in the file
+    _ = pd.read_csv(paths.data / file, delimiter="\s+", skiprows=3, names=names)
+    
+    # split at the NO NO NO line
+    split_at = _[_["E [keV]"]=="NO"].index.values
+
+    d = _[:split_at[0]], _[split_at[0]+1:split_at[1]], _[split_at[1]+1:split_at[2]]
+    r = _[split_at[2]+1:split_at[3]], _[split_at[3]+1:split_at[4]], _[split_at[4]+1:]
+
+    # make a figure with two subplots, one for data and model, and one for residuals
+    fig, ax = plt.subplots(nrows=2, ncols=1, sharex=True, figsize=(7,7),
+                           gridspec_kw={'height_ratios': [2.5, 1],
+                                        'wspace':0, 'hspace':0})
+    
+    for din, rin, c, label, m in zip(d, r, ["olive", "orange", "blue"],
+                                     ["MOS1", "MOS2", "PN"],
+                                     ["o", "s", "x"]):
+
+        data = din.astype(float)
+        resid = rin.drop("model [counts/s/keV]", axis=1).astype(float)
+
+         # model
+        ax[0].stairs(data["model [counts/s/keV]"], 
+                    np.append(data["E [keV]"].values-data["dE"].values, 
+                            data["E [keV]"].values[-1] + data["dE"].values[-1]), 
+                    color="white", edgecolor=c, linewidth=2, label=label)
+
+        # flux with errors
+        ax[0].errorbar(data["E [keV]"], data["flux [counts/s/keV]"],
+                    yerr=data["e_flux"], fmt=m, markersize=6, alpha=1, c=c)
+
+        # residuals
+        ax[1].errorbar(resid["E [keV]"], resid["flux [counts/s/keV]"],
+                    yerr=resid["e_flux"], fmt=m, markersize=6, c=c)
+        
+      # zero line for residuals
+    ax[1].axhline(0,c="k")
+
+    # labels
+    ax[1].set_xlabel("E [keV]")
+    ax[0].set_ylabel("flux [counts/s/keV]")
[…]

### 2023-04-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
+\begin{figure}[ht!]
+    \script{FIGURE_data_resid.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/joint_onlyflare_data_resid.png}
+        \caption{
+         Flare only joint Double APEC spectrum fit.
+        }
+        \label{fig:spec_joint_onlyflare}
+    \end{centering}
+\end{figure}
+
+\begin{figure}[ht!]
+    \script{FIGURE_data_resid.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/joint_noflare_data_resid.png}
+        \caption{
+         Joint Double APEC spectrum fit with flare excluded.
+        }
+        \label{fig:spec_joint_noflare}
+    \end{centering}
+\end{figure}
+
+\begin{figure}[ht!]
+    \script{FIGURE_data_resid.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/joint_all_data_resid.png}
+        \caption{
+         Joint Double APEC spectrum fit.
+        }
+        \label{fig:spec_joint_all}
+    \end{centering}
+\end{figure}
+
 spectral analysis
@@
     \begin{centering}
-        \includegraphics[width=0.75\linewidth]{figures/lx_lbol.png}
+        \includegraphics[width=\linewidth]{figures/lx_lbol.png}
         \caption{

### 2023-04-19 src/scripts/FIGURE_data_resid.py
more data, more plots

@@ -155,8 +155,8 @@ def plot_data_resid_1(file):
 if __name__ == "__main__":
 
     # filename
-    files = ["pn_onlyflare1.txt", "joint_onlyflare.txt", "joint_noflare.txt",
-             "joint_all.txt"]
+    files = ["pn_onlyflare.txt", "joint_onlyflare.txt", "joint_noflare.txt",
+             "joint_all.txt", 'mos1.txt', 'mos2.txt', 'pn.txt', "pn_noflare.txt"]
 
     # plot
     for file in files:

### 2023-04-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\begin{figure}[ht!]
+\begin{figure}
     \script{FIGURE_data_resid.py}
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/pn_onlyflare1_data_resid.png}
+        \includegraphics[width=\linewidth]{figures/pn_onlyflare_data_resid.png}
         \caption{
@@
 
-\begin{figure}[ht!]
+\begin{figure}
+    \script{FIGURE_data_resid.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/pn_onlyflare_data_resid.png}
+        \caption{
+         No flare PN Double apec spectrum fit.
+        }
+        \label{fig:spec_pn_noflare}
+    \end{centering}
+\end{figure}
+
+\begin{figure}
+    \script{FIGURE_data_resid.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/pn_data_resid.png}
+        \caption{
+         PN Double apec spectrum fit.
+        }
+        \label{fig:spec_pn}
+    \end{centering}
+\end{figure}
+
+\begin{figure}
+    \script{FIGURE_data_resid.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/mos1_data_resid.png}
+        \caption{
+         MOS1 Double apec spectrum fit.
+        }
+        \label{fig:spec_mos1}
+    \end{centering}
+\end{figure}
+
+\begin{figure}
+    \script{FIGURE_data_resid.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/mos2_data_resid.png}
+        \caption{
+         MOS2 Double apec spectrum fit.
+        }
+        \label{fig:spec_mos2}
+    \end{centering}
+\end{figure}
+
+\begin{figure}
     \script{FIGURE_data_resid.py}
@@
 
-\begin{figure}[ht!]
+\begin{figure}
     \script{FIGURE_data_resid.py}

### 2023-04-19 src/scripts/TABLE_specfits.py
added table with fit results

@@ -1,0 +1,161 @@
+"""
+Python 3.8 - UTF-8
+
+X-ray Loops
+Ekaterina Ilin, 2023
+MIT License
+
+---
+
+This script reads in the XPEC fit results and derived parameters and produces
+a LaTeX table.
+"""
+
+import pandas as pd
+import numpy as np
+import paths
+
+def convert_to_scinote(series, rel_err=1e-2):
+    """Convert a series to scientific notation.
+
+    Parameters
+    ----------
+    series : pd.Series
+        The series to convert.
+
+    Returns
+    -------
+    pd.Series
+        The converted series of strings.
+    """
+    return series.apply(lambda x: f"{x:.1e}" if np.abs(x)<rel_err else f"{x:.1f}").astype(str)
+
+
+
+def tex_up_low(val, err, err2):
+    """Convert a value and two errors into a LaTeX string.
+
+    Parameters
+    ----------
+    val : float
+        The value.
+    err : float
+        The first error.
+    err2 : float
+        The second error.
+    
+    Returns
+    -------
+    str
+        The LaTeX string.
+    """
+
+    return ("$" +
+            convert_to_scinote(val) +
+            "^{" + 
+            convert_to_scinote(err) + 
+            "}_{" +
+            convert_to_scinote(err2) + "}$")
+
+def tex_one_err(val, err):
+    """Convert a value and one error into a LaTeX string.
+
+    Parameters
+    ----------
+    val : float
+        The value.
+    err : float
+        The error.
+
+    Returns
+    -------
+    str
+        The LaTeX string.
+    """
+
+    return ("$" +
+            convert_to_scinote(val) +
+            "[" +
+            convert_to_scinote(err) +
+            "]$")
+
+if __name__ == "__main__":
+
+    # read in the data
+    df = pd.read_csv(paths.data / "apec2_results.csv")
+    print(df.shape)
+    dfjoint = pd.read_csv(paths.data / "apec2_joint_results.csv")
+    print(dfjoint.shape)
+    # append the joint results
+    df = pd.concat([df, dfjoint], ignore_index=True)
+    print(df.shape)
+
+    # values with one error
+    cols2 = [("Lx_erg_s","Lx_erg_s_err"),
[…]

### 2023-04-19 src/scripts/TABLE_specfits.py
up and low limits were the wrong way

@@ -28,7 +28,7 @@ def convert_to_scinote(series, rel_err=1e-2):
     pd.Series
         The converted series of strings.
     """
-    return series.apply(lambda x: f"{x:.1e}" if np.abs(x)<rel_err else f"{x:.1f}").astype(str)
+    return series.apply(lambda x: f"{x:.1e}" if np.abs(x)<rel_err else f"{x:.2f}").astype(str)
 
 
 
@@ -124,10 +124,10 @@ if __name__ == "__main__":
         del df[err]
 
     # convert to latex strings
-    for col, up, low in cols_mid_low_high:
+    for col, low, up in cols_mid_low_high:
         
         newname = mapcolnames[col]
-        df[newname] = tex_up_low(df[col], df[up]-df[col], df[col]-df[low])
+        df[newname] = tex_up_low(df[col], df[up]-df[col], df[low]-df[col])
         del df[col]
         del df[up]
         del df[low]

### 2023-04-24 src/scripts/FIGURE_tess_ffd.py
TESS LCs and FFD

@@ -1,0 +1,94 @@
+"""
+Python 3.8 - UTF-8
+
+X-ray Loops
+Ekaterina Ilin, 2023
+MIT License
+
+---
+
+This script reads in the flare table and FFD fitting results, and plots the FFD
+and power law.
+"""
+
+
+import paths
+
+import numpy as np
+import pandas as pd
+
+from altaipony.ffd import FFD
+
+import matplotlib.pyplot as plt
+
+from astropy.modeling import models
+from astropy import units as u
+
+if __name__ == "__main__":
+
+    # stellar Teff
+    teff = 2680
+
+    # stellar radius
+    radius = 0.145
+
+    # TESS band response
+    tessresp = pd.read_csv(paths.data / "TESS_response.csv")
+
+    # blackbody
+    bb = models.BlackBody(temperature=teff * u.K)
+
+    # blackbody flux in TESS band
+    bbtess = bb(tessresp["WAVELENGTH"].values * u.nm)  * tessresp["PASSBAND"].values
+
+    # flux in TESS band
+    fstar = np.trapz(bbtess.value, tessresp["WAVELENGTH"].values)
+
+    # flux in all bands
+    wav = np.linspace(5,3e4, 20000)
+    fall = np.trapz(bb(wav * u.nm).value, wav)
+
+    # luminosity in TESS band
+    lumtess = (fstar / fall * bb.bolometric_flux * np.pi * (radius * u.Rsun)**2).to("erg/s")
+
+    # get FFD values
+    ffd_vals = pd.read_csv(paths.data / "tess_ffd.csv")
+
+    # get flares
+    df = pd.read_csv(paths.data / "tess_flares.csv")
+
+    # convert ED to E
+    df["ed_rec"] = df["ed_rec"] * lumtess
+    df["ed_rec_err"] = df["ed_rec_err"] * lumtess
+
+    # make FFD
+    ffd = FFD(df, tot_obs_time=df["tot_obs_time"].values[0])
+    ffd.alpha = ffd_vals["alpha"].values[0]
+    ffd.beta = ffd_vals["beta"].values[0] * lumtess**(ffd.alpha -1)
+    ffd.alpha_low_err = ffd_vals["alpha_low_err"].values[0]
+    ffd.alpha_up_err = ffd_vals["alpha_up_err"].values[0]
+    ed, freq, counts = ffd.ed_and_freq()
+
+    # plot
[…]

### 2023-04-24 src/scripts/FIGURE_tess_lcs.py

@@ -1,0 +1,56 @@
+"""
+Python 3.8 - UTF-8
+
+X-ray Loops
+Ekaterina Ilin, 2023
+MIT License
+
+---
+
+This script reads in the de-trended TESS light curves and plots them in a
+three-panel figure.
+"""
+
+
+from astropy.table import Table
+import matplotlib.pyplot as plt
+import paths
+
+import numpy as np
+
+
+if __name__ == "__main__":
+
+    # define sectors
+    sectors = 12, 37, 39
+
+    # read in light curves
+    lcrs = [Table.read(paths.data / f"tic277_tess_detrended_{s}.fits") for s in sectors]
+
+    # make three subplots with two light curves each, one showing the flux the other the
+    # detrended flux
+    fig, axes = plt.subplots(3, 1, figsize=(10, 9))
+
+    # loop over axes, light curves and sectors
+    for ax, lcr, sector in zip(axes, lcrs, sectors):
+
+        # plot the un-detrended and detrended flux
+        ax.plot(lcr['TIME'], lcr['FLUX']/np.nanmedian(lcr["FLUX"]), 'k.',
+                label=f"sector {sector}", ms=1.5)
+        ax.plot(lcr['TIME'], lcr['DETRENDED_FLUX']/np.nanmedian(lcr["DETRENDED_FLUX"]) +0.2, 
+                'r.', ms=1.5)
+        
+        # layout
+        ax.set_ylabel("flux")
+        ax.set_ylabel("normalized flux")
+        ax.set_ylim(0.9, 1.5)
+        ax.set_xlim(lcr["TIME"][0], lcr["TIME"][-1])
+
+        # add sector number
+        ax.text(0.02, 0.9, f"Sector {sector}", transform=ax.transAxes)
+
+    # shared x-axis
+    axes[-1].set_xlabel("time [BJD - 2457000]")
+
+    plt.tight_layout()
+    plt.savefig(paths.figures / "tess_lcs.png", dpi=250)

### 2023-04-24 src/scripts/FIGURE_tess_ffd.py
make ffds

@@ -23,9 +23,55 @@ import matplotlib.pyplot as plt
 
 from astropy.modeling import models
 from astropy import units as u
+from astropy.constants import sigma_sb
+
+
+def flare_factor(teff, radius, wav, resp,  tflare=10000):
+    """Calculate the flare energy factor in ergs.
+
+    Parameters
+    ----------
+    teff : float
+        Stellar effective temperature in Kelvin.
+    radius : float
+        Stellar radius in solar radii.
+    wav : array
+        Array of wavelengths in nanometers.
+    resp : array
+        Array of bandpass responses.
+     tflare : float
+        Flare temperature in Kelvin.
+    
+    Returns
+    -------
+    factor : float
+        Flare energy factor in ergs/s.
+    """
+
+    # blackbody
+    bb = models.BlackBody(temperature=teff * u.K)
+
+    # blackbody flux in TESS band
+    bbwavs = bb(wav * u.nm)  * resp
+
+    fluxs = np.trapz(bbwavs.value, wav)
+
+    # blackbody
+    bb = models.BlackBody(temperature=tflare * u.K)
+
+    # blackbody flux in TESS band
+    bbwavf = bb(wav * u.nm)  * resp
+
+    fluxf = np.trapz(bbwavf.value, wav)
+
+    ratio = fluxs / fluxf
+
+    factor = ratio * np.pi * (radius * u.R_sun) ** 2 * sigma_sb * (tflare * u.K)**4
+
+    return factor.to("erg/s")
+
 
 if __name__ == "__main__":
-
     # stellar Teff
     teff = 2680
 
@@ -35,21 +81,7 @@ if __name__ == "__main__":
     # TESS band response
     tessresp = pd.read_csv(paths.data / "TESS_response.csv")
 
-    # blackbody
-    bb = models.BlackBody(temperature=teff * u.K)
-
-    # blackbody flux in TESS band
-    bbtess = bb(tessresp["WAVELENGTH"].values * u.nm)  * tessresp["PASSBAND"].values
-
-    # flux in TESS band
-    fstar = np.trapz(bbtess.value, tessresp["WAVELENGTH"].values)
-
-    # flux in all bands
-    wav = np.linspace(5,3e4, 20000)
-    fall = np.trapz(bb(wav * u.nm).value, wav)
-
-    # luminosity in TESS band
[…]

### 2023-04-24 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Title
-\title{Corona of an M7 fast rotator with a high latitude flare}
+\title{Case study: The corona of an M7 fast rotator with a high latitude flare}
 
@@
 
+Age unknown ...
+\citep{schneider2018hazmat} classify it as field star, but its FUV and NUV fluxes are on the high end for field stars and would be compatible with the NUV/FUV emission from younger open cluster members as well
+
 \begin{table}
@@
          SpT & M7 [1]\\
+         age & \\
          Rotation period & $273.618 \pm 0.007\,$min [1]\\
@@
 
+From the proposal: $F_X=2.7^{-13}$ erg/s/cm$^2$, $L_X=6\cdot10^27$ erg/s (update with new eROSITA data) -- 1 oom above the measured XMM values.
+\begin{table*}
+% \movetableright=-20mm
+\footnotesize
+    \script{TABLE_specfits.py}
+    \caption{XSPEC fits to PN, MOS1/MOS2 and joint data also.}
+    \input{output/specfit.tex}
+        \label{tab:specfit}
+\end{table*}
+
 \begin{figure}
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/pn_onlyflare_data_resid.png}
+        \includegraphics[width=\linewidth]{figures/pn_noflare_data_resid.png}
         \caption{
@@
 \section{Discussion}
+\citep{pass2022constraints} on rotational evol of late Ms
+
+FFD power law fit converged after 16000 steps on $\alpha = 1.77_{-0.23}^{+0.29}$.
+
+$\log_10(R_{31.5}$, that is, the rate of flares per day with energies above $\log E_{flare}=10^{31.5}\,$erg, is ca. $-0.7$, consistent with other late M dwarfs~\citep{medina2022galactic}.
+
+\begin{figure}
+    \script{FIGURE_tess_ffd.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/tess_ffd.png}
+        \caption{
+         Flare frequency distribution and power law fit.
+        }
+        \label{fig:ffd}
+    \end{centering}
+\end{figure}
 

### 2023-04-24 src/scripts/FIGURE_lightcurves.py
fixed paths

@@ -21,12 +21,12 @@ import paths
 if __name__ == "__main__":
 
     # read in optical data
-    om = pd.read_csv("../data/timeseries.csv")
+    om = pd.read_csv(paths.data / "timeseries.csv")
     om.time = om.time / 3600. / 24.
     om.rate = om.rate / np.nanmedian(om.rate)
 
     # read in X-ray data
-    xray = pd.read_csv("../data/stacked_xray_lightcurve.csv")
+    xray = pd.read_csv(paths.data / "stacked_xray_lightcurve.csv")
     xray.time = xray.time / 3600. / 24.
 
     # make the figure

### 2023-06-12 src/scripts/FIGURE_lx_lbol.py
added OM flare to FFD

@@ -59,6 +59,7 @@ if __name__ == "__main__":
     plt.figure(figsize=(6.5,5))
     
     # TIC 277 data
+    df = df[df["label"] == "PN (all)"]
     plt.errorbar(df.Rossby, df.Lx_Lbol, 
                 xerr=[df.Rossby-df.Rossby_low,df.Rossby_high-df.Rossby], 
                 yerr=df.e_Lx_Lbol, fmt='o', label="TIC 277", alpha=1, c="k")

### 2023-06-12 src/scripts/FIGURE_tess_ffd.py

@@ -72,55 +72,76 @@ def flare_factor(teff, radius, wav, resp,  tflare=10000):
 
 
 if __name__ == "__main__":
-    # stellar Teff
-    teff = 2680
 
-    # stellar radius
-    radius = 0.145
+    tic_teff_rad = [(277539431, 2680, 0.145),
+                    (237880881, 3060, 0.275),
+                    (452922110, 2680, 0.137),
+                    (44984200, 2810, 0.145)]
 
-    # TESS band response
-    tessresp = pd.read_csv(paths.data / "TESS_response.csv")
+    for tic, teff, radius in tic_teff_rad:
 
-    factor = flare_factor(2680, 0.145, tessresp["WAVELENGTH"].values, tessresp["PASSBAND"].values)
+        # TESS band response
+        tessresp = pd.read_csv(paths.data / "TESS_response.csv")
 
-    # get FFD values
-    ffd_vals = pd.read_csv(paths.data / "tess_ffd.csv")
+        factor = flare_factor(teff, radius, tessresp["WAVELENGTH"].values, tessresp["PASSBAND"].values)
 
-    # get flares
-    df = pd.read_csv(paths.data / "tess_flares.csv")
+        # get FFD values
+        ffd_vals = pd.read_csv(paths.data / "tess_ffd.csv")
+        ffd_vals = ffd_vals[ffd_vals["TIC"] == tic]
 
-    # convert ED to E
-    df["ed_rec"] = df["ed_rec"] * factor
-    df["ed_rec_err"] = df["ed_rec_err"] * factor
+        # get flares
+        df = pd.read_csv(paths.data / "tess_flares.csv")
+        df = df[df["TIC"] == tic]
 
-    # make FFD
-    ffd = FFD(df, tot_obs_time=df["tot_obs_time"].values[0])
-    ffd.alpha = ffd_vals["alpha"].values[0]
-    ffd.beta = ffd_vals["beta"].values[0] * factor**(ffd.alpha -1)
-    ffd.alpha_low_err = ffd_vals["alpha_low_err"].values[0]
-    ffd.alpha_up_err = ffd_vals["alpha_up_err"].values[0]
-    ed, freq, counts = ffd.ed_and_freq()
+        # convert ED to E
+        df["ed_rec"] = df["ed_rec"] * factor
+        df["ed_rec_err"] = df["ed_rec_err"] * factor
 
-    # plot
[…]

### 2023-06-15 src/scripts/FIGURE_lx_lbol.py
updated calculations for r315

@@ -1,3 +1,17 @@
+"""
+Python 3.8 - UTF-8
+
+X-ray Loops
+Ekaterina Ilin, 2023
+MIT License
+
+---
+
+This script plots the Lx/Lbol vs Rossby number for TIC 277 and 
+the Wright et al. 2011 sample.
+"""
+
+
 import pandas as pd
 import matplotlib.pyplot as plt
 import paths

### 2023-06-15 src/scripts/FIGURE_r315_comparison.py

@@ -1,0 +1,112 @@
+"""
+Python 3.8 - UTF-8
+
+X-ray Loops
+Ekaterina Ilin, 2023
+MIT License
+
+---
+
+This script plots the FFD results from Medina et al. 2020, Murray et al. 2022,
+Trappist-1, and the four stars with high latitude flares from Ilin et al. 2021.
+"""
+
+import paths
+
+import pandas as pd
+import numpy as np
+
+from astropy.table import Table
+
+import matplotlib.pyplot as plt
+
+import adjustText as aT
+
+if __name__ == "__main__":
+
+    # get Medina et al table
+    df = Table.read(paths.data / "medina2020.fit")
+    df = df.to_pandas()
+
+    # get FFD values
+    ffd_vals = pd.read_csv(paths.data / "tess_ffd.csv")
+
+    # get rotation periods
+    ilin2021 = pd.read_csv(paths.data / "ilin2021updated_w_Rossby_Lbol.csv")
+
+    ffd_vals = ffd_vals.merge(ilin2021, on="TIC")
+
+    # calculate r315
+    r315s = []
+    for tic, ff in ffd_vals.groupby("TIC"):
+
+        # energy beta
+        with open(paths.data / f"{tic}_energy_beta.txt", "r") as f:
+            beta = float(f.read())
+
+        f315 = beta / (ff.alpha - 1) * (10**31.5)**(-ff.alpha +1)
+        r315s.append(np.log10(f315).values[0])
+
+    ffd_vals["r315"] = r315s
+
+
+    # TRAPPIST-1
+    trapp_r315 = np.log10(10**(-2.5) * 24) # paudel 2018
+    trapp_ro = 0.047 # roettenbacher 2017
+    trapp_rot = 3.3 
+
+    # data from Murray 2022 Table 1
+    murray = [("M4-M5",np.log10(.055),"dashed"),
+            ("M6", np.log10(0.112),"dotted"),
+            ("M7", np.log10(0.06),"solid")]
+
+
+    # make plot
+    plt.figure(figsize=(8,4))
+
+    # add Murray+2022
+    for label, r, sty in murray:
+        plt.axhline(r, label=label + " (Murray+2022)", linestyle=sty, c="darkgrey",zorder=-10)
+
+
+    # color code Medina+2020 by mass
+    color_marker_label = [("palegreen","o",0.1,0.12),
+                        ("yellow","d",0.12,0.15),
[…]

### 2023-06-15 src/scripts/FIGURE_tess_ffd.py

@@ -77,6 +77,8 @@ if __name__ == "__main__":
                     (237880881, 3060, 0.275),
                     (452922110, 2680, 0.137),
                     (44984200, 2810, 0.145)]
+    
+    
 
     for tic, teff, radius in tic_teff_rad:
 
@@ -126,7 +128,7 @@ if __name__ == "__main__":
             plt.plot(x, pl, linestyle="dashed", c= "grey")
 
         # FFD
-        plt.scatter(ed, freq, c="k")
+        plt.scatter(ed, freq, c="k", label="TESS")
 
         # power law
         ffd.plot_powerlaw(ax, c="olive", label=label)

### 2023-06-15 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Title
-\title{Case study: The corona of an M7 fast rotator with a high latitude flare}
+\title{Does a fully convective star need a special corona to produce a polar flare?}
 
 % Author list
-\author{Ekaterina Ilin, Katja Poppenh\"ager}
+\author{Ekaterina Ilin, Katja Poppenh\"ager, Beate Stelzer}
 
@@
 \begin{abstract}
-    Small scale fields
+    TIC 277 is currently the star with the highest latitude flare located to date.
+In 2020, TESS observed this rapidly rotating M7 dwarf produce a flare at 81 deg latitude.
+This is in stark contrast to solar flares that occur much closer to the equator, typically below 30 deg.
+Does a star require a special corona to produce polar flares?
 
-    late M dwarfs
-
-    XMM-Newton observations
-
-    Flare
-
-    Anything unusual about the dwarf
+    We obtained 36 ks ob XMM-Newton observations, and determined that the X-ray luminosity $L_X$ is slightly lower but not far below the $L_X$ of other stars with similar spectral types a rotation rates.
+We constructed flare frequency distributions using the three TESS Sectors, with a power law slope of $XXXX\pm XXXX$, and flare rate above $\log_{10}E=31.5\,$erg of XXXX, again typical of similar stars.
+Our results suggest that the star's magnetic activity is typical for its rotation rate and mass.
+However, the rotation period of 4.56 hours is on the short end, which could indicate that polar flares on these stars are a feature of youth.
 
@@
 \label{sec:intro}
+The smallest stars are crawling with rocky planets.
+The mass of the disk decreases such that giant planet are no longer efficiently formed, leaving room for smaller objects in the orbit.
+Terrestrial planets in the habitable zones of these stars are a factor of XXXX more likely than around higher mass stars.
+However, to be truly habitable in an Earth-like sense, more conditions need to be met than simply the possibility of surface water in its liquid form.
+One of them is the space weather of the host, that is, the high energy radiation and particles that force the planet's atmosphere.
+If too high, the atmosphere may blow off.
+If too energetic, life may not survive.
+If not energetic enough, life may not emerge in the first place.
+So, what is the space weather of planets around these smallest, and most abundant, stars, like?
 
-Small scale fields of late M dwarfs
+At the lowest masses, a star's life unfolds much slower than for a Sun.
+While the Sun's magnetic activity, and with it the flares, winds, coronal mass ejections and energetic particle eruption that make up space weather, decline rapidly over a few hundred megayears, a fully convective M dwarf stays highly active for gigayears.
+Add that the zone where liquid water can reside is much closer to these faint stars, and the result is a planet that may be exposed to violent space weather conditions for a good fraction of the age of the universe.
 
-We know little about the corona of late M dwarfs
+Space weather originates from the stellar corona.
+The transition from star to brown dwarf is characterized by increasing rotation speed and decreasing coronal emission.
+The simultaneous surge of radio emission indicates a transition from stellar corona to planet-like magnetosphere at the bottom of the main sequence.
+Yet, brown dwarfs are regularly detected with energetic flares regardless of their low X-ray luminosity.
+How these flares are produced in the apparent absence of a solar-type corona is unclear.
+An intermittent corona is conceivable, similarly XYZ.
+Perhaps, the rapid rotation of these low-mass objects plays a role, as it governs the dynamo process within.
 
-Planets around late M dwarfs exists, see TRAPPIST-1 or LHS 3154~\citep{stefansson2023extreme}
+\citet{ilin2021giant} searched fully convective dwarfs observed in TESS, and found four flares on four rapidly rotating ($P<10\,$h) M5-M7 dwarfs.
+These flares' durations spanned multiple rotation periods of each star, so that the flares were observed to rotate in and out view.
+The shape of this modulation together with known inclination (combining $P$ and $v \sin i$ from high resolution spectroscopy) allowed them to determine the latitudes of these flares.
+Surprisingly, these flares were found closer to the pole than to the equator, in contrast to the Sun, where flares are usually found below 30 deg latitude.
+Their results indicate a preference for flares at high latitudes in these star, as a chance finding was unlikely ($0.1\%$).
+If flares and associated particle eruptions have a latitude preference, the space weather of a planet in an aligned orbit would be less severe than previously thought.
+%Are high flaring latitudes  an effect of the dynamo at high rotation speeds?
+Or is there something special about these stars' coronae that allowed them to produce flares at these latitudes?
 
-TIC 277 -- edge-on rotation, high-latitude flare
+%Planets around late M dwarfs exists, see TRAPPIST-1 or LHS 3154~\citep{stefansson2023extreme}
 
-Age unknown ...
-\citep{schneider2018hazmat} classify it as field star, but its FUV and NUV fluxes are on the high end for field stars and would be compatible with the NUV/FUV emission from younger open cluster members as well
+TIC 277539431, TIC 277 for short, is one out of four stars with a high-latitude flare in its TESS observations~\citep{ilin2021giant}.
+In fact, it showed the highest latitude flare known to date at $81\pm1\,$deg.
+TIC 277 is an M7 dwarf with a rotation period of 4.56 hours.
+Its age is unknown.
+\citet{schneider2018hazmat} classify it as a field star, but its FUV and NUV fluxes are on the high end for field stars and would be compatible with the NUV/FUV emission from younger open cluster members, as well.
+Is its high flaring latitude an effect of the dynamo at high rotation speeds? Or is there something special about this star's corona that led it to produce are flare at these latitudes?
+
+In this work, we follow TIC 277 with XMM-Newton for 36 ks, derive its coronal properties, and investigate if producing a high-latitude flare requires an unusual corona, or if it is rather typical of stars of this spectral type and rotation.
 
@@
          Inclination & $87.0^{+2.0}_{-2.4}\,$ deg [1]\\
-         Radius & $0.145\pm0.004\,R_\odot$ [1]\\
-         TESS flaring rate $\beta$ &  \\\hline
+         Radius & $0.145\pm0.004\,R_\odot$ [1]\\\hline
 
@@
 
+\subsection{XMM-Newton}
 XMM-Newton
@@
 MOS1/2 and PN, as well as OM data
+
+\subsection{TESS}
+
+\section{Methods}
+\label{sec:methods}
+
+\subsection{Flux and luminosity}
+
+\subsection{Spectral fitting}
+
+\subsection{Flares}
+
+\subsubsection{XMM-Newton flare}
+
+X-ray energy of $8.4\pm0.7\cdot10^{30}\,$erg
+
+\subsubsection{TESS flares}
 
@@
 
-\subsection{Rotational variability}
+% \subsection{Rotational variability}
 
-\begin{figure*}[ht!]
-    \script{FIGURE_periodograms_xray_and_optical.py}
-    \begin{centering}
-        \includegraphics[width=0.75\linewidth]{figures/periodogram_xmm_stacked.png}
-        \caption{
-         MOS and PN light curves stacked Lomb-Scargle periodogram.
-        }
-        \label{fig:coherence_hist}
-    \end{centering}
-\end{figure*}
+% \begin{figure*}[ht!]
+%     \script{FIGURE_periodograms_xray_and_optical.py}
+%     \begin{centering}
+%         \includegraphics[width=0.75\linewidth]{figures/periodogram_xmm_stacked.png}
+%         \caption{
+%          MOS and PN light curves stacked Lomb-Scargle periodogram.
+%         }
+%         \label{fig:coherence_hist}
+%     \end{centering}
+% \end{figure*}
 
-\begin{figure*}[ht!]
-    \script{FIGURE_periodograms_xray_and_optical.py}
-    \begin{centering}
-        \includegraphics[width=0.75\linewidth]{figures/periodogram_om.png}
-        \caption{
-         OM light curve Lomb-Scargle periodogram.
-        }
-        \label{fig:coherence_hist}
-    \end{centering}
-\end{figure*}
+% \begin{figure*}[ht!]
+%     \script{FIGURE_periodograms_xray_and_optical.py}
+%     \begin{centering}
+%         \includegraphics[width=0.75\linewidth]{figures/periodogram_om.png}
+%         \caption{
+%          OM light curve Lomb-Scargle periodogram.
+%         }
+%         \label{fig:coherence_hist}
+%     \end{centering}
+% \end{figure*}
 
-periodograms
+% periodograms
 
-present simple model and upper limits from the data
+% present simple model and upper limits from the data
+
+% how large of a loop would that be? convert relative amplitude to flux, convert flux to emission measure, convert emission measure to loop size?
 
@@
 
-\subsection{Flare}
+\subsection{Flares}
 
@@
 
-\section{Discussion}
-\citep{pass2022constraints} on rotational evol of late Ms
-
 FFD power law fit converged after 16000 steps on $\alpha = 1.77_{-0.23}^{+0.29}$.
 
-$\log_10(R_{31.5}$, that is, the rate of flares per day with energies above $\log E_{flare}=10^{31.5}\,$erg, is ca. $-0.7$, consistent with other late M dwarfs~\citep{medina2022galactic}.
+\section{Discussion}
+\label{sec:discussion}
+% \citep{pass2022constraints} on rotational evol of late Ms
+
+Is the magnetic activity of the star with the highest known flare latitude known to date is any way unusual?
+The star is rotating rapidly, even for the typically fast rotating late M dwarfs~\citep{medina2022galactic}.
+This could lead to an unusual manifestation of the stellar dynamo that produces high latitude flares that other dynamo states cannot generate.
+With TESS and XMM-Newton, we can now evaluate if this suspected exceptionality is reflected in the star's magnetic activity and coronal properties.
+We compare its flaring activity, coronal luminosity and temperatures to stars of similar spectral type and rotation rates
+
+\subsection{Flaring activity}
+\label{sec:discussion:flares}
+
+$\log_10(R_{31.5}$, that is, the rate of flares per day with energies above $\log E_{flare}=10^{31.5}\,$erg, is ca. $-1.XXXX$, consistent with other saturated late M dwarfs~\citep{medina2022galactic}.
+
+Compared to the X-ray luminosities of M7 dwarfs among the late M dwarfs observed by eROSITA in \citet{stelzer2022first}, TIC 277 appears mildly underluminous.
+This could be an effect of its fast rotation, and consequent supersaturation as in \citet{magaudda2022first}, but we lack the rotation periods to confirm this.
 
@@
 
+\begin{figure}
+    \script{FIGURE_r315_comparison.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/r315_prot.png}
+        \caption{
+         R31.5 against rotation period for fully convective M dwarfs.
+        }
+        \label{fig:r315}
+    \end{centering}
+\end{figure}
+
+\subsection{X-ray luminosity}
+\label{sec:discussion:xraylum}
+
 \begin{figure}[ht!]
@@
 
+The $L_X/L_{\rm bol}$ distributions in \citet{cook2014trends, wright2016solartype} grow sparse around $Ro\sim10^{-3}$.
+Using eROSITA data, \citet{magaudda2022first} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend with decreasing $Ro$, pointing toward a supersaturation of the corona at fast rotation speeds, as has been conjectured before~\citep{jeffries2011investigating,ramsay2020tess}.
+\citet{jeffries2011investigating} suggest that the cause for supersaturation is centrifugal stripping of the corona, but \citet{reiners2022magnetism} show that if magnetic field increases, X-ray increases likewise with no sign of a break.
+They conclude that the trend is rather due to the underlying dynamo.
+
+\cite{brown2023coronal}
+
+\cite{johnstone2012soft}
+
+\subsection{Coronal temperature}
+\label{sec:discussion:xraytemp}
+
 \section{Summary and Conclusions}
 
-\section{Acknowledgements}
+\section*{Acknowledgements}
 \citep{lightkurvecollaboration2018lightkurve}
+Based on observations obtained with XMM-Newton, an ESA science mission with instruments and contributions directly funded by ESA Member States and NASA.
 
-\section{Data Availability}
+DLR
+\section*{Data Availability}
 

### 2023-06-15 src/scripts/FIGURE_tess_ffd.py
adjusted ffd figure size

@@ -108,7 +108,7 @@ if __name__ == "__main__":
         ed, freq, counts = ffd.ed_and_freq()
 
         # plot
-        fig, ax = plt.subplots(1,1,figsize=(5,4))
+        fig, ax = plt.subplots(1,1,figsize=(5.5,4.5))
 
         # make label
         label = (fr"$\alpha =$ {ffd.alpha:.1f} (+{ffd.alpha_up_err:.1f} "

### 2023-06-15 src/scripts/FIGURE_lx_lbol.py
fixed rossby

@@ -74,8 +74,8 @@ if __name__ == "__main__":
     
     # TIC 277 data
     df = df[df["label"] == "PN (all)"]
-    plt.errorbar(df.Rossby, df.Lx_Lbol, 
-                xerr=[df.Rossby-df.Rossby_low,df.Rossby_high-df.Rossby], 
+    plt.errorbar(df.Rossby_wright18, df.Lx_Lbol, 
+                xerr=[df.Rossby_wright18-df.Rossby_low_wright18,df.Rossby_high_wright18-df.Rossby_wright18], 
                 yerr=df.e_Lx_Lbol, fmt='o', label="TIC 277", alpha=1, c="k")
     
     # Wright et al. 2011 data

### 2023-06-29 src/scripts/FIGURE_lightcurves.py
updated FFD looks

@@ -26,16 +26,17 @@ if __name__ == "__main__":
     om.rate = om.rate / np.nanmedian(om.rate)
 
     # read in X-ray data
-    xray = pd.read_csv(paths.data / "stacked_xray_lightcurve.csv")
-    xray.time = xray.time / 3600. / 24.
+    xray = pd.read_csv(paths.data / "corrected_merged_epic_lc.csv")
+    xray["TIME"] = xray["TIME"] / 3600. / 24.
 
     # make the figure
 
     fix, ax = plt.subplots(2, 1, sharex=True, figsize=(8, 6),
                         gridspec_kw={'height_ratios': [2, 3],
                                             'wspace':0, 'hspace':0})
-    ax[0].scatter(om.time, om.rate, label="OM", s=2)
-    ax[1].plot(xray.time, xray.normalized_flux + 1., label="PN + MOS1 + MOS2", c="olive")
+    ax[0].scatter(om.time, om.rate, label="OM", s=2, alpha=0.8)
+    ax[1].errorbar(xray["TIME"], xray["RATE"], yerr=xray["ERROR"],
+                   label="PN + MOS1 + MOS2", c="olive", alpha=0.8, lw=1.5)
     
     
     # add inset
@@ -45,20 +46,20 @@ if __name__ == "__main__":
     axins.set_xlim(om.time[l], om.time[r])
     axins.set_ylim(0, 3)
     patch, lines = ax[0].indicate_inset_zoom(axins)
-    # [line.set(visible=False) for line in lines]
-
+   
 
     # layout
     ax[0].set_ylim(0,3)
+    ax[0].axhline(1, c="grey")
 
     for a in ax:
-        a.axhline(1, c="grey")
-        a.set_xlim(om.time.min(), xray.time.max())
+        
+        a.set_xlim(om.time.min(), xray["TIME"].max())
         a.legend(loc=1, frameon=True)
 
 
     ax[0].set_ylabel("normalized flux")
-    ax[1].set_ylabel("normalized flux")
+    ax[1].set_ylabel("background subtracted flux <10 keV [cts/s]")
     ax[1].set_xlabel("time [days]")
 
     plt.tight_layout()

### 2023-06-29 src/scripts/FIGURE_tess_ffd.py

@@ -134,11 +134,11 @@ if __name__ == "__main__":
         ffd.plot_powerlaw(ax, c="olive", label=label)
 
         # layout
-        plt.legend(loc=1, frameon=False)
+        plt.legend(loc=1, frameon=False, fontsize=12)
         plt.xscale("log")
         plt.yscale("log")
-        plt.xlabel(r"$E_{flare}$ [erg]")
-        plt.ylabel(r"flares per day above $E_{flare}$")
+        plt.xlabel(r"$E_{flare}$ [erg]", fontsize=12)
+        plt.ylabel(r"flares per day above $E_{flare}$", fontsize=12)
         plt.tight_layout()
 
         # save to file

### 2023-06-29 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 The smallest stars are crawling with rocky planets.
-The mass of the disk decreases such that giant planet are no longer efficiently formed, leaving room for smaller objects in the orbit.
-Terrestrial planets in the habitable zones of these stars are a factor of XXXX more likely than around higher mass stars.
-However, to be truly habitable in an Earth-like sense, more conditions need to be met than simply the possibility of surface water in its liquid form.
-One of them is the space weather of the host, that is, the high energy radiation and particles that force the planet's atmosphere.
-If too high, the atmosphere may blow off.
-If too energetic, life may not survive.
+The disk mass around the smallest stars decreases such that giant planets are no longer efficiently formed, leaving room for smaller objects in the orbit~(XXXX).
+Terrestrial planets in the habitable zones of these stars are a factor of XXXX more likely than around higher mass stars~\citep{hardegree-ullman2019kepler}.
+However, to be habitable in an Earth-like sense, merely the possibility of surface water in its liquid form does not suffice.
+Space weather of the host, that is, the high energy radiation and particles that force the planet's atmosphere, is another important factor.
+If the energetic particle flux is too high, the atmosphere may blow off.
 If not energetic enough, life may not emerge in the first place.
-So, what is the space weather of planets around these smallest, and most abundant, stars, like?
 
-At the lowest masses, a star's life unfolds much slower than for a Sun.
+At the lowest masses, a star's evolution unfolds much slower than for a Sun.
 While the Sun's magnetic activity, and with it the flares, winds, coronal mass ejections and energetic particle eruption that make up space weather, decline rapidly over a few hundred megayears, a fully convective M dwarf stays highly active for gigayears.
@@
 An intermittent corona is conceivable, similarly XYZ.
-Perhaps, the rapid rotation of these low-mass objects plays a role, as it governs the dynamo process within.
+Perhaps, the rapid rotation of these low-mass objects plays a role, as it governs the dynamo process within, and consequently influences the complex process of flux emergence~\citep{weber2023understanding}.
 
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/tess_ffd.png}
+        \includegraphics[width=\linewidth]{figures/277539431_tess_ffd.png}
         \caption{

### 2023-06-29 src/scripts/FIGURE_data_resid.py
added uncertainties to the FFD and spectra

@@ -156,7 +156,10 @@ if __name__ == "__main__":
 
     # filename
     files = ["pn_onlyflare.txt", "joint_onlyflare.txt", "joint_noflare.txt",
-             "joint_all.txt", 'mos1.txt', 'mos2.txt', 'pn.txt', "pn_noflare.txt"]
+             "joint_all.txt", 'mos1.txt', 'mos2.txt', 'pn.txt', "pn_noflare.txt",
+             "joint_3vapec_fe06.txt", "pn_all_vapec_fe06.txt", 
+             "pn_onlyflare_3vapec_fe06.txt", "joint_vapec_fe06.txt", 
+             "pn_noflare_vapec_fe06.txt", "pn_onlyflare_vapec_fe06.txt"]
 
     # plot
     for file in files:

### 2023-06-29 src/scripts/FIGURE_tess_ffd.py

@@ -21,54 +21,54 @@ from altaipony.ffd import FFD
 
 import matplotlib.pyplot as plt
 
-from astropy.modeling import models
-from astropy import units as u
-from astropy.constants import sigma_sb
+# from astropy.modeling import models
+# from astropy import units as u
+# from astropy.constants import sigma_sb
 
 
-def flare_factor(teff, radius, wav, resp,  tflare=10000):
-    """Calculate the flare energy factor in ergs.
+# def flare_factor(teff, radius, wav, resp,  tflare=10000):
+#     """Calculate the flare energy factor in ergs.
 
-    Parameters
-    ----------
-    teff : float
-        Stellar effective temperature in Kelvin.
-    radius : float
-        Stellar radius in solar radii.
-    wav : array
-        Array of wavelengths in nanometers.
-    resp : array
-        Array of bandpass responses.
-     tflare : float
-        Flare temperature in Kelvin.
+#     Parameters
+#     ----------
+#     teff : float
+#         Stellar effective temperature in Kelvin.
+#     radius : float
+#         Stellar radius in solar radii.
+#     wav : array
+#         Array of wavelengths in nanometers.
+#     resp : array
+#         Array of bandpass responses.
+#      tflare : float
+#         Flare temperature in Kelvin.
     
-    Returns
-    -------
-    factor : float
-        Flare energy factor in ergs/s.
-    """
+#     Returns
+#     -------
+#     factor : float
+#         Flare energy factor in ergs/s.
+#     """
 
-    # blackbody
-    bb = models.BlackBody(temperature=teff * u.K)
+#     # blackbody
+#     bb = models.BlackBody(temperature=teff * u.K)
 
-    # blackbody flux in TESS band
-    bbwavs = bb(wav * u.nm)  * resp
+#     # blackbody flux in TESS band
+#     bbwavs = bb(wav * u.nm)  * resp
 
-    fluxs = np.trapz(bbwavs.value, wav)
+#     fluxs = np.trapz(bbwavs.value, wav)
 
-    # blackbody
-    bb = models.BlackBody(temperature=tflare * u.K)
+#     # blackbody
[…]

### 2023-06-30 src/scripts/FIGURE_r315_comparison.py
adjusted figure, added new fits to table

@@ -89,12 +89,12 @@ if __name__ == "__main__":
     # plot our values and annotate with SpT
     txts = []
     for i, d in ffd_vals.iterrows():
-
-        plt.scatter([d.Prot_days],[d.r315],c="k",s=35,marker="s")
-        plt.scatter([d.Prot_days],[d.r315],c="magenta",s=20,marker="s")
-        
-        txts.append(plt.text( d.Prot_days, d.r315, d.SpT,
-                        fontsize=11,  color="k"))
+        if d.TIC == 277539431:
+            plt.scatter([d.Prot_days],[d.r315],c="k",s=35,marker="s")
+            plt.scatter([d.Prot_days],[d.r315],c="magenta",s=20,marker="s")
+            
+            txts.append(plt.text( d.Prot_days, d.r315, "TIC 277",
+                            fontsize=11,  color="k"))
         
     # legend and layout    
     plt.legend(loc=(1.05,0.3), frameon=False, )

### 2023-06-30 src/scripts/TABLE_specfits.py

@@ -82,13 +82,20 @@ def tex_one_err(val, err):
 if __name__ == "__main__":
 
     # read in the data
-    df = pd.read_csv(paths.data / "apec2_results.csv")
+    df = pd.read_csv(paths.data / "apec_apec_results.csv")
+    df["model"] = "APEC+APEC"
     print(df.shape)
-    dfjoint = pd.read_csv(paths.data / "apec2_joint_results.csv")
-    print(dfjoint.shape)
+    dfjoint0 = pd.read_csv(paths.data / "vapec_vapec_results.csv")
+    dfjoint0["model"] = "VAPEC+VAPEC"
+    dfjoint = pd.read_csv(paths.data / "apec_apec_joint.csv")
+    dfjoint["model"] = "APEC+APEC"
+    dfjoint2 = pd.read_csv(paths.data / "vapec_vapec_joint.csv")
+    dfjoint2["model"] = "VAPEC+VAPEC"
+
+    print(dfjoint.columns)
     # append the joint results
-    df = pd.concat([df, dfjoint], ignore_index=True)
-    print(df.shape)
+    df = pd.concat([df, dfjoint0, dfjoint, dfjoint2], ignore_index=True)
+    print(df.columns)
 
     # values with one error
     cols2 = [("Lx_erg_s","Lx_erg_s_err"),
@@ -136,13 +143,14 @@ if __name__ == "__main__":
     # delete xtra columns
     del df["detector"]
     del df["cut"]
-    del df["Rossby"]
-    del df["Rossby_high"]
-    del df["Rossby_low"]
+
+    for col in df.columns:
+        if "Rossby" in col:
+            del df[col]
 
 
     # convert to LaTeX
-    df = df.sort_values("label")
+    df = df.sort_values("model")
     string = df.to_latex(escape=False,index=False)
 
     # layout

### 2023-06-30 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \subsection{Spectral fitting}
+We fitted a two-temperature VAPEC+VAPEC model to the joint PN and MOS spectra, using solar abundances from GRSA, but adjusting to an Fe/0=.6 ratio, which is more typical of M dwarfs.
+An APEC+APEC model leads to similar results with a XXXX Chi-square value.
 
@@
 
-From the proposal: $F_X=2.7^{-13}$ erg/s/cm$^2$, $L_X=6\cdot10^27$ erg/s (update with new eROSITA data) -- 1 oom above the measured XMM values.
+% From the proposal: $F_X=2.7^{-13}$ erg/s/cm$^2$, $L_X=6\cdot10^27$ erg/s (update with new eROSITA data) -- 1 oom above the measured XMM values.
 \begin{table*}
@@
 
-APEC+APEC model, grsa abundances
-
 \subsection{Flares}

### 2023-07-03 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\subsection{Flux and luminosity}
-
 \subsection{Spectral fitting}
-We fitted a two-temperature VAPEC+VAPEC model to the joint PN and MOS spectra, using solar abundances from GRSA, but adjusting to an Fe/0=.6 ratio, which is more typical of M dwarfs.
-An APEC+APEC model leads to similar results with a XXXX Chi-square value.
+We use XSPEC~\citep{arnaud1996xspec} to fit a two-temperature additive APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/0=.6 ratio, which is more typical of M dwarfs~\citep{wood2018chandra}.
+From the same fit, but to PN data only, we derived the flux and X-ray luminosity in the 0.2-12 keV band.
 
@@
 
-X-ray energy of $8.4\pm0.7\cdot10^{30}\,$erg
+We fitted the soft X-ray light curve using constant flux with a flare, applying the~\citet{davenport2014kepler} optical flare model that captures the fast rise and exponential decay of the flare seen with XMM-Newton.
+From the fit, we extracted the baseline flux against which we measured the equivalent duration $ED$ of the flare
 
+ED EQUATION
+
+We then converted ED to flare energy in the 0.2-12 keV band by multiplying with $L_X$.
+
+Lacking calibration for the white light filter in OM, we used the effective width of the passband to extract the conversion factor for a 10kK flare to obtain the corrected $ED$, which we could multiply with $L_{\rm bol}$ to arrive at the bolometric flare energy.
+$L_{\rm bol}$ was derived using G and J magnitudes, converting J to V mag, then using the bolometric correction from
 \subsubsection{TESS flares}
+
+Similar to the OM flare, we used the TESS response curve to obtain the conversion factor in the redder optical filter of TESS, and convert to bolometric flare energy.
 
@@
 
-% \subsection{Rotational variability}
-
-% \begin{figure*}[ht!]
-%     \script{FIGURE_periodograms_xray_and_optical.py}
-%     \begin{centering}
-%         \includegraphics[width=0.75\linewidth]{figures/periodogram_xmm_stacked.png}
-%         \caption{
-%          MOS and PN light curves stacked Lomb-Scargle periodogram.
-%         }
-%         \label{fig:coherence_hist}
-%     \end{centering}
-% \end{figure*}
-
-% \begin{figure*}[ht!]
-%     \script{FIGURE_periodograms_xray_and_optical.py}
-%     \begin{centering}
-%         \includegraphics[width=0.75\linewidth]{figures/periodogram_om.png}
-%         \caption{
-%          OM light curve Lomb-Scargle periodogram.
-%         }
-%         \label{fig:coherence_hist}
-%     \end{centering}
-% \end{figure*}
-
-% periodograms
-
-% present simple model and upper limits from the data
-
-% how large of a loop would that be? convert relative amplitude to flux, convert flux to emission measure, convert emission measure to loop size?
-
 \subsection{Coronal temperature and luminosity}
@@
 % From the proposal: $F_X=2.7^{-13}$ erg/s/cm$^2$, $L_X=6\cdot10^27$ erg/s (update with new eROSITA data) -- 1 oom above the measured XMM values.
-\begin{table*}
-% \movetableright=-20mm
-\footnotesize
-    \script{TABLE_specfits.py}
-    \caption{XSPEC fits to PN, MOS1/MOS2 and joint data also.}
-    \input{output/specfit.tex}
-        \label{tab:specfit}
-\end{table*}
+
+% \begin{table*}
+% % \movetableright=-20mm
+% \footnotesize
+%     \script{TABLE_specfits.py}
+%     \caption{XSPEC fits to PN, MOS1/MOS2 and joint data also.}
+%     \input{output/specfit.tex}
+%         \label{tab:specfit}
+% \end{table*}
+
+% \begin{figure}
+%     \script{FIGURE_data_resid.py}
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{figures/pn_onlyflare_data_resid.png}
+%         \caption{
+%          Flare only PN Double apec spectrum fit.
+%         }
+%         \label{fig:spec_pn_onlyflare}
+%     \end{centering}
+% \end{figure}
+
+% \begin{figure}
+%     \script{FIGURE_data_resid.py}
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{figures/pn_noflare_data_resid.png}
+%         \caption{
+%          No flare PN Double apec spectrum fit.
+%         }
+%         \label{fig:spec_pn_noflare}
+%     \end{centering}
+% \end{figure}
+
+% \begin{figure}
+%     \script{FIGURE_data_resid.py}
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{figures/pn_data_resid.png}
+%         \caption{
+%          PN Double apec spectrum fit.
+%         }
+%         \label{fig:spec_pn}
+%     \end{centering}
+% \end{figure}
+
+% \begin{figure}
+%     \script{FIGURE_data_resid.py}
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{figures/mos1_data_resid.png}
+%         \caption{
+%          MOS1 Double apec spectrum fit.
+%         }
+%         \label{fig:spec_mos1}
+%     \end{centering}
+% \end{figure}
+
+% \begin{figure}
+%     \script{FIGURE_data_resid.py}
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{figures/mos2_data_resid.png}
+%         \caption{
+%          MOS2 Double apec spectrum fit.
+%         }
+%         \label{fig:spec_mos2}
+%     \end{centering}
+% \end{figure}
+
+% \begin{figure}
+%     \script{FIGURE_data_resid.py}
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{figures/joint_onlyflare_data_resid.png}
+%         \caption{
+%          Flare only joint Double APEC spectrum fit.
+%         }
+%         \label{fig:spec_joint_onlyflare}
+%     \end{centering}
+% \end{figure}
+
+% \begin{figure}
+%     \script{FIGURE_data_resid.py}
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{figures/joint_noflare_data_resid.png}
+%         \caption{
+%          Joint Double APEC spectrum fit with flare excluded.
+%         }
+%         \label{fig:spec_joint_noflare}
+%     \end{centering}
+% \end{figure}
+
+% \begin{figure}
+%     \script{FIGURE_data_resid.py}
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{figures/joint_all_data_resid.png}
+%         \caption{
+%          Joint Double APEC spectrum fit.
+%         }
+%         \label{fig:spec_joint_all}
+%     \end{centering}
+% \end{figure}
 
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/pn_onlyflare_data_resid.png}
+        \includegraphics[width=\linewidth]{figures/joint_vapec_fe06_data_resid.png}
         \caption{
-         Flare only PN Double apec spectrum fit.
-        }
-        \label{fig:spec_pn_onlyflare}
-    \end{centering}
-\end{figure}
-
-\begin{figure}
-    \script{FIGURE_data_resid.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/pn_noflare_data_resid.png}
-        \caption{
-         No flare PN Double apec spectrum fit.
-        }
-        \label{fig:spec_pn_noflare}
-    \end{centering}
-\end{figure}
-
-\begin{figure}
-    \script{FIGURE_data_resid.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/pn_data_resid.png}
-        \caption{
-         PN Double apec spectrum fit.
-        }
-        \label{fig:spec_pn}
-    \end{centering}
-\end{figure}
-
-\begin{figure}
-    \script{FIGURE_data_resid.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/mos1_data_resid.png}
-        \caption{
-         MOS1 Double apec spectrum fit.
-        }
-        \label{fig:spec_mos1}
-    \end{centering}
-\end{figure}
-
-\begin{figure}
-    \script{FIGURE_data_resid.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/mos2_data_resid.png}
-        \caption{
-         MOS2 Double apec spectrum fit.
-        }
-        \label{fig:spec_mos2}
-    \end{centering}
-\end{figure}
-
-\begin{figure}
-    \script{FIGURE_data_resid.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/joint_onlyflare_data_resid.png}
-        \caption{
-         Flare only joint Double APEC spectrum fit.
-        }
-        \label{fig:spec_joint_onlyflare}
-    \end{centering}
-\end{figure}
-
-\begin{figure}
-    \script{FIGURE_data_resid.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/joint_noflare_data_resid.png}
-        \caption{
-         Joint Double APEC spectrum fit with flare excluded.
-        }
-        \label{fig:spec_joint_noflare}
-    \end{centering}
-\end{figure}
-
-\begin{figure}
-    \script{FIGURE_data_resid.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/joint_all_data_resid.png}
-        \caption{
-         Joint Double APEC spectrum fit.
+         Joint Double APEC spectrum fit with Fe/0$=0.6$.
         }
@@
 
-optical and X-ray combined
+X-ray energy of $8.4\pm0.7\cdot10^{30}\,$erg
+OM energy of XXXX
+
+Ratio of OM and EPIC flare compared to \citet{guarcello2019simultaneous} simultaneous K2 and XMM flares
+
+In TESS, we found 8 flares
 
 FFD power law fit converged after 16000 steps on $\alpha = 1.77_{-0.23}^{+0.29}$.
+
+The OM flare energy fits the extension of TESS' flare frequency distributions to lower energies within uncertainties on the power law fit.
+Roughly one flare per day of the order of the observed one or above can therefore be expected in future X-ray observations of TIC 277.
 
@@
 
-The $L_X/L_{\rm bol}$ distributions in \citet{cook2014trends, wright2016solartype} grow sparse around $Ro\sim10^{-3}$.
-Using eROSITA data, \citet{magaudda2022first} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend with decreasing $Ro$, pointing toward a supersaturation of the corona at fast rotation speeds, as has been conjectured before~\citep{jeffries2011investigating,ramsay2020tess}.
+Using eROSITA data, \citet{magaudda2022first} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend with decreasing rotation period, pointing toward a supersaturation of the corona at fast rotation speeds, as has been conjectured before~\citep{jeffries2011investigating,ramsay2020tess}.
 \citet{jeffries2011investigating} suggest that the cause for supersaturation is centrifugal stripping of the corona, but \citet{reiners2022magnetism} show that if magnetic field increases, X-ray increases likewise with no sign of a break.
@@
 \label{sec:discussion:xraytemp}
+
+3 and 13 MK
 

### 2023-07-03 src/scripts/FIGURE_data_resid.py

@@ -39,8 +39,8 @@ def plot_data_resid_3(file):
                                         'wspace':0, 'hspace':0})
     
     for din, rin, c, label, m in zip(d, r, ["olive", "orange", "blue"],
-                                     ["MOS1", "MOS2", "PN"],
-                                     ["o", "s", "x"]):
+                                           ["PN", "MOS1", "MOS2"],
+                                           ["o", "s", "x"]):
 
         data = din.astype(float)
         resid = rin.drop("model [counts/s/keV]", axis=1).astype(float)
@@ -59,21 +59,23 @@ def plot_data_resid_3(file):
         ax[1].errorbar(resid["E [keV]"], resid["flux [counts/s/keV]"],
                     yerr=resid["e_flux"], fmt=m, markersize=6, c=c)
         
+        
       # zero line for residuals
     ax[1].axhline(0,c="k")
 
     # labels
     ax[1].set_xlabel("E [keV]")
-    ax[0].set_ylabel("flux [counts/s/keV]")
+    ax[0].set_ylabel("flux [counts / s / keV]")
     ax[1].set_ylabel("residuals")
     ax[0].set_ylim(0,)
     ax[0].legend(loc=1, frameon=False, fontsize=12)
 
     # x limits
     for a in ax:
-        first = data["E [keV]"].values[0] - data["dE"].values[0]
-        last = data["E [keV]"].values[-1] + data["dE"].values[-1]
-        a.set_xlim(first, last)
+        # first = data["E [keV]"].values[0] - data["dE"].values[0]
+        # last = data["E [keV]"].values[-1] + data["dE"].values[-1]
+        a.set_xlim(0.25, 5)
+
         a.set_xscale("log")
 
 

### 2023-07-07 src/scripts/FIGURE_mcmc_results.py
MCMC results for double VAPEC fit

@@ -1,0 +1,172 @@
+"""
+Python 3.8 - UTF-8
+
+X-ray Loops
+Ekaterina Ilin, 2023
+MIT License
+
+---
+
+This script reads in the chains from the MCMC analysis, makes corner plots, 
+extarct the 16, 50, and 84 percentiles, calculates the weighted mean coronal
+T, and plot diagnostic plots, i.e.,
+
+- T1 vs T2
+- mean weighted T vs. ratio of emission measures
+"""
+
+
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+from astropy.table import Table
+
+import paths
+import corner
+
+if __name__ == "__main__":
+
+    # MCMC RESULTS -------------------------------------------------------------
+
+    # read in the MCMC chains and plot them, extract percentiles and save to file
+
+    subsets = [("full data set", "chain_joint_vapec_feo06.fits","x"),
+            ("no flare", "chain_joint_vapec_feo06_noflare.fits","o"),
+            ("flare only", "chain_joint_vapec_feo06_flareonly.fits","d"),]
+
+    res = dict()
+
+    for subset, fn, symb in subsets:
+
+        # read in the chain
+        df = Table.read(paths.data / fn, format='fits').to_pandas()
+
+        # convert units
+        df["norm__16"] = df["norm__16"] * 1e6 # now in 10^-6
+        df["norm__32"] = df["norm__32"] * 1e6 # now in 10^-6
+        df["kT__1"] = df["kT__1"] * 11.604525 # now in MK
+        df["kT__17"] = df["kT__17"] * 11.604525 # now in MK
+
+        # make a corner plot
+        corner.corner(df[["kT__1", "norm__16", "kT__17", "norm__32"]].values,
+                    labels=[r"$T_1$ [MK]", r"$norm_1 \cdot 10^6$",
+                            r"$T_2$ [MK]",r"$norm_2  \cdot 10^6$"],
+                    quantiles=[0.16, 0.5, 0.84],
+                    show_titles=True,
+                    title_kwargs={"fontsize": 12},)
+                
+        # write to file
+        plt.tight_layout()
+        figname = "corner" + fn[5:-4] + ".png"
+        plt.savefig(paths.figures / figname, dpi=300)
+
+        # caclulate the 0.16, 0.5, 0.84 quantiles and add them to a new dataframe
+        q = np.quantile(df[["kT__1","norm__16","kT__17","norm__32"]].values,
[…]

### 2023-07-07 src/scripts/TABLE_specfits.py

@@ -81,89 +81,59 @@ def tex_one_err(val, err):
 
 if __name__ == "__main__":
 
-    # read in the data
-    df = pd.read_csv(paths.data / "apec_apec_results.csv")
-    df["model"] = "APEC+APEC"
-    print(df.shape)
-    dfjoint0 = pd.read_csv(paths.data / "vapec_vapec_results.csv")
-    dfjoint0["model"] = "VAPEC+VAPEC"
-    dfjoint = pd.read_csv(paths.data / "apec_apec_joint.csv")
-    dfjoint["model"] = "APEC+APEC"
-    dfjoint2 = pd.read_csv(paths.data / "vapec_vapec_joint.csv")
-    dfjoint2["model"] = "VAPEC+VAPEC"
+    df = pd.read_csv(paths.data / "mcmc_results.csv", index_col=0)
 
-    print(dfjoint.columns)
-    # append the joint results
-    df = pd.concat([df, dfjoint0, dfjoint, dfjoint2], ignore_index=True)
-    print(df.columns)
+    # remove extra columns
+    del df["symb"]
+    del df["norm_ratio"]
+    del df["e_norm_ratio"]
 
-    # values with one error
-    cols2 = [("Lx_erg_s","Lx_erg_s_err"),
-            ("Lx_Lbol","e_Lx_Lbol") ]
+    for c in df.columns:
+        if "_err" in c:
+            del df[c]
 
-    # values with up and lower 90% errors
-    cols_mid_low_high = [("flux_erg_s_cm2","flux_erg_s_cm2_low","flux_erg_s_cm2_high"),
-                        ("T_MK_1","T_MK_1_low","T_MK_1_high"),
-                        ("T_MK_5","T_MK_5_low","T_MK_5_high"),]
-    
     # rename columns
-    mapcolnames = {"flux_erg_s_cm2":r"$10^{-14} F_X$ [erg/s/cm$^2$]",
-                    "T_MK_1":r"$T_1$ [MK]",
-                    "T_MK_5":r"$T_2$ [MK]",
-                    "Lx_erg_s":r"$10^{26} L_X$ [erg/s]",
-                    "Lx_Lbol":r"$10^{-4} L_X/L_{bol} $"}
-
-    # convert to new units
-    df["Lx_Lbol"] = 1e4 * df["Lx_Lbol"]
[…]

### 2023-07-07 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-In this work, we follow TIC 277 with XMM-Newton for 36 ks, derive its coronal properties, and investigate if producing a high-latitude flare requires an unusual corona, or if it is rather typical of stars of this spectral type and rotation.
+In this work, we follow TIC 277 with XMM-Newton for 36 ks and use optical monitoring from the Transiting Exoplanet Survey Satellite~(TESS,~\citealt{ricker2015transiting}) to measure its coronal properties, and investigate if producing a high-latitude flare requires an unusual corona, or if it is rather typical of stars of this spectral type and rotation.
 
@@
 \subsection{XMM-Newton}
-XMM-Newton
+XMM-Newton observed TIC 277 for 36 ks, i.e.
+two full rotation periods of the star, on August 5, 2022 (Proposal 090120; PI: E.
+Ilin).
+We used the EPIC instruments aboard XMM-Newton, i.e., MOS1/2 and PN, that cover the soft X-ray band between 0.2-12 keV.
 
-uninterrupted two rotation periods
+\subsubsection{Time series with EPIC and OM}
 
-MOS1/2 and PN, as well as OM data
+ We extracted the events time series from EPIC using XMM SAS version 20 using the \texttt{evselect} task.
+We selected a circular source region with 20 arcsec radius on all detectors, and a circular source-free background region 120 and 90 arcsec in MOS and PN, respectively.
+We used the \texttt{epiclccorr} task in XMM SAS to correct for energy and time dependent loss of events\footnote{\url{https://heasarc.gsfc.nasa.gov/docs/xmm/sas/help/epiclccorr/node3.html}, accessed on July 3, 2023}, and extract a soft X-ray time series with a time binning of 200 s.
+
+Simultaneously, we used the optical monitor (OM) aboard XMM-Newton, to monitor TIC 277 in its white light filter at $10$\,s cadence.
+The OM white light filter is centered on $XXXX\AA$, with an equivalent width of $XXXX\AA$.
+We used the \texttt{omfchain} in XMM SAS to extract the 8 individual light curves.
+
+Fig.~\ref{fig:lightcurves} shows the time series using the combined PN and MOS time series together with the OM light curve.
+
+\begin{figure*}[ht!]
+    \script{FIGURE_lightcurves.py}
+    \begin{centering}
+        \includegraphics[width=0.75\linewidth]{figures/lightcurves.png}
+        \caption{
+         Top panel: Optical Monitoring (OM) light curve. Bottom panel: X-ray (PN, MOS1 and MOS2 combined) light curve. 
+        }
+        \label{fig:lightcurves}
+    \end{centering}
+\end{figure*}
+
+\subsubsection{EPIC Spectra}
+From the EPIC observations, we extracted spectra using \texttt{evselect,epatplot,backscale,rmfgen,arfgen}, and \texttt{grppha}.
 
 \subsection{TESS}
+
+Since 2018, TESS has been supplying publicly available red-optical high-precision time series photometry in an ongoing all-sky survey.
+Each uninterrupted observing Sector provides an approximately 27 day long light curve in a broad $XXXX$-$XXXX\AA$ band pass.
+
+We used 5 Sectors of optical monitoring of TIC 277 in 2-minute cadence by TESS.
+Sector 12 was observed in May 2019, Sectors 37 and 39 in April and June 2021, and Sectors 64 and 65 in April and May 2023.
+
+We used AltaiPony~\citep{ilin2021altaipony} to de-trend the light curves using the custom de-trending function \texttt{custom\_detrending}\footnote{see also the documentation online at \url{https://altaipony.readthedocs.io/en/latest/tutorials/detrend.html}, accessed on July 3, 2023}, detailed in~\citet{ilin2022searching}.
+The procedure includes 3rd order spline fitting of coarse trends, followed by iterative sine fitting to remove rotational modulation, and the two Savitzky-Golay filters with a 6h and 3h window, respectively to remove aperiodic short term trends, while masking outliers above $>2.5\sigma$, as well as the six-fold of consecutive orginally masked data points following the orginal mask, and finally, additional 3 adjacent data point before and after the so expanded mask.
+
+In the de-trended light curves, we searched for flare candidates as at least 3 consecutive $>3\sigma$ outliers above the noise level in the masked light curve, filling in the noise level inside the masked areas with the mean of the noise levels adjecent to the mask.
+To the series of outliers, we kept adding data points to the flare candidate until a data point falls below $XXXX\sigma$.
 
@@
 \subsection{Spectral fitting}
-We use XSPEC~\citep{arnaud1996xspec} to fit a two-temperature additive APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/0=.6 ratio, which is more typical of M dwarfs~\citep{wood2018chandra}.
+We use XSPEC version 12.12.1~\citep{arnaud1996xspec} to fit a two-temperature additive APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/0=.6 ratio, which is more typical of M dwarfs~\citep{wood2018chandra}.
 From the same fit, but to PN data only, we derived the flux and X-ray luminosity in the 0.2-12 keV band.
@@
 
-\begin{figure*}[ht!]
-    \script{FIGURE_lightcurves.py}
-    \begin{centering}
-        \includegraphics[width=0.75\linewidth]{figures/lightcurves.png}
-        \caption{
-         Top panel: Optical Monitoring (OM) light curve. Bottom panel: X-ray (PN, MOS1 and MOS2 combined) light curve. 
-        }
-        \label{fig:lightcurves}
-    \end{centering}
-\end{figure*}
-
 show spectra with fits and residuals
@@
 
-% \begin{table*}
-% % \movetableright=-20mm
-% \footnotesize
-%     \script{TABLE_specfits.py}
-%     \caption{XSPEC fits to PN, MOS1/MOS2 and joint data also.}
-%     \input{output/specfit.tex}
-%         \label{tab:specfit}
-% \end{table*}
+\begin{table*}
+% \movetableright=-20mm
+\footnotesize
+    \script{TABLE_specfits.py}
+    \caption{XSPEC fits to PN, MOS1/MOS2 and joint data also.}
+    \input{output/mcmc_specfit.tex}
+        \label{tab:specfit}
+\end{table*}
 
@@
 
-\begin{figure}
-    \script{FIGURE_r315_comparison.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/r315_prot.png}
-        \caption{
-         R31.5 against rotation period for fully convective M dwarfs.
-        }
-        \label{fig:r315}
-    \end{centering}
-\end{figure}
+% \begin{figure}
+%     \script{FIGURE_r315_comparison.py}
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{figures/r315_prot.png}
+%         \caption{
+%          R31.5 against rotation period for fully convective M dwarfs.
+%         }
+%         \label{fig:r315}
+%     \end{centering}
+% \end{figure}
 
@@
 
-\begin{figure}[ht!]
-    \script{FIGURE_lx_lbol.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/lx_lbol.png}
-        \caption{
-         Rossby number versus $L_x/L_{bol}$.
-        }
-        \label{fig:lxlbol}
-    \end{centering}
-\end{figure}
+% \begin{figure}[ht!]
+%     \script{FIGURE_lx_lbol.py}
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{figures/lx_lbol.png}
+%         \caption{
+%          Rossby number versus $L_x/L_{bol}$.
+%         }
+%         \label{fig:lxlbol}
+%     \end{centering}
+% \end{figure}
 

### 2023-07-07 src/scripts/TABLE_specfits.py
layout

@@ -95,7 +95,7 @@ if __name__ == "__main__":
     # rename columns
     mapcolnames = {'T1': r'$T_{cool}$ [MK]',
                    'T2': r'$T_{hot}$ [MK]',
-                   'norm1': r'$norm_{cool}$',
+                   'norm1': r'$10^6$ norm$_{\rm cool}$',
                    'norm2': r'$norm_{hot}$',
                    'weighted_mean_T': r'$T_{mean}$ [MK]'}
     

### 2023-07-07 src/scripts/FIGURE_mcmc_results.py
update burn-in and values

@@ -31,8 +31,8 @@ if __name__ == "__main__":
     # read in the MCMC chains and plot them, extract percentiles and save to file
 
     subsets = [("full data set", "chain_joint_vapec_feo06.fits","x"),
-            ("no flare", "chain_joint_vapec_feo06_noflare.fits","o"),
-            ("flare only", "chain_joint_vapec_feo06_flareonly.fits","d"),]
+            ("quiescent", "chain_joint_vapec_feo06_noflare.fits","o"),
+            ("flaring", "chain_joint_vapec_feo06_flareonly.fits","d"),]
 
     res = dict()
 
@@ -40,6 +40,10 @@ if __name__ == "__main__":
 
         # read in the chain
         df = Table.read(paths.data / fn, format='fits').to_pandas()
+
+        
+        # discard the first 5000 steps
+        df = df.iloc[5000:]
 
         # convert units
         df["norm__16"] = df["norm__16"] * 1e6 # now in 10^-6

### 2023-07-07 src/scripts/TABLE_specfits.py

@@ -93,11 +93,11 @@ if __name__ == "__main__":
             del df[c]
 
     # rename columns
-    mapcolnames = {'T1': r'$T_{cool}$ [MK]',
-                   'T2': r'$T_{hot}$ [MK]',
+    mapcolnames = {'T1': r'$T_{\rm cool}$ [MK]',
+                   'T2': r'$T_{\rm hot}$ [MK]',
                    'norm1': r'$10^6$ norm$_{\rm cool}$',
-                   'norm2': r'$norm_{hot}$',
-                   'weighted_mean_T': r'$T_{mean}$ [MK]'}
+                   'norm2': r'$10^6$ norm$_{\rm hot}$',
+                   'weighted_mean_T': r'$T_{\rm mean}$ [MK]'}
     
     # convert to LaTeX
     for col in ["T1", "T2", "norm1", "norm2"]:

### 2023-07-07 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 We used AltaiPony~\citep{ilin2021altaipony} to de-trend the light curves using the custom de-trending function \texttt{custom\_detrending}\footnote{see also the documentation online at \url{https://altaipony.readthedocs.io/en/latest/tutorials/detrend.html}, accessed on July 3, 2023}, detailed in~\citet{ilin2022searching}.
-The procedure includes 3rd order spline fitting of coarse trends, followed by iterative sine fitting to remove rotational modulation, and the two Savitzky-Golay filters with a 6h and 3h window, respectively to remove aperiodic short term trends, while masking outliers above $>2.5\sigma$, as well as the six-fold of consecutive orginally masked data points following the orginal mask, and finally, additional 3 adjacent data point before and after the so expanded mask.
+The procedure includes 3rd order spline fitting of coarse trends, followed by iterative sine fitting to remove rotational modulation, and the two Savitzky-Golay filters with a 6h and 3h window, respectively to remove aperiodic short term trends, while masking outliers above $>2.5\sigma$, as well as the six-fold of consecutive orginally masked data points following the original mask, and finally, additional 3 adjacent data point before and after the so expanded mask.
 
@@
 We use XSPEC version 12.12.1~\citep{arnaud1996xspec} to fit a two-temperature additive APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/0=.6 ratio, which is more typical of M dwarfs~\citep{wood2018chandra}.
-From the same fit, but to PN data only, we derived the flux and X-ray luminosity in the 0.2-12 keV band.
+We analyzed both the full data set, and the quiescent and flaring parts separately.
+In the full data set, we used the \texttt{bayes} method in XSPEC to assign constant priors within the model's hard limits, and sample the uncertainties using the Markov-Chain-Monte-Carlo (MCMC) method using \texttt{chain} for a total of 30000 steps, discarding the first 5000 as the burn-in phase.
+From the same fit, but using PN data only, we derived the flux and X-ray luminosity in the 0.2-12 keV band.
+For the quiescent and flaring portions of the observations, we chose the same procedure, but chose the Gaussian prior distributions for the two coronal temperatures using the full data set as ground truth, and assigning the largest of the upper and lower uncertainties as standard deviation.
 
@@
 
-\begin{table*}
+From the XMM-Newton EPIC spectra, we could measure the coronal properties of TIC in both its quiescent, and flaring states~(Table~\ref{tab:specfit}).
+Using the combined quiescent and flaring data, we found a two-temperature VAPEC model to fit better than a one-temperature solution and no worse than a three-temperature solution.
+The resulting fit~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about $3.1\,$MK and $13.4\,$MK, respectively.
+The emission measure (norm) weighted mean temperature was $8\,$MK.
+With these temperatures as ground truth, we also fitted two-temperature VAPEC models to the flaring and quiescent portions of the observations separately.
+The coronal temperatures did not update significantly using either subset, although the quiescent spectrum appeared slightly cooler overall.
+However, the emission measure weighted mean temperature was about $1.5\,$MK cooler in the quiescent spectrum, and $\sim3.3\,$MK hotter in the flaring spectrum.
+This suggests that the flaring corona emits the hotter component, while the cooler component corresponds to the quiescent corona.
+In this interpretation, the presence of a hot component in the quiescent spectrum indicates suggests that TIC 277 is flaring below the detection threshold, consistent with an extrapolation of the flare frequency distribution to lower temperatures~(see the following Section, and Fig.~\ref{fig:ffd}).
+
+From the PN spectra, we measured a soft X-ray flux of $F_X=XXXX$, and a luminosity $L_X=XXXX$ in the $0.2-2\,$keV band. In the broader $0.2-10\,$keV band, we measured XXXX and XXXX.
+ \begin{table}
 % \movetableright=-20mm
-\footnotesize
+\centering
     \script{TABLE_specfits.py}
-    \caption{XSPEC fits to PN, MOS1/MOS2 and joint data also.}
+    \caption{XSPEC fits EPIC spectra for different subsets of observations.}
     \input{output/mcmc_specfit.tex}
         \label{tab:specfit}
-\end{table*}
+\end{table}
 
@@
 
-spectral analysis
-
 \subsection{Flares}

### 2023-07-18 src/scripts/FIGURE_tess_lcs.py
all tess sectors included

@@ -22,35 +22,42 @@ import numpy as np
 if __name__ == "__main__":
 
     # define sectors
-    sectors = 12, 37, 39
+    sectors = 12, 37, 39, 64, 65
 
     # read in light curves
-    lcrs = [Table.read(paths.data / f"tic277_tess_detrended_{s}.fits") for s in sectors]
+    lcrs = [Table.read(paths.data / f"tic277539431_tess_detrended_{s}.fits") for s in sectors]
 
     # make three subplots with two light curves each, one showing the flux the other the
     # detrended flux
-    fig, axes = plt.subplots(3, 1, figsize=(10, 9))
+    fig, axes = plt.subplots(5, 1, figsize=(13, 15))
 
     # loop over axes, light curves and sectors
     for ax, lcr, sector in zip(axes, lcrs, sectors):
 
         # plot the un-detrended and detrended flux
         ax.plot(lcr['TIME'], lcr['FLUX']/np.nanmedian(lcr["FLUX"]), 'k.',
-                label=f"sector {sector}", ms=1.5)
+                ms=1.5)
         ax.plot(lcr['TIME'], lcr['DETRENDED_FLUX']/np.nanmedian(lcr["DETRENDED_FLUX"]) +0.2, 
-                'r.', ms=1.5)
+                'r.', ms=1.5,)
         
         # layout
-        ax.set_ylabel("flux")
-        ax.set_ylabel("normalized flux")
+        ax.set_ylabel("normalized flux", fontsize=13)
         ax.set_ylim(0.9, 1.5)
         ax.set_xlim(lcr["TIME"][0], lcr["TIME"][-1])
 
         # add sector number
         ax.text(0.02, 0.9, f"Sector {sector}", transform=ax.transAxes)
 
+        # create custom legend that shows the flux and detrended flux
+        # increased size of symbols
+        if sector == 12:
+                ax.legend(["flux", "detrended flux"], loc="upper right",
+                          markerscale=15, frameon=False, fontsize=13)
+
+            
+
     # shared x-axis
-    axes[-1].set_xlabel("time [BJD - 2457000]")
+    axes[-1].set_xlabel("time [BJD - 2457000]", fontsize=13)
 
     plt.tight_layout()
     plt.savefig(paths.figures / "tess_lcs.png", dpi=250)

### 2023-07-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \author{Ekaterina Ilin, Katja Poppenh\"ager, Beate Stelzer}
+
+%\affiliation
 
@@
          Parameter & Value  \\\hline
-         Distance & $13.70\pm0.11\,$pc [2] \\
-         $G$ mag & \\
-         $G_{BP}$ mag & \\
-         $G_{BP}$ mag & \\
-         $K_s$ mag & $9.666 \pm 0.024$ [3]\\
-         $J$ mag & $10.630 \pm 0.023$ [3] \\
+         Distance $d$ & $13.70\pm0.11\,$pc [2] \\
+         % $G$ mag & $14.7189 \pm 0.0011$ [4]\\
+         % $G_{BP}$ mag & \\
+         % $G_{BP}$ mag & \\
+         % $K_s$ mag & $9.666 \pm 0.024$ [3]\\
+         % $J$ mag & $10.630 \pm 0.023$ [3] \\
          SpT & M7 [1]\\
-         age & \\
-         Rotation period & $273.618 \pm 0.007\,$min [1]\\
+         $T_{eff}$ & $2680^{60}_{-50}\,$ K [5]\\
+         % age & \\
+         Rotation period $P$ & $273.618 \pm 0.007\,$min [1]\\
          $v\sin i$ & $38.6\pm1.0\,$km/s [1] \\
-         Inclination & $87.0^{+2.0}_{-2.4}\,$ deg [1]\\
-         Radius & $0.145\pm0.004\,R_\odot$ [1]\\\hline
+         Inclination $i$ & $87.0^{+2.0}_{-2.4}\,$ deg [1]\\
+         Radius $R$ & $0.145\pm0.004\,R_\odot$ [1]\\\hline
 
@@
     \newline\footnotesize
-    [1] \citet{ilin2021giant}, [2] \citet{bailer-jones2018estimating}, [3] 2MASS, \citet{skrutskie2006two}
+    [1] \citet{ilin2021giant}, 
+    [2] \citet{bailer-jones2018estimating}, 
+    % [3] 2MASS, \citet{skrutskie2006two}, 
+    % [4] \citet{collaboration2016gaia,gaiacollaboration2023gaia}, 
+    [5] \citet{pecaut2013intrinsic}
     \label{tab:modelparams}
@@
 \subsection{XMM-Newton}
-XMM-Newton observed TIC 277 for 36 ks, i.e.
+\label{sec:obs:xmm}
+XMM-Newton is an X-ray telescope, which launched into orbit 1999, equipped with six science instruments that operate simultaneously, among them the European Photon Imaging Camera with three cameras, PN, MOS1 and MOS2; and the Optical Monitor (OM) that provides near-UV and optical photometry, among other capabilities.
+We used XMM-Newton to observe TIC 277 for 36 ks, i.e.
 two full rotation periods of the star, on August 5, 2022 (Proposal 090120; PI: E.
 Ilin).
-We used the EPIC instruments aboard XMM-Newton, i.e., MOS1/2 and PN, that cover the soft X-ray band between 0.2-12 keV.
+We used the EPIC instruments aboard XMM-Newton, i.e., MOS1/2 and PN, that cover the soft X-ray band between 0.2-12 keV, as well as the OM using its broad band white light filter.
 
@@
 \subsubsection{EPIC Spectra}
-From the EPIC observations, we extracted spectra using \texttt{evselect,epatplot,backscale,rmfgen,arfgen}, and \texttt{grppha}.
+From the EPIC observations, we used standard spectral extraction procedure for point sources.
+We extracted spectra using \texttt{evselect} to select all events in the circular event and background regions,\texttt{epatplot} to check for pile-up of multiple photons in one exposure, \texttt{backscale} to calculate the area of the source region, and \texttt{rmfgen} and \texttt{arfgen} to convolve the observed energies with the instrument response, and produce the spectra.
+Finally, we use \texttt{grppha} to rebin the data in the spectra to obtain at least 15 counts per bin.
+To prduce spectra for the quiescent and flaring portions of the observations separately, we used \texttt{tabgtigen} to select the time intervals to pass to \texttt{evselect}.
 
 \subsection{TESS}
-
+\label{sec:obs:tess}
 Since 2018, TESS has been supplying publicly available red-optical high-precision time series photometry in an ongoing all-sky survey.
@@
 
-\subsection{Spectral fitting}
+\subsection{Spectral fitting: EPIC}
 We use XSPEC version 12.12.1~\citep{arnaud1996xspec} to fit a two-temperature additive APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/0=.6 ratio, which is more typical of M dwarfs~\citep{wood2018chandra}.
@@
 
-\subsection{Flares}
-
-\subsubsection{XMM-Newton flare}
+\subsection{Flare energies: TESS, OM, and EPIC}
 
 We fitted the soft X-ray light curve using constant flux with a flare, applying the~\citet{davenport2014kepler} optical flare model that captures the fast rise and exponential decay of the flare seen with XMM-Newton.
-From the fit, we extracted the baseline flux against which we measured the equivalent duration $ED$ of the flare
+From the fit, we extracted the baseline flux against which we measured the equivalent duration $ED$ of the flare, i.e., the flare flux $F_{\rm flare}$, divided by the median flux $F_0$ of the star, integrated over the flare duration~\citep{gershberg1972results}:
 
-ED EQUATION
+\begin{equation}
+\label{eq:ED}
+ED=\displaystyle \int \mathrm dt\, \frac{F_{\rm flare}(t)}{F_0}.
+\end{equation}
 
-We then converted ED to flare energy in the 0.2-12 keV band by multiplying with $L_X$.
+We then converted $ED$ to flare energy in the $0.2-XXXX\,$keV band by multiplying with $L_X$.
 
-Lacking calibration for the white light filter in OM, we used the effective width of the passband to extract the conversion factor for a 10kK flare to obtain the corrected $ED$, which we could multiply with $L_{\rm bol}$ to arrive at the bolometric flare energy.
-$L_{\rm bol}$ was derived using G and J magnitudes, converting J to V mag, then using the bolometric correction from
-\subsubsection{TESS flares}
+For the OM flare, we derived the bolometric flare energy following the procedure in~\citet{shibayama2013superflares}.
+We used the effective width of $347\,$nm around a mid-wavelength of $406\,$nm of the pass as the response curve $R_{\lambda, \rm OM}$ band to extract a ratio $f_{\rm OM}$ of stellar to flare flux of $1.4\cdot 10^{-4}$ for a $10,000\,$K blackbody flare\footnote{A flare temperature of $10,000\,$K is a typical approximation for energetic M dwarf flares~\citep{kowalski2013timeresolved, howard2020evryflarea}.}:
 
-Similar to the OM flare, we used the TESS response curve to obtain the conversion factor in the redder optical filter of TESS, and convert to bolometric flare energy.
+\begin{equation}
+    f_{\rm OM} = \frac{\displaystyle\int R_{\lambda, \rm OM} B_{\lambda(T_{\rm eff})}  d\lambda}{\displaystyle\int R_{\lambda, \rm OM} B_{\lambda(T_{\rm eff})} d\lambda} 
+    \label{eq:ratio}
+\end{equation}
+
+With this, we can calculate the bolometric flare energy as:
+
+\begin{equation}
+    E_{\rm flare} = ED \cdot \pi R^2 \cdot \sigma_{kB} T_{\rm eff}^4 \cdot f_{\rm OM}
+    \label{eq:eflare}
+\end{equation}
+
+% To obtain the corrected $ED$, we then multiplied with $L_{\rm bol}$ to arrive at the bolometric flare energy.
+$L_{\rm bol}$ was derived using G and J magnitudes by converting J to V magnitude, then using the bolometric correction from \citep{mann2015how,mann2016erratum}.
+
+Analogously to Eq.~\ref{eq:ratio} for the OM flare, we used the TESS response curve to obtain a flux ratio $f_{\rm TESS}$ of $6.3\cdot10^{-3}$ in the redder optical filter of TESS, and convert $ED$ to bolometric flare energy using Eq.~\ref{eq:eflare}.
+Besides the $ED$, \texttt{altaipony} also yields flare start and end times, defined as the difference between the first and last data point above the flare detection criterion defined in Section~\ref{sec:obs:tess}; and relative amplitudes $a$.
 
@@
 The coronal temperatures did not update significantly using either subset, although the quiescent spectrum appeared slightly cooler overall.
-However, the emission measure weighted mean temperature was about $1.5\,$MK cooler in the quiescent spectrum, and $\sim3.3\,$MK hotter in the flaring spectrum.
+However, the emission measure weighted mean temperature was about $1.5\,$MK cooler in the quiescent spectrum, and $\sim3.2\,$MK hotter in the flaring spectrum.
 This suggests that the flaring corona emits the hotter component, while the cooler component corresponds to the quiescent corona.
-In this interpretation, the presence of a hot component in the quiescent spectrum indicates suggests that TIC 277 is flaring below the detection threshold, consistent with an extrapolation of the flare frequency distribution to lower temperatures~(see the following Section, and Fig.~\ref{fig:ffd}).
+In this interpretation, the presence of a hot component in the quiescent spectrum indicates that TIC 277 is flaring below the detection threshold, consistent with an extrapolation of the flare frequency distribution to lower temperatures~(see the following Section, and Fig.~\ref{fig:ffd}).
+
+Not using the full set of observations as ground truth resulted in an unconstrained hot coronal component.
+This can be interpreted as due to a lack of X-ray flux above 5 keV.
+Hot flares with temperatures of >XXXX MK have been measured in other stars.
+On the Sun, the highest flare associated coronal temperatures are around 35 MK~\citep{kay2003soft}.
 
@@
 \label{sec:discussion:xraytemp}
+A third, very hot, $>40\,$MK component could only be poorly constrained due to a lack of sensitivity to flux above $5\,$keV.
+DS Tuc A, a young G dwarf, has been observed flaring with a 74 MK in the flare peaks~\citep{pillitteri2022xraya}, and PMS stars can exceed 100 MK~\citep{getman2008xray, getman2021xray}.
+The Sun can heat the corona up to over 30 MK during flares~\citep{benz2016flare}.
+A flare on AD Leo had a peak temperature of about 37 MK~\citep{stelzer2022great}.
+Other flares are well-described with two temperatures~\citep{behr2023muscles, williams2015simultaneous}.
+We suggest that the poorly constrained third component could stem from the low energy tail of the hard X-ray emission at >10 keV.
 
 3 and 13 MK
+
+The 13 MK component represents thermal bremsstrahlung radiation from hot plasma that evaporates from the chromosphere into the coronal flaring loops~\citep{benz2016flare}, which is supported by the observed delay between the optical and the X-ray flare~\citep{hawley2003multiwavelength}.
+The optical flare originates from the lower chromosphere, which is heated by the precipitating particles.
+This heating then causes evaporation and lights up the flare loop.
+%In reality, the corona emits at a range of temperatures decreasing after the flare peak~\citep[e.g., ][]{stelzer2006simultaneous}. Or is it just that the hot contribution decreases?
 
@@
 \section*{Acknowledgements}
-\citep{lightkurvecollaboration2018lightkurve}
-Based on observations obtained with XMM-Newton, an ESA science mission with instruments and contributions directly funded by ESA Member States and NASA.
-
-DLR
+This work made use of the open source python software packages \texttt{lightkurve}\citep{lightkurvecollaboration2018lightkurve}, astropy~\citep{robitaille2013astropy}, numpy~(XXXX), pandas~\citep{reback2022pandasdev}, altaipony~\citep{ilin2021altaipony}, and XXXX.
+This work has made use of data from the European Space Agency (ESA) mission
+{\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia}
+Data Processing and Analysis Consortium (DPAC,
+\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC
+has been provided by national institutions, in particular the institutions
+participating in the {\it Gaia} Multilateral Agreement.
+This work is based on observations obtained with XMM-Newton, an ESA science mission with instruments and contributions directly funded by ESA Member States and NASA.
+E.I.
+acknowledges funding from the Deutsche Luft- und Raumfahrtgesellschaft (FKZ 50 OR 2209, project \textit{X-ray Loops}).
+K.P.
+acknowledges funding from XXXX.
 \section*{Data Availability}
+All data used in this study are publicly available in their respective archives.
+This study used the reproducibility software \href{https://github.com/showyourwork/showyourwork}{showyourwork}
+\citep{luger2021mappinga} to create all figures, format all tables, and compile the manuscript.
+Each figure links to the dataset stored on Zenodo (link to be added after acceptance), and to the script that produced the figure in the git repository~\href{https://github.com/ekaterinailin/xmm_for_tic277}{github.com/ekaterinailin/xmm\_for\_tic277}.
+The data analysis scripts can be found under \href{https://github.com/ekaterinailin/tic277}{github.com/ekaterinailin/tic277}.
 

### 2023-07-19 src/scripts/FIGURE_tess_ffd.py
add flare table, format stuff in FFDs

@@ -80,7 +80,7 @@ if __name__ == "__main__":
     
     
 
-    for tic, teff, radius in tic_teff_rad:
+    for tic, teff, radius in tic_teff_rad[:1]:
 
         print(tic)
 
@@ -112,6 +112,7 @@ if __name__ == "__main__":
         high_alpa, high_beta = ffd.alpha_up_err + ffd.alpha, ffd.beta_up_err + ffd.beta
         low_alpa, low_beta = ffd.alpha - ffd.alpha_low_err, ffd.beta - ffd.beta_low_err
         print(ffd.beta, high_beta, low_beta)
+        print(ffd.beta_low_err)
         ed, freq, counts = ffd.ed_and_freq()
 
         # plot

### 2023-07-19 src/scripts/TABLE_tess_flares.py

@@ -1,0 +1,53 @@
+"""
+Python 3.8 - UTF-8
+
+X-ray Loops
+Ekaterina Ilin, 2023
+MIT License
+
+---
+
+This script reads in the TESS flare table and produces a LaTeX table.
+"""
+
+import pandas as pd
+import paths
+
+if __name__ == "__main__":
+
+    # read in the flare table
+    df = pd.read_csv(paths.data / "tess_flares.csv")
+
+    # select the columns we want, sort by time
+    sel = df[["tstart", "ampl_rec", 'ed_rec', 'ed_rec_err', 'Sector']].sort_values("tstart")
+
+    # convert to 10^30 erg
+    sel["ed_rec"] = sel["ed_rec"] / 1e30
+    sel["ed_rec_err"] = sel["ed_rec_err"] / 1e30
+
+    # format
+    sel[r"$E_{\rm bol}$ [$10^{30}$ erg]"] = (sel["ed_rec"].round(2).astype(str) + 
+                                            " [" + 
+                                            sel["ed_rec_err"].round(2).astype(str) + 
+                                            "]")
+    # delete unused columns
+    del sel["ed_rec_err"]
+    del sel["ed_rec"]
+
+    # rename columns
+    sel = sel.rename(columns={"tstart": r"$t_{s}$ [BJD - 2457000]", "ampl_rec": r"$a$"})
+
+    # select columns in the correct order
+    sel = sel[[r"$t_{s}$ [BJD - 2457000]", r"$a$", r"$E_{\rm bol}$ [$10^{30}$ erg]", "Sector"]]
+
+    # convert to LaTeX
+    string = sel.to_latex(index=False, escape=False)
+
+    # layout
+    string = string.replace("midrule","hline")
+    string = string.replace("toprule","hline")
+    string = string.replace("bottomrule","hline")
+
+    # write to file
+    with open(paths.output / "tess_flares.tex", "w") as f:
+        f.write(string)

### 2023-07-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \begin{abstract}
-    TIC 277 is currently the star with the highest latitude flare located to date.
+    TIC 277 is the star with the highest latitude flare located to date.
 In 2020, TESS observed this rapidly rotating M7 dwarf produce a flare at 81 deg latitude.
@@
 
-    We obtained 36 ks ob XMM-Newton observations, and determined that the X-ray luminosity $L_X$ is slightly lower but not far below the $L_X$ of other stars with similar spectral types a rotation rates.
+    We studied five Sectors of TESS observations, and obtained 36 ks ob XMM-Newton observations, to find that TIC 277's corona does not differ significantly from other stars with its spectral type and fast rotation period ($P_{\rm rot})$ with respect to X-ray luminosity, coronal temperatures, and flaring activity.
+Additionally, XMM-Newton captured a simultaneous X-ray and optical flare on TIC 277 with a bolometric energy of XXXX and a soft X-ray energy of XXXX in the 0.2-2 keV band.
+Combining the 18 flares found in TESS with the flare from XMM-Newton, we inferred that TIC 277 produces flares at or above this energy approximately once every 10 h.
+We suggest that TIC 277's X-ray and flaring activity is representative of other rapidly rotating late M dwarfs, but that further measurements of localized flares are needed to understand the conditions under which high latitude flares in these stars can occur.
+
+%    and determined that the X-ray luminosity $L_X$ is slightly lower but not far below the $L_X$ of other stars with similar spectral types a rotation rates.
 We constructed flare frequency distributions using the three TESS Sectors, with a power law slope of $XXXX\pm XXXX$, and flare rate above $\log_{10}E=31.5\,$erg of XXXX, again typical of similar stars.
@@
 Simultaneously, we used the optical monitor (OM) aboard XMM-Newton, to monitor TIC 277 in its white light filter at $10$\,s cadence.
-The OM white light filter is centered on $XXXX\AA$, with an equivalent width of $XXXX\AA$.
+The OM white light filter is centered on $4,060\AA$, with an equivalent width of $3,470\AA$.
 We used the \texttt{omfchain} in XMM SAS to extract the 8 individual light curves.
@@
 Since 2018, TESS has been supplying publicly available red-optical high-precision time series photometry in an ongoing all-sky survey.
-Each uninterrupted observing Sector provides an approximately 27 day long light curve in a broad $XXXX$-$XXXX\AA$ band pass.
+Each uninterrupted observing Sector provides an approximately 27 day long light curve in a broad $6,000$-$10,000\AA$ band pass.
 
 We used 5 Sectors of optical monitoring of TIC 277 in 2-minute cadence by TESS.
-Sector 12 was observed in May 2019, Sectors 37 and 39 in April and June 2021, and Sectors 64 and 65 in April and May 2023.
+Sector 12 was observed in May 2019, Sectors 37 and 39 in April and June 2021, and Sectors 64 and 65 in April and May 2023~(Fig.~\ref{fig:tess_lcs}).
 
 We used AltaiPony~\citep{ilin2021altaipony} to de-trend the light curves using the custom de-trending function \texttt{custom\_detrending}\footnote{see also the documentation online at \url{https://altaipony.readthedocs.io/en/latest/tutorials/detrend.html}, accessed on July 3, 2023}, detailed in~\citet{ilin2022searching}.
-The procedure includes 3rd order spline fitting of coarse trends, followed by iterative sine fitting to remove rotational modulation, and the two Savitzky-Golay filters with a 6h and 3h window, respectively to remove aperiodic short term trends, while masking outliers above $>2.5\sigma$, as well as the six-fold of consecutive orginally masked data points following the original mask, and finally, additional 3 adjacent data point before and after the so expanded mask.
+The procedure includes 3rd order spline fitting of coarse trends, followed by iterative sine fitting to remove rotational modulation, and the two Savitzky-Golay filters with a 6h and 3h window, respectively to remove aperiodic short term trends, while masking outliers above $>2.5\sigma$, as well as the six-fold of consecutive originally masked data points following the original mask, and finally, additional 3 adjacent data point before and after the so expanded mask.
 
-In the de-trended light curves, we searched for flare candidates as at least 3 consecutive $>3\sigma$ outliers above the noise level in the masked light curve, filling in the noise level inside the masked areas with the mean of the noise levels adjecent to the mask.
-To the series of outliers, we kept adding data points to the flare candidate until a data point falls below $XXXX\sigma$.
+In the de-trended light curves, we searched for flare candidates as at least 3 consecutive $>3\sigma$ outliers above the noise level in the masked light curve, filling in the noise level inside the masked areas with the mean of the noise levels adjacent to the mask.
+To the series of outliers, we kept adding data points to the flare candidate until a data point falls below $2\sigma$.
+
+\begin{figure*}
+    \script{FIGURE_tess_lcs.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/tess_lcs.png}
+        \caption{
+         Normalized TESS light curves.
+Black and red dots show the light curve with and without rotational variability and trends, respectively.
+The de-trended light curve is offset by $0.2$.
+The large, rotationally modulated, flare, localized at about 81 deg latitude by~\citet{ilin2021giant}, appears in the second half of the top panel~(Sector 12).}
+        \label{fig:tess_lcs}
+    \end{centering}
+\end{figure*}
 
@@
 
-We then converted $ED$ to flare energy in the $0.2-XXXX\,$keV band by multiplying with $L_X$.
+We then converted $ED$ to flare energy in the $0.2-2\,$keV band by multiplying with $L_X$.
 
@@
 
-Figure~\ref{fig:lightcurves} shows the time series in X-ray and optical.
-
-show spectra with fits and residuals
+With the XMM-Newton and TESS observations, we could constrain the coronal and flaring properties of TIC 277.
+We derive its coronal temperature and luminosity from the EPIC spectrum~(Section~\ref{sec:res:XrayTL}).
+We also measure the flare energies in X-ray from the EPIC instruments, and optical from the OM, and compare them to the flare frequency distributions obtained from the 18 detected flares in TESS~(Section~\ref{sec:res:flares}).
 
 \subsection{Coronal temperature and luminosity}
-
+\label{sec:res:XrayTL}
 % From the proposal: $F_X=2.7^{-13}$ erg/s/cm$^2$, $L_X=6\cdot10^27$ erg/s (update with new eROSITA data) -- 1 oom above the measured XMM values.
@@
 This can be interpreted as due to a lack of X-ray flux above 5 keV.
-Hot flares with temperatures of >XXXX MK have been measured in other stars.
+Hot flares with temperatures of >50 MK have been measured on DS Tuc A~\citep{pillitteri2011xray}.
 On the Sun, the highest flare associated coronal temperatures are around 35 MK~\citep{kay2003soft}.
@@
 \subsection{Flares}
+\label{sec:res:flares}
 
-X-ray energy of $8.4\pm0.7\cdot10^{30}\,$erg
-OM energy of XXXX
+\begin{table}
+% \movetableright=-20mm
+\centering
+    \script{TABLE_tess_flares.py}
+    \caption{Flares detected with TESS. $a$ is the relative amplitude, and $E$ is the bolometric flare energy assuming a 10,000 K blackbody emission from the flare.
+    \input{output/tess_flares.tex}
+        \label{tab:flares}
+\end{table}
 
-Ratio of OM and EPIC flare compared to \citet{guarcello2019simultaneous} simultaneous K2 and XMM flares
+In TESS, we found a total of 18 flares in the five Sectors, or equivalently, 125 days of observing at 2-min cadence. We fit the flare frequency distribution (FFD) with a power law of the form
 
-In TESS, we found 8 flares
+\begin{equation}
+    f(E) \mathrm{d} E = \beta \cdot E^{-\alpha} \mathrm{d} E.
+\end{equation}
+following the procedure in~\citep{ilin2021flares}, implemented as \texttt{FFD.fit\_powerlaw} using the \textit{mcmc} argument in \texttt{AltaiPony}~\citep{ilin2021altaipony}, which is using the posterior distribution derived in \citet{wheatland2004bayesian}, and sampling the posterior distribution with the \texttt{emcee} package~\citep{foreman-mackey2013emcee}.
 
-FFD power law fit converged after 16000 steps on $\alpha = 1.77_{-0.23}^{+0.29}$.
+The FFD power law fit converged after 13,500 steps on a slope $\alpha = 1.9_{-0.2}^{+0.2}$, and $\log_{10}\beta=27.7_{-6.0}^{+7.1}\,\mathrm{d}^{-1}$.
+The frequency $R_{31.5}$ of flares per day above $\log_{10} E = 31.5\,\mathrm{erg}$ is about $-1.1$.
+The slope is typical of other flaring stars, regardless of spectral type a rotation period~(XXXX).
+$R_{31.5}$ is typical of flaring M dwarfs, in particular typical of late M dwarfs in the saturated activity regime, where $R_{31.5}$ becomes independent of rotation period~\citep{medina2020flare,murray2022study}.
 
-The OM flare energy fits the extension of TESS' flare frequency distributions to lower energies within uncertainties on the power law fit.
-Roughly one flare per day of the order of the observed one or above can therefore be expected in future X-ray observations of TIC 277.
+For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of $6.5\pm0.7\cdot10^{30}\,$erg in the $0.2-5\,$keV band.
+The energy of the corresponding OM flare is $1.9\pm0.4\cdot 10^{30}\,$erg, which fits the extension of TESS' flare frequency distributions to lower energies within uncertainties on the power law fit~(Fig.~\ref{fig:ffd}).
+About one flare every $10\,$h of the order of the observed one or above can therefore be expected in future X-ray observations of TIC~277, assuming all optical flares above this energy in the OM data correspond to X-ray flares of similar or higher energy to the observed.
+\citet{guarcello2019simultaneous} find moderate correlation between flare energies in Kepler/K2 data and XMM-Newton, but the majority of their sampled energies are above $10^{33}\,$erg both in X-ray and optical.
+\citet{kuznetsov2021stellar} compare primary Kepler mission observations with XMM-Newton to find a similar result.
+
+\begin{figure}
+    \script{FIGURE_tess_ffd.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/277539431_tess_ffd.png}
+        \caption{
+         Cumulative flare frequency distribution and power law fit.
+        }
+        \label{fig:ffd}
+    \end{centering}
+\end{figure}
 
@@
 Is the magnetic activity of the star with the highest known flare latitude known to date is any way unusual?
-The star is rotating rapidly, even for the typically fast rotating late M dwarfs~\citep{medina2022galactic}.
+Our findings indicate that TIC 277's coronal temperature~(Section~\ref{sec:discussion:xraytemp}), X-ray luminosity~(Section~\ref{sec:discussion:xraylum}), and flaring activity~(Section~\ref{sec:discussion:flares}) are typical of stars with its rotation rate and mass.
+TIC 277 is rotating rapidly, even for the typically fast rotating late M dwarfs~\citep{medina2022galactic}.
 This could lead to an unusual manifestation of the stellar dynamo that produces high latitude flares that other dynamo states cannot generate.
 With TESS and XMM-Newton, we can now evaluate if this suspected exceptionality is reflected in the star's magnetic activity and coronal properties.
-We compare its flaring activity, coronal luminosity and temperatures to stars of similar spectral type and rotation rates
+
+\subsection{Coronal temperature}
+\label{sec:discussion:xraytemp}
+
+We find that the quiescent corona of TIC 277 is best described by two components of 3 MK and 12 MK with an emission-measure weighted temperature of about 6.5 MK, which rises to 11 MK during the flare~(Table~\ref{tab:specfit}).
+The hotter 13 MK component represents thermal bremsstrahlung radiation from hot plasma that evaporates from the chromosphere into the coronal flaring loops~\citep{benz2016flare}, which is supported by the observed delay between the optical and the X-ray flare~\citep{hawley2003multiwavelength}.
+The optical flare originates from the lower chromosphere, which is heated by the precipitating particles.
+This heating then causes evaporation of particles into the flare loop.
+
+A third, very hot, $>40\,$MK component could only be poorly constrained due to a lack of sensitivity to flux above $5\,$keV.
+DS Tuc A, a young G dwarf, has been observed flaring with a >50 MK temperature in the flare peaks~\citep{pillitteri2022xraya}, and PMS stars can exceed 100 MK~\citep{getman2008xray, getman2021xray}.
+The Sun can heat the corona up to over 30 MK during flares~\citep{benz2016flare}.
+A flare on AD Leo had a peak temperature of about 37 MK~\citep{stelzer2022great}.
+Other M dwarf flares are well-described with two temperatures~\citep{behr2023muscles, magaudda2022firsta}.
+The corona of NLTT 33370 AB, an M7 binary with a rotation period of about $3.8\,$h (close to TIC 277's $4.56\,$h) is combined of a 3.1 MK and 14 MK component~\citep{williams2015simultaneous}.
+
+\textcolor{red}{the poorly constrained third component could stem from the low energy tail of the hard X-ray emission at >10 keV.}
+
+%In reality, the corona emits at a range of temperatures decreasing after the flare peak~\citep[e.g., ][]{stelzer2006simultaneous}. Or is it just that the hot contribution decreases?
+
+\subsection{X-ray luminosity}
+\label{sec:discussion:xraylum}
+
+Using eROSITA data, \citet{magaudda2022first} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend with decreasing rotation period, pointing toward a supersaturation of the corona at fast rotation speeds, as has been conjectured before~\citep{jeffries2011investigating,ramsay2020tess}.
+\citet{jeffries2011investigating} suggest that the cause for supersaturation is centrifugal stripping of the corona, but \citet{reiners2022magnetism} show that if magnetic field increases, X-ray XXXX increases likewise with no sign of a break.
+They conclude that the trend is rather due to the underlying dynamo.
+
+\cite{brown2023coronal}
+
+\cite{johnstone2012soft}
+
+Compared to the X-ray luminosities of M7 dwarfs among the late M dwarfs observed by eROSITA in \citet{stelzer2022first}, TIC 277 appears mildly underluminous.
+This could be an effect of its fast rotation, and consequent supersaturation as in \citet{magaudda2022first}, but we lack the rotation periods to confirm this.
 
@@
 
-$\log_10(R_{31.5}$, that is, the rate of flares per day with energies above $\log E_{flare}=10^{31.5}\,$erg, is ca. $-1.XXXX$, consistent with other saturated late M dwarfs~\citep{medina2022galactic}.
-
-Compared to the X-ray luminosities of M7 dwarfs among the late M dwarfs observed by eROSITA in \citet{stelzer2022first}, TIC 277 appears mildly underluminous.
-This could be an effect of its fast rotation, and consequent supersaturation as in \citet{magaudda2022first}, but we lack the rotation periods to confirm this.
-
-\begin{figure}
-    \script{FIGURE_tess_ffd.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/277539431_tess_ffd.png}
-        \caption{
-         Flare frequency distribution and power law fit.
-        }
-        \label{fig:ffd}
-    \end{centering}
-\end{figure}
+$\log_10(R_{31.5}$, that is, the rate of flares per day with energies above $\log E_{flare}=10^{31.5}\,$erg, is ca.
+$-1.1$, consistent with other saturated late M dwarfs~\citep{medina2020flare, murray2022study}.
 
@@
 
-\subsection{X-ray luminosity}
-\label{sec:discussion:xraylum}
+\section{Summary and Conclusions}
+We investigated whether TIC 277, a rapidly rotating M7 dwarf with a flare localized at 81 deg latitude, shows exceptional coronal or flaring properties to produce said flare so close to the rotational pole.
+We obtained 36 ks of EPIC and OM observations with XMM-Newton, and studied the five Sectors of red-optical 27-day light curves provided by TESS.
+We found an X-ray luminosity of XXXX, and a mean coronal temperature of XXXX MK, typical of rapidly rotating late M dwarfs.
+The flaring activity is also typical of late M dwarfs in the saturated regime, with a flare rate $f(> log 10 Eflare = 31.5 erg)\approx -1.1$.
 
-% \begin{figure}[ht!]
-%     \script{FIGURE_lx_lbol.py}
-%     \begin{centering}
-%         \includegraphics[width=\linewidth]{figures/lx_lbol.png}
-%         \caption{
-%          Rossby number versus $L_x/L_{bol}$.
-%         }
-%         \label{fig:lxlbol}
-%     \end{centering}
-% \end{figure}
+XMM-Newton also captured a simultaneous X-ray and optical flare with its EPIC and OM instruments, respectively.
+The flare energy and rate of the OM flare are consistent with the flare frequency distribution of the 18 detected TESS flares, indicating that X-ray flares on TIC 277 can be observed with XMM-Newton about once every $10\,$h.
+The bolometric and soft-X-ray flare energies are within an order of magnitude within each other, which places them among other stars with simultaneous X-ray and optical flare energies~\citep{guarcello2019simultaneous, kuznetsov2021stellar}, as well as the Sun~(XXXX).
 
-Using eROSITA data, \citet{magaudda2022first} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend with decreasing rotation period, pointing toward a supersaturation of the corona at fast rotation speeds, as has been conjectured before~\citep{jeffries2011investigating,ramsay2020tess}.
-\citet{jeffries2011investigating} suggest that the cause for supersaturation is centrifugal stripping of the corona, but \citet{reiners2022magnetism} show that if magnetic field increases, X-ray increases likewise with no sign of a break.
-They conclude that the trend is rather due to the underlying dynamo.
-
-\cite{brown2023coronal}
-
-\cite{johnstone2012soft}
-
-\subsection{Coronal temperature}
-\label{sec:discussion:xraytemp}
-A third, very hot, $>40\,$MK component could only be poorly constrained due to a lack of sensitivity to flux above $5\,$keV.
-DS Tuc A, a young G dwarf, has been observed flaring with a 74 MK in the flare peaks~\citep{pillitteri2022xraya}, and PMS stars can exceed 100 MK~\citep{getman2008xray, getman2021xray}.
-The Sun can heat the corona up to over 30 MK during flares~\citep{benz2016flare}.
-A flare on AD Leo had a peak temperature of about 37 MK~\citep{stelzer2022great}.
-Other flares are well-described with two temperatures~\citep{behr2023muscles, williams2015simultaneous}.
-We suggest that the poorly constrained third component could stem from the low energy tail of the hard X-ray emission at >10 keV.
-
-3 and 13 MK
-
-The 13 MK component represents thermal bremsstrahlung radiation from hot plasma that evaporates from the chromosphere into the coronal flaring loops~\citep{benz2016flare}, which is supported by the observed delay between the optical and the X-ray flare~\citep{hawley2003multiwavelength}.
-The optical flare originates from the lower chromosphere, which is heated by the precipitating particles.
-This heating then causes evaporation and lights up the flare loop.
-%In reality, the corona emits at a range of temperatures decreasing after the flare peak~\citep[e.g., ][]{stelzer2006simultaneous}. Or is it just that the hot contribution decreases?
-
-\section{Summary and Conclusions}
+While TIC 277's coronal and flaring properties do not appear uncommon, this does not mean that the corona has no influence on the high latitude occurrence of flares.
+TIC 277 is a fully convective rapid rotator with a rotation period of only $4.56\,$h.
+The detected high-latitude flare in TESS~\citep{ilin2021giant} may hence be a product of high-latitude flux emergence driven by its fast rotation~\citep{weber2016modeling,weber2017suppression}.
+However, the type of magnetic dynamo driven by such fast rotation may produce a peculiarly structured corona~(XXXX), which then appears common within the mass-rotation regime of TIC 277.
+Less than ten flare latitudes on stars other than the Sun are known to date~\citep{wolter2008doppler, ilin2021giant, johnson2021simultaneous}.
+Localization of more flares on the surfaces of a broader range of stars will allow us to discriminate the role of dynamo and flux emergence theories from that of coronal properties in the production of active latitudes.
 

### 2023-07-20 src/scripts/FIGURE_data_resid.py
more layout, added time slice for flare portion, new spectrum

@@ -161,7 +161,8 @@ if __name__ == "__main__":
              "joint_all.txt", 'mos1.txt', 'mos2.txt', 'pn.txt', "pn_noflare.txt",
              "joint_3vapec_fe06.txt", "pn_all_vapec_fe06.txt", 
              "pn_onlyflare_3vapec_fe06.txt", "joint_vapec_fe06.txt", 
-             "pn_noflare_vapec_fe06.txt", "pn_onlyflare_vapec_fe06.txt"]
+             "pn_noflare_vapec_fe06.txt", "pn_onlyflare_vapec_fe06.txt",
+             "joint_chain_fit.txt"]
 
     # plot
     for file in files:

### 2023-07-20 src/scripts/FIGURE_lightcurves.py

@@ -38,6 +38,9 @@ if __name__ == "__main__":
     ax[1].errorbar(xray["TIME"], xray["RATE"], yerr=xray["ERROR"],
                    label="PN + MOS1 + MOS2", c="olive", alpha=0.8, lw=1.5)
     
+    # add vertical filled area
+    ax[1].axvspan(776075500 / 3600. / 24., 776080000 / 3600. / 24.,
+                  color="grey", alpha=0.2)
     
     # add inset
     axins = ax[1].inset_axes([0.3, 0.55, 0.4, 0.5])

### 2023-07-20 src/scripts/FIGURE_r315_comparison.py

@@ -47,7 +47,9 @@ if __name__ == "__main__":
         f315 = beta / (ff.alpha - 1) * (10**31.5)**(-ff.alpha +1)
         r315s.append(np.log10(f315).values[0])
 
+
     ffd_vals["r315"] = r315s
+    print(ffd_vals)    
 
 
     # TRAPPIST-1

### 2023-07-20 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \subsection{Spectral fitting: EPIC}
-We use XSPEC version 12.12.1~\citep{arnaud1996xspec} to fit a two-temperature additive APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/0=.6 ratio, which is more typical of M dwarfs~\citep{wood2018chandra}.
+We use XSPEC version 12.12.1~\citep{arnaud1996xspec} to fit a two-temperature (2T) additive APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/0$=0.6$ ratio, which is more typical of M dwarfs~\citep{wood2018chandra}.
 We analyzed both the full data set, and the quiescent and flaring parts separately.
+Neither subset could be adequately fit with a single temperature component (1T), and a 3T model did not improve the fit compared to the 2T-model either.
 In the full data set, we used the \texttt{bayes} method in XSPEC to assign constant priors within the model's hard limits, and sample the uncertainties using the Markov-Chain-Monte-Carlo (MCMC) method using \texttt{chain} for a total of 30000 steps, discarding the first 5000 as the burn-in phase.
 From the same fit, but using PN data only, we derived the flux and X-ray luminosity in the 0.2-12 keV band.
-For the quiescent and flaring portions of the observations, we chose the same procedure, but chose the Gaussian prior distributions for the two coronal temperatures using the full data set as ground truth, and assigning the largest of the upper and lower uncertainties as standard deviation.
+For the quiescent and flaring portions of the observations, we chose the same procedure, but used Gaussian prior distributions for the two coronal temperatures using the full data set as ground truth, and assigning the largest of the upper and lower uncertainties as standard deviation.
 
@@
 
-From the PN spectra, we measured a soft X-ray flux of $F_X=XXXX$, and a luminosity $L_X=XXXX$ in the $0.2-2\,$keV band. In the broader $0.2-10\,$keV band, we measured XXXX and XXXX.
+From the PN spectra, we measured a soft X-ray flux of $F_X=XXXX$, and a luminosity $L_X=XXXX$ in the $0.2-2\,$keV band. 
  \begin{table}
@@
 
-% \begin{figure}
-%     \script{FIGURE_data_resid.py}
-%     \begin{centering}
-%         \includegraphics[width=\linewidth]{figures/pn_onlyflare_data_resid.png}
-%         \caption{
-%          Flare only PN Double apec spectrum fit.
-%         }
-%         \label{fig:spec_pn_onlyflare}
-%     \end{centering}
-% \end{figure}
-
-% \begin{figure}
-%     \script{FIGURE_data_resid.py}
-%     \begin{centering}
-%         \includegraphics[width=\linewidth]{figures/pn_noflare_data_resid.png}
-%         \caption{
-%          No flare PN Double apec spectrum fit.
-%         }
-%         \label{fig:spec_pn_noflare}
-%     \end{centering}
-% \end{figure}
-
-% \begin{figure}
-%     \script{FIGURE_data_resid.py}
-%     \begin{centering}
-%         \includegraphics[width=\linewidth]{figures/pn_data_resid.png}
-%         \caption{
-%          PN Double apec spectrum fit.
-%         }
-%         \label{fig:spec_pn}
-%     \end{centering}
-% \end{figure}
-
-% \begin{figure}
-%     \script{FIGURE_data_resid.py}
-%     \begin{centering}
-%         \includegraphics[width=\linewidth]{figures/mos1_data_resid.png}
-%         \caption{
-%          MOS1 Double apec spectrum fit.
-%         }
-%         \label{fig:spec_mos1}
-%     \end{centering}
-% \end{figure}
-
-% \begin{figure}
-%     \script{FIGURE_data_resid.py}
-%     \begin{centering}
-%         \includegraphics[width=\linewidth]{figures/mos2_data_resid.png}
-%         \caption{
-%          MOS2 Double apec spectrum fit.
-%         }
-%         \label{fig:spec_mos2}
-%     \end{centering}
-% \end{figure}
-
-% \begin{figure}
-%     \script{FIGURE_data_resid.py}
-%     \begin{centering}
-%         \includegraphics[width=\linewidth]{figures/joint_onlyflare_data_resid.png}
-%         \caption{
-%          Flare only joint Double APEC spectrum fit.
-%         }
-%         \label{fig:spec_joint_onlyflare}
-%     \end{centering}
-% \end{figure}
-
-% \begin{figure}
-%     \script{FIGURE_data_resid.py}
-%     \begin{centering}
-%         \includegraphics[width=\linewidth]{figures/joint_noflare_data_resid.png}
-%         \caption{
-%          Joint Double APEC spectrum fit with flare excluded.
-%         }
-%         \label{fig:spec_joint_noflare}
-%     \end{centering}
-% \end{figure}
-
-% \begin{figure}
-%     \script{FIGURE_data_resid.py}
-%     \begin{centering}
-%         \includegraphics[width=\linewidth]{figures/joint_all_data_resid.png}
-%         \caption{
-%          Joint Double APEC spectrum fit.
-%         }
-%         \label{fig:spec_joint_all}
-%     \end{centering}
-% \end{figure}
-
 \begin{figure}
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/joint_vapec_fe06_data_resid.png}
+        \includegraphics[width=\linewidth]{figures/joint_chain_fit_data_resid.png}
         \caption{
@@
     \script{TABLE_tess_flares.py}
-    \caption{Flares detected with TESS. $a$ is the relative amplitude, and $E$ is the bolometric flare energy assuming a 10,000 K blackbody emission from the flare.
+    \caption{Flares detected with TESS. $a$ is the relative amplitude, and $E_{\rm bol}$ is the bolometric flare energy assuming a 10,000 K blackbody emission from the flare.}
     \input{output/tess_flares.tex}
@@
 A flare on AD Leo had a peak temperature of about 37 MK~\citep{stelzer2022great}.
-Other M dwarf flares are well-described with two temperatures~\citep{behr2023muscles, magaudda2022firsta}.
+Other M dwarf flares are well-described with two temperatures~\citep{robrade2010quiescent, behr2023muscles, magaudda2022firsta}.
 The corona of NLTT 33370 AB, an M7 binary with a rotation period of about $3.8\,$h (close to TIC 277's $4.56\,$h) is combined of a 3.1 MK and 14 MK component~\citep{williams2015simultaneous}.
-
-\textcolor{red}{the poorly constrained third component could stem from the low energy tail of the hard X-ray emission at >10 keV.}
+However, flares on Proxima Cen (M5.5, 85 d rotation~XXXX), also appear with temperatures of 2-3 MK and 10-20 MK~\citep{gudel2004flares,fuhrmeister2022high, howard2022mouse}.
 

### 2023-07-20 src/scripts/FIGURE_data_resid.py
removed unused and added all four figures and two tables to bash script

@@ -157,12 +157,12 @@ def plot_data_resid_1(file):
 if __name__ == "__main__":
 
     # filename
-    files = ["pn_onlyflare.txt", "joint_onlyflare.txt", "joint_noflare.txt",
-             "joint_all.txt", 'mos1.txt', 'mos2.txt', 'pn.txt', "pn_noflare.txt",
-             "joint_3vapec_fe06.txt", "pn_all_vapec_fe06.txt", 
-             "pn_onlyflare_3vapec_fe06.txt", "joint_vapec_fe06.txt", 
-             "pn_noflare_vapec_fe06.txt", "pn_onlyflare_vapec_fe06.txt",
-             "joint_chain_fit.txt"]
+    # ["pn_onlyflare.txt", "joint_onlyflare.txt", "joint_noflare.txt",
+    #          "joint_all.txt", 'mos1.txt', 'mos2.txt', 'pn.txt', "pn_noflare.txt",
+    #          "joint_3vapec_fe06.txt", "pn_all_vapec_fe06.txt", 
+    #          "pn_onlyflare_3vapec_fe06.txt", "joint_vapec_fe06.txt", 
+    #          "pn_noflare_vapec_fe06.txt", "pn_onlyflare_vapec_fe06.txt",
+    files =  ["joint_chain_fit.txt"]
 
     # plot
     for file in files:

### 2023-07-20 src/scripts/FIGURE_lx_lbol.py

@@ -1,95 +1,0 @@
-"""
-Python 3.8 - UTF-8
-
-X-ray Loops
-Ekaterina Ilin, 2023
-MIT License
-
----
-
-This script plots the Lx/Lbol vs Rossby number for TIC 277 and 
-the Wright et al. 2011 sample.
-"""
-
-
-import pandas as pd
-import matplotlib.pyplot as plt
-import paths
-
-from astropy.table import Table
-
-def tau_wright2018(V, Ks, err=False, eV=None, eKs=None):
-    """Convective turnover time from Wright et al. 2018 using
-    Eq. 5 from that paper.
-
-    Parameters
-    ----------
-    V : float
-        The V magnitude of the star.
-    Ks : float
-        The Ks magnitude of the star.
-    err : bool, optional
-        If True, return the error on the Rossby number.
-        The default is False.
-    eV : float, optional
-        The error on the V magnitude of the star.
-        The default is None.
-    eKs : float, optional
-        The error on the Ks magnitude of the star.
-        The default is None.
-
-    Returns 
-    -------
-    tau : float
-        The convective turnover time of the star.
-    tau_err_high : float
-        The upper error on the convective turnover time.
-    tau_err_low : float
-        The lower error on the convective turnover time.
-
-    """
-
-    tau = 0.64 + 0.25 * (V - Ks)
-
-    if err:
-        tau_err_high = 0.74 + 0.33 * (V + eV - Ks + eKs)
-        tau_err_low = 0.54 + 0.17 * (V - eV - Ks - eKs)
-        return 10**tau, 10**tau_err_high, 10**tau_err_low
-    else:
-        return 10**tau
-
-
-
-if __name__ == "__main__":
-
-    # read Wright et al. 2011 data
-    wright2011 = Table.read(paths.data / "wright2011.fit")
-    wright2011["tau"] = tau_wright2018(wright2011["Vmag"], wright2011["Vmag"] - wright2011["V-K"])
-    wright2011["Ro"] = wright2011["Prot"] / wright2011["tau"]   
-
-    # read in TIC 277 data
-    df = pd.read_csv(paths.data / "apec2_results.csv")
-
-    plt.figure(figsize=(6.5,5))
-    
-    # TIC 277 data
-    df = df[df["label"] == "PN (all)"]
[…]

### 2023-07-20 src/scripts/paper_figures_and_tables.sh

@@ -1,0 +1,7 @@
+python FIGURE_data_resid.py
+python FIGURE_lightcurves.py
+python FIGURE_tess_ffd.py
+python FIGURE_tess_lc.py
+
+python TABLE_specfits.py
+python TABLE_tess_flares.py

### 2023-07-20 src/scripts/TABLE_tess_flares.py
bump up energy conversion by one oom

@@ -22,11 +22,11 @@ if __name__ == "__main__":
     sel = df[["tstart", "ampl_rec", 'ed_rec', 'ed_rec_err', 'Sector']].sort_values("tstart")
 
     # convert to 10^30 erg
-    sel["ed_rec"] = sel["ed_rec"] / 1e30
-    sel["ed_rec_err"] = sel["ed_rec_err"] / 1e30
+    sel["ed_rec"] = sel["ed_rec"] / 1e31
+    sel["ed_rec_err"] = sel["ed_rec_err"] / 1e31
 
     # format
-    sel[r"$E_{\rm bol}$ [$10^{30}$ erg]"] = (sel["ed_rec"].round(2).astype(str) + 
+    sel[r"$E_{\rm bol}$ [$10^{31}$ erg]"] = (sel["ed_rec"].round(2).astype(str) + 
                                             " [" + 
                                             sel["ed_rec_err"].round(2).astype(str) + 
                                             "]")
@@ -38,7 +38,7 @@ if __name__ == "__main__":
     sel = sel.rename(columns={"tstart": r"$t_{s}$ [BJD - 2457000]", "ampl_rec": r"$a$"})
 
     # select columns in the correct order
-    sel = sel[[r"$t_{s}$ [BJD - 2457000]", r"$a$", r"$E_{\rm bol}$ [$10^{30}$ erg]", "Sector"]]
+    sel = sel[[r"$t_{s}$ [BJD - 2457000]", r"$a$", r"$E_{\rm bol}$ [$10^{31}$ erg]", "Sector"]]
 
     # convert to LaTeX
     string = sel.to_latex(index=False, escape=False)

### 2023-08-08 src/scripts/VALUES.py
added all values to output

@@ -1,0 +1,164 @@
+"""
+Python 3.8 - UTF-8
+
+X-ray Loops
+Ekaterina Ilin, 2023
+MIT License
+
+---
+
+This script reads in high level results and produces latex values in the output
+folder.
+
+Values covered:
+- FFD slope
+- FFD intercept
+- R31.5
+
+- Lx
+- Fx
+- Lx/Lbol
+
+- T1 and T2
+- quiescient mean T
+- flaring mean T
+
+- flare energy OM and EPIC
+"""
+
+import pandas as pd
+import numpy as np
+import paths
+
+if __name__ == "__main__":
+    
+
+    # FFD alpha beta -----------------------------------------------------------
+
+    df = pd.read_csv(paths.data / "tess_ffd.csv")
+    row = df[df.TIC == 277539431].iloc[0]
+
+    # convert alpha and beta results to latex strings with 
+    # upper and lower uncertainties for beta convert to log10(beta)
+    alpha, beta = row["alpha"], row["beta"]
+    beta_low_err, beta_up_err = row["beta_low_err"], row["beta_up_err"]
+
+    alpha_str = f"${alpha:.2f}_{{-{row['alpha_low_err']:.2f}}}^{{+{row['alpha_up_err']:.2f}}}$"
+    beta_str = f"${np.log10(beta):.2f}_{{-{np.log10(beta_low_err):.2f}}}^{{+{np.log10(beta_up_err):.2f}}}$"
+
+    with open(paths.output / "tess_ffd_alpha.tex", "w") as f:
+        f.write(f"{alpha_str}")
+
+    with open(paths.output / "tess_ffd_beta.tex", "w") as f:
+        f.write(f"{beta_str}")
+
+    # R31.5 --------------------------------------------------------------------
+
+    r315 = beta / (alpha - 1) * (10**31.5)**(-alpha +1)
+    r315 = np.log10(r315)
+
+    with open(paths.output / "R315.tex", "w") as f:
+        f.write(f"{r315:.2f}")
+
+    # Lx -----------------------------------------------------------------------
+
+    df = pd.read_csv(paths.data / "joint_vapec_chain_fits.csv")
+
+    # select full data set
+    row = df[df.subset == "spec"].iloc[0]
+
+    # get Lx and make latex string with a single uncertainty
[…]

### 2023-08-08 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 %\affiliation
+\newcommand{\T1}{\input{output/T1.tex}}
+\newcommand{\T2}{\input{output/T1.tex}}
 
@@
         \caption{
-         Top panel: Optical Monitoring (OM) light curve. Bottom panel: X-ray (PN, MOS1 and MOS2 combined) light curve. 
+         Top panel: Optical Monitoring (OM) light curve.
+Bottom panel: X-ray (PN, MOS1 and MOS2 combined) light curve.
+The grey shaded portion defines the flare-only subset of the observations (see Section~\ref{sec:methods:epic} and Table~\ref{tab:specfit}).
         }
@@
 \subsection{Spectral fitting: EPIC}
+\label{sec:methods:epic}
 We use XSPEC version 12.12.1~\citep{arnaud1996xspec} to fit a two-temperature (2T) additive APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/0$=0.6$ ratio, which is more typical of M dwarfs~\citep{wood2018chandra}.
@@
 Using the combined quiescent and flaring data, we found a two-temperature VAPEC model to fit better than a one-temperature solution and no worse than a three-temperature solution.
-The resulting fit~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about $3.1\,$MK and $13.4\,$MK, respectively.
-The emission measure (norm) weighted mean temperature was $8\,$MK.
+The resulting fit~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about $\T1\,$MK and $\T2\,$MK, respectively.
 With these temperatures as ground truth, we also fitted two-temperature VAPEC models to the flaring and quiescent portions of the observations separately.
@@
     \script{TABLE_specfits.py}
-    \caption{XSPEC fits EPIC spectra for different subsets of observations.}
+    \caption{XSPEC fits to EPIC spectra for different subsets of observations.}
     \input{output/mcmc_specfit.tex}
@@
         \caption{
-         Joint Double APEC spectrum fit with Fe/0$=0.6$.
+         Best-fit joint Double APEC spectrum fit with Fe/0$=0.6$.
         }
@@
 \citet{kuznetsov2021stellar} compare primary Kepler mission observations with XMM-Newton to find a similar result.
+We note that the X-ray flare light curve follows the typical fast-rise exponential decay shape, and is therefore unlikely eclipsed by the star~\citep{johnstone2012soft}.
 
@@
 \label{sec:discussion:xraytemp}
+Low mass stars' coronae are not uniform.
+Magnetic structures create a range of temperatures, starting at about 1 MK by definition and rising up to 30 MK in flares, in the heterogeneous Solar corona~\citep{vaiana1978recent}.
+Stellar point source X-ray observations are a superposition of various temperatures and densities.
+The resolution of different temperature component depends on the apparent brightness and integration time of the spectrum.
+Nearby stars like Proxima Cen can be modeled with a range of temperatures without overfitting~\citep[e.g.,][]{gudel2004flares}, while one can often only capture one dominant temperature in fainter objects~(XXXX).
 
-We find that the quiescent corona of TIC 277 is best described by two components of 3 MK and 12 MK with an emission-measure weighted temperature of about 6.5 MK, which rises to 11 MK during the flare~(Table~\ref{tab:specfit}).
+We find that the quiescent corona of TIC 277 is best described by two components, similar to many other active M dwarfs~\citep{wargelin2008xray, robrade2010quiescent, behr2023muscles, magaudda2022firsta}.
+We measure temperatures of 3 MK and 12 MK with an emission-measure weighted average of about 6.5 MK, which rises to 11 MK during the flare~(Table~\ref{tab:specfit}).
 The hotter 13 MK component represents thermal bremsstrahlung radiation from hot plasma that evaporates from the chromosphere into the coronal flaring loops~\citep{benz2016flare}, which is supported by the observed delay between the optical and the X-ray flare~\citep{hawley2003multiwavelength}.
@@
 This heating then causes evaporation of particles into the flare loop.
+A third, very hot, $>40\,$MK component could only be poorly constrained due to a lack of sensitivity to flux above $5\,$keV.
 
-A third, very hot, $>40\,$MK component could only be poorly constrained due to a lack of sensitivity to flux above $5\,$keV.
-DS Tuc A, a young G dwarf, has been observed flaring with a >50 MK temperature in the flare peaks~\citep{pillitteri2022xraya}, and PMS stars can exceed 100 MK~\citep{getman2008xray, getman2021xray}.
+The range of 6.5-11 MK for the emission measure weighted temperature of TIC 277 places it squarely in the saturated regime with other mid M dwarfs down to spectral type M6~\citep{wright2018stellar, magaudda2020relation, stelzer2022first,robrade2005xray,raassen2003xray,paudel2021simultaneous, foster2020corona}.
+Flares on the slowly rotating Proxima Cen (M5.5, 85 d rotation~XXXX), appear with a range of temperatures at 2-3 MK and 10-20 MK~\citep{gudel2004flares, fuhrmeister2011multiwavelength, fuhrmeister2022high, howard2022mouse} placing it in the transition region between saturated and unsaturated regimes.
+Only at rotation periods beyond 100 days, flaring activity declines to yield coronal temperatures $<2\,$MK~\citep{wright2018stellar, foster2020corona}.
+For a more direct comparison, there are only few late M dwarfs beyond spectral type M6 with a spectrally resolved corona.
+Their intrinsic faintness and gradual decline in coronal emission make them hard to detect~\citep{berger2010, cook2014trends, stelzer2022first}.
+One example is NLTT 33370 AB, an M7 binary with a rotation period of about $3.8\,$h (close to TIC 277's $4.56\,$h and of the same spectral type) is combined of a 3.1 MK and 14 MK component~\citep{williams2015simultaneous}.
+Another is TRAPPIST-1~(M8, $P_{\rm rot} = 1.4\,$d~XXXX).
+Its corona appears cooler, with a 1.74 MK and a 9.6 MK component~\citep{wheatley2017strong}, and an emission measure weighted temperature of 5.36 MK~\citep{brown2023coronal}.
+
+To summarize, considering the relatively low two-temperature resolution of TIC~277's X-ray spectrum and the dearth of X-ray spectra for $>M7$ dwarfs to compare with, it shows a coronal temperature make-up common for mid-to-late M dwarf in the saturated activity regime.
+
+%DS Tuc A, a young G dwarf, has been observed flaring with a >50 MK temperature in the flare peaks~\citep{pillitteri2022xraya}, and PMS stars can exceed 100 MK~\citep{getman2008xray, getman2021xray}.
 The Sun can heat the corona up to over 30 MK during flares~\citep{benz2016flare}.
-A flare on AD Leo had a peak temperature of about 37 MK~\citep{stelzer2022great}.
-Other M dwarf flares are well-described with two temperatures~\citep{robrade2010quiescent, behr2023muscles, magaudda2022firsta}.
-The corona of NLTT 33370 AB, an M7 binary with a rotation period of about $3.8\,$h (close to TIC 277's $4.56\,$h) is combined of a 3.1 MK and 14 MK component~\citep{williams2015simultaneous}.
-However, flares on Proxima Cen (M5.5, 85 d rotation~XXXX), also appear with temperatures of 2-3 MK and 10-20 MK~\citep{gudel2004flares,fuhrmeister2022high, howard2022mouse}.
-
-%In reality, the corona emits at a range of temperatures decreasing after the flare peak~\citep[e.g., ][]{stelzer2006simultaneous}. Or is it just that the hot contribution decreases?
+A flare on AD Leo (M3.5) had a peak temperature of about 37 MK~\citep{stelzer2022great}.
+CC Eri, a K7 dwarf, was described with a 3T model~\citep{crespo-chacon2007xray} with energies about $10^{33}\,$erg and peak temperatures of 30 MK.
 
@@
 \label{sec:discussion:xraylum}
+Similar to its temperature make-up, while TIC~277's coronal luminosity of XXXX is somewhat lower than the $L_X/L_{\rm bol}\sim 10^{-3}$ average for saturated regime, it is not an outlier to other M dwarfs in this activity regime.
 
-Using eROSITA data, \citet{magaudda2022first} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend with decreasing rotation period, pointing toward a supersaturation of the corona at fast rotation speeds, as has been conjectured before~\citep{jeffries2011investigating,ramsay2020tess}.
+Using eROSITA data, \citet{magaudda2022first} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend in $L_X/L_{\rm bol}$ with decreasing rotation period, pointing toward a supersaturation of the corona at fast rotation speeds, as has been conjectured before~\citep{jeffries2011investigating,ramsay2020tess}.
 \citet{jeffries2011investigating} suggest that the cause for supersaturation is centrifugal stripping of the corona, but \citet{reiners2022magnetism} show that if magnetic field increases, X-ray XXXX increases likewise with no sign of a break.
 They conclude that the trend is rather due to the underlying dynamo.
+If TIC~277 is indeed underluminous, these results suggest that its dynamo produces a weaker magnetic field than that of other saturated M dwarfs, which results in a slightly lower X-ray luminosity.
+However, since both the temperature structure and flaring activity of TIC~277 are in line with saturated activity, this conjecture currently lacks evidence from other activity indicators.
+Moreover, late M dwarfs are underrepresented in these studies.
 
-\cite{brown2023coronal}
+\citet{stelzer2022first} investigated late M and L dwarfs with eROSITA to show that $L_X/L_{\rm bol}$ increases from the full convection boundary (M3) towards later spectral types, but begins to decrease again from M7 downward.
+In that study, TIC~277 is at the lower boundary of the $L_X/L_{\rm bol}$ distribution for M7 dwarfs.
+At the same time, strong magnetic fields are found across the board in brown dwarfs~(XXXX).
+While it is not clear whether M7 falls in the stellar regime of~\citet{reiners2022magnetism} or into that latter brown dwarf realm, or has features of both, it appears at the lower bound regardless.
+The small number of supersaturated stars in \citet{magaudda2022first} tentatively suggests that this may be a common phenomenon among mid-to-late M dwarfs when they are quickly rotating.
 
-\cite{johnstone2012soft}
-
-Compared to the X-ray luminosities of M7 dwarfs among the late M dwarfs observed by eROSITA in \citet{stelzer2022first}, TIC 277 appears mildly underluminous.
+%Compared to the X-ray luminosities of M7 dwarfs among the late M dwarfs observed by eROSITA in \citet{stelzer2022first}, TIC 277 appears mildly underluminous.
 This could be an effect of its fast rotation, and consequent supersaturation as in \citet{magaudda2022first}, but we lack the rotation periods to confirm this.
 
-\subsection{Flaring activity}
+\subsection{Flaring activity -- a young TRAPPIST-1}
 \label{sec:discussion:flares}
 
-$\log_10(R_{31.5}$, that is, the rate of flares per day with energies above $\log E_{flare}=10^{31.5}\,$erg, is ca.
+$\log_{10}(R_{31.5}$, that is, the rate of flares per day with energies above $\log E_{flare}=10^{31.5}\,$erg, is ca.
 $-1.1$, consistent with other saturated late M dwarfs~\citep{medina2020flare, murray2022study}.
+It is similar to TRAPPIST-1's flaring rate~(XXXX), which has been investigated for its effects on the seven terrestrial planets in its orbit.
+
+Recent observations of TRAPPIST-1 b~(XXXX) with the James Webb Space Telescope~(XXXX) find the planet mostly stripped of an atmosphere, which the authors attribute to the intense flare driven X-ray radiation and energetic particle exposure at the short orbital periods of the innermost planet.
+For the habitable zone planets TRAPPIST-1 e and f, simulations suggest that flares alter their chemical composition, in particular ozone, nitric oxide and hydroxide, but that the effects would be below the noise floor of JWST~\citep{chen2021persistence}.
+However, given the uncertain age of TRAPPIST-1~(XXXX), the temperate planets e and f may have been entirely stripped of their atmospheres in the past gigayear(s)~\citep{garraffo2017threatening}.
+
+TIC~277's rapid rotation is indicative of a younger age than TRAPPIST, although it has not been attributed to any star cluster so far~(XXXX).
+A correct understanding of whatever the upcoming observations of TRAPPIST-1 e will yield -- detection or non-detection of an atmosphere -- will depend strongly on its activity history.
+Studies of its younger counterparts, late M dwarfs like TIC~277, are required to empirically constrain cumulative effects of atmospheric forcing of planets.
+If, as this work suggests, TIC~277 is typical of stars of its spectral type and rotation rate, it may represent the youth of TRAPPIST-1.
+And if its energetic flares commonly occur at high latitudes, as suggested by~\citet{ilin2021giant}, future models of energetic particle exposure of habitable zone planets may have to include age dependent latitudes of particle eruption to reproduce observations.
 
@@
 \section*{Acknowledgements}
+ E.I.
+acknowledges funding from the Deutsche Luft- und Raumfahrtgesellschaft (FKZ 50 OR 2209, project \textit{X-ray Loops}).
+K.P.
+acknowledges funding from the German \textit{Leibniz-Gemeinschaft} under project number P67/2018.
 This work made use of the open source python software packages \texttt{lightkurve}\citep{lightkurvecollaboration2018lightkurve}, astropy~\citep{robitaille2013astropy}, numpy~(XXXX), pandas~\citep{reback2022pandasdev}, altaipony~\citep{ilin2021altaipony}, and XXXX.
@@
 This work is based on observations obtained with XMM-Newton, an ESA science mission with instruments and contributions directly funded by ESA Member States and NASA.
-E.I.
-acknowledges funding from the Deutsche Luft- und Raumfahrtgesellschaft (FKZ 50 OR 2209, project \textit{X-ray Loops}).
-K.P.
-acknowledges funding from XXXX.
 \section*{Data Availability}

### 2023-08-08 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \begin{document}
+
+\newcommand{\Tcool}{\input{output/T1.tex}}
+\newcommand{\Thot}{\input{output/T2.tex}}
 
@@
 %\affiliation
-\newcommand{\T1}{\input{output/T1.tex}}
-\newcommand{\T2}{\input{output/T1.tex}}
+\author[0000-0002-6299-7542]{Ekaterina Ilin}
+\affiliation{Leibniz Insitute for Astrophysics Potsdam (AIP), An der Sternwarte 16, 14482 Potsdam, Germany}
+
+\author[0000-0003-1231-2194]{Katja Poppenh\"ager}
+\affiliation{Leibniz Insitute for Astrophysics Potsdam (AIP), An der Sternwarte 16, 14482 Potsdam, Germany}
+\affiliation{Institute for Physics and Astronomy, University of Potsdam, Karl-Liebknecht-Strasse 24/25, 14476 Potsdam, Germany}
+
+\author[]{Beate Stelzer}
+\affiliation{Institut für Astronomie & Astrophysik, Eberhard-Karls-Universität Tübingen, Sand 1, 72076 Tübingen, Germany}
+\affiliation{INAF – Osservatorio Astronomico di Palermo, Piazza del Parlamento 1, 90134 Palermo, Italy}
 
@@
 Using the combined quiescent and flaring data, we found a two-temperature VAPEC model to fit better than a one-temperature solution and no worse than a three-temperature solution.
-The resulting fit~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about $\T1\,$MK and $\T2\,$MK, respectively.
+The resulting fit~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about \Tcool and \Thot, respectively.
 With these temperatures as ground truth, we also fitted two-temperature VAPEC models to the flaring and quiescent portions of the observations separately.

### 2023-08-08 src/scripts/VALUES.py
added Lbol and units

@@ -44,7 +44,7 @@ if __name__ == "__main__":
     beta_low_err, beta_up_err = row["beta_low_err"], row["beta_up_err"]
 
     alpha_str = f"${alpha:.2f}_{{-{row['alpha_low_err']:.2f}}}^{{+{row['alpha_up_err']:.2f}}}$"
-    beta_str = f"${np.log10(beta):.2f}_{{-{np.log10(beta_low_err):.2f}}}^{{+{np.log10(beta_up_err):.2f}}}$"
+    beta_str = f"${np.log10(beta):.2f}_{{-{np.log10(beta_low_err):.2f}}}^{{+{np.log10(beta_up_err):.2f}}}" + r"\,\mathrm{d}^{-1}$"
 
     with open(paths.output / "tess_ffd_alpha.tex", "w") as f:
         f.write(f"{alpha_str}")
@@ -58,7 +58,7 @@ if __name__ == "__main__":
     r315 = np.log10(r315)
 
     with open(paths.output / "R315.tex", "w") as f:
-        f.write(f"{r315:.2f}")
+        f.write(f"${r315:.2f}" + r"\,\mathrm{d}^{-1}$")
 
     # Lx -----------------------------------------------------------------------
 
@@ -82,11 +82,16 @@ if __name__ == "__main__":
     with open(paths.output / "epic_flux.tex", "w") as f:
         f.write(f"{flux_str}")
 
-    # L_X / L_bol --------------------------------------------------------------
+    # L_bol and L_X / L_bol --------------------------------------------------------------
 
     ilin2021 = pd.read_csv(paths.data / "ilin2021updated_w_Rossby_Lbol.csv")
     row = ilin2021[ilin2021.TIC == 277539431].iloc[0]
     Lbol, eLbol = row.Lbol_erg_s, row.eLbol_erg_s
+
+    Lbolstr = fr"${Lbol/1e30:.1e} \pm {eLbol/1e30:.0e}" + r" \times 10^{30}\,\rm{erg s}^{-1}$"
+
+    with open(paths.output / "Lbol.tex", "w") as f:
+        f.write(f"{Lbolstr}")
 
     lxlbol, elxlbol = (Lx / Lbol, 
                     Lx / Lbol * np.sqrt(Lxerr**2 / (Lx**2) + eLbol**2 / (Lbol**2)))
[…]

### 2023-08-08 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \newcommand{\Thot}{\input{output/T2.tex}}
+\newcommand{\eom}{\input{output/om_flare.tex}}
+\newcommand{\eepic}{\input{output/epic_flare.tex}}
+\newcommand{\Tqmean}{\input{output/Tqmean.tex}}
+\newcommand{\Tfmean}{\input{output/Tfmean.tex}}
+\newcommand{\FX}{\input{output/epic_flux.tex}}
+\newcommand{\LX}{\input{output/epic_Lx.tex}}
+\newcommand{\Lbol}{\input{output/Lbol.tex}}
+\newcommand{\ffdalpha}{\input{output/tess_ffd_alpha}}
+\newcommand{\ffdbeta}{\input{output/tess_ffd_beta}}
+\newcommand{\ffdr}{\input{output/R315}}
 
@@
 
-\author[]{Beate Stelzer}
-\affiliation{Institut für Astronomie & Astrophysik, Eberhard-Karls-Universität Tübingen, Sand 1, 72076 Tübingen, Germany}
+\author{Beate Stelzer}
+\affiliation{Institut für Astronomie \& Astrophysik, Eberhard-Karls-Universität Tübingen, Sand 1, 72076 Tübingen, Germany}
 \affiliation{INAF – Osservatorio Astronomico di Palermo, Piazza del Parlamento 1, 90134 Palermo, Italy}
@@
 
-    We studied five Sectors of TESS observations, and obtained 36 ks ob XMM-Newton observations, to find that TIC 277's corona does not differ significantly from other stars with its spectral type and fast rotation period ($P_{\rm rot})$ with respect to X-ray luminosity, coronal temperatures, and flaring activity.
-Additionally, XMM-Newton captured a simultaneous X-ray and optical flare on TIC 277 with a bolometric energy of XXXX and a soft X-ray energy of XXXX in the 0.2-2 keV band.
+    We studied five Sectors of TESS observations, and obtained 36 ks ob XMM-Newton observations, to find that TIC 277's corona does not differ significantly from other low mass stars on the saturated activity branch with respect to X-ray luminosity, coronal temperatures, and flaring activity.
+Additionally, XMM-Newton captured a simultaneous X-ray and optical flare on TIC 277 with a bolometric energy of \eom and a soft X-ray energy of \eepic in the 0.2-2 keV band.
 Combining the 18 flares found in TESS with the flare from XMM-Newton, we inferred that TIC 277 produces flares at or above this energy approximately once every 10 h.
@@
          Inclination $i$ & $87.0^{+2.0}_{-2.4}\,$ deg [1]\\
-         Radius $R$ & $0.145\pm0.004\,R_\odot$ [1]\\\hline
+         Radius $R$ & $0.145\pm0.004\,R_\odot$ [1]\\
+         $L_{\rm bol}$ & \Lbol \\\hline
 
@@
 The coronal temperatures did not update significantly using either subset, although the quiescent spectrum appeared slightly cooler overall.
-However, the emission measure weighted mean temperature was about $1.5\,$MK cooler in the quiescent spectrum, and $\sim3.2\,$MK hotter in the flaring spectrum.
+However, the emission measure weighted mean temperature change dfrom \Tqmean in the quiescent spectrum to \Tfmean in the flaring spectrum.
 This suggests that the flaring corona emits the hotter component, while the cooler component corresponds to the quiescent corona.
@@
 
-From the PN spectra, we measured a soft X-ray flux of $F_X=XXXX$, and a luminosity $L_X=XXXX$ in the $0.2-2\,$keV band. 
+From the PN spectra, we measured a soft X-ray flux of \FX, and a luminosity \LX in the $0.2-2\,$keV band. 
  \begin{table}
@@
 
-The FFD power law fit converged after 13,500 steps on a slope $\alpha = 1.9_{-0.2}^{+0.2}$, and $\log_{10}\beta=27.7_{-6.0}^{+7.1}\,\mathrm{d}^{-1}$.
-The frequency $R_{31.5}$ of flares per day above $\log_{10} E = 31.5\,\mathrm{erg}$ is about $-1.1$.
-The slope is typical of other flaring stars, regardless of spectral type a rotation period~(XXXX).
+The FFD power law fit converged after 13,500 steps on a slope $\alpha =$\ffdalpha.
+The frequency $R_{31.5}$ of flares per day above $\log_{10} E = 31.5\,\mathrm{erg}$ is about $\log_{10}R_{31.5}=$\ffdr.
+The slope is typical of other flaring stars, regardless of spectral type a rotation period~(see, e.g., Fig.~13 in \citealt{ilin2021flares}).
 $R_{31.5}$ is typical of flaring M dwarfs, in particular typical of late M dwarfs in the saturated activity regime, where $R_{31.5}$ becomes independent of rotation period~\citep{medina2020flare,murray2022study}.
 
-For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of $6.5\pm0.7\cdot10^{30}\,$erg in the $0.2-5\,$keV band.
-The energy of the corresponding OM flare is $1.9\pm0.4\cdot 10^{30}\,$erg, which fits the extension of TESS' flare frequency distributions to lower energies within uncertainties on the power law fit~(Fig.~\ref{fig:ffd}).
+For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of \eepic in the $0.2-5\,$keV band.
+The energy of the corresponding OM flare is \eom, which fits the extension of TESS' flare frequency distributions to lower energies within uncertainties on the power law fit~(Fig.~\ref{fig:ffd}).
 About one flare every $10\,$h of the order of the observed one or above can therefore be expected in future X-ray observations of TIC~277, assuming all optical flares above this energy in the OM data correspond to X-ray flares of similar or higher energy to the observed.
@@
 \label{sec:discussion:xraylum}
-Similar to its temperature make-up, while TIC~277's coronal luminosity of XXXX is somewhat lower than the $L_X/L_{\rm bol}\sim 10^{-3}$ average for saturated regime, it is not an outlier to other M dwarfs in this activity regime.
+Similar to its temperature make-up, while TIC~277's coronal luminosity relative to bolometric luminosity of $L_X/L_{\rm bol}=\LXLbol$ is somewhat lower than the $L_X/L_{\rm bol}\sim 10^{-3}$ average in the saturated regime~\citep{wright2018stellar}, it is not an outlier to other M dwarfs in this activity regime.
 
-Using eROSITA data, \citet{magaudda2022first} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend in $L_X/L_{\rm bol}$ with decreasing rotation period, pointing toward a supersaturation of the corona at fast rotation speeds, as has been conjectured before~\citep{jeffries2011investigating,ramsay2020tess}.
+In the Hyades and Praesepe, \citet{nunez2022factory} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend in $L_X/L_{\rm bol}$ with decreasing rotation period for M3.5-M7 dwarfs, tentatively pointing toward a supersaturation of the corona at fast rotation speeds, as has been conjectured before~\citep{jeffries2011investigating,ramsay2020tess}.
 \citet{jeffries2011investigating} suggest that the cause for supersaturation is centrifugal stripping of the corona, but \citet{reiners2022magnetism} show that if magnetic field increases, X-ray XXXX increases likewise with no sign of a break.
@@
 While it is not clear whether M7 falls in the stellar regime of~\citet{reiners2022magnetism} or into that latter brown dwarf realm, or has features of both, it appears at the lower bound regardless.
-The small number of supersaturated stars in \citet{magaudda2022first} tentatively suggests that this may be a common phenomenon among mid-to-late M dwarfs when they are quickly rotating.
+The small number of supersaturated stars in \citet{nunez2022factory}  can only tentatively suggest that this is a common phenomenon among mid-to-late M dwarfs when they are quickly rotating.
 

### 2023-08-08 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \newcommand{\Lbol}{\input{output/Lbol.tex}}
+\newcommand{\LXLbol}{\input{output/lxlbol}}
 \newcommand{\ffdalpha}{\input{output/tess_ffd_alpha}}
@@
 \label{sec:discussion:xraylum}
-Similar to its temperature make-up, while TIC~277's coronal luminosity relative to bolometric luminosity of $L_X/L_{\rm bol}=\LXLbol$ is somewhat lower than the $L_X/L_{\rm bol}\sim 10^{-3}$ average in the saturated regime~\citep{wright2018stellar}, it is not an outlier to other M dwarfs in this activity regime.
+Similar to its temperature make-up, while TIC~277's coronal luminosity relative to bolometric luminosity of $L_X/L_{\rm bol}=$\LXLbol is somewhat lower than the $L_X/L_{\rm bol}\sim 10^{-3}$ average in the saturated regime~\citep{wright2018stellar}, it is not an outlier to other M dwarfs in this activity regime.
 

### 2023-08-08 src/scripts/VALUES.py
add Lbol

@@ -88,7 +88,7 @@ if __name__ == "__main__":
     row = ilin2021[ilin2021.TIC == 277539431].iloc[0]
     Lbol, eLbol = row.Lbol_erg_s, row.eLbol_erg_s
 
-    Lbolstr = fr"${Lbol/1e30:.1e} \pm {eLbol/1e30:.0e}" + r" \times 10^{30}\,\rm{erg s}^{-1}$"
+    Lbolstr = fr"${Lbol/1e30:.1f} \pm {eLbol/1e30:.1f}" + r" \times 10^{30}\,\rm{erg s}^{-1}$"
 
     with open(paths.output / "Lbol.tex", "w") as f:
         f.write(f"{Lbolstr}")

### 2023-08-08 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \label{sec:discussion:xraylum}
-Similar to its temperature make-up, while TIC~277's coronal luminosity relative to bolometric luminosity of $L_X/L_{\rm bol}=$\LXLbol is somewhat lower than the $L_X/L_{\rm bol}\sim 10^{-3}$ average in the saturated regime~\citep{wright2018stellar}, it is not an outlier to other M dwarfs in this activity regime.
+Similar to its temperature make-up, while TIC~277's coronal luminosity relative to bolometric luminosity of $L_X/L_{\rm bol}=$\LXLbol is somewhat lower than the $L_X/L_{\rm bol}\sim 10^{-3}$ average in the saturated regime~\citep{wright2011stellaractivityrotation,wright2018stellar}, it is not an outlier to other M dwarfs in this activity regime.
 
 In the Hyades and Praesepe, \citet{nunez2022factory} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend in $L_X/L_{\rm bol}$ with decreasing rotation period for M3.5-M7 dwarfs, tentatively pointing toward a supersaturation of the corona at fast rotation speeds, as has been conjectured before~\citep{jeffries2011investigating,ramsay2020tess}.
-\citet{jeffries2011investigating} suggest that the cause for supersaturation is centrifugal stripping of the corona, but \citet{reiners2022magnetism} show that if magnetic field increases, X-ray XXXX increases likewise with no sign of a break.
+\citet{jeffries2011investigating} suggest that the cause for supersaturation is centrifugal stripping of the corona, but \citet{reiners2022magnetism} show that if the averae surface magnetic field strength increases, $L_X/L_{\rm bol}$ increases likewise with no sign of a break, up to $10^{-3}$.
 They conclude that the trend is rather due to the underlying dynamo.
 If TIC~277 is indeed underluminous, these results suggest that its dynamo produces a weaker magnetic field than that of other saturated M dwarfs, which results in a slightly lower X-ray luminosity.
-However, since both the temperature structure and flaring activity of TIC~277 are in line with saturated activity, this conjecture currently lacks evidence from other activity indicators.
+However, since the flaring activity of TIC~277 is in line with saturated activity, this conjecture currently lacks evidence from other activity indicators~\citep[see][, for supersaturation in flaring activity in fully convective M dwarfs]{ramsay2020tess}.
 Moreover, late M dwarfs are underrepresented in these studies.

### 2023-08-08 src/scripts/paper_figures_and_tables.sh
added the values script and fixed a typo

@@ -1,7 +1,9 @@
 python FIGURE_data_resid.py
 python FIGURE_lightcurves.py
 python FIGURE_tess_ffd.py
-python FIGURE_tess_lc.py
+python FIGURE_tess_lcs.py
 
 python TABLE_specfits.py
 python TABLE_tess_flares.py
+
+python VALUES.py

### 2023-08-08 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \begin{table}
-
+% \footnotesize
+\centering
     \caption{TIC 277539431 (WISEA J105515.71-735611.3)}
@@
          % $J$ mag & $10.630 \pm 0.023$ [3] \\
-         SpT & M7 [1]\\
-         $T_{eff}$ & $2680^{60}_{-50}\,$ K [5]\\
+         Spectral type & M7 [1]\\
+         Effective temperature $T_{\rm eff}$ & $2680^{60}_{-50}\,$ K [3]\\
          % age & \\
-         Rotation period $P$ & $273.618 \pm 0.007\,$min [1]\\
-         $v\sin i$ & $38.6\pm1.0\,$km/s [1] \\
+         Rotation period $P_{\rm rot}$ & $273.618 \pm 0.007\,$min [1]\\
+         Projected rotation velocity $v\sin i$ & $38.6\pm1.0\,$km/s [1] \\
          Inclination $i$ & $87.0^{+2.0}_{-2.4}\,$ deg [1]\\
          Radius $R$ & $0.145\pm0.004\,R_\odot$ [1]\\
-         $L_{\rm bol}$ & \Lbol \\\hline
+         Bol. luminosity $L_{\rm bol}$ & \Lbol \\\hline
 
@@
     % [4] \citet{collaboration2016gaia,gaiacollaboration2023gaia}, 
-    [5] \citet{pecaut2013intrinsic}
+    [3] \citet{pecaut2013intrinsic}
     \label{tab:modelparams}
@@
 For the OM flare, we derived the bolometric flare energy following the procedure in~\citet{shibayama2013superflares}.
-We used the effective width of $347\,$nm around a mid-wavelength of $406\,$nm of the pass as the response curve $R_{\lambda, \rm OM}$ band to extract a ratio $f_{\rm OM}$ of stellar to flare flux of $1.4\cdot 10^{-4}$ for a $10,000\,$K blackbody flare\footnote{A flare temperature of $10,000\,$K is a typical approximation for energetic M dwarf flares~\citep{kowalski2013timeresolved, howard2020evryflarea}.}:
+We used the throughput curve for the white light filter scaled to unity at the peak of transmission as an optimistic response curve $R_{\lambda, \rm OM}$, given that the degradation of this filter's response is poorly constrained.
+From that, we extracted a ratio $f_{\rm OM}$ of stellar to flare flux of $1.4\cdot 10^{-4}$ for a $10,000\,$K blackbody flare\footnote{A flare temperature of $10,000\,$K is a typical approximation for energetic M dwarf flares~\citep{kowalski2013timeresolved, howard2020evryflarea}.}:
 
@@
 \end{equation}
+
+We note that due to the optimistic response curve assumption the resulting $E_{\rm flare}$ is likely higher in reality.
 

### 2023-08-09 src/scripts/VALUES.py
improved lxlbol

@@ -65,7 +65,7 @@ if __name__ == "__main__":
     df = pd.read_csv(paths.data / "joint_vapec_chain_fits.csv")
 
     # select full data set
-    row = df[df.subset == "spec"].iloc[0]
+    row = df[df.subset == "noflare"].iloc[0]
 
     # get Lx and make latex string with a single uncertainty
     Lx, Lxerr = row.Lx_erg_s, row.Lx_erg_s_err
@@ -96,9 +96,10 @@ if __name__ == "__main__":
     lxlbol, elxlbol = (Lx / Lbol, 
                     Lx / Lbol * np.sqrt(Lxerr**2 / (Lx**2) + eLbol**2 / (Lbol**2)))
 
-    lxlbolstr = fr"${lxlbol:.1e}X\pm {elxlbol:.0e}$ "
-    lxlbolstr = lxlbolstr.replace("e-0","\\times 10^{-")
-    lxlbolstr = lxlbolstr.replace(r"$ ","}$").replace(r"X","}")
+
+    lxlbolstr = fr"${lxlbol*1e4:.1f}\pm {elxlbol*1e4:.1f}" + r"\times 10^{-4}$ "
+
+    print(lxlbolstr)
 
     with open(paths.output / "lxlbol.tex", "w") as f:
         f.write(rf"{lxlbolstr}")

### 2023-08-09 src/scripts/VALUES.py
futher improvement to formatting

@@ -57,8 +57,12 @@ if __name__ == "__main__":
     r315 = beta / (alpha - 1) * (10**31.5)**(-alpha +1)
     r315 = np.log10(r315)
 
+    r315str = f"${r315:.2f}" + r"\,\mathrm{d}^{-1}$"
+
+    print(r315str)
+
     with open(paths.output / "R315.tex", "w") as f:
-        f.write(f"${r315:.2f}" + r"\,\mathrm{d}^{-1}$")
+        f.write(r315str)
 
     # Lx -----------------------------------------------------------------------
 
@@ -69,7 +73,9 @@ if __name__ == "__main__":
 
     # get Lx and make latex string with a single uncertainty
     Lx, Lxerr = row.Lx_erg_s, row.Lx_erg_s_err
-    Lx_str = fr"${Lx:.2f} \pm {Lxerr:.2f}\,\rm erg/\rm s$"
+    Lx_str = fr"${Lx/1e26:.2f} \pm {Lxerr/1e26:.2f}" + r"\times 10^{26}\,\rm{erg s}^{-1}$"
+
+    print(Lx_str)
 
     with open(paths.output / "epic_Lx.tex", "w") as f:
         f.write(f"{Lx_str}")
@@ -77,7 +83,9 @@ if __name__ == "__main__":
     # Fx -----------------------------------------------------------------------
 
     flux, fluxerr = row.flux_erg_s_cm2, row.flux_erg_s_cm2_err
-    flux_str = f"${flux:.2f} \pm {fluxerr:.2f}\,\rm erg/\rm cm^2/\rm s$"
+    flux_str = fr"${flux*1e14:.1f} \pm {fluxerr*1e14:.1f}" + r"\times 10^{-14} \,\rm{erg}\, \rm{cm}^{-2} \rm{s}^{-1}$"
+
+    print(flux_str)
 
     with open(paths.output / "epic_flux.tex", "w") as f:
         f.write(f"{flux_str}")

### 2023-08-09 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Title
-\title{Does a fully convective star need a special corona to produce a polar flare?}
+\title{Fully convective star that produced a near-polar flare shows a typical corona}
 
-% Author list
-\author{Ekaterina Ilin, Katja Poppenh\"ager, Beate Stelzer}
+% Author list with  \affiliation
 
-%\affiliation
 \author[0000-0002-6299-7542]{Ekaterina Ilin}
@@
 \begin{abstract}
-    TIC 277 is the star with the highest latitude flare located to date.
-In 2020, TESS observed this rapidly rotating M7 dwarf produce a flare at 81 deg latitude.
+    In 2020, TESS observed a rapidly rotating M7 dwarf produce a flare at 81 deg latitude, the highest latitude flare located to date.
 This is in stark contrast to solar flares that occur much closer to the equator, typically below 30 deg.
-Does a star require a special corona to produce polar flares?
-
-    We studied five Sectors of TESS observations, and obtained 36 ks ob XMM-Newton observations, to find that TIC 277's corona does not differ significantly from other low mass stars on the saturated activity branch with respect to X-ray luminosity, coronal temperatures, and flaring activity.
+The circumstances that allow flares at high latitudes to occur are poorly understood.
+We studied five Sectors of TESS observations, and obtained 36 ks of XMM-Newton observations, to find that TIC 277's corona does not differ significantly from other low mass stars on the saturated activity branch with respect to X-ray luminosity, coronal temperatures, and flaring activity.
 Additionally, XMM-Newton captured a simultaneous X-ray and optical flare on TIC 277 with a bolometric energy of \eom and a soft X-ray energy of \eepic in the 0.2-2 keV band.
 Combining the 18 flares found in TESS with the flare from XMM-Newton, we inferred that TIC 277 produces flares at or above this energy approximately once every 10 h.
-We suggest that TIC 277's X-ray and flaring activity is representative of other rapidly rotating late M dwarfs, but that further measurements of localized flares are needed to understand the conditions under which high latitude flares in these stars can occur.
+We conclude that TIC 277's X-ray and flaring activity is representative of other rapidly rotating late M dwarfs, but that further measurements of localized flares are needed to understand the conditions under which high latitude flares in these stars can occur.
 
@@
 \label{sec:intro}
-The smallest stars are crawling with rocky planets.
-The disk mass around the smallest stars decreases such that giant planets are no longer efficiently formed, leaving room for smaller objects in the orbit~(XXXX).
-Terrestrial planets in the habitable zones of these stars are a factor of XXXX more likely than around higher mass stars~\citep{hardegree-ullman2019kepler}.
-However, to be habitable in an Earth-like sense, merely the possibility of surface water in its liquid form does not suffice.
-Space weather of the host, that is, the high energy radiation and particles that force the planet's atmosphere, is another important factor.
-If the energetic particle flux is too high, the atmosphere may blow off.
-If not energetic enough, life may not emerge in the first place.
+Most M dwarfs hosts terrestrial planets with a sizable fraction orbiting at instellations where liquid water can exist~\citep{dressing2015occurrence, hardegree-ullman2019kepler, ment2023occurrence}.
+However, to be habitable in an Earth-like manner, merely the possibility of surface water in its liquid form does not suffice.
+Space weather of the host, that is, the high energy radiation and particles that force the planet's atmosphere affects hospitability for life, for M dwarfs in particular~\citep{airapetian2020impact}.
+If the energetic (particle) flux is too high, the atmosphere may blow off~(XXXX).
+If not energetic enough, life may not emerge in the first place~\citep{rimmer2018origin}.
 
-At the lowest masses, a star's evolution unfolds much slower than for a Sun.
-While the Sun's magnetic activity, and with it the flares, winds, coronal mass ejections and energetic particle eruption that make up space weather, decline rapidly over a few hundred megayears, a fully convective M dwarf stays highly active for gigayears.
-Add that the zone where liquid water can reside is much closer to these faint stars, and the result is a planet that may be exposed to violent space weather conditions for a good fraction of the age of the universe.
+In M dwarfs, stellar activity evolution on the main sequence unfolds much slower than for a FGK star.
+While for the latter, magnetic activity, and with it the flares, winds, coronal mass ejections and energetic particle eruptions that make up their space weather, decline rapidly over a few hundred megayears, a fully convective M dwarf stays highly active for gigayears~\citep{XXXX}.
+Moreover, the zone where liquid water can reside is much closer to these faint stars.
+The result is a planet that may be exposed to violent space weather conditions for a good fraction of the age of the universe.
 
-Space weather originates from the stellar corona.
+Stellar space weather originates from the stellar corona.
 The transition from star to brown dwarf is characterized by increasing rotation speed and decreasing coronal emission.
-The simultaneous surge of radio emission indicates a transition from stellar corona to planet-like magnetosphere at the bottom of the main sequence.
-Yet, brown dwarfs are regularly detected with energetic flares regardless of their low X-ray luminosity.
-How these flares are produced in the apparent absence of a solar-type corona is unclear.
-An intermittent corona is conceivable, similarly XYZ.
-Perhaps, the rapid rotation of these low-mass objects plays a role, as it governs the dynamo process within, and consequently influences the complex process of flux emergence~\citep{weber2023understanding}.
+The simultaneous surge of radio emission indicates a transition from stellar corona to planet-like magnetosphere at the bottom of the main sequence~(XXXX).
+Yet, brown dwarfs are regularly detected with energetic flares regardless of their low X-ray luminosity~(XXXX).
+How these flares are produced in the apparent absence of a solar-type corona is unclear~\citep{weber2023understanding}.%.
+An intermittent corona is conceivable, similarly XXXX.
+Perhaps, the rapid rotation of these low-mass objects plays a role, as it governs the dynamo process within, and consequently influences the complex process of flux emergence
 
-\citet{ilin2021giant} searched fully convective dwarfs observed in TESS, and found four flares on four rapidly rotating ($P<10\,$h) M5-M7 dwarfs.
-These flares' durations spanned multiple rotation periods of each star, so that the flares were observed to rotate in and out view.
+\citet{ilin2021giant} searched fully convective dwarfs that straddle the transition from star to brown dwarf with TESS.
+They found four flares on four rapidly rotating ($P<10\,$h) M5-M7 dwarfs that lasted for multiple rotation periods of each star.
+The flares were observed to rotate in and out view.
 The shape of this modulation together with known inclination (combining $P$ and $v \sin i$ from high resolution spectroscopy) allowed them to determine the latitudes of these flares.
-Surprisingly, these flares were found closer to the pole than to the equator, in contrast to the Sun, where flares are usually found below 30 deg latitude.
-Their results indicate a preference for flares at high latitudes in these star, as a chance finding was unlikely ($0.1\%$).
-If flares and associated particle eruptions have a latitude preference, the space weather of a planet in an aligned orbit would be less severe than previously thought.
+Surprisingly, these flares were found significantly closer to the pole than to the equator, in contrast to the Sun, where flares are usually found below 30 deg latitude.
+Their results indicate a preference for flares at high latitudes in these stars, as a chance finding was unlikely ($\sim 0.1\%$).
+If flares and associated particle eruptions indeed have a latitude preference, the space weather of a planet in an aligned orbit would be less severe than previously thought.
 %Are high flaring latitudes  an effect of the dynamo at high rotation speeds?
@@
 
-TIC 277539431, TIC 277 for short, is one out of four stars with a high-latitude flare in its TESS observations~\citep{ilin2021giant}.
-In fact, it showed the highest latitude flare known to date at $81\pm1\,$deg.
-TIC 277 is an M7 dwarf with a rotation period of 4.56 hours.
-Its age is unknown.
-\citet{schneider2018hazmat} classify it as a field star, but its FUV and NUV fluxes are on the high end for field stars and would be compatible with the NUV/FUV emission from younger open cluster members, as well.
-Is its high flaring latitude an effect of the dynamo at high rotation speeds? Or is there something special about this star's corona that led it to produce are flare at these latitudes?
+TIC 277539431, TIC 277 for short, is one out of four stars with a high-latitude flare in its TESS observations~\citep[][Table \ref{tab:starparams}]{ilin2021giant}.
+In fact, it showed the highest latitude flare known to date at $\sim81\,$deg.
+TIC 277 is an M7 dwarf with an extremely short rotation period of 4.56 hours, detected in the rotational modulation of its light curve.
+While the fast rotation and late spectral type could be indicative of a brown dwarf~(), the detection of flares suggests the presence of a stellar corona~().%The star's age is unknown.
+\citet{schneider2018hazmat} classify it as a field star, but its FUV and NUV fluxes are on the high end for field stars and would, if not as high, be compatible with the NUV/FUV emission from younger open cluster members.
+%Is its high flaring latitude an effect of the dynamo at high rotation speeds? Or is there something special about this star's corona that led it to produce are flare at these latitudes? 
+In this work, we follow TIC 277 with XMM-Newton for 36 ks and use optical monitoring from the Transiting Exoplanet Survey Satellite~(TESS,~\citealt{ricker2015transiting}, Section~\ref{sec:obs}) to measure its coronal and flaring properties, respectively~(Section~\ref{sec:methods}).
+We contrast our results~(Section~\ref{sec:results}) with the literature to investigate if TIC~277 behaves like a low mass star, brown dwarf, or something in between~(Section~\ref{sec:discussion}), and conclude in Section~\ref{sec:summary}.
 
-In this work, we follow TIC 277 with XMM-Newton for 36 ks and use optical monitoring from the Transiting Exoplanet Survey Satellite~(TESS,~\citealt{ricker2015transiting}) to measure its coronal properties, and investigate if producing a high-latitude flare requires an unusual corona, or if it is rather typical of stars of this spectral type and rotation.
+%producing a high-latitude flare requires an unusual corona, or if it is rather typical of stars of this spectral type and rotation.
 
@@
     [3] \citet{pecaut2013intrinsic}
-    \label{tab:modelparams}
+    \label{tab:starparams}
 \end{table}
@@
 \section{Observations}
+\label{sec:obs}
 
@@
 For the OM flare, we derived the bolometric flare energy following the procedure in~\citet{shibayama2013superflares}.
-We used the throughput curve for the white light filter scaled to unity at the peak of transmission as an optimistic response curve $R_{\lambda, \rm OM}$, given that the degradation of this filter's response is poorly constrained.
+We used the throughput curve for the white light filter scaled to unity at the peak of transmission as an optimistic response curve $R_{\lambda, \rm OM}$, given that the degradation of this filter's response is poorly constrained\footnote{XMM-Newton User's Handbook 3.5.3.1, \url{https://xmm-tools.cosmos.esa.int/external/xmm_user_support/documentation/uhb/omfilters.html}, accessed Aug 8, 2023}.
 From that, we extracted a ratio $f_{\rm OM}$ of stellar to flare flux of $1.4\cdot 10^{-4}$ for a $10,000\,$K blackbody flare\footnote{A flare temperature of $10,000\,$K is a typical approximation for energetic M dwarf flares~\citep{kowalski2013timeresolved, howard2020evryflarea}.}:
@@
 
-We note that due to the optimistic response curve assumption the resulting $E_{\rm flare}$ is likely higher in reality.
+We note that due to the optimistic response curve assumption, the resulting $E_{\rm flare}$ is likely higher in reality.
 
@@
 \section{Results}
-
+\label{sec:results}
 With the XMM-Newton and TESS observations, we could constrain the coronal and flaring properties of TIC 277.
@@
 The resolution of different temperature component depends on the apparent brightness and integration time of the spectrum.
-Nearby stars like Proxima Cen can be modeled with a range of temperatures without overfitting~\citep[e.g.,][]{gudel2004flares}, while one can often only capture one dominant temperature in fainter objects~(XXXX).
+Nearby stars like Proxima Cen can be modeled with a range of temperatures without overfitting~\citep[e.g.,][]{gudel2004flares, drake2020pointing}, while one can often only capture one dominant temperature in fainter objects~(XXXX).
 
@@
 \label{sec:discussion:xraylum}
-Similar to its temperature make-up, while TIC~277's coronal luminosity relative to bolometric luminosity of $L_X/L_{\rm bol}=$\LXLbol is somewhat lower than the $L_X/L_{\rm bol}\sim 10^{-3}$ average in the saturated regime~\citep{wright2011stellaractivityrotation,wright2018stellar}, it is not an outlier to other M dwarfs in this activity regime.
+TIC~277's coronal luminosity relative to bolometric luminosity of $L_X/L_{\rm bol}=$\LXLbol in quiescence is about an order of magnitude lower than the $L_X/L_{\rm bol}\sim 10^{-3}$ average in the saturated regime of partly and fully convective M dwarfs~\citep{wright2011stellaractivityrotation,wright2018stellar}.
+In the Hyades and Praesepe, \citet{nunez2022factory} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend in $L_X/L_{\rm bol}$ with decreasing rotation period for M3.5-M7 dwarfs.
+This tentatively points toward a supersaturation of the corona at fast rotation speeds, as has been suggested before for M dwarfs~\citep{jeffries2011investigating}, but also earlier type stars~\citep{prosser1996rosat, argiroffi2016supersaturation} and even red giants~\citep{dixon2020rotationally}.
+None of the M dwarf studies probe below spectral type M6.
+At later spectral type, a decline in coronal activity is expected as the magnetosphere transitions from a stellar to a planetary one~\citep{pineda2017panchromatic}.
+\citet{stelzer2022first} investigated late M and L dwarfs with eROSITA to show that $L_X/L_{\rm bol}$ increases from the full convection boundary (M3) towards later spectral types~\citep{magaudda2022firsta}, but begins to decrease again from M7 downward.
+In a compilation together with previous work from \citet{stelzer2012, cook2014trends, deluca2020extras, williams2014trends}, TIC~277 appears at the lower boundary of the $L_X/L_{\rm bol}$ distribution for M7 dwarfs.
+However, the rotation periods are unknown for most of these late M dwarfs.
+Sometime projected rotation periods are known~\citep{cook2014trends}, but true periods are only detected in a handful of these stars: TRAPPIST-1 is similar to TIC~277 with $L_X/L_{\rm bol}\sim 10^{-3}$ at its slower 1.4 d rotation period.
 
-In the Hyades and Praesepe, \citet{nunez2022factory} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend in $L_X/L_{\rm bol}$ with decreasing rotation period for M3.5-M7 dwarfs, tentatively pointing toward a supersaturation of the corona at fast rotation speeds, as has been conjectured before~\citep{jeffries2011investigating,ramsay2020tess}.
-\citet{jeffries2011investigating} suggest that the cause for supersaturation is centrifugal stripping of the corona, but \citet{reiners2022magnetism} show that if the averae surface magnetic field strength increases, $L_X/L_{\rm bol}$ increases likewise with no sign of a break, up to $10^{-3}$.
-They conclude that the trend is rather due to the underlying dynamo.
-If TIC~277 is indeed underluminous, these results suggest that its dynamo produces a weaker magnetic field than that of other saturated M dwarfs, which results in a slightly lower X-ray luminosity.
+At the same time, strong magnetic fields are found across the board in brown dwarfs~(XXXX).
+
+If TIC~277 is indeed systematically underluminous, these results suggest that its dynamo produces a weaker magnetic field than that of other saturated M dwarfs, which results in a slightly lower X-ray luminosity.
 However, since the flaring activity of TIC~277 is in line with saturated activity, this conjecture currently lacks evidence from other activity indicators~\citep[see][, for supersaturation in flaring activity in fully convective M dwarfs]{ramsay2020tess}.
-Moreover, late M dwarfs are underrepresented in these studies.
+As often, late M dwarfs are underrepresented in these studies.
 
-\citet{stelzer2022first} investigated late M and L dwarfs with eROSITA to show that $L_X/L_{\rm bol}$ increases from the full convection boundary (M3) towards later spectral types, but begins to decrease again from M7 downward.
-In that study, TIC~277 is at the lower boundary of the $L_X/L_{\rm bol}$ distribution for M7 dwarfs.
-At the same time, strong magnetic fields are found across the board in brown dwarfs~(XXXX).
-While it is not clear whether M7 falls in the stellar regime of~\citet{reiners2022magnetism} or into that latter brown dwarf realm, or has features of both, it appears at the lower bound regardless.
+% While it is not clear whether M7 falls in the stellar regime of~\citet{reiners2022magnetism} or into that latter brown dwarf realm, or has features of both, it appears at the lower bound regardless.
 The small number of supersaturated stars in \citet{nunez2022factory}  can only tentatively suggest that this is a common phenomenon among mid-to-late M dwarfs when they are quickly rotating.
@@
 \section{Summary and Conclusions}
-We investigated whether TIC 277, a rapidly rotating M7 dwarf with a flare localized at 81 deg latitude, shows exceptional coronal or flaring properties to produce said flare so close to the rotational pole.
+\label{sec:summary}
+We investigated whether TIC 277, a rapidly rotating M7 dwarf with a flare localized at 81 deg latitude, shows exceptional coronal or flaring properties to produce this flare so close to the rotational pole.
 We obtained 36 ks of EPIC and OM observations with XMM-Newton, and studied the five Sectors of red-optical 27-day light curves provided by TESS.
-We found an X-ray luminosity of XXXX, and a mean coronal temperature of XXXX MK, typical of rapidly rotating late M dwarfs.
-The flaring activity is also typical of late M dwarfs in the saturated regime, with a flare rate $f(> log 10 Eflare = 31.5 erg)\approx -1.1$.
+We found an X-ray luminosity of \LX, and a mean quiescent coronal temperature of \Tqmean, typical of rapidly rotating fully convective M dwarfs.
+The flaring activity is also typical of late M dwarfs in the saturated regime, with a flare rate $f(> log 10 Eflare = 31.5 erg)\approx$\ffdr.
 
@@
 The flare energy and rate of the OM flare are consistent with the flare frequency distribution of the 18 detected TESS flares, indicating that X-ray flares on TIC 277 can be observed with XMM-Newton about once every $10\,$h.
-The bolometric and soft-X-ray flare energies are within an order of magnitude within each other, which places them among other stars with simultaneous X-ray and optical flare energies~\citep{guarcello2019simultaneous, kuznetsov2021stellar}, as well as the Sun~(XXXX).
+The bolometric and soft-X-ray flare energies are within an order of magnitude within each other, which places them among other stars with simultaneous X-ray and optical flare detections, although, typically, the white light emission dominates over soft X-ray~\citep{guarcello2019simultaneous, kuznetsov2021stellar}.
 
-While TIC 277's coronal and flaring properties do not appear uncommon, this does not mean that the corona has no influence on the high latitude occurrence of flares.
-TIC 277 is a fully convective rapid rotator with a rotation period of only $4.56\,$h.
+While TIC 277's overall coronal and flaring properties do not appear uncommon, this does not mean that all activity saturated fully convective dwarfs produce high-latitude flares.
+TIC 277 is an extreme rapid rotator with a rotation period of only $4.56\,$h.
 The detected high-latitude flare in TESS~\citep{ilin2021giant} may hence be a product of high-latitude flux emergence driven by its fast rotation~\citep{weber2016modeling,weber2017suppression}.
-However, the type of magnetic dynamo driven by such fast rotation may produce a peculiarly structured corona~(XXXX), which then appears common within the mass-rotation regime of TIC 277.
+This process may in turn produce a spatially structured corona that is not resolved in our observations, so that it appears typical of slower rotators on the saturated activity branch.
 Less than ten flare latitudes on stars other than the Sun are known to date~\citep{wolter2008doppler, ilin2021giant, johnson2021simultaneous}.

### 2023-08-28 src/scripts/FIGURE_data_resid.py
improved fig3 fontsize and added a minus to slope value

@@ -64,10 +64,19 @@ def plot_data_resid_3(file):
     ax[1].axhline(0,c="k")
 
     # labels
-    ax[1].set_xlabel("E [keV]")
-    ax[0].set_ylabel("flux [counts / s / keV]")
-    ax[1].set_ylabel("residuals")
+        # labels
+    ax[1].set_xlabel("E [keV]", fontsize=14)
+    ax[0].set_ylabel("flux [counts/s/keV]", fontsize=14)
+    ax[1].set_ylabel("residuals [counts/s/keV]", fontsize=14)
+
+    # increase tick size labels
+    for a in ax:
+        a.tick_params(axis='both', which='major', labelsize=13)
+
+    # y limits
     ax[0].set_ylim(0,)
+
+    # legend
     ax[0].legend(loc=1, frameon=False, fontsize=12)
 
     # x limits
@@ -135,9 +144,12 @@ def plot_data_resid_1(file):
     ax[1].axhline(0,c="k")
 
     # labels
-    ax[1].set_xlabel("E [keV]")
-    ax[0].set_ylabel("flux [counts/s/keV]")
-    ax[1].set_ylabel("residuals")
+    ax[1].set_xlabel("E [keV]", fontsize=15)
+    ax[0].set_ylabel("flux [counts/s/keV]", fontsize=15)
+    ax[1].set_ylabel("residuals [counts/s/keV]", fontsize=15)
+
+    # increase tick size labels
+    ax[0].tick_params(axis='both', which='major', labelsize=15)
 
     # x limits
     for a in ax:

### 2023-08-28 src/scripts/VALUES.py

@@ -43,7 +43,7 @@ if __name__ == "__main__":
     alpha, beta = row["alpha"], row["beta"]
     beta_low_err, beta_up_err = row["beta_low_err"], row["beta_up_err"]
 
-    alpha_str = f"${alpha:.2f}_{{-{row['alpha_low_err']:.2f}}}^{{+{row['alpha_up_err']:.2f}}}$"
+    alpha_str = f"$-{alpha:.2f}_{{-{row['alpha_up_err']:.2f}}}^{{+{row['alpha_low_err']:.2f}}}$"
     beta_str = f"${np.log10(beta):.2f}_{{-{np.log10(beta_low_err):.2f}}}^{{+{np.log10(beta_up_err):.2f}}}" + r"\,\mathrm{d}^{-1}$"
 
     with open(paths.output / "tess_ffd_alpha.tex", "w") as f:

### 2023-08-28 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Title
-\title{Fully convective star that produced a near-polar flare shows a typical corona}
+\title{The corona of a fully convective star with a near-polar flare}
 
@@
 \begin{abstract}
-    In 2020, TESS observed a rapidly rotating M7 dwarf produce a flare at 81 deg latitude, the highest latitude flare located to date.
+    In 2020, the Transiting Exoplanet Survey Satellite (TESS) observed a rapidly rotating M7 dwarf produce a flare at 81 deg latitude, the highest latitude flare located to date.
 This is in stark contrast to solar flares that occur much closer to the equator, typically below 30 deg.
 The circumstances that allow flares at high latitudes to occur are poorly understood.
-We studied five Sectors of TESS observations, and obtained 36 ks of XMM-Newton observations, to find that TIC 277's corona does not differ significantly from other low mass stars on the saturated activity branch with respect to X-ray luminosity, coronal temperatures, and flaring activity.
-Additionally, XMM-Newton captured a simultaneous X-ray and optical flare on TIC 277 with a bolometric energy of \eom and a soft X-ray energy of \eepic in the 0.2-2 keV band.
-Combining the 18 flares found in TESS with the flare from XMM-Newton, we inferred that TIC 277 produces flares at or above this energy approximately once every 10 h.
-We conclude that TIC 277's X-ray and flaring activity is representative of other rapidly rotating late M dwarfs, but that further measurements of localized flares are needed to understand the conditions under which high latitude flares in these stars can occur.
+We studied five Sectors of TESS monitoring, and obtained 36 ks of XMM-Newton observations, to find that TIC 277's corona does not differ significantly from other low mass stars on the saturated activity branch with respect to coronal temperatures and flaring activity, but is underluminous in soft X-ray emission by about an order of magnitude.
+Combining the lack of X-ray flux, the high latitude flare, the star's viewing geometry, and the otherwise typical stellar corona, we suggest that the migration of flux emergence to the poles in rapid rotators can explain the observations of TIC~277.
+
+    %Additionally, XMM-Newton captured a simultaneous X-ray and optical flare on TIC 277 with a bolometric energy of \eom and a soft X-ray energy of \eepic in the 0.2-2 keV band.
+In combination with the TESS data, we infer that TIC 277 produces flares at or above this energy approximately once a day.
+    %Our results indicate that TIC 277 is underluminous in X-ray copX-ray and flaring activity is representative of other rapidly rotating late M dwarfs, but that further measurements of localized flares are needed to understand the conditions under which high latitude flares in these stars can occur.
 
@@
 \label{sec:intro}
-Most M dwarfs hosts terrestrial planets with a sizable fraction orbiting at instellations where liquid water can exist~\citep{dressing2015occurrence, hardegree-ullman2019kepler, ment2023occurrence}.
+Most M dwarfs hosts terrestrial planets, with a sizable fraction orbiting at instellations where liquid water can exist~\citep{dressing2015occurrence, hardegree-ullman2019kepler, ment2023occurrence}.
 However, to be habitable in an Earth-like manner, merely the possibility of surface water in its liquid form does not suffice.
-Space weather of the host, that is, the high energy radiation and particles that force the planet's atmosphere affects hospitability for life, for M dwarfs in particular~\citep{airapetian2020impact}.
-If the energetic (particle) flux is too high, the atmosphere may blow off~(XXXX).
+Space weather of the host, that is, the high energy radiation and particles that force the planet's atmosphere affects hospitality for life, for M dwarfs in particular~\citep{airapetian2020impact}.
+If the energetic (particle) flux is too high, the atmosphere may blow off~\citep[e.g., ][]{lammer2003atmospheric, garraffo2017threatening, ketzer2022influence} and water may be lost from the surface~\citep{doamaral2022contribution}.
 If not energetic enough, life may not emerge in the first place~\citep{rimmer2018origin}.
 
-In M dwarfs, stellar activity evolution on the main sequence unfolds much slower than for a FGK star.
-While for the latter, magnetic activity, and with it the flares, winds, coronal mass ejections and energetic particle eruptions that make up their space weather, decline rapidly over a few hundred megayears, a fully convective M dwarf stays highly active for gigayears~\citep{XXXX}.
+In M dwarfs, stellar activity evolution on the main sequence unfolds much slower than for an FGK star.
+While for the latter, magnetic activity, and with it the flares, winds, coronal mass ejections and energetic particle eruptions that make up their space weather, decline rapidly over a few hundred megayears, a fully convective M dwarf stays highly active for gigayears~\citep{johnstone2021active, medina2022galactic}.
 Moreover, the zone where liquid water can reside is much closer to these faint stars.
@@
 The transition from star to brown dwarf is characterized by increasing rotation speed and decreasing coronal emission.
-The simultaneous surge of radio emission indicates a transition from stellar corona to planet-like magnetosphere at the bottom of the main sequence~(XXXX).
-Yet, brown dwarfs are regularly detected with energetic flares regardless of their low X-ray luminosity~(XXXX).
-How these flares are produced in the apparent absence of a solar-type corona is unclear~\citep{weber2023understanding}.%.
+The simultaneous surge of radio emission indicates a transition from stellar corona to planet-like magnetosphere at the bottom of the main sequence~\citep{zarka1998auroral,pineda2017panchromatic}.
+Yet, brown dwarfs are regularly detected with energetic flares regardless of their expected low X-ray luminosity~\citep[e.g., ][]{hambaryan2004new, gizis2013kepler, paudel2020k2, schmidt2019largest}.
+How these flares are produced in the apparent absence of a solar-type corona is unclear~\citep{mullan2018frequencies}.%.
 An intermittent corona is conceivable, similarly XXXX.
@@
 \citet{ilin2021giant} searched fully convective dwarfs that straddle the transition from star to brown dwarf with TESS.
-They found four flares on four rapidly rotating ($P<10\,$h) M5-M7 dwarfs that lasted for multiple rotation periods of each star.
-The flares were observed to rotate in and out view.
-The shape of this modulation together with known inclination (combining $P$ and $v \sin i$ from high resolution spectroscopy) allowed them to determine the latitudes of these flares.
+They found four flares on four rapidly rotating ($P_{\rm rot}<10\,$h) M5-M7 dwarfs that lasted for multiple rotation periods of each star.
+The flares were observed to rotate in and out of view.
+The shape of this modulation together with known inclination (combining $P_{\rm rot}$ and $v \sin i$ from high resolution spectroscopy) allowed them to determine the latitudes of these flares.
 Surprisingly, these flares were found significantly closer to the pole than to the equator, in contrast to the Sun, where flares are usually found below 30 deg latitude.
@@
 
-TIC 277539431, TIC 277 for short, is one out of four stars with a high-latitude flare in its TESS observations~\citep[][Table \ref{tab:starparams}]{ilin2021giant}.
-In fact, it showed the highest latitude flare known to date at $\sim81\,$deg.
-TIC 277 is an M7 dwarf with an extremely short rotation period of 4.56 hours, detected in the rotational modulation of its light curve.
-While the fast rotation and late spectral type could be indicative of a brown dwarf~(), the detection of flares suggests the presence of a stellar corona~().%The star's age is unknown.
+TIC 277539431, TIC 277 for short, showed the highest latitude flare known to date at $\sim81\,$deg~\citep[][Table \ref{tab:starparams}]{ilin2021giant}.
+It is an M7 dwarf with an extremely short rotation period of $4.56\,$h.
+While the fast rotation and late spectral type could be indicative of a brown dwarf, the detection of flares suggests the presence of a stellar corona.
+%The star's age is unknown.
 \citet{schneider2018hazmat} classify it as a field star, but its FUV and NUV fluxes are on the high end for field stars and would, if not as high, be compatible with the NUV/FUV emission from younger open cluster members.
 %Is its high flaring latitude an effect of the dynamo at high rotation speeds? Or is there something special about this star's corona that led it to produce are flare at these latitudes?
-In this work, we follow TIC 277 with XMM-Newton for 36 ks and use optical monitoring from the Transiting Exoplanet Survey Satellite~(TESS,~\citealt{ricker2015transiting}, Section~\ref{sec:obs}) to measure its coronal and flaring properties, respectively~(Section~\ref{sec:methods}).
-We contrast our results~(Section~\ref{sec:results}) with the literature to investigate if TIC~277 behaves like a low mass star, brown dwarf, or something in between~(Section~\ref{sec:discussion}), and conclude in Section~\ref{sec:summary}.
+
+In this work, we follow TIC 277 with XMM-Newton for 36 ks and use optical monitoring from TESS~(Section~\ref{sec:obs}) to measure its coronal and flaring properties, respectively~(Section~\ref{sec:methods}).
+We contrast our results~(Section~\ref{sec:results}) with the literature to investigate if TIC~277 behaves like a low mass star or brown dwarf, investigate scenarios that are consistent with high latitude flare occurrence ~(Section~\ref{sec:discussion}), and conclude in Section~\ref{sec:summary}.
 
@@
          Parameter & Value  \\\hline
-         Distance $d$ & $13.70\pm0.11\,$pc [2] \\
+
          % $G$ mag & $14.7189 \pm 0.0011$ [4]\\
@@
          Spectral type & M7 [1]\\
+         Distance $d$ & $13.70\pm0.11\,$pc [2] \\
          Effective temperature $T_{\rm eff}$ & $2680^{60}_{-50}\,$ K [3]\\
@@
 \label{sec:obs:xmm}
-XMM-Newton is an X-ray telescope, which launched into orbit 1999, equipped with six science instruments that operate simultaneously, among them the European Photon Imaging Camera with three cameras, PN, MOS1 and MOS2; and the Optical Monitor (OM) that provides near-UV and optical photometry, among other capabilities.
+XMM-Newton is an X-ray telescope, which launched into orbit in 1999, equipped with six science instruments that operate simultaneously, among them the European Photon Imaging Camera with three cameras, PN, MOS1 and MOS2; and the Optical Monitor (OM) that provides near-UV and optical photometry, among other capabilities.
 We used XMM-Newton to observe TIC 277 for 36 ks, i.e.
@@
 Ilin).
-We used the EPIC instruments aboard XMM-Newton, i.e., MOS1/2 and PN, that cover the soft X-ray band between 0.2-12 keV, as well as the OM using its broad band white light filter.
+We used the EPIC instruments aboard XMM-Newton, i.e., MOS1/2 and PN, that cover the soft X-ray band from $0.2$ and up to $12\,$keV, as well as the OM using its broad band white light filter.
 
@@
 
- We extracted the events time series from EPIC using XMM SAS version 20 using the \texttt{evselect} task.
+Fig.~\ref{fig:lightcurves} shows the time series using the combined PN and MOS time series together with the OM light curve.
+We extracted the events time series from EPIC using XMM SAS version 20~\footnote{"Users Guide to the XMM-Newton Science Analysis System", Issue 18.0, 2023 (ESA: XMM-Newton SOC)} using the \texttt{evselect} task.
 We selected a circular source region with 20 arcsec radius on all detectors, and a circular source-free background region 120 and 90 arcsec in MOS and PN, respectively.
-We used the \texttt{epiclccorr} task in XMM SAS to correct for energy and time dependent loss of events\footnote{\url{https://heasarc.gsfc.nasa.gov/docs/xmm/sas/help/epiclccorr/node3.html}, accessed on July 3, 2023}, and extract a soft X-ray time series with a time binning of 200 s.
+We used the \texttt{epiclccorr} task in XMM SAS to correct for energy and time dependent loss of events\footnote{\url{https://heasarc.gsfc.nasa.gov/docs/xmm/sas/help/epiclccorr/node3.html}, accessed on July 3, 2023}, and extract a soft X-ray time series with a time binning of $200\,$s.
 
 Simultaneously, we used the optical monitor (OM) aboard XMM-Newton, to monitor TIC 277 in its white light filter at $10$\,s cadence.
-The OM white light filter is centered on $4,060\AA$, with an equivalent width of $3,470\AA$.
+The OM white light filter is centered on $4,060\,$\AA, with an equivalent width of $3,470\,$\AA.
 We used the \texttt{omfchain} in XMM SAS to extract the 8 individual light curves.
-
-Fig.~\ref{fig:lightcurves} shows the time series using the combined PN and MOS time series together with the OM light curve.
 
@@
 \subsubsection{EPIC Spectra}
-From the EPIC observations, we used standard spectral extraction procedure for point sources.
-We extracted spectra using \texttt{evselect} to select all events in the circular event and background regions,\texttt{epatplot} to check for pile-up of multiple photons in one exposure, \texttt{backscale} to calculate the area of the source region, and \texttt{rmfgen} and \texttt{arfgen} to convolve the observed energies with the instrument response, and produce the spectra.
+From the EPIC observations, we used the standard spectral extraction procedure for point sources.
+We extracted spectra using \texttt{evselect} to select all events in the circular event and background regions, \texttt{epatplot} to check for pile-up of multiple photons in one exposure, \texttt{backscale} to calculate the area of the source region, and \texttt{rmfgen} and \texttt{arfgen} to convolve the observed energies with the instrument response, and produce the spectra.
 Finally, we use \texttt{grppha} to rebin the data in the spectra to obtain at least 15 counts per bin.
@@
 
-We used 5 Sectors of optical monitoring of TIC 277 in 2-minute cadence by TESS.
+We used 5 Sectors of optical monitoring of TIC 277 in 2-minute cadence.
 Sector 12 was observed in May 2019, Sectors 37 and 39 in April and June 2021, and Sectors 64 and 65 in April and May 2023~(Fig.~\ref{fig:tess_lcs}).
@@
 We used AltaiPony~\citep{ilin2021altaipony} to de-trend the light curves using the custom de-trending function \texttt{custom\_detrending}\footnote{see also the documentation online at \url{https://altaipony.readthedocs.io/en/latest/tutorials/detrend.html}, accessed on July 3, 2023}, detailed in~\citet{ilin2022searching}.
-The procedure includes 3rd order spline fitting of coarse trends, followed by iterative sine fitting to remove rotational modulation, and the two Savitzky-Golay filters with a 6h and 3h window, respectively to remove aperiodic short term trends, while masking outliers above $>2.5\sigma$, as well as the six-fold of consecutive originally masked data points following the original mask, and finally, additional 3 adjacent data point before and after the so expanded mask.
+The procedure includes 3rd order spline fitting of coarse trends, followed by iterative sine fitting to remove rotational modulation, and two Savitzky-Golay filters~\citep{savitzky1964smoothing} with a 6h and 3h window, respectively to remove aperiodic short term trends, while masking a.
+outliers above $>2.5\sigma$, b.
+the six-fold of consecutive originally masked data points following the original mask, and c.
+additional three adjacent data points before and after the so expanded mask.
 
@@
 We analyzed both the full data set, and the quiescent and flaring parts separately.
-Neither subset could be adequately fit with a single temperature component (1T), and a 3T model did not improve the fit compared to the 2T-model either.
+Neither subset could be adequately fit with a single temperature component (1T), and a 3T model did not improve the fit compared to the 2T-model.
 In the full data set, we used the \texttt{bayes} method in XSPEC to assign constant priors within the model's hard limits, and sample the uncertainties using the Markov-Chain-Monte-Carlo (MCMC) method using \texttt{chain} for a total of 30000 steps, discarding the first 5000 as the burn-in phase.
-From the same fit, but using PN data only, we derived the flux and X-ray luminosity in the 0.2-12 keV band.
-For the quiescent and flaring portions of the observations, we chose the same procedure, but used Gaussian prior distributions for the two coronal temperatures using the full data set as ground truth, and assigning the largest of the upper and lower uncertainties as standard deviation.
+From the same fit, but using PN data only, we derived the flux $F_X$ and X-ray luminosity $L_X$ in the soft $0.2-2\,$keV band.
+For the quiescent and flaring portions of the observations, we chose the same procedure, but used Gaussian prior distributions for the two coronal temperatures using the full data set as ground truth, and assigning the largest difference between the 50th, and 16th and 84th percentiles as standard deviation.
 
@@
 We used the throughput curve for the white light filter scaled to unity at the peak of transmission as an optimistic response curve $R_{\lambda, \rm OM}$, given that the degradation of this filter's response is poorly constrained\footnote{XMM-Newton User's Handbook 3.5.3.1, \url{https://xmm-tools.cosmos.esa.int/external/xmm_user_support/documentation/uhb/omfilters.html}, accessed Aug 8, 2023}.
-From that, we extracted a ratio $f_{\rm OM}$ of stellar to flare flux of $1.4\cdot 10^{-4}$ for a $10,000\,$K blackbody flare\footnote{A flare temperature of $10,000\,$K is a typical approximation for energetic M dwarf flares~\citep{kowalski2013timeresolved, howard2020evryflarea}.}:
+From that, we extracted a ratio $f_{\rm OM}$ of stellar to flare flux of $1.4\cdot 10^{-4}$ for a $10,000\,$K blackbody flare\footnote{A flare temperature of $T_{\rm flare}=10,000\,$K is a typical approximation for energetic M dwarf flares~\citep{kowalski2013timeresolved, howard2020evryflarea}.}:
 
 \begin{equation}
-    f_{\rm OM} = \frac{\displaystyle\int R_{\lambda, \rm OM} B_{\lambda(T_{\rm eff})}  d\lambda}{\displaystyle\int R_{\lambda, \rm OM} B_{\lambda(T_{\rm eff})} d\lambda} 
+    f_{\rm OM} = \frac{\displaystyle\int R_{\lambda, \rm OM} B_{\lambda(T_{\rm eff})}  d\lambda}{\displaystyle\int R_{\lambda, \rm OM} B_{\lambda(T_{\rm flare})} d\lambda} 
     \label{eq:ratio}
@@
 Analogously to Eq.~\ref{eq:ratio} for the OM flare, we used the TESS response curve to obtain a flux ratio $f_{\rm TESS}$ of $6.3\cdot10^{-3}$ in the redder optical filter of TESS, and convert $ED$ to bolometric flare energy using Eq.~\ref{eq:eflare}.
-Besides the $ED$, \texttt{altaipony} also yields flare start and end times, defined as the difference between the first and last data point above the flare detection criterion defined in Section~\ref{sec:obs:tess}; and relative amplitudes $a$.
+Besides the $ED$, \texttt{AltaiPony} also yields start and end time, defined as the difference between the first and last data point above the flare detection criterion defined in Section~\ref{sec:obs:tess}; and relative amplitude $a$, for each flare.
 
@@
 From the XMM-Newton EPIC spectra, we could measure the coronal properties of TIC in both its quiescent, and flaring states~(Table~\ref{tab:specfit}).
-Using the combined quiescent and flaring data, we found a two-temperature VAPEC model to fit better than a one-temperature solution and no worse than a three-temperature solution.
 The resulting fit~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about \Tcool and \Thot, respectively.
-With these temperatures as ground truth, we also fitted two-temperature VAPEC models to the flaring and quiescent portions of the observations separately.
+Assuming these temperatures as ground truth, we also fitted 2T APEC models to the flaring and quiescent portions of the observations separately.
 The coronal temperatures did not update significantly using either subset, although the quiescent spectrum appeared slightly cooler overall.
-However, the emission measure weighted mean temperature change dfrom \Tqmean in the quiescent spectrum to \Tfmean in the flaring spectrum.
-This suggests that the flaring corona emits the hotter component, while the cooler component corresponds to the quiescent corona.
+However, the emission measure weighted mean temperature $T_{\rm mean}$ changed from \Tqmean in the quiescent spectrum to \Tfmean in the flaring spectrum.
+From the quiescent PN spectra, we then also calculated soft X-ray flux of $F_X$\FX, and a luminosity $L_X=$\LX in the $0.2$-$2\,$keV band.
+%This suggests that the flaring corona emits the hotter component, while the cooler component corresponds to the quiescent corona.
 In this interpretation, the presence of a hot component in the quiescent spectrum indicates that TIC 277 is flaring below the detection threshold, consistent with an extrapolation of the flare frequency distribution to lower temperatures~(see the following Section, and Fig.~\ref{fig:ffd}).
 
-Not using the full set of observations as ground truth resulted in an unconstrained hot coronal component.
-This can be interpreted as due to a lack of X-ray flux above 5 keV.
-Hot flares with temperatures of >50 MK have been measured on DS Tuc A~\citep{pillitteri2011xray}.
-On the Sun, the highest flare associated coronal temperatures are around 35 MK~\citep{kay2003soft}.
+We note that not using the full set of observations as ground truth resulted in an unconstrained hot coronal component.
+This can be interpreted as due to a lack of X-ray flux above $5\,$keV.
+Hot flares with temperatures of $>50\,$MK have been measured on DS~Tuc~A~\citep{pillitteri2011xray}.
+On the Sun, the highest flare associated coronal temperatures are around $35\,$MK~\citep{kay2003soft}.
 
-From the PN spectra, we measured a soft X-ray flux of \FX, and a luminosity \LX in the $0.2-2\,$keV band. 
  \begin{table}
@@
         \caption{
-         Best-fit joint Double APEC spectrum fit with Fe/0$=0.6$.
+         Best-fit joint two-temperature APEC spectral fit with Fe/O$=0.6$. The top panel shows the spectra taken with the different EPIC instruments. The bottom panel shows the residuals to the fit.
         }
@@
     \script{TABLE_tess_flares.py}
-    \caption{Flares detected with TESS. $a$ is the relative amplitude, and $E_{\rm bol}$ is the bolometric flare energy assuming a 10,000 K blackbody emission from the flare.}
+    \caption{Flares detected with TESS.
+$t_s$ is the starting time of the flare, $a$ is its relative amplitude, and $E_{\rm bol}$ is the bolometric energy assuming a 10,000 K blackbody emission from the flare.}
     \input{output/tess_flares.tex}
@@
 
-The FFD power law fit converged after 13,500 steps on a slope $\alpha =$\ffdalpha.
-The frequency $R_{31.5}$ of flares per day above $\log_{10} E = 31.5\,\mathrm{erg}$ is about $\log_{10}R_{31.5}=$\ffdr.
-The slope is typical of other flaring stars, regardless of spectral type a rotation period~(see, e.g., Fig.~13 in \citealt{ilin2021flares}).
+The FFD power law fit converged after 13,500 steps on a slope $\alpha =$\ffdalpha\unskip.
+The frequency $R_{31.5}$ of flares per day above $\log_{10} E_{\rm flare} = 31.5\,\mathrm{erg}$ is about $\log_{10}R_{31.5}=$\ffdr\unskip.
+The slope is typical of other flaring stars, regardless of spectral type and rotation period~(see, e.g., Fig.~13 in \citealt{ilin2021flares}).
 $R_{31.5}$ is typical of flaring M dwarfs, in particular typical of late M dwarfs in the saturated activity regime, where $R_{31.5}$ becomes independent of rotation period~\citep{medina2020flare,murray2022study}.
@@
 For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of \eepic in the $0.2-5\,$keV band.
-The energy of the corresponding OM flare is \eom, which fits the extension of TESS' flare frequency distributions to lower energies within uncertainties on the power law fit~(Fig.~\ref{fig:ffd}).
-About one flare every $10\,$h of the order of the observed one or above can therefore be expected in future X-ray observations of TIC~277, assuming all optical flares above this energy in the OM data correspond to X-ray flares of similar or higher energy to the observed.
-\citet{guarcello2019simultaneous} find moderate correlation between flare energies in Kepler/K2 data and XMM-Newton, but the majority of their sampled energies are above $10^{33}\,$erg both in X-ray and optical.
+The energy of the corresponding OM flare is \eom\unskip.
+We use the 36ks observing baseline of XMM-Newton to calculate a flare rate in OM.
+Fig.~\ref{fig:ffd} illustrates that it is roughly consistent, if slightly higher, than the extension of TESS' flare frequency distributions to lower energies.
+About one flare per day of the order of the observed one or above can therefore be expected in future X-ray observations of TIC~277.
+However, this assumes that all optical flares above this energy in the optical data correspond to detected X-ray flares in the EPIC instruments, which is not always the case~\citep{paudel2021simultaneous}.
+\citet{guarcello2019simultaneous, kuznetsov2021stellar} find moderate correlation between flare energies in Kepler/K2 data and XMM-Newton, but the majority of their sampled flare energies are two orders of magnitude above the flare in this work, both in X-ray and optical.
 \citet{kuznetsov2021stellar} compare primary Kepler mission observations with XMM-Newton to find a similar result.
@@
         \caption{
-         Cumulative flare frequency distribution and power law fit.
+         Cumulative flare frequency distribution of TESS flares (black dots), with a power law fit (green line), and uncertainties (dotted lines).
+The rate of OM flares (gray square) is calculated from the observing baseline of XMM-Newton.
         }
@@
 
-Is the magnetic activity of the star with the highest known flare latitude known to date is any way unusual?
-Our findings indicate that TIC 277's coronal temperature~(Section~\ref{sec:discussion:xraytemp}), X-ray luminosity~(Section~\ref{sec:discussion:xraylum}), and flaring activity~(Section~\ref{sec:discussion:flares}) are typical of stars with its rotation rate and mass.
-TIC 277 is rotating rapidly, even for the typically fast rotating late M dwarfs~\citep{medina2022galactic}.
-This could lead to an unusual manifestation of the stellar dynamo that produces high latitude flares that other dynamo states cannot generate.
-With TESS and XMM-Newton, we can now evaluate if this suspected exceptionality is reflected in the star's magnetic activity and coronal properties.
+TIC~277's spectral type and rotation places it in the middle of the transition from star to brown dwarf.
+Its coronal properties are therefore not known a priori.
+Our findings indicate that TIC 277's coronal temperature~(Section~\ref{sec:discussion:xraytemp}), and flaring activity~(Section~\ref{sec:discussion:flares}) are typical of saturated fully convective M dwarf stars.
+However, TIC 277 is rotating rapidly, even for the typically fast rotating late M dwarfs~\citep{medina2022galactic}, which is more typical of brown dwarfs.
+This combination could lead to an unusual manifestation of the stellar dynamo that produces high latitude flares that other dynamo states cannot generate~\citep{weber2016modeling, weber2023understanding}.
+TIC 277 is underluminous in X-ray by about an order of magnitude compared to other activity saturated M dwarfs, which, in combination with the detection of a high latitude flare, may be an indication of polar updraft migration that is suspected to occur in rapid rotators~(Section~\ref{sec:discussion:xraylum}).
 
@@
 Stellar point source X-ray observations are a superposition of various temperatures and densities.
-The resolution of different temperature component depends on the apparent brightness and integration time of the spectrum.
-Nearby stars like Proxima Cen can be modeled with a range of temperatures without overfitting~\citep[e.g.,][]{gudel2004flares, drake2020pointing}, while one can often only capture one dominant temperature in fainter objects~(XXXX).
+The resolution of different temperature components depends on the apparent brightness and integration time of the spectrum.
+Nearby stars like Proxima Cen can be modeled with a range of temperatures~\citep[e.g.,][]{gudel2004flares, drake2020pointing}, while one can often only capture few dominant temperatures in fainter objects.
 
 We find that the quiescent corona of TIC 277 is best described by two components, similar to many other active M dwarfs~\citep{wargelin2008xray, robrade2010quiescent, behr2023muscles, magaudda2022firsta}.
-We measure temperatures of 3 MK and 12 MK with an emission-measure weighted average of about 6.5 MK, which rises to 11 MK during the flare~(Table~\ref{tab:specfit}).
-The hotter 13 MK component represents thermal bremsstrahlung radiation from hot plasma that evaporates from the chromosphere into the coronal flaring loops~\citep{benz2016flare}, which is supported by the observed delay between the optical and the X-ray flare~\citep{hawley2003multiwavelength}.
+The hotter component represents thermal bremsstrahlung radiation from hot plasma that evaporates from the chromosphere into the coronal flaring loops~\citep{benz2016flare}, which is supported by the observed delay between the optical and the X-ray flare~\citep[Fig.~\ref{fig:lightcurves}, ][]{hawley2003multiwavelength}.
 The optical flare originates from the lower chromosphere, which is heated by the precipitating particles.
 This heating then causes evaporation of particles into the flare loop.
-A third, very hot, $>40\,$MK component could only be poorly constrained due to a lack of sensitivity to flux above $5\,$keV.
+%A third, very hot, possibly $>40\,$MK component could only be poorly constrained due to a lack of sensitivity to flux above $5\,$keV, and its short lifetime during the peak of the flare.
 
 The range of 6.5-11 MK for the emission measure weighted temperature of TIC 277 places it squarely in the saturated regime with other mid M dwarfs down to spectral type M6~\citep{wright2018stellar, magaudda2020relation, stelzer2022first,robrade2005xray,raassen2003xray,paudel2021simultaneous, foster2020corona}.
-Flares on the slowly rotating Proxima Cen (M5.5, 85 d rotation~XXXX), appear with a range of temperatures at 2-3 MK and 10-20 MK~\citep{gudel2004flares, fuhrmeister2011multiwavelength, fuhrmeister2022high, howard2022mouse} placing it in the transition region between saturated and unsaturated regimes.
+Flares on the slowly rotating Proxima Cen (M5.5, $P_{\rm rot} = 83\,$d,~\citealt{anglada-escude2016terrestrial}), appear with a range of temperatures at 2-3 MK and 10-20 MK~\citep{gudel2004flares, fuhrmeister2011multiwavelength, fuhrmeister2022high, howard2022mouse} placing it in the transition region between saturated and unsaturated regimes.
 Only at rotation periods beyond 100 days, flaring activity declines to yield coronal temperatures $<2\,$MK~\citep{wright2018stellar, foster2020corona}.
-For a more direct comparison, there are only few late M dwarfs beyond spectral type M6 with a spectrally resolved corona.
-Their intrinsic faintness and gradual decline in coronal emission make them hard to detect~\citep{berger2010, cook2014trends, stelzer2022first}.
-One example is NLTT 33370 AB, an M7 binary with a rotation period of about $3.8\,$h (close to TIC 277's $4.56\,$h and of the same spectral type) is combined of a 3.1 MK and 14 MK component~\citep{williams2015simultaneous}.
-Another is TRAPPIST-1~(M8, $P_{\rm rot} = 1.4\,$d~XXXX).
+
+There are only few late M dwarfs beyond spectral type M6 with a spectrally resolved corona.
+Their intrinsic faintness and gradual decline in coronal emission toward the brown dwarf regime make them hard to detect~\citep{berger2010simultaneous, cook2014trends, stelzer2022first}.
+One example is NLTT 33370 AB, an M7 binary with a rotation period of about $3.8\,$h (close to TIC 277's $4.56\,$h, and of the same spectral type) is combined of a 3.1 MK and 14 MK component~\citep{williams2015simultaneous}.
+Another is TRAPPIST-1~(M8, $P_{\rm rot} = 3.3\,$d~\citep{luger2017sevenplanet}).
 Its corona appears cooler, with a 1.74 MK and a 9.6 MK component~\citep{wheatley2017strong}, and an emission measure weighted temperature of 5.36 MK~\citep{brown2023coronal}.
 
-To summarize, considering the relatively low two-temperature resolution of TIC~277's X-ray spectrum and the dearth of X-ray spectra for $>M7$ dwarfs to compare with, it shows a coronal temperature make-up common for mid-to-late M dwarf in the saturated activity regime.
+Bearing in mind the relatively low two-temperature resolution of TIC~277's X-ray spectrum and the dearth of X-ray spectra for stars at the bottom of the main sequence, it shows a coronal temperature make-up common for mid-to-late M dwarf in the saturated activity regime that it will possibly keep for gigayears until it spins down to very low rotation rates~\citep{medina2022galactic, engle2023living}.
 
@@
 CC Eri, a K7 dwarf, was described with a 3T model~\citep{crespo-chacon2007xray} with energies about $10^{33}\,$erg and peak temperatures of 30 MK.
+\subsection{Flaring activity}
+\label{sec:discussion:flares}
+
+TIC~277's flare rate is consistent with other saturated fully convective dwarfs~\citep{medina2020flare, murray2022study}, and so is the energy distribution with a power law slope of \ffdalpha.
+It is in fact comparable to TRAPPIST-1's flaring behavior~\citep{paudel2018k2}, which has been investigated for its effects on the seven terrestrial planets in its orbit.
+
+Recent observations of TRAPPIST-1 b~\citep{greene2023thermal, ih2023constraining} with the James Webb Space Telescope find the planet mostly stripped of an atmosphere, which the authors attribute to the intense flare driven X-ray radiation and energetic particle exposure at the short orbital periods of the innermost planet.
+For the habitable zone planets, TRAPPIST-1 e and f, simulations suggest that flares alter their chemical composition, in particular ozone, nitric oxide and hydroxide, but that the effects would be below the noise floor of JWST~\citep{chen2021persistence}.
+However, given the uncertain age of TRAPPIST-1~\citep{burgasser2017age, gonzales2019reanalysis, birky2021improved}, the temperate planets e and f may have been entirely stripped of their atmospheres in the past gigayear(s)~\citep{garraffo2017threatening}.
+
+TIC~277's rapid rotation and UV flux is indicative of a younger age than TRAPPIST-1, although it has not been attributed to any star cluster so far~\citep{schneider2018hazmat}.
+Interpreting the upcoming observations of TRAPPIST-1 e with JWST will most likely involve its activity history.
+Studies of its younger counterparts, late M dwarfs like TIC~277, are required to empirically constrain cumulative effects of atmospheric forcing of planets.
+If its energetic flares commonly occur at high latitudes~\citep{ilin2021giant}, future models of energetic particle exposure of habitable zone planets may have to include age dependent latitudes of particle eruption to reproduce observations.
 
@@
 In the Hyades and Praesepe, \citet{nunez2022factory} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend in $L_X/L_{\rm bol}$ with decreasing rotation period for M3.5-M7 dwarfs.
-This tentatively points toward a supersaturation of the corona at fast rotation speeds, as has been suggested before for M dwarfs~\citep{jeffries2011investigating}, but also earlier type stars~\citep{prosser1996rosat, argiroffi2016supersaturation} and even red giants~\citep{dixon2020rotationally}.
-None of the M dwarf studies probe below spectral type M6.
+This tentatively points toward a supersaturation of the corona at fast rotation speeds, as has been suggested before for M~dwarfs~\citep{jeffries2011investigating}, but also earlier type stars~\citep{prosser1996rosat, argiroffi2016supersaturation} and even red giants~\citep{dixon2020rotationally}.
+However, none of the aforementioned M~dwarf studies probe below spectral type M6.
+
 At later spectral type, a decline in coronal activity is expected as the magnetosphere transitions from a stellar to a planetary one~\citep{pineda2017panchromatic}.
-\citet{stelzer2022first} investigated late M and L dwarfs with eROSITA to show that $L_X/L_{\rm bol}$ increases from the full convection boundary (M3) towards later spectral types~\citep{magaudda2022firsta}, but begins to decrease again from M7 downward.
-In a compilation together with previous work from \citet{stelzer2012, cook2014trends, deluca2020extras, williams2014trends}, TIC~277 appears at the lower boundary of the $L_X/L_{\rm bol}$ distribution for M7 dwarfs.
+\citet{stelzer2022first} investigated late M and L dwarfs with eROSITA to show that $L_X/L_{\rm bol}$ increases towards later spectral types~\citep{magaudda2022firsta}, but begins to decrease again from M7 downward.
+In their compilation together with previous work from \citet{stelzer2012ultracool, cook2014trends, deluca2020extras, williams2014trends}, TIC~277 appears at the lower boundary of the $L_X/L_{\rm bol}$ distribution for M7 dwarfs.
 However, the rotation periods are unknown for most of these late M dwarfs.
-Sometime projected rotation periods are known~\citep{cook2014trends}, but true periods are only detected in a handful of these stars: TRAPPIST-1 is similar to TIC~277 with $L_X/L_{\rm bol}\sim 10^{-3}$ at its slower 1.4 d rotation period.
+Sometime projected rotation periods are known~\citep{cook2014trends}, but true periods are only detected in a handful of these stars, which we list in Table~\ref{tab:latmlxlbol}.
+TIC~277 appears at the lower end of this small sample.
 
-At the same time, strong magnetic fields are found across the board in brown dwarfs~(XXXX).
+\begin{table}[]
+    \centering
+    \caption{Late M dwarfs with known rotation period and X-ray luminosity.}
+    \begin{tabular}{lccc}
+        ID & SpT &$P_{\rm rot}$ [d] & $\log_{10} \frac{L_X}{L_{\rm bol}} $\\\hline
+        TIC 277 (this work) & M7 & 0.19 & $-4$ \\
+        TRAPPIST-1 & M8 & 3.3 & $-4$ [1]\\
+        LHS 248 & M6.5 & 0.46 & $-3.9$ [2] \\
+        NLTT 33370 AB & M7 & 0.16 & $-3.5$ [3] \\
+        LP 412-31 & M8 & 0.61 & $-3.1$ [4]\\\hline
+    \end{tabular}
+    \flushleft
+     \footnotesize
+    [1] \citet{brown2023coronal}, 
+    [2] \citet{cook2014trends}, 
+    [3] \citet{williams2015simultaneous}, 
+    [4] \citet{stelzer2006simultaneous}    
+    %3.3 d from \citep{luger2017sevenplanet}
+    \label{tab:latmlxlbol}
+\end{table}
 
-If TIC~277 is indeed systematically underluminous, these results suggest that its dynamo produces a weaker magnetic field than that of other saturated M dwarfs, which results in a slightly lower X-ray luminosity.
-However, since the flaring activity of TIC~277 is in line with saturated activity, this conjecture currently lacks evidence from other activity indicators~\citep[see][, for supersaturation in flaring activity in fully convective M dwarfs]{ramsay2020tess}.
-As often, late M dwarfs are underrepresented in these studies.
+%At the same time, strong magnetic fields are found across the board in brown dwarfs~(XXXX).
+
+If TIC~277 is indeed systematically underluminous, these results could mean that its dynamo produces a weaker magnetic field than that of other saturated M dwarfs.
+However, since the flaring activity and coronal temperature of TIC~277 are in line with saturated activity, this conjecture lacks evidence from these other activity indicators.
+An alternative explanation, that would at the same time explain the occurrence of a high latitude flare in TESS, is polar updraft migration~\citep{stepien2001rosat}.
+Polar updraft migration implies that at high rotation rates, flux emergence becomes more efficient near the poles than at the equator~\citep{weber2016modeling}, draining the equatorial regions of magnetic flux required to produce a corona there.
+As a result, the X-ray luminosity diminishes.
+Since TIC~277 is seen nearly equator-on~(Table~\ref{tab:starparams}), polar updraft could explain its fainter corona.
+Considering the high-latitude flare, viewing geometry, star-like flaring behavior and coronal temperatures together, we favor this scenario over a brown dwarf transition explanation, that would invoke a diminishing corona due to the overall cooler, more neutral atmosphere.
+
+%As often, late M dwarfs are underrepresented in these studies.
 
@@
 This could be an effect of its fast rotation, and consequent supersaturation as in \citet{magaudda2022first}, but we lack the rotation periods to confirm this.
-
-\subsection{Flaring activity -- a young TRAPPIST-1}
-\label{sec:discussion:flares}
-
-$\log_{10}(R_{31.5}$, that is, the rate of flares per day with energies above $\log E_{flare}=10^{31.5}\,$erg, is ca.
-$-1.1$, consistent with other saturated late M dwarfs~\citep{medina2020flare, murray2022study}.
-It is similar to TRAPPIST-1's flaring rate~(XXXX), which has been investigated for its effects on the seven terrestrial planets in its orbit.
-
-Recent observations of TRAPPIST-1 b~(XXXX) with the James Webb Space Telescope~(XXXX) find the planet mostly stripped of an atmosphere, which the authors attribute to the intense flare driven X-ray radiation and energetic particle exposure at the short orbital periods of the innermost planet.
-For the habitable zone planets TRAPPIST-1 e and f, simulations suggest that flares alter their chemical composition, in particular ozone, nitric oxide and hydroxide, but that the effects would be below the noise floor of JWST~\citep{chen2021persistence}.
-However, given the uncertain age of TRAPPIST-1~(XXXX), the temperate planets e and f may have been entirely stripped of their atmospheres in the past gigayear(s)~\citep{garraffo2017threatening}.
-
-TIC~277's rapid rotation is indicative of a younger age than TRAPPIST, although it has not been attributed to any star cluster so far~(XXXX).
-A correct understanding of whatever the upcoming observations of TRAPPIST-1 e will yield -- detection or non-detection of an atmosphere -- will depend strongly on its activity history.
-Studies of its younger counterparts, late M dwarfs like TIC~277, are required to empirically constrain cumulative effects of atmospheric forcing of planets.
-If, as this work suggests, TIC~277 is typical of stars of its spectral type and rotation rate, it may represent the youth of TRAPPIST-1.
-And if its energetic flares commonly occur at high latitudes, as suggested by~\citet{ilin2021giant}, future models of energetic particle exposure of habitable zone planets may have to include age dependent latitudes of particle eruption to reproduce observations.
 
@@
 \label{sec:summary}
-We investigated whether TIC 277, a rapidly rotating M7 dwarf with a flare localized at 81 deg latitude, shows exceptional coronal or flaring properties to produce this flare so close to the rotational pole.
+We investigated whether TIC 277, a rapidly rotating M7 dwarf with a flare localized at 81 deg latitude, shows unusual coronal or flaring properties that could explain the occurrence of this flare so close to the rotational pole.
 We obtained 36 ks of EPIC and OM observations with XMM-Newton, and studied the five Sectors of red-optical 27-day light curves provided by TESS.
-We found an X-ray luminosity of \LX, and a mean quiescent coronal temperature of \Tqmean, typical of rapidly rotating fully convective M dwarfs.
-The flaring activity is also typical of late M dwarfs in the saturated regime, with a flare rate $f(> log 10 Eflare = 31.5 erg)\approx$\ffdr.
+We found a mean quiescent coronal temperature of \Tqmean, and a flaring rate $f(> \log_{10} E_{\rm flare, bol} = 31.5\,\rm{erg})\approx$\ffdr with energy distribution with a slope of \ffdalpha, all typical of fully convective M dwarfs in the saturated regime.
+We also detected a simultaneous X-ray and optical flare with an energy of \eepic in the $0.2-5\,$keV band.
+In combination with the 18 detected TESS flares, we estimate that X-ray flares on TIC~277 can be observed with XMM-Newton at least about once a day.
+%The bolometric and soft-X-ray flare energies are within an order of magnitude within each other, which places them among other stars with simultaneous X-ray and optical flare detections, although, typically, the white light emission dominates over soft X-ray~\citep{guarcello2019simultaneous, kuznetsov2021stellar}.
 
-XMM-Newton also captured a simultaneous X-ray and optical flare with its EPIC and OM instruments, respectively.
-The flare energy and rate of the OM flare are consistent with the flare frequency distribution of the 18 detected TESS flares, indicating that X-ray flares on TIC 277 can be observed with XMM-Newton about once every $10\,$h.
-The bolometric and soft-X-ray flare energies are within an order of magnitude within each other, which places them among other stars with simultaneous X-ray and optical flare detections, although, typically, the white light emission dominates over soft X-ray~\citep{guarcello2019simultaneous, kuznetsov2021stellar}.
-
-While TIC 277's overall coronal and flaring properties do not appear uncommon, this does not mean that all activity saturated fully convective dwarfs produce high-latitude flares.
+An indication of an unusual corona stems from its X-ray luminosity relative to bolometric luminosity $L_X/L_{\rm bol}=$\LXLbol, which is about an order of magnitude lower than is typical for saturated activity M dwarfs.
 TIC 277 is an extreme rapid rotator with a rotation period of only $4.56\,$h.
 The detected high-latitude flare in TESS~\citep{ilin2021giant} may hence be a product of high-latitude flux emergence driven by its fast rotation~\citep{weber2016modeling,weber2017suppression}.
-This process may in turn produce a spatially structured corona that is not resolved in our observations, so that it appears typical of slower rotators on the saturated activity branch.
-Less than ten flare latitudes on stars other than the Sun are known to date~\citep{wolter2008doppler, ilin2021giant, johnson2021simultaneous}.
-Localization of more flares on the surfaces of a broader range of stars will allow us to discriminate the role of dynamo and flux emergence theories from that of coronal properties in the production of active latitudes.
+This effect is closely related to polar updraft migration, which suppresses coronal emission in equatorial regions~\citep{stepien2001rosat}.
+However, less than ten flare latitudes on stars other than the Sun are known to date~\citep{wolter2008doppler, ilin2021giant, johnson2021simultaneous}.
+Localization of more flares on the surfaces of a broader range of stars will allow us to resolve if polar updraft indeed occurs in these stars.
+
+%If TIC~277 is relatively young, as its NUV/FUV fluxes at the high end for field stars suggest~\citep{schneider2018hazmat}, the space weather it creates may be representative of that in the youth of similar very low mass planet hosts.
+These hosts, such as M8 type TRAPPIST-1~\citep{gillon2017seven}, or M6 type LP 890-9~\citep{delrez2022two}, are the most favorable targets for transmission spectroscopy of terrestrial planet atmospheres for the JWST.
+As the makeup of their planets' atmospheres strongly depends on the high energy radiation and particle environment created by their host star~\citep{airapetian2020impact}, their activity history as traced by TIC~277 is crucial for the interpretation of present~(XXXX) and future observations.
 
@@
 acknowledges funding from the German \textit{Leibniz-Gemeinschaft} under project number P67/2018.
-This work made use of the open source python software packages \texttt{lightkurve}\citep{lightkurvecollaboration2018lightkurve}, astropy~\citep{robitaille2013astropy}, numpy~(XXXX), pandas~\citep{reback2022pandasdev}, altaipony~\citep{ilin2021altaipony}, and XXXX.
+This work made use of the open source python software packages \texttt{lightkurve}\citep{lightkurvecollaboration2018lightkurve}, \texttt{astropy}~\citep{robitaille2013astropy}, \texttt{numpy}~\citep{harris2020array}, \texttt{pandas}~\citep{reback2022pandasdev}, \texttt{matplotlib}~\citep{hunter2007matplotlib}, \texttt{emcee}~\citep{foreman-mackey2013emcee}, \texttt{scipy}~\citep{mckinney2010data}, and \texttt{altaipony}~\citep{ilin2021altaipony}.
 This work has made use of data from the European Space Agency (ESA) mission
@@
 \citep{luger2021mappinga} to create all figures, format all tables, and compile the manuscript.
-Each figure links to the dataset stored on Zenodo (link to be added after acceptance), and to the script that produced the figure in the git repository~\href{https://github.com/ekaterinailin/xmm_for_tic277}{github.com/ekaterinailin/xmm\_for\_tic277}.
+Each figure links to the dataset stored on Zenodo (link to be added after acceptance), and to the script that produced the figure in the git repository on GitHub:~\href{https://github.com/ekaterinailin/xmm_for_tic277}{github.com/ekaterinailin/xmm\_for\_tic277}.
 The data analysis scripts can be found under \href{https://github.com/ekaterinailin/tic277}{github.com/ekaterinailin/tic277}.

### 2023-09-15 src/scripts/VALUES.py
added quiescent and flaring luminosities in EPIC

@@ -68,7 +68,7 @@ if __name__ == "__main__":
 
     df = pd.read_csv(paths.data / "joint_vapec_chain_fits.csv")
 
-    # select full data set
+    # select quiescent data set
     row = df[df.subset == "noflare"].iloc[0]
 
     # get Lx and make latex string with a single uncertainty
@@ -77,7 +77,20 @@ if __name__ == "__main__":
 
     print(Lx_str)
 
-    with open(paths.output / "epic_Lx.tex", "w") as f:
+    with open(paths.output / "epic_Lx_quiet.tex", "w") as f:
+        f.write(f"{Lx_str}")
+
+
+    # select flaring data set
+    row = df[df.subset == "onlyflare"].iloc[0]
+
+    # get Lx and make latex string with a single uncertainty
+    Lx, Lxerr = row.Lx_erg_s, row.Lx_erg_s_err
+    Lx_str = fr"${Lx/1e27:.2f} \pm {Lxerr/1e27:.2f}" + r"\times 10^{27}\,\rm{erg s}^{-1}$"
+
+    print(Lx_str)
+
+    with open(paths.output / "epic_Lx_flaring.tex", "w") as f:
         f.write(f"{Lx_str}")
 
     # Fx -----------------------------------------------------------------------

### 2023-09-15 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \newcommand{\FX}{\input{output/epic_flux.tex}}
-\newcommand{\LX}{\input{output/epic_Lx.tex}}
+\newcommand{\LXquiet}{\input{output/epic_Lx_quiet.tex}}
+\newcommand{\LXflaring}{\input{output/epic_Lx_flaring.tex}}
 \newcommand{\Lbol}{\input{output/Lbol.tex}}
@@
 
-We fitted the soft X-ray light curve using constant flux with a flare, applying the~\citet{davenport2014kepler} optical flare model that captures the fast rise and exponential decay of the flare seen with XMM-Newton.
-From the fit, we extracted the baseline flux against which we measured the equivalent duration $ED$ of the flare, i.e., the flare flux $F_{\rm flare}$, divided by the median flux $F_0$ of the star, integrated over the flare duration~\citep{gershberg1972results}:
+For the OM flare, we median flux of the lightcurve as the quiescent flux $F_0$, against which we measured the equivalent duration $ED$ of the flare, i.e., the flare flux $F_{\rm flare}$, divided by $F_0$, integrated over the flare duration~\citep{gershberg1972results}:
 
@@
 
-We then converted $ED$ to flare energy in the $0.2-2\,$keV band by multiplying with $L_X$.
-
-For the OM flare, we derived the bolometric flare energy following the procedure in~\citet{shibayama2013superflares}.
+We then converted $ED$ to flare energy using the bolometric flare energy following the procedure in~\citet{shibayama2013superflares}.
 We used the throughput curve for the white light filter scaled to unity at the peak of transmission as an optimistic response curve $R_{\lambda, \rm OM}$, given that the degradation of this filter's response is poorly constrained\footnote{XMM-Newton User's Handbook 3.5.3.1, \url{https://xmm-tools.cosmos.esa.int/external/xmm_user_support/documentation/uhb/omfilters.html}, accessed Aug 8, 2023}.
@@
 
+For the X-ray flare, we use the flare duration $\Delta t$ marked in the bottom panel in Fig.~\ref{fig:lightcurves}, and multiply it by the flare luminosity in the EPIC data.
+The flare luminosity is equal to the difference between the quiescent and flaring X-ray luminosities in the 0.2-2 keV band, i.e.:
+
+\begin{equation}
+    E_{X, \mathrm{flare}} = \left(L_{X,\mathrm{flaring}} -  L_{X,\mathrm{quiescent}}\right) \times \Delta t
+    \label{eq:xrayflare}
+\end{equation}
+
 \section{Results}
@@
 However, the emission measure weighted mean temperature $T_{\rm mean}$ changed from \Tqmean in the quiescent spectrum to \Tfmean in the flaring spectrum.
-From the quiescent PN spectra, we then also calculated soft X-ray flux of $F_X$\FX, and a luminosity $L_X=$\LX in the $0.2$-$2\,$keV band.
+From the quiescent PN spectra, we then also calculated soft X-ray flux of $F_X$\FX, and a luminosity $L_X=$\LXquiet in the $0.2$-$2\,$keV band.
 %This suggests that the flaring corona emits the hotter component, while the cooler component corresponds to the quiescent corona.
@@
 
-For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of \eepic in the $0.2-5\,$keV band.
+For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of \eepic in the $0.2-2\,$keV band, using $L_{X,\rm quiescent}=$\LXquiet\unskip, $L_{X,\rm flaring}=$\LXflaring\unskip, and $\Delta t =$\xraydur\unskip in Eq.~\ref{eq:xrayflare}.
 The energy of the corresponding OM flare is \eom\unskip.

### 2023-09-15 src/scripts/FIGURE_lightcurves.py
updated flare duration to the marked value

@@ -41,7 +41,7 @@ if __name__ == "__main__":
     # add vertical filled area
     ax[1].axvspan(776075500 / 3600. / 24., 776080000 / 3600. / 24.,
                   color="grey", alpha=0.2)
-    
+
     # add inset
     axins = ax[1].inset_axes([0.3, 0.55, 0.4, 0.5])
     l, r = 600, 750

### 2023-09-15 src/scripts/VALUES.py

@@ -189,3 +189,15 @@ if __name__ == "__main__":
     with open(paths.output / "om_flare.tex", "w") as f:
         f.write(omstr)
 
+    # flare duration in EPIC ---------------------------------------------------
+
+    tstart, tstop = epic.tstart, epic.tstop
+
+    dur = tstop - tstart
+
+    durstr = f"${dur/1e3:.1f}\,$ks"
+
+    print(durstr)
+
+    with open(paths.output / "epic_flare_dur.tex", "w") as f:
+        f.write(durstr)

### 2023-09-15 src/scripts/TABLE_specfits.py
added luminosities

@@ -83,10 +83,27 @@ if __name__ == "__main__":
 
     df = pd.read_csv(paths.data / "mcmc_results.csv", index_col=0)
 
+    # load luminosity from PN data
+    lxs = pd.read_csv(paths.data / "joint_vapec_chain_fits.csv")
+
+    lxs[r"$L_X$ [$10^{26}$ erg s$^{-1}$]"] = lxs.apply(lambda x: fr"${x.Lx_erg_s/1e26:.2f} [{x.Lx_erg_s_err/1e26:.2f}]$", axis=1) 
+
+    # rename subset in lxs as follows: spec to full data set, no flare to quiescent, onlyflare to flaring
+    lxs["subset"] = lxs["subset"].replace({"spec": "full data set", "noflare": "quiescent", "onlyflare": "flaring"})
+
+
+    # merge the subset, and LX column in lxs with df
+    df = df.reset_index()
+    df = df.merge(lxs[["subset", r"$L_X$ [$10^{26}$ erg s$^{-1}$]"]], right_on="subset", left_on="index")
+    df = df.set_index("index")
+
     # remove extra columns
     del df["symb"]
     del df["norm_ratio"]
     del df["e_norm_ratio"]
+    del df["subset"]
+
+    print(df.index)
 
     for c in df.columns:
         if "_err" in c:
@@ -126,6 +143,7 @@ if __name__ == "__main__":
     string = string.replace("midrule","hline")
     string = string.replace("toprule","hline")
     string = string.replace("bottomrule","hline")
+    string = string.replace("index","")
    
 
     # write to file

### 2023-09-15 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \newcommand{\eepic}{\input{output/epic_flare.tex}}
+\newcommand{\xraydur}{\input{output/epic_flare_dur.tex}}
 \newcommand{\Tqmean}{\input{output/Tqmean.tex}}
@@
 
-For the X-ray flare, we use the flare duration $\Delta t$ marked in the bottom panel in Fig.~\ref{fig:lightcurves}, and multiply it by the flare luminosity in the EPIC data.
+For the X-ray flare energy, we multiply flare duration $\Delta t$ by the flare luminosity in the EPIC data.
 The flare luminosity is equal to the difference between the quiescent and flaring X-ray luminosities in the 0.2-2 keV band, i.e.:
@@
 \begin{equation}
-    E_{X, \mathrm{flare}} = \left(L_{X,\mathrm{flaring}} -  L_{X,\mathrm{quiescent}}\right) \times \Delta t
+    E_{X, \mathrm{flare}} = \left(L_{X,\mathrm{flaring}} -  L_{X,\mathrm{quiescent}}\right) \cdot \Delta t
     \label{eq:xrayflare}
@@
 However, the emission measure weighted mean temperature $T_{\rm mean}$ changed from \Tqmean in the quiescent spectrum to \Tfmean in the flaring spectrum.
-From the quiescent PN spectra, we then also calculated soft X-ray flux of $F_X$\FX, and a luminosity $L_X=$\LXquiet in the $0.2$-$2\,$keV band.
+From the quiescent PN spectra, we then also calculated soft X-ray flux of $F_X=$\FX, and a luminosity $L_X=$\LXquiet in the $0.2$-$2\,$keV band.
 %This suggests that the flaring corona emits the hotter component, while the cooler component corresponds to the quiescent corona.
@@
     \script{TABLE_specfits.py}
-    \caption{XSPEC fits to EPIC spectra for different subsets of observations.}
+    \caption{XSPEC fits to EPIC spectra for different subsets of observations. Fluxes and luminosities are given in the $0.2-2\,$keV band.}
     \input{output/mcmc_specfit.tex}
@@
 
-For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of \eepic in the $0.2-2\,$keV band, using $L_{X,\rm quiescent}=$\LXquiet\unskip, $L_{X,\rm flaring}=$\LXflaring\unskip, and $\Delta t =$\xraydur\unskip in Eq.~\ref{eq:xrayflare}.
+For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of \eepic in the $0.2-2\,$keV band, using $L_{X,\rm quiescent}=$\LXquiet\unskip, $L_{X,\rm flaring}=$\LXflaring\unskip, and $\Delta t =$\xraydur\unskip (see Fig.~\ref{fig:lightcurves}) in Eq.~\ref{eq:xrayflare}.
 The energy of the corresponding OM flare is \eom\unskip.

### 2023-09-18 src/scripts/TABLE_specfits.py
added flux to table

@@ -86,15 +86,26 @@ if __name__ == "__main__":
     # load luminosity from PN data
     lxs = pd.read_csv(paths.data / "joint_vapec_chain_fits.csv")
 
-    lxs[r"$L_X$ [$10^{26}$ erg s$^{-1}$]"] = lxs.apply(lambda x: fr"${x.Lx_erg_s/1e26:.2f} [{x.Lx_erg_s_err/1e26:.2f}]$", axis=1) 
+    # get flux and luminosity
+    g = lambda x: fr"${x.flux_erg_s_cm2/1e-14:.2e} [{x.flux_erg_s_cm2_err/1e-14:.2e}]$"
+    lxs[r"$F_X$ [$10^{-14}$ erg s$^{-1}$ cm$^{-2}$]"] = lxs.apply(g, axis=1)
 
-    # rename subset in lxs as follows: spec to full data set, no flare to quiescent, onlyflare to flaring
-    lxs["subset"] = lxs["subset"].replace({"spec": "full data set", "noflare": "quiescent", "onlyflare": "flaring"})
+    g = lambda x: fr"${x.Lx_erg_s/1e26:.2f} [{x.Lx_erg_s_err/1e26:.2f}]$"
+    lxs[r"$L_X$ [$10^{26}$ erg s$^{-1}$]"] = lxs.apply(g, axis=1) 
 
+    # rename subset in lxs as follows: spec to full data set, 
+    # no flare to quiescent, onlyflare to flaring
+    lxs["subset"] = lxs["subset"].replace({"spec": "full data set", 
+                                           "noflare": "quiescent", 
+                                           "onlyflare": "flaring"})
 
     # merge the subset, and LX column in lxs with df
     df = df.reset_index()
-    df = df.merge(lxs[["subset", r"$L_X$ [$10^{26}$ erg s$^{-1}$]"]], right_on="subset", left_on="index")
+    df = df.merge(lxs[["subset",
+                       r"$L_X$ [$10^{26}$ erg s$^{-1}$]",
+                       r"$F_X$ [$10^{-14}$ erg s$^{-1}$ cm$^{-2}$]"]], 
+                       right_on="subset", left_on="index")
+    
     df = df.set_index("index")
 
     # remove extra columns
@@ -102,8 +113,6 @@ if __name__ == "__main__":
     del df["norm_ratio"]
     del df["e_norm_ratio"]
     del df["subset"]
-
-    print(df.index)
 
     for c in df.columns:
         if "_err" in c:
[…]

### 2023-09-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \begin{abstract}
-    In 2020, the Transiting Exoplanet Survey Satellite (TESS) observed a rapidly rotating M7 dwarf produce a flare at 81 deg latitude, the highest latitude flare located to date.
-This is in stark contrast to solar flares that occur much closer to the equator, typically below 30 deg.
+    In 2020, the \textit{Transiting Exoplanet Survey Satellite} (\textit{TESS}) observed a rapidly rotating M7 dwarf, TIC~277539431, produce a flare at $81^{\circ}$ latitude, the highest latitude flare located to date.
+This is in stark contrast to solar flares that occur much closer to the equator, typically below $30^{\circ}$.
 The circumstances that allow flares at high latitudes to occur are poorly understood.
-We studied five Sectors of TESS monitoring, and obtained 36 ks of XMM-Newton observations, to find that TIC 277's corona does not differ significantly from other low mass stars on the saturated activity branch with respect to coronal temperatures and flaring activity, but is underluminous in soft X-ray emission by about an order of magnitude.
-Combining the lack of X-ray flux, the high latitude flare, the star's viewing geometry, and the otherwise typical stellar corona, we suggest that the migration of flux emergence to the poles in rapid rotators can explain the observations of TIC~277.
+We studied five Sectors of \textit{TESS} monitoring, and obtained 36 ks of \textit{XMM-Newton} observations, to find that TIC~277539431's corona does not differ significantly from other low mass stars on the saturated activity branch with respect to coronal temperatures and flaring activity, but is underluminous in soft X-ray emission by about an order of magnitude.
+Combining the lack of X-ray flux, the high latitude flare, the star's viewing geometry, and the otherwise typical stellar corona, we suggest that the migration of flux emergence to the poles in rapid rotators can explain the observations of TIC~277539431.
 
@@
 However, to be habitable in an Earth-like manner, merely the possibility of surface water in its liquid form does not suffice.
-Space weather of the host, that is, the high energy radiation and particles that force the planet's atmosphere affects hospitality for life, for M dwarfs in particular~\citep{airapetian2020impact}.
-If the energetic (particle) flux is too high, the atmosphere may blow off~\citep[e.g., ][]{lammer2003atmospheric, garraffo2017threatening, ketzer2022influence} and water may be lost from the surface~\citep{doamaral2022contribution}.
+Space weather of the host, that is, the high energy radiation and particles that impact the planet's atmosphere, affect its hospitality for life, for M dwarfs in particular~\citep{airapetian2020impact}.
+If the energetic (photon or particle) flux is too high, the atmosphere may blow off~\citep[e.g., ][]{lammer2003atmospheric, garraffo2017threatening, ketzer2022influence} and water may be lost from the surface~\citep{doamaral2022contribution}.
 If not energetic enough, life may not emerge in the first place~\citep{rimmer2018origin}.
 
-In M dwarfs, stellar activity evolution on the main sequence unfolds much slower than for an FGK star.
+In M dwarfs, stellar activity evolution on the main sequence unfolds much more slowly than for an FGK star.
 While for the latter, magnetic activity, and with it the flares, winds, coronal mass ejections and energetic particle eruptions that make up their space weather, decline rapidly over a few hundred megayears, a fully convective M dwarf stays highly active for gigayears~\citep{johnstone2021active, medina2022galactic}.
@@
 The simultaneous surge of radio emission indicates a transition from stellar corona to planet-like magnetosphere at the bottom of the main sequence~\citep{zarka1998auroral,pineda2017panchromatic}.
-Yet, brown dwarfs are regularly detected with energetic flares regardless of their expected low X-ray luminosity~\citep[e.g., ][]{hambaryan2004new, gizis2013kepler, paudel2020k2, schmidt2019largest}.
+Yet, brown dwarfs are regularly detected with energetic flares regardless of their expected low X-ray luminosity~\citep[e.g., ][]{hambaryan2004new, stelzer2006simultaneous, gizis2013kepler, paudel2020k2, schmidt2019largest}.
 How these flares are produced in the apparent absence of a solar-type corona is unclear~\citep{mullan2018frequencies}.%.
@@
 
-\citet{ilin2021giant} searched fully convective dwarfs that straddle the transition from star to brown dwarf with TESS.
+\citet{ilin2021giant} searched fully convective dwarfs that straddle the transition from star to brown dwarf with \textit{TESS}.
 They found four flares on four rapidly rotating ($P_{\rm rot}<10\,$h) M5-M7 dwarfs that lasted for multiple rotation periods of each star.
@@
 The shape of this modulation together with known inclination (combining $P_{\rm rot}$ and $v \sin i$ from high resolution spectroscopy) allowed them to determine the latitudes of these flares.
-Surprisingly, these flares were found significantly closer to the pole than to the equator, in contrast to the Sun, where flares are usually found below 30 deg latitude.
+Surprisingly, these flares were found significantly closer to the pole than to the equator, in contrast to the Sun, where flares are usually found below $30^{\circ}$ latitude.
 Their results indicate a preference for flares at high latitudes in these stars, as a chance finding was unlikely ($\sim 0.1\%$).
@@
 
-TIC 277539431, TIC 277 for short, showed the highest latitude flare known to date at $\sim81\,$deg~\citep[][Table \ref{tab:starparams}]{ilin2021giant}.
+TIC 277539431, TIC 277 for short, showed the highest latitude flare known to date at $\sim81^{\circ}$~\citep[][Table \ref{tab:starparams}]{ilin2021giant}.
 It is an M7 dwarf with an extremely short rotation period of $4.56\,$h.
@@
 
-In this work, we follow TIC 277 with XMM-Newton for 36 ks and use optical monitoring from TESS~(Section~\ref{sec:obs}) to measure its coronal and flaring properties, respectively~(Section~\ref{sec:methods}).
+In this work, we follow TIC 277 with \textit{XMM-Newton} for 36 ks and use optical monitoring from \textit{TESS}~(Section~\ref{sec:obs}) to measure its coronal and flaring properties, respectively~(Section~\ref{sec:methods}).
 We contrast our results~(Section~\ref{sec:results}) with the literature to investigate if TIC~277 behaves like a low mass star or brown dwarf, investigate scenarios that are consistent with high latitude flare occurrence ~(Section~\ref{sec:discussion}), and conclude in Section~\ref{sec:summary}.
@@
 \begin{table}
-% \footnotesize
+\footnotesize
 \centering
-    \caption{TIC 277539431 (WISEA J105515.71-735611.3)}
+    \caption{Stellar properties of TIC 277539431}
     \begin{tabular}{ll}\hline 
@@
          % $J$ mag & $10.630 \pm 0.023$ [3] \\
+         Alternative ID & WISEA J105515.71-735611.3 \\
          Spectral type & M7 [1]\\
@@
          Rotation period $P_{\rm rot}$ & $273.618 \pm 0.007\,$min [1]\\
-         Projected rotation velocity $v\sin i$ & $38.6\pm1.0\,$km/s [1] \\
-         Inclination $i$ & $87.0^{+2.0}_{-2.4}\,$ deg [1]\\
+         Projected rot. velocity $v\sin i$ & $38.6\pm1.0\,$km/s [1] \\
+         Inclination $i$ & $(87.0^{+2.0}_{-2.4})^{\circ}$ [1]\\
          Radius $R$ & $0.145\pm0.004\,R_\odot$ [1]\\
-         Bol. luminosity $L_{\rm bol}$ & \Lbol \\\hline
+         Bol. luminosity $L_{\rm bol}$ & \Lbol [*]\\\hline
 
@@
     \newline\footnotesize
+    \flushleft
     [1] \citet{ilin2021giant}, 
@@
     [3] \citet{pecaut2013intrinsic}
+    [*] see Section~\ref{sec:methods:flareenergies}
     \label{tab:starparams}
@@
 \label{sec:obs:xmm}
-XMM-Newton is an X-ray telescope, which launched into orbit in 1999, equipped with six science instruments that operate simultaneously, among them the European Photon Imaging Camera with three cameras, PN, MOS1 and MOS2; and the Optical Monitor (OM) that provides near-UV and optical photometry, among other capabilities.
-We used XMM-Newton to observe TIC 277 for 36 ks, i.e.
+\textit{XMM-Newton} is an X-ray telescope, which launched into orbit in 1999, equipped with six science instruments that operate simultaneously, among them the European Photon Imaging Camera with three cameras, PN, MOS1 and MOS2; and the Optical Monitor (OM) that provides near-UV and optical photometry, among other capabilities.
+We used \textit{XMM-Newton} to observe TIC 277 for 36 ks, i.e.
 two full rotation periods of the star, on August 5, 2022 (Proposal 090120; PI: E.
 Ilin).
-We used the EPIC instruments aboard XMM-Newton, i.e., MOS1/2 and PN, that cover the soft X-ray band from $0.2$ and up to $12\,$keV, as well as the OM using its broad band white light filter.
+We used the EPIC instruments aboard \textit{XMM-Newton}, i.e., MOS1/2 and PN, that cover the soft X-ray band from $0.2$ and up to $12\,$keV, as well as the OM using its broad band white light filter.
 
@@
 Fig.~\ref{fig:lightcurves} shows the time series using the combined PN and MOS time series together with the OM light curve.
-We extracted the events time series from EPIC using XMM SAS version 20~\footnote{"Users Guide to the XMM-Newton Science Analysis System", Issue 18.0, 2023 (ESA: XMM-Newton SOC)} using the \texttt{evselect} task.
+We extracted the events time series from EPIC using XMM SAS version 20~\footnote{"Users Guide to the \textit{XMM-Newton} Science Analysis System", Issue 18.0, 2023 (ESA: \textit{XMM-Newton} SOC)} using the \texttt{evselect} task.
 We selected a circular source region with 20 arcsec radius on all detectors, and a circular source-free background region 120 and 90 arcsec in MOS and PN, respectively.
@@
 
-Simultaneously, we used the optical monitor (OM) aboard XMM-Newton, to monitor TIC 277 in its white light filter at $10$\,s cadence.
+Simultaneously, we used the optical monitor (OM) aboard \textit{XMM-Newton}, to monitor TIC 277 in its white light filter at $10$\,s cadence.
 The OM white light filter is centered on $4,060\,$\AA, with an equivalent width of $3,470\,$\AA.
 We used the \texttt{omfchain} in XMM SAS to extract the 8 individual light curves.
+In each light curve, OM observed uninterruptedly for $4390\,$s, followed by a short gap of $318\,$s before the next one.
 
@@
 \subsubsection{EPIC Spectra}
-From the EPIC observations, we used the standard spectral extraction procedure for point sources.
-We extracted spectra using \texttt{evselect} to select all events in the circular event and background regions, \texttt{epatplot} to check for pile-up of multiple photons in one exposure, \texttt{backscale} to calculate the area of the source region, and \texttt{rmfgen} and \texttt{arfgen} to convolve the observed energies with the instrument response, and produce the spectra.
+From the EPIC observations, we used the standard spectral extraction procedure for point sources, as specified in the XMM SAS handbook.
+We extracted spectra using \texttt{evselect} to select events in the circular event and background regions, \texttt{epatplot} to check for pile-up of multiple photons in one exposure, \texttt{backscale} to calculate the area of the source region, and \texttt{rmfgen} and \texttt{arfgen} to convolve the observed energies with the instrument response, and produce the spectra.
 Finally, we use \texttt{grppha} to rebin the data in the spectra to obtain at least 15 counts per bin.
@@
 \label{sec:obs:tess}
-Since 2018, TESS has been supplying publicly available red-optical high-precision time series photometry in an ongoing all-sky survey.
-Each uninterrupted observing Sector provides an approximately 27 day long light curve in a broad $6,000$-$10,000\AA$ band pass.
+Since 2018, \textit{TESS} has been supplying publicly available red-optical high-precision time series photometry in an ongoing all-sky survey.
+Each uninterrupted observing Sector provides an approximately 27 day long light curve in a broad $6,000$-$10,000$\AA band pass.
 
@@
 We used AltaiPony~\citep{ilin2021altaipony} to de-trend the light curves using the custom de-trending function \texttt{custom\_detrending}\footnote{see also the documentation online at \url{https://altaipony.readthedocs.io/en/latest/tutorials/detrend.html}, accessed on July 3, 2023}, detailed in~\citet{ilin2022searching}.
-The procedure includes 3rd order spline fitting of coarse trends, followed by iterative sine fitting to remove rotational modulation, and two Savitzky-Golay filters~\citep{savitzky1964smoothing} with a 6h and 3h window, respectively to remove aperiodic short term trends, while masking a.
-outliers above $>2.5\sigma$, b.
-the six-fold of consecutive originally masked data points following the original mask, and c.
-additional three adjacent data points before and after the so expanded mask.
+The procedure includes 3rd order spline fitting of coarse trends, followed by iterative sine fitting to remove rotational modulation, and two Savitzky-Golay filters~\citep{savitzky1964smoothing} with a 6h and 3h window, respectively to remove aperiodic short term trends, while masking potential flare candidates in each step.
+The masking takes three steps each time.
+First, all outliers above $>2.5\sigma$ are masked.
+Second, the six-fold of the number of data points masked in the first step are also masked.
+Third, three adjacent data points before and after the so expanded mask are added to the mask.
 
-In the de-trended light curves, we searched for flare candidates as at least 3 consecutive $>3\sigma$ outliers above the noise level in the masked light curve, filling in the noise level inside the masked areas with the mean of the noise levels adjacent to the mask.
+In the de-trended light curves, we searched for flare candidates as at least 3 consecutive $>3\sigma$ outliers above the noise level, filling in the noise level inside the previously masked areas with the mean of the noise levels adjacent to the mask.
 To the series of outliers, we kept adding data points to the flare candidate until a data point falls below $2\sigma$.
@@
         \caption{
-         Normalized TESS light curves.
+         Normalized \textit{TESS} light curves.
 Black and red dots show the light curve with and without rotational variability and trends, respectively.
 The de-trended light curve is offset by $0.2$.
-The large, rotationally modulated, flare, localized at about 81 deg latitude by~\citet{ilin2021giant}, appears in the second half of the top panel~(Sector 12).}
+The large, rotationally modulated, flare, localized at about $81^{\circ}$ latitude by~\citet{ilin2021giant}, appears in the second half of the top panel~(Sector 12).}
         \label{fig:tess_lcs}
@@
 \label{sec:methods:epic}
+The stellar corona can be described as an optically thin thermal plasma that can consist of one or multiple temperature components that represent different regions.
 We use XSPEC version 12.12.1~\citep{arnaud1996xspec} to fit a two-temperature (2T) additive APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/0$=0.6$ ratio, which is more typical of M dwarfs~\citep{wood2018chandra}.
@@
 \subsection{Flare energies: TESS, OM, and EPIC}
-
+\label{sec:methods:flareenergies}
 For the OM flare, we median flux of the lightcurve as the quiescent flux $F_0$, against which we measured the equivalent duration $ED$ of the flare, i.e., the flare flux $F_{\rm flare}$, divided by $F_0$, integrated over the flare duration~\citep{gershberg1972results}:
@@
 We then converted $ED$ to flare energy using the bolometric flare energy following the procedure in~\citet{shibayama2013superflares}.
-We used the throughput curve for the white light filter scaled to unity at the peak of transmission as an optimistic response curve $R_{\lambda, \rm OM}$, given that the degradation of this filter's response is poorly constrained\footnote{XMM-Newton User's Handbook 3.5.3.1, \url{https://xmm-tools.cosmos.esa.int/external/xmm_user_support/documentation/uhb/omfilters.html}, accessed Aug 8, 2023}.
+We used the throughput curve for the white light filter scaled to unity at the peak of transmission as an optimistic response curve $R_{\lambda, \rm OM}$, given that the degradation of this filter's response is poorly constrained\footnote{\textit{XMM-Newton} User's Handbook 3.5.3.1, \url{https://xmm-tools.cosmos.esa.int/external/xmm_user_support/documentation/uhb/omfilters.html}, accessed Aug 8, 2023}.
 From that, we extracted a ratio $f_{\rm OM}$ of stellar to flare flux of $1.4\cdot 10^{-4}$ for a $10,000\,$K blackbody flare\footnote{A flare temperature of $T_{\rm flare}=10,000\,$K is a typical approximation for energetic M dwarf flares~\citep{kowalski2013timeresolved, howard2020evryflarea}.}:
@@
 
-Analogously to Eq.~\ref{eq:ratio} for the OM flare, we used the TESS response curve to obtain a flux ratio $f_{\rm TESS}$ of $6.3\cdot10^{-3}$ in the redder optical filter of TESS, and convert $ED$ to bolometric flare energy using Eq.~\ref{eq:eflare}.
+Analogously to Eq.~\ref{eq:ratio} for the OM flare, we used the \textit{TESS} response curve to obtain a flux ratio $f_{\rm TESS}$ of $6.3\cdot10^{-3}$ in the redder optical filter of \textit{TESS}, and convert $ED$ to bolometric flare energy using Eq.~\ref{eq:eflare}.
 Besides the $ED$, \texttt{AltaiPony} also yields start and end time, defined as the difference between the first and last data point above the flare detection criterion defined in Section~\ref{sec:obs:tess}; and relative amplitude $a$, for each flare.
@@
 \label{sec:results}
-With the XMM-Newton and TESS observations, we could constrain the coronal and flaring properties of TIC 277.
+With the \textit{XMM-Newton} and \textit{TESS} observations, we could constrain the coronal and flaring properties of TIC 277.
 We derive its coronal temperature and luminosity from the EPIC spectrum~(Section~\ref{sec:res:XrayTL}).
-We also measure the flare energies in X-ray from the EPIC instruments, and optical from the OM, and compare them to the flare frequency distributions obtained from the 18 detected flares in TESS~(Section~\ref{sec:res:flares}).
+We also measure the flare energies in X-ray from the EPIC instruments, and optical from the OM, and compare them to the flare frequency distributions obtained from the 18 detected flares in \textit{TESS}~(Section~\ref{sec:res:flares}).
 
@@
 
-From the XMM-Newton EPIC spectra, we could measure the coronal properties of TIC in both its quiescent, and flaring states~(Table~\ref{tab:specfit}).
+From the \textit{XMM-Newton} EPIC spectra, we could measure the coronal properties of TIC~277 in both its quiescent, and flaring states~(Table~\ref{tab:specfit}).
 The resulting fit~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about \Tcool and \Thot, respectively.
 Assuming these temperatures as ground truth, we also fitted 2T APEC models to the flaring and quiescent portions of the observations separately.
-The coronal temperatures did not update significantly using either subset, although the quiescent spectrum appeared slightly cooler overall.
-However, the emission measure weighted mean temperature $T_{\rm mean}$ changed from \Tqmean in the quiescent spectrum to \Tfmean in the flaring spectrum.
+The coronal temperatures did not update significantly using either subset.
+However, the emission measure weighted mean temperature $T_{\rm mean}$ changed from \Tqmean in the quiescent spectrum to \Tfmean in the flaring spectrum, since the hot component became more much dominant during the flare.
 From the quiescent PN spectra, we then also calculated soft X-ray flux of $F_X=$\FX, and a luminosity $L_X=$\LXquiet in the $0.2$-$2\,$keV band.
@@
 
-We note that not using the full set of observations as ground truth resulted in an unconstrained hot coronal component.
+We note that not using the full set of observations as ground truth resulted in an unconstrained hot coronal component during the flare.
 This can be interpreted as due to a lack of X-ray flux above $5\,$keV.
@@
     \script{TABLE_tess_flares.py}
-    \caption{Flares detected with TESS.
+    \caption{Flares detected with \textit{TESS}.
 $t_s$ is the starting time of the flare, $a$ is its relative amplitude, and $E_{\rm bol}$ is the bolometric energy assuming a 10,000 K blackbody emission from the flare.}
@@
 
-In TESS, we found a total of 18 flares in the five Sectors, or equivalently, 125 days of observing at 2-min cadence. We fit the flare frequency distribution (FFD) with a power law of the form
+In \textit{TESS}, we found a total of 18 flares in the five Sectors, or equivalently, 125 days of observing at 2-min cadence. We fit the flare frequency distribution (FFD) with a power law of the form
 
@@
 The energy of the corresponding OM flare is \eom\unskip.
-We use the 36ks observing baseline of XMM-Newton to calculate a flare rate in OM.
-Fig.~\ref{fig:ffd} illustrates that it is roughly consistent, if slightly higher, than the extension of TESS' flare frequency distributions to lower energies.
+We use the 36ks observing baseline of \textit{XMM-Newton} to calculate a flare rate in OM.
+Fig.~\ref{fig:ffd} illustrates that it is roughly consistent, if slightly higher, than the extension of \textit{TESS}' flare frequency distributions to lower energies.
 About one flare per day of the order of the observed one or above can therefore be expected in future X-ray observations of TIC~277.
 However, this assumes that all optical flares above this energy in the optical data correspond to detected X-ray flares in the EPIC instruments, which is not always the case~\citep{paudel2021simultaneous}.
-\citet{guarcello2019simultaneous, kuznetsov2021stellar} find moderate correlation between flare energies in Kepler/K2 data and XMM-Newton, but the majority of their sampled flare energies are two orders of magnitude above the flare in this work, both in X-ray and optical.
-\citet{kuznetsov2021stellar} compare primary Kepler mission observations with XMM-Newton to find a similar result.
-We note that the X-ray flare light curve follows the typical fast-rise exponential decay shape, and is therefore unlikely eclipsed by the star~\citep{johnstone2012soft}.
+\citet{guarcello2019simultaneous, kuznetsov2021stellar} find moderate correlation between flare energies in Kepler/K2 data and \textit{XMM-Newton}, but the majority of their sampled flare energies are two orders of magnitude above the flare in this work, both in X-ray and optical.
+\citet{kuznetsov2021stellar} compare primary Kepler mission observations with \textit{XMM-Newton} to find a similar result.
+We note that the X-ray flare light curve follows the typical fast-rise exponential decay shape, and is therefore unlikely to be eclipsed by the star~\citep{johnstone2012soft}.
 
@@
         \caption{
-         Cumulative flare frequency distribution of TESS flares (black dots), with a power law fit (green line), and uncertainties (dotted lines).
-The rate of OM flares (gray square) is calculated from the observing baseline of XMM-Newton.
+         Cumulative flare frequency distribution of \textit{TESS} flares (black dots), with a power law fit (green line), and uncertainties (dotted lines).
+The rate of OM flares (gray square) is calculated from the observing baseline of \textit{XMM-Newton}.
         }
@@
 Its coronal properties are therefore not known a priori.
-Our findings indicate that TIC 277's coronal temperature~(Section~\ref{sec:discussion:xraytemp}), and flaring activity~(Section~\ref{sec:discussion:flares}) are typical of saturated fully convective M dwarf stars.
+Our findings indicate that TIC 277's coronal temperature~(Section~\ref{sec:discussion:xraytemp}) and flaring activity~(Section~\ref{sec:discussion:flares}) are typical of saturated fully convective M dwarf stars.
 However, TIC 277 is rotating rapidly, even for the typically fast rotating late M dwarfs~\citep{medina2022galactic}, which is more typical of brown dwarfs.
@@
 
-TIC~277's flare rate is consistent with other saturated fully convective dwarfs~\citep{medina2020flare, murray2022study}, and so is the energy distribution with a power law slope of \ffdalpha.
+TIC~277's flare rate is consistent with other saturated fully convective dwarfs~\citep{medina2020flare, murray2022study}, and so is the energy distribution with a power law slope of \ffdalpha\unskip.
 It is in fact comparable to TRAPPIST-1's flaring behavior~\citep{paudel2018k2}, which has been investigated for its effects on the seven terrestrial planets in its orbit.
@@
 However, the rotation periods are unknown for most of these late M dwarfs.
-Sometime projected rotation periods are known~\citep{cook2014trends}, but true periods are only detected in a handful of these stars, which we list in Table~\ref{tab:latmlxlbol}.
+Sometimes projected rotation periods are known~\citep{cook2014trends}, but true periods are only detected in a handful of these stars, which we list in Table~\ref{tab:latmlxlbol}.
 TIC~277 appears at the lower end of this small sample.
@@
 However, since the flaring activity and coronal temperature of TIC~277 are in line with saturated activity, this conjecture lacks evidence from these other activity indicators.
-An alternative explanation, that would at the same time explain the occurrence of a high latitude flare in TESS, is polar updraft migration~\citep{stepien2001rosat}.
+An alternative explanation, that would at the same time explain the occurrence of a high latitude flare in \textit{TESS}, is polar updraft migration~\citep{stepien2001rosat}.
 Polar updraft migration implies that at high rotation rates, flux emergence becomes more efficient near the poles than at the equator~\citep{weber2016modeling}, draining the equatorial regions of magnetic flux required to produce a corona there.
@@
 \label{sec:summary}
-We investigated whether TIC 277, a rapidly rotating M7 dwarf with a flare localized at 81 deg latitude, shows unusual coronal or flaring properties that could explain the occurrence of this flare so close to the rotational pole.
-We obtained 36 ks of EPIC and OM observations with XMM-Newton, and studied the five Sectors of red-optical 27-day light curves provided by TESS.
+We investigated whether TIC 277, a rapidly rotating M7 dwarf with a flare localized at $81^{\circ}$ latitude, shows unusual coronal or flaring properties that could explain the occurrence of this flare so close to the rotational pole.
+We obtained 36 ks of EPIC and OM observations with \textit{XMM-Newton}, and studied the five Sectors of red-optical 27-day light curves provided by \textit{TESS}.
 We found a mean quiescent coronal temperature of \Tqmean, and a flaring rate $f(> \log_{10} E_{\rm flare, bol} = 31.5\,\rm{erg})\approx$\ffdr with energy distribution with a slope of \ffdalpha, all typical of fully convective M dwarfs in the saturated regime.
-We also detected a simultaneous X-ray and optical flare with an energy of \eepic in the $0.2-5\,$keV band.
-In combination with the 18 detected TESS flares, we estimate that X-ray flares on TIC~277 can be observed with XMM-Newton at least about once a day.
+We also detected a simultaneous X-ray and optical flare with an energy of \eepic in the $0.2-2\,$keV band, and bolometric energy of \eom from the OM observations.
+In combination with the 18 detected \textit{TESS} flares, we estimate that X-ray flares on TIC~277 can be observed with \textit{XMM-Newton} at least about once a day.
 %The bolometric and soft-X-ray flare energies are within an order of magnitude within each other, which places them among other stars with simultaneous X-ray and optical flare detections, although, typically, the white light emission dominates over soft X-ray~\citep{guarcello2019simultaneous, kuznetsov2021stellar}.
@@
 TIC 277 is an extreme rapid rotator with a rotation period of only $4.56\,$h.
-The detected high-latitude flare in TESS~\citep{ilin2021giant} may hence be a product of high-latitude flux emergence driven by its fast rotation~\citep{weber2016modeling,weber2017suppression}.
+The detected high-latitude flare in \textit{TESS}~\citep{ilin2021giant} may hence be a product of high-latitude flux emergence driven by its fast rotation~\citep{weber2016modeling,weber2017suppression}.
 This effect is closely related to polar updraft migration, which suppresses coronal emission in equatorial regions~\citep{stepien2001rosat}.
@@
 acknowledges funding from the German \textit{Leibniz-Gemeinschaft} under project number P67/2018.
+This work is based on observations obtained with \textit{XMM-Newton}, an ESA science mission with instruments and contributions directly funded by ESA Member States and NASA.
 This work made use of the open source python software packages \texttt{lightkurve}\citep{lightkurvecollaboration2018lightkurve}, \texttt{astropy}~\citep{robitaille2013astropy}, \texttt{numpy}~\citep{harris2020array}, \texttt{pandas}~\citep{reback2022pandasdev}, \texttt{matplotlib}~\citep{hunter2007matplotlib}, \texttt{emcee}~\citep{foreman-mackey2013emcee}, \texttt{scipy}~\citep{mckinney2010data}, and \texttt{altaipony}~\citep{ilin2021altaipony}.
@@
 participating in the {\it Gaia} Multilateral Agreement.
-This work is based on observations obtained with XMM-Newton, an ESA science mission with instruments and contributions directly funded by ESA Member States and NASA.
+
 \section*{Data Availability}

### 2023-09-18 src/scripts/TABLE_specfits.py
formatting

@@ -87,7 +87,7 @@ if __name__ == "__main__":
     lxs = pd.read_csv(paths.data / "joint_vapec_chain_fits.csv")
 
     # get flux and luminosity
-    g = lambda x: fr"${x.flux_erg_s_cm2/1e-14:.2e} [{x.flux_erg_s_cm2_err/1e-14:.2e}]$"
+    g = lambda x: fr"${x.flux_erg_s_cm2/1e-14:.2f} [{x.flux_erg_s_cm2_err/1e-14:.2f}]$"
     lxs[r"$F_X$ [$10^{-14}$ erg s$^{-1}$ cm$^{-2}$]"] = lxs.apply(g, axis=1)
 
     g = lambda x: fr"${x.Lx_erg_s/1e26:.2f} [{x.Lx_erg_s_err/1e26:.2f}]$"

### 2023-09-18 src/scripts/TABLE_specfits.py
more formatting of table

@@ -28,7 +28,7 @@ def convert_to_scinote(series, rel_err=1e-2):
     pd.Series
         The converted series of strings.
     """
-    return series.apply(lambda x: f"{x:.1e}" if np.abs(x)<rel_err else f"{x:.2f}").astype(str)
+    return series.apply(lambda x: f"{x:.1e}" if np.abs(x)<rel_err else f"{x:.1f}").astype(str)
 
 
 
@@ -87,10 +87,10 @@ if __name__ == "__main__":
     lxs = pd.read_csv(paths.data / "joint_vapec_chain_fits.csv")
 
     # get flux and luminosity
-    g = lambda x: fr"${x.flux_erg_s_cm2/1e-14:.2f} [{x.flux_erg_s_cm2_err/1e-14:.2f}]$"
+    g = lambda x: fr"${x.flux_erg_s_cm2/1e-14:.1f} [{x.flux_erg_s_cm2_err/1e-14:.1f}]$"
     lxs[r"$F_X$ [$10^{-14}$ erg s$^{-1}$ cm$^{-2}$]"] = lxs.apply(g, axis=1)
 
-    g = lambda x: fr"${x.Lx_erg_s/1e26:.2f} [{x.Lx_erg_s_err/1e26:.2f}]$"
+    g = lambda x: fr"${x.Lx_erg_s/1e26:.1f} [{x.Lx_erg_s_err/1e26:.1f}]$"
     lxs[r"$L_X$ [$10^{26}$ erg s$^{-1}$]"] = lxs.apply(g, axis=1) 
 
     # rename subset in lxs as follows: spec to full data set, 

### 2023-09-18 src/scripts/FIGURE_data_resid.py
improved x-axis labeling and formatting of units

@@ -66,12 +66,13 @@ def plot_data_resid_3(file):
     # labels
         # labels
     ax[1].set_xlabel("E [keV]", fontsize=14)
-    ax[0].set_ylabel("flux [counts/s/keV]", fontsize=14)
-    ax[1].set_ylabel("residuals [counts/s/keV]", fontsize=14)
+    ax[0].set_ylabel(r"flux [counts s$^{-1}\,$keV$^{-1}$]", fontsize=13)
+    ax[1].set_ylabel(r"residuals [counts s$^{-1}\,$keV$^{-1}$]", fontsize=13)
 
     # increase tick size labels
     for a in ax:
         a.tick_params(axis='both', which='major', labelsize=13)
+    
 
     # y limits
     ax[0].set_ylim(0,)
@@ -87,6 +88,8 @@ def plot_data_resid_3(file):
 
         a.set_xscale("log")
 
+    ax[1].set_xticks(ticks=[0.25,0.4,0.6,0.8,1,2,3,4,5],
+                     labels=[0.25,0.4,0.6,0.8,1,2,3,4,5])
 
     # layout
     plt.tight_layout()
@@ -145,8 +148,8 @@ def plot_data_resid_1(file):
 
     # labels
     ax[1].set_xlabel("E [keV]", fontsize=15)
-    ax[0].set_ylabel("flux [counts/s/keV]", fontsize=15)
-    ax[1].set_ylabel("residuals [counts/s/keV]", fontsize=15)
+    ax[0].set_ylabel(r"flux [counts s$^{-1}$ keV^$^{-1}$]", fontsize=15)
+    ax[1].set_ylabel(r"residuals [counts s$^{-1}$ keV^$^{-1}$]", fontsize=15)
 
     # increase tick size labels
     ax[0].tick_params(axis='both', which='major', labelsize=15)

### 2023-09-18 src/scripts/TABLE_tess_flares.py
flare table energy remove one digit

@@ -26,9 +26,9 @@ if __name__ == "__main__":
     sel["ed_rec_err"] = sel["ed_rec_err"] / 1e31
 
     # format
-    sel[r"$E_{\rm bol}$ [$10^{31}$ erg]"] = (sel["ed_rec"].round(2).astype(str) + 
+    sel[r"$E_{\rm bol}$ [$10^{31}$ erg]"] = (sel["ed_rec"].round(1).astype(str) + 
                                             " [" + 
-                                            sel["ed_rec_err"].round(2).astype(str) + 
+                                            sel["ed_rec_err"].round(1).astype(str) + 
                                             "]")
     # delete unused columns
     del sel["ed_rec_err"]

### 2023-09-19 src/scripts/FIGURE_lightcurves.py
updated figure and lxlbol

@@ -61,9 +61,9 @@ if __name__ == "__main__":
         a.legend(loc=1, frameon=True)
 
 
-    ax[0].set_ylabel("normalized flux")
-    ax[1].set_ylabel("background subtracted flux <10 keV [cts/s]")
-    ax[1].set_xlabel("time [days]")
+    ax[0].set_ylabel("normalized flux", fontsize=11)
+    ax[1].set_ylabel(r"background subtracted flux [counts s$^-1$]", fontsize=11)
+    ax[1].set_xlabel("time [d]", fontsize=13)
 
     plt.tight_layout()
     plt.savefig(paths.figures / "lightcurves.png", dpi=300)

### 2023-09-19 src/scripts/VALUES.py

@@ -85,13 +85,13 @@ if __name__ == "__main__":
     row = df[df.subset == "onlyflare"].iloc[0]
 
     # get Lx and make latex string with a single uncertainty
-    Lx, Lxerr = row.Lx_erg_s, row.Lx_erg_s_err
-    Lx_str = fr"${Lx/1e27:.2f} \pm {Lxerr/1e27:.2f}" + r"\times 10^{27}\,\rm{erg s}^{-1}$"
+    Lxq, Lxerrq = row.Lx_erg_s, row.Lx_erg_s_err
+    Lx_strq = fr"${Lx/1e27:.2f} \pm {Lxerr/1e27:.2f}" + r"\times 10^{27}\,\rm{erg s}^{-1}$"
 
-    print(Lx_str)
+    print(Lx_strq)
 
     with open(paths.output / "epic_Lx_flaring.tex", "w") as f:
-        f.write(f"{Lx_str}")
+        f.write(f"{Lx_strq}")
 
     # Fx -----------------------------------------------------------------------
 

### 2023-09-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Fig.~\ref{fig:lightcurves} shows the time series using the combined PN and MOS time series together with the OM light curve.
-We extracted the events time series from EPIC using XMM SAS version 20~\footnote{"Users Guide to the \textit{XMM-Newton} Science Analysis System", Issue 18.0, 2023 (ESA: \textit{XMM-Newton} SOC)} using the \texttt{evselect} task.
+We extracted the events time series from EPIC using XMM SAS version 20\footnote{"Users Guide to the \textit{XMM-Newton} Science Analysis System", Issue 18.0, 2023 (ESA: \textit{XMM-Newton} SOC)} using the \texttt{evselect} task in the full $0-10\,$keV band.
 We selected a circular source region with 20 arcsec radius on all detectors, and a circular source-free background region 120 and 90 arcsec in MOS and PN, respectively.
@@
          Top panel: Optical Monitoring (OM) light curve.
-Bottom panel: X-ray (PN, MOS1 and MOS2 combined) light curve.
+Bottom panel: Background-subtracted X-ray light curve using the flux in the entire $0-10\,$keV range  (PN, MOS1 and MOS2 combined).
 The grey shaded portion defines the flare-only subset of the observations (see Section~\ref{sec:methods:epic} and Table~\ref{tab:specfit}).
@@
 % \movetableright=-20mm
+\footnotesize
 \centering
@@
 
-We find that the quiescent corona of TIC 277 is best described by two components, similar to many other active M dwarfs~\citep{wargelin2008xray, robrade2010quiescent, behr2023muscles, magaudda2022firsta}.
-The hotter component represents thermal bremsstrahlung radiation from hot plasma that evaporates from the chromosphere into the coronal flaring loops~\citep{benz2016flare}, which is supported by the observed delay between the optical and the X-ray flare~\citep[Fig.~\ref{fig:lightcurves}, ][]{hawley2003multiwavelength}.
+We find that the quiescent corona of TIC 277 is best described by two components, similar to many other active M dwarfs, in line with the standard picture for stellar flares~\citep{wargelin2008xray, robrade2010quiescent, behr2023muscles, magaudda2022firsta}.
+Both the cooler and the hotter component represent thermal radiation from hot plasma that evaporates from the chromosphere into the corona.
+The hotter component is driven by particles precipitating downward during flares that then heat the chromosphere to high temperatures, and cause additional evaporation into the coronal flaring loops~\citep{benz2016flare}.
+In our osbervation, this interpretation is supported by the observed delay between the optical and the X-ray flare~\citep[Fig.~\ref{fig:lightcurves}, ][]{hawley2003multiwavelength}.
 The optical flare originates from the lower chromosphere, which is heated by the precipitating particles.
-This heating then causes evaporation of particles into the flare loop.
+The heating then causes evaporation of particles into the flare loop.
+In the cooler component, the source of the more quiescent coronal heating is an open question, even for the Sun, where the respective contributions of reconnection and magnetohydrodynamical waves are under investigation~\citep{vandoorsselaere2020coronal}.
 %A third, very hot, possibly $>40\,$MK component could only be poorly constrained due to a lack of sensitivity to flux above $5\,$keV, and its short lifetime during the peak of the flare.
@@
 Flares on the slowly rotating Proxima Cen (M5.5, $P_{\rm rot} = 83\,$d,~\citealt{anglada-escude2016terrestrial}), appear with a range of temperatures at 2-3 MK and 10-20 MK~\citep{gudel2004flares, fuhrmeister2011multiwavelength, fuhrmeister2022high, howard2022mouse} placing it in the transition region between saturated and unsaturated regimes.
-Only at rotation periods beyond 100 days, flaring activity declines to yield coronal temperatures $<2\,$MK~\citep{wright2018stellar, foster2020corona}.
+Only at rotation periods beyond 100 days, flaring activity declines to yield coronal temperatures $<2\,$MK in M3-M6 dwarfs~\citep{wright2018stellar, foster2020corona}.
 
@@
 
-Recent observations of TRAPPIST-1 b~\citep{greene2023thermal, ih2023constraining} with the James Webb Space Telescope find the planet mostly stripped of an atmosphere, which the authors attribute to the intense flare driven X-ray radiation and energetic particle exposure at the short orbital periods of the innermost planet.
+%Recent observations of TRAPPIST-1 b~\citep{greene2023thermal, ih2023constraining} with the James Webb Space Telescope find the planet mostly stripped of an atmosphere, which the authors attribute to the intense flare driven X-ray radiation and energetic particle exposure at the short orbital periods of the innermost planet.
 For the habitable zone planets, TRAPPIST-1 e and f, simulations suggest that flares alter their chemical composition, in particular ozone, nitric oxide and hydroxide, but that the effects would be below the noise floor of JWST~\citep{chen2021persistence}.
@@
 TIC~277's rapid rotation and UV flux is indicative of a younger age than TRAPPIST-1, although it has not been attributed to any star cluster so far~\citep{schneider2018hazmat}.
-Interpreting the upcoming observations of TRAPPIST-1 e with JWST will most likely involve its activity history.
+Interpreting the upcoming observations of TRAPPIST-1 e, one of the rocky planets in the habitable zone of the system, with the James Webb Space Telescope will most likely involve its activity history.
 Studies of its younger counterparts, late M dwarfs like TIC~277, are required to empirically constrain cumulative effects of atmospheric forcing of planets.
@@
 
-\begin{table}[]
+\begin{table*}[]
     \centering
     \caption{Late M dwarfs with known rotation period and X-ray luminosity.}
-    \begin{tabular}{lccc}
-        ID & SpT &$P_{\rm rot}$ [d] & $\log_{10} \frac{L_X}{L_{\rm bol}} $\\\hline
-        TIC 277 (this work) & M7 & 0.19 & $-4$ \\
-        TRAPPIST-1 & M8 & 3.3 & $-4$ [1]\\
-        LHS 248 & M6.5 & 0.46 & $-3.9$ [2] \\
-        NLTT 33370 AB & M7 & 0.16 & $-3.5$ [3] \\
-        LP 412-31 & M8 & 0.61 & $-3.1$ [4]\\\hline
+    \begin{tabular}{lcccc}
+        ID & SpT &$P_{\rm rot}$ [d] & $\log_{10} \frac{L_X}{L_{\rm bol}} $ & energy range [keV]\\\hline
+        TIC 277 (this work) & M7 & 0.19 & $-4$ & $0.2-2$ \\
+        TRAPPIST-1 [1] & M8 & 3.3 & $-4$ & $0.3-10$\\
+        LHS 248 [2] & M6.5 & 0.46 & $-3.9$ & $0.2-2$\\
+        NLTT 33370 AB [3] & M7 & 0.16 & $-3.5$ & $0.2-2$ \\
+        LP 412-31  [4] & M8 & 0.61 & $-3.1$& \textcolor{red}{soft X-ray band?}\\\hline
     \end{tabular}
-    \flushleft
      \footnotesize
+     \vspace{0.1cm}\newline
     [1] \citet{brown2023coronal}, 
@@
     \label{tab:latmlxlbol}
-\end{table}
+\end{table*}
 

### 2023-09-25 src/scripts/FIGURE_LXLBol_rot.py
added figure in discussion

@@ -1,0 +1,62 @@
+"""
+Python 3.8 - UTF-8
+
+X-ray Loops
+Ekaterina Ilin, 2023
+MIT License
+
+---
+
+This script pull the data from Wright et al. 2011 and 2016, and compares with
+late M dwarfs in Table 4.
+"""
+
+
+import matplotlib.pyplot as plt
+import pandas as pd
+import paths
+
+
+if __name__ == "__main__":
+
+    d_ = pd.read_csv(paths.data / 'wright2016.csv')
+
+    d = d_.sort_values(by='V-K(mag)', ascending=True)
+    df = d_[d_["V-K(mag)"] > 5]
+    dff = d_[d_["V-K(mag)"] < 5]
+
+    fig, ax = plt.subplots(figsize=(7, 5))
+
+    plt.scatter(dff["Rotation_period(days)"],  dff["log(Lx/Lbol)"], c="silver", 
+                alpha=1, s=8,
+                label=r"Wright et al. 2011, 2016 ($<$M3.5)")
+
+
+    plt.scatter(df["Rotation_period(days)"], df["log(Lx/Lbol)"], s=30,
+                cmap='viridis', c=df["V-K(mag)"], alpha=0.9, marker='o',
+                label=r"Wright et al. 2011, 2016 (M3.5 - M5.5)")
+
+    plt.colorbar(label="V-K [mag]", ax=ax)
+
+    s, ss = 80, 300
+
+    plt.scatter([0.19], [-4.04], s=ss, c='k', marker='*')
+    plt.scatter([3.3], [-4], s=ss/1.5, c='k', marker='d') 
+    plt.scatter([0.46], [-3.9], s=ss/1.5, c='k', marker='d')
+    plt.scatter([0.16], [-3.5], s=ss/1.5, c='k', marker='d')
+    plt.scatter([0.61], [-3.1], s=ss/1.5, c='k', marker='d')
+
+    plt.scatter([0.19], [-4.04], s=s, c='r', marker='*', label="TIC 277 (M7)")
+    plt.scatter([3.3], [-4], s=s/1.5, c='dimgrey', marker='d', label="TRAPPIST-1 (M8)") 
+    plt.scatter([0.46], [-3.9], s=s/1.5, c='cyan', marker='d', label="LHS 248 (M6.5)")
+    plt.scatter([0.16], [-3.5], s=s/1.5, c='brown', marker='d', label="NLTT 33370 AB (M7)")
[…]

### 2023-09-25 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \label{sec:discussion:xraylum}
-TIC~277's coronal luminosity relative to bolometric luminosity of $L_X/L_{\rm bol}=$\LXLbol in quiescence is about an order of magnitude lower than the $L_X/L_{\rm bol}\sim 10^{-3}$ average in the saturated regime of partly and fully convective M dwarfs~\citep{wright2011stellaractivityrotation,wright2018stellar}.
+
+\begin{figure}
+    \script{FIGURE_LXLBol_rot.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/lx_lbol.png}
+        \caption{$L_X/L_{\rm bol}$ over rotation period of stars of different spectral types.
+For early spetral types and fully convective mid-M dwarfs, we show the results from~    \citet{wright2011stellaractivityrotation, wright2016solartype}.
+Late M dwarfs from the literature are marked with diamonds, and are detailed in Table~\ref{tab:latmlxlbol}.
+TIC 277 is shown as red star.
+These few late M dwarfs with known rotation period and X-ray luminosity show a tentative decline in $L_X/L_{\rm bol}$ relative to earlier type, but also fully convective dwarfs, which might be due to polar updraft migration~(see Section \ref{sec:discussion:xraylum}).
+        }
+        \label{fig:lx_lbol}
+    \end{centering}
+\end{figure}
+
+TIC~277's coronal luminosity relative to bolometric luminosity of $L_X/L_{\rm bol}=$\LXLbol in quiescence is about an order of magnitude lower than the $L_X/L_{\rm bol}\sim 10^{-3}$ average in the saturated regime of partly and fully convective M dwarfs~\citep[Fig.~\ref{fig:lx_lbol}][]{wright2011stellaractivityrotation,wright2018stellar}.
 In the Hyades and Praesepe, \citet{nunez2022factory} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend in $L_X/L_{\rm bol}$ with decreasing rotation period for M3.5-M7 dwarfs.

### 2023-10-03 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2023-09-19 src/tex/ms.tex
@@
 \label{sec:discussion:xraylum}
-TIC~277's coronal luminosity relative to bolometric luminosity of $L_X/L_{\rm bol}=$\LXLbol in quiescence is about an order of magnitude lower than the $L_X/L_{\rm bol}\sim 10^{-3}$ average in the saturated regime of partly and fully convective M dwarfs~\citep{wright2011stellaractivityrotation,wright2018stellar}.
+
+\begin{figure}
+    \script{FIGURE_LXLBol_rot.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/lx_lbol.png}
+        \caption{$L_X/L_{\rm bol}$ over rotation period of stars of different spectral types.
+For partially convective dwarfs, and fully convective mid-M dwarfs, we show the results from~\citet{wright2011stellaractivityrotation, wright2016solartype}, color-coded by $V-K$ magnitude.
+Late M dwarfs from the literature are marked with diamonds, and are detailed in Table~\ref{tab:latmlxlbol}.
+TIC~277 is shown as red star.
+The few late M dwarfs with known rotation period and X-ray luminosity show a tentative decline in $L_X/L_{\rm bol}$ relative to earlier fully convective dwarfs, which might be due to polar updraft migration~(see Section \ref{sec:discussion:xraylum}).
+        }
+        \label{fig:lx_lbol}
+    \end{centering}
+\end{figure}
+
+TIC~277's coronal luminosity relative to bolometric luminosity of $L_X/L_{\rm bol}=$\LXLbol in quiescence is about an order of magnitude lower than the $L_X/L_{\rm bol}\sim 10^{-3}$ average in the saturated regime of partly and fully convective M dwarfs~\citep[Fig.~\ref{fig:lx_lbol},][]{wright2011stellaractivityrotation,wright2016solartype}.
 In the Hyades and Praesepe, \citet{nunez2022factory} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend in $L_X/L_{\rm bol}$ with decreasing rotation period for M3.5-M7 dwarfs.

### 2023-11-17 src/scripts/FIGURE_LXLBol_rot.py
added EM with error prop to Table

@@ -36,7 +36,7 @@ if __name__ == "__main__":
                 cmap='viridis', c=df["V-K(mag)"], alpha=0.9, marker='o',
                 label=r"Wright et al. 2011, 2016 (M3.5 - M5.5)")
 
-    plt.colorbar(label="V-K [mag]", ax=ax)
+    plt.colorbar(label=r"$V-K$ [mag]", ax=ax)
 
     s, ss = 80, 300
 

### 2023-11-17 src/scripts/FIGURE_mcmc_results.py

@@ -45,16 +45,27 @@ if __name__ == "__main__":
         # discard the first 5000 steps
         df = df.iloc[5000:]
 
+        norm_to_EM = lambda x: np.log10(1e14 * np.pi * 4. * (13.7 * 3.08567758 * 1e18)**2 * x) #unit cm^-3
+        l10 = np.log(10)
+
+        def norm_to_EM_err(norm, norm_err, d, derr):
+            # print(norm, norm_err, d, derr, (norm_err**2 / (norm * l10)**2 + 4 * derr**2 / (l10 * d)**2))
+            return np.sqrt(norm_err**2 / (norm * l10)**2 + 4 * derr**2 / (l10 * d)**2)
+
         # convert units
+        df["EM1"] = norm_to_EM(df["norm__16"])
+        df["EM2"] = norm_to_EM(df["norm__32"])
+
         df["norm__16"] = df["norm__16"] * 1e6 # now in 10^-6
         df["norm__32"] = df["norm__32"] * 1e6 # now in 10^-6
+        
         df["kT__1"] = df["kT__1"] * 11.604525 # now in MK
         df["kT__17"] = df["kT__17"] * 11.604525 # now in MK
 
         # make a corner plot
-        corner.corner(df[["kT__1", "norm__16", "kT__17", "norm__32"]].values,
-                    labels=[r"$T_1$ [MK]", r"$norm_1 \cdot 10^6$",
-                            r"$T_2$ [MK]",r"$norm_2  \cdot 10^6$"],
+        corner.corner(df[["kT__1", "norm__16", "EM1", "kT__17", "norm__32", "EM2"]].values,
+                    labels=[r"$T_1$ [MK]", r"$norm_1 \cdot 10^6$", r"$\log_{10} EM_1$ [cm$^{-3}$]",
+                            r"$T_2$ [MK]",r"$norm_2  \cdot 10^6$", r"$\log_{10} EM_2$ [cm$^{-3}$]"],
                     quantiles=[0.16, 0.5, 0.84],
                     show_titles=True,
                     title_kwargs={"fontsize": 12},)
@@ -65,14 +76,17 @@ if __name__ == "__main__":
         plt.savefig(paths.figures / figname, dpi=300)
 
         # caclulate the 0.16, 0.5, 0.84 quantiles and add them to a new dataframe
-        q = np.quantile(df[["kT__1","norm__16","kT__17","norm__32"]].values,
[…]

### 2023-11-17 src/scripts/TABLE_specfits.py

@@ -123,6 +123,8 @@ if __name__ == "__main__":
                    'T2': r'$T_{\rm hot}$ [MK]',
                    'norm1': r'$10^6$ norm$_{\rm cool}$',
                    'norm2': r'$10^6$ norm$_{\rm hot}$',
+                   'EM1_50' : r'$\log_{10} EM_{\rm cool}$ [cm$^{-3}$]',
+                   'EM2_50' : r'$\log_{10} EM_{\rm hot}$ [cm$^{-3}$]',
                    'weighted_mean_T': r'$T_{\rm mean}$ [MK]'}
     
     # convert to LaTeX
@@ -143,10 +145,11 @@ if __name__ == "__main__":
         del df[f"{col}_lowerr"]
 
 
-    col = "weighted_mean_T"
-    df[mapcolnames[col]] = tex_one_err(df[col], df[f"e_{col}"])
-    del df[col]
-    del df[f"e_{col}"]
+    for col in ["weighted_mean_T", "EM1_50", "EM2_50"]:
+        df[mapcolnames[col]] = tex_one_err(df[col], df[f"e_{col}"])
+
+        del df[col]
+        del df[f"e_{col}"]
 
     string = df.T.to_latex(escape=False,index=True)
 

### 2023-11-17 src/scripts/TABLE_specfits.py
removed 16 84 percentiles changed order of lines in table

@@ -145,11 +145,16 @@ if __name__ == "__main__":
         del df[f"{col}_lowerr"]
 
 
-    for col in ["weighted_mean_T", "EM1_50", "EM2_50"]:
+    for col in ["EM1_50", "EM2_50","weighted_mean_T"]:
         df[mapcolnames[col]] = tex_one_err(df[col], df[f"e_{col}"])
 
         del df[col]
         del df[f"e_{col}"]
+
+    del df["EM1_16"]
+    del df["EM1_84"]
+    del df["EM2_16"]
+    del df["EM2_84"]
 
     string = df.T.to_latex(escape=False,index=True)
 

### 2023-11-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 The circumstances that allow flares at high latitudes to occur are poorly understood.
-We studied five Sectors of \textit{TESS} monitoring, and obtained 36 ks of \textit{XMM-Newton} observations, to find that TIC~277539431's corona does not differ significantly from other low mass stars on the saturated activity branch with respect to coronal temperatures and flaring activity, but is underluminous in soft X-ray emission by about an order of magnitude.
+We studied five Sectors of \textit{TESS} monitoring, and obtained 36 ks of \textit{XMM-Newton} observations, to find that TIC~277539431's corona does not differ significantly from other low mass stars on the saturated activity branch with respect to coronal temperatures and flaring activity, but is underluminous, in soft X-ray emission by about an order of magnitude compared to M dwarfs on the saturated activity branch.
 Combining the lack of X-ray flux, the high latitude flare, the star's viewing geometry, and the otherwise typical stellar corona, we suggest that the migration of flux emergence to the poles in rapid rotators can explain the observations of TIC~277539431.
@@
 
-We used AltaiPony~\citep{ilin2021altaipony} to de-trend the light curves using the custom de-trending function \texttt{custom\_detrending}\footnote{see also the documentation online at \url{https://altaipony.readthedocs.io/en/latest/tutorials/detrend.html}, accessed on July 3, 2023}, detailed in~\citet{ilin2022searching}.
-The procedure includes 3rd order spline fitting of coarse trends, followed by iterative sine fitting to remove rotational modulation, and two Savitzky-Golay filters~\citep{savitzky1964smoothing} with a 6h and 3h window, respectively to remove aperiodic short term trends, while masking potential flare candidates in each step.
+\section{Methods}
+\label{sec:methods}
+In this work, we analyse the coronal properties and flaring behavior of the late M dwarf TIC~277.
+From the X-ray observations with XMM-Newton, we obtain a soft X-ray spectrum, which we model with two thermal emitter components~(Section~\ref{sec:methods:epic}).
+In TESS' optical time series photometry, we remove the variability introduced by the star's rotation, and find and characterize the flares in the so de-trended light curves~(Section~\ref{sec:methods:tess}).
+XMM-Newton captured a flare simultaneously in both the PN and MOS instruments in X-ray, and marginally the OM instrument in the optical.
+We derive its soft X-ray energy using the luminosity derived from X-ray spectrum, and its bolometric energy from the OM data.
+For the following comparison between optical and X-ray flaring activity, we also calculate the bolometric energies of the flares detected in TESS~(Section~\ref{sec:methods:flareenergies}).
+
+\subsection{Spectral fitting: EPIC}
+\label{sec:methods:epic}
+The stellar corona can be described as an optically thin thermal plasma that can consist of one or multiple temperature components that represent different regions.
+We use XSPEC version 12.12.1~\citep{arnaud1996xspec} to fit a two-temperature (2T) additive APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/0$=0.6$ ratio, which is more typical of M dwarfs~\citep{wood2018chandra}.
+We analyzed both the full data set, and the quiescent and flaring parts separately.
+Neither subset could be adequately fit with a single temperature component (1T), and a 3T model did not improve the fit compared to the 2T-model.
+In the full data set, we used the \texttt{bayes} method in XSPEC to assign constant priors within the model's hard limits, and sample the uncertainties using the Markov-Chain-Monte-Carlo (MCMC) method using \texttt{chain} for a total of 30000 steps, discarding the first 5000 as the burn-in phase.
+From the same fit, but using PN data only, we derived the flux $F_X$ and X-ray luminosity $L_X$ in the soft $0.2-2\,$keV band.
+For the quiescent and flaring portions of the observations, we chose the same procedure, but used Gaussian prior distributions for the two coronal temperatures from the full data set, where we assigned the largest difference between the 50th, and 16th and 84th percentiles as standard deviation.
+
+\subsection{Lightcurve de-trending and flare finding: TESS}
+\label{sec:methods:tess}
+We used AltaiPony~\citep{ilin2021altaipony} to de-trend the light curves using the de-trending function \texttt{custom\_detrending}\footnote{see also the documentation online at \url{https://altaipony.readthedocs.io/en/latest/tutorials/detrend.html}, accessed on July 3, 2023}, detailed in~\citet{ilin2022searching}.
+The procedure begins with a 3rd order spline fitting of coarse trends, followed by iterative sine fitting to remove rotational modulation, and two Savitzky-Golay filters~\citep{savitzky1964smoothing} with a 6h and 3h window, respectively to remove aperiodic short term trends, while masking potential flare candidates in each step.
 The masking takes three steps each time.
@@
 \end{figure*}
-
-\section{Methods}
-\label{sec:methods}
-
-\subsection{Spectral fitting: EPIC}
-\label{sec:methods:epic}
-The stellar corona can be described as an optically thin thermal plasma that can consist of one or multiple temperature components that represent different regions.
-We use XSPEC version 12.12.1~\citep{arnaud1996xspec} to fit a two-temperature (2T) additive APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/0$=0.6$ ratio, which is more typical of M dwarfs~\citep{wood2018chandra}.
-We analyzed both the full data set, and the quiescent and flaring parts separately.
-Neither subset could be adequately fit with a single temperature component (1T), and a 3T model did not improve the fit compared to the 2T-model.
-In the full data set, we used the \texttt{bayes} method in XSPEC to assign constant priors within the model's hard limits, and sample the uncertainties using the Markov-Chain-Monte-Carlo (MCMC) method using \texttt{chain} for a total of 30000 steps, discarding the first 5000 as the burn-in phase.
-From the same fit, but using PN data only, we derived the flux $F_X$ and X-ray luminosity $L_X$ in the soft $0.2-2\,$keV band.
-For the quiescent and flaring portions of the observations, we chose the same procedure, but used Gaussian prior distributions for the two coronal temperatures using the full data set as ground truth, and assigning the largest difference between the 50th, and 16th and 84th percentiles as standard deviation.
 
@@
         \caption{
-         Best-fit joint two-temperature APEC spectral fit with Fe/O$=0.6$. The top panel shows the spectra taken with the different EPIC instruments. The bottom panel shows the residuals to the fit.
+         Best-fit joint two-temperature APEC spectral fit with Fe/O$=0.6$ for the full data set.
+The top panel shows the spectra taken with the different EPIC instruments.
+The bottom panel shows the residuals to the fit.
         }
@@
 This combination could lead to an unusual manifestation of the stellar dynamo that produces high latitude flares that other dynamo states cannot generate~\citep{weber2016modeling, weber2023understanding}.
-TIC 277 is underluminous in X-ray by about an order of magnitude compared to other activity saturated M dwarfs, which, in combination with the detection of a high latitude flare, may be an indication of polar updraft migration that is suspected to occur in rapid rotators~(Section~\ref{sec:discussion:xraylum}).
+TIC 277 is underluminous, or supersaturated, in X-ray by about an order of magnitude compared to other activity saturated M dwarfs, which, in combination with the detection of a high latitude flare, may be an indication of polar updraft migration that is suspected to occur in rapid rotators~(Section~\ref{sec:discussion:xraylum}).
 

### 2023-12-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 The circumstances that allow flares at high latitudes to occur are poorly understood.
-We studied five Sectors of \textit{TESS} monitoring, and obtained 36 ks of \textit{XMM-Newton} observations, to find that TIC~277539431's corona does not differ significantly from other low mass stars on the saturated activity branch with respect to coronal temperatures and flaring activity, but is underluminous, in soft X-ray emission by about an order of magnitude compared to M dwarfs on the saturated activity branch.
-Combining the lack of X-ray flux, the high latitude flare, the star's viewing geometry, and the otherwise typical stellar corona, we suggest that the migration of flux emergence to the poles in rapid rotators can explain the observations of TIC~277539431.
+We studied five Sectors of \textit{TESS} monitoring, and obtained 36 ks of \textit{XMM-Newton} observations, to find that TIC~277539431's corona does not differ significantly from other low mass stars on the \textcolor{red}{canonical saturated activity branch with respect to coronal temperatures and flaring activity, but shows lower luminosity in soft X-ray emission by about an order of magnitude, consistent with other late M dwarfs.} Combining the lack of X-ray flux, the high latitude flare, the star's viewing geometry, and the otherwise typical stellar corona, we suggest that the migration of flux emergence to the poles in rapid rotators can explain the observations of TIC~277539431.
 
@@
 In combination with the TESS data, we infer that TIC 277 produces flares at or above this energy approximately once a day.
-    %Our results indicate that TIC 277 is underluminous in X-ray copX-ray and flaring activity is representative of other rapidly rotating late M dwarfs, but that further measurements of localized flares are needed to understand the conditions under which high latitude flares in these stars can occur.
+    %Our results indicate that TIC 277 is faint in X-ray copX-ray and flaring activity is representative of other rapidly rotating late M dwarfs, but that further measurements of localized flares are needed to understand the conditions under which high latitude flares in these stars can occur.
 
@@
 
-In M dwarfs, stellar activity evolution on the main sequence unfolds much more slowly than for an FGK star.
-While for the latter, magnetic activity, and with it the flares, winds, coronal mass ejections and energetic particle eruptions that make up their space weather, decline rapidly over a few hundred megayears, a fully convective M dwarf stays highly active for gigayears~\citep{johnstone2021active, medina2022galactic}.
+In M dwarfs, stellar activity evolution on the main sequence unfolds much more slowly than for \textcolor{red}{FGK stars}.
+While for the latter, magnetic activity, and with it the flares, winds, coronal mass ejections and energetic particle eruptions that make up their space weather, decline rapidly over a few hundred megayears, a fully convective M dwarf stays highly active for gigayears~\citep{magaudda2020relation, johnstone2021active, medina2022galactic}.
 Moreover, the zone where liquid water can reside is much closer to these faint stars.
@@
 The transition from star to brown dwarf is characterized by increasing rotation speed and decreasing coronal emission.
-The simultaneous surge of radio emission indicates a transition from stellar corona to planet-like magnetosphere at the bottom of the main sequence~\citep{zarka1998auroral,pineda2017panchromatic}.
+\textcolor{red}{At the same time, the observed surge} surge of radio emission indicates a transition from stellar corona to planet-like magnetosphere at the bottom of the main sequence~\citep{zarka1998auroral,pineda2017panchromatic}.
 Yet, brown dwarfs are regularly detected with energetic flares regardless of their expected low X-ray luminosity~\citep[e.g., ][]{hambaryan2004new, stelzer2006simultaneous, gizis2013kepler, paudel2020k2, schmidt2019largest}.
@@
 The flares were observed to rotate in and out of view.
-The shape of this modulation together with known inclination (combining $P_{\rm rot}$ and $v \sin i$ from high resolution spectroscopy) allowed them to determine the latitudes of these flares.
-Surprisingly, these flares were found significantly closer to the pole than to the equator, in contrast to the Sun, where flares are usually found below $30^{\circ}$ latitude.
-Their results indicate a preference for flares at high latitudes in these stars, as a chance finding was unlikely ($\sim 0.1\%$).
+The shape of this modulation together with the known inclination (combining $P_{\rm rot}$ and $v \sin i$ from high resolution spectroscopy) allowed us to determine the latitudes of these flares.
+These flares were found significantly closer to the pole than to the equator, in contrast to the Sun, where flares are usually found below $30^{\circ}$ latitude.
+Our results indicate a preference for flares at high latitudes in these stars, as a chance finding \textcolor{red}{of these latitudes among latitudinally equidistributed flares was } unlikely ($\sim 0.1\%$).
 If flares and associated particle eruptions indeed have a latitude preference, the space weather of a planet in an aligned orbit would be less severe than previously thought.
@@
 
-TIC 277539431, TIC 277 for short, showed the highest latitude flare known to date at $\sim81^{\circ}$~\citep[][Table \ref{tab:starparams}]{ilin2021giant}.
+\textcolor{red}{Among the studied objects, }TIC 277539431, TIC 277 for short, showed the highest latitude flare known to date at $\sim81^{\circ}$~\citep[][Table \ref{tab:starparams}]{ilin2021giant}.
 It is an M7 dwarf with an extremely short rotation period of $4.56\,$h.
-While the fast rotation and late spectral type could be indicative of a brown dwarf, the detection of flares suggests the presence of a stellar corona.
-%The star's age is unknown.
+While the fast rotation and late spectral type could be indicative of \textcolor{red}{a diminishing corona, the detection of flares suggests otherwise.} %The star's age is unknown.
 \citet{schneider2018hazmat} classify it as a field star, but its FUV and NUV fluxes are on the high end for field stars and would, if not as high, be compatible with the NUV/FUV emission from younger open cluster members.
@@
 In this work, we follow TIC 277 with \textit{XMM-Newton} for 36 ks and use optical monitoring from \textit{TESS}~(Section~\ref{sec:obs}) to measure its coronal and flaring properties, respectively~(Section~\ref{sec:methods}).
-We contrast our results~(Section~\ref{sec:results}) with the literature to investigate if TIC~277 behaves like a low mass star or brown dwarf, investigate scenarios that are consistent with high latitude flare occurrence ~(Section~\ref{sec:discussion}), and conclude in Section~\ref{sec:summary}.
+We contrast our results~(Section~\ref{sec:results}) with the literature to investigate if TIC~277 behaves like a low mass star or brown dwarf, investigate scenarios that are consistent with high latitude flare occurrence ~(Section~\ref{sec:discussion}), and \textcolor{red}{summarize our results} in Section~\ref{sec:summary}.
 
@@
 \label{sec:obs:xmm}
-\textit{XMM-Newton} is an X-ray telescope, which launched into orbit in 1999, equipped with six science instruments that operate simultaneously, among them the European Photon Imaging Camera with three cameras, PN, MOS1 and MOS2; and the Optical Monitor (OM) that provides near-UV and optical photometry, among other capabilities.
+\textit{XMM-Newton} is an X-ray telescope, which \textcolor{red}{was} launched into orbit in 1999, equipped with six science instruments that operate simultaneously, among them the European Photon Imaging Camera with three cameras, PN, MOS1 and MOS2; and the Optical Monitor (OM) that provides near-UV and optical photometry, among other capabilities.
 We used \textit{XMM-Newton} to observe TIC 277 for 36 ks, i.e.
@@
 Ilin).
-We used the EPIC instruments aboard \textit{XMM-Newton}, i.e., MOS1/2 and PN, that cover the soft X-ray band from $0.2$ and up to $12\,$keV, as well as the OM using its broad band white light filter.
+We used the EPIC instruments aboard \textit{XMM-Newton}, i.e., MOS1/2 and PN, that cover the soft X-ray band from $0.2$ and up to $12\,$keV, as well as the OM using its white light filter.
 
@@
 Fig.~\ref{fig:lightcurves} shows the time series using the combined PN and MOS time series together with the OM light curve.
-We extracted the events time series from EPIC using XMM SAS version 20\footnote{"Users Guide to the \textit{XMM-Newton} Science Analysis System", Issue 18.0, 2023 (ESA: \textit{XMM-Newton} SOC)} using the \texttt{evselect} task in the full $0-10\,$keV band.
+We extracted the events time series from EPIC using XMM SAS version 20\footnote{"Users Guide to the \textit{XMM-Newton} Science Analysis System", Issue 18.0, 2023 (ESA: \textit{XMM-Newton} SOC)} using the \texttt{evselect} task in the full $0.2-10\,$keV band.
 We selected a circular source region with 20 arcsec radius on all detectors, and a circular source-free background region 120 and 90 arcsec in MOS and PN, respectively.
-We used the \texttt{epiclccorr} task in XMM SAS to correct for energy and time dependent loss of events\footnote{\url{https://heasarc.gsfc.nasa.gov/docs/xmm/sas/help/epiclccorr/node3.html}, accessed on July 3, 2023}, and extract a soft X-ray time series with a time binning of $200\,$s.
+We used the \texttt{epiclccorr} task in XMM SAS to correct for energy and time dependent loss of events\footnote{\url{https://heasarc.gsfc.nasa.gov/docs/xmm/sas/help/epiclccorr/node3.html}, accessed on July 3, 2023}, and extracted a soft X-ray time series with a time binning of $200\,$s.
 
-Simultaneously, we used the optical monitor (OM) aboard \textit{XMM-Newton}, to monitor TIC 277 in its white light filter at $10$\,s cadence.
-The OM white light filter is centered on $4,060\,$\AA, with an equivalent width of $3,470\,$\AA.
+Simultaneously, we used the Optical Monitor (OM) aboard \textit{XMM-Newton}, to monitor TIC 277 in its white light filter at $10$\,s cadence.
+The OM white light filter is centered on $4060\,$\AA, with an equivalent width of $3470\,$\AA.
+\textcolor{red}{In each light curve, OM observed uninterruptedly for $4390\,$s, followed by a short gap of $318\,$s before the next one.
 We used the \texttt{omfchain} in XMM SAS to extract the 8 individual light curves.
-In each light curve, OM observed uninterruptedly for $4390\,$s, followed by a short gap of $318\,$s before the next one.
+}
 
@@
         \caption{
-         Top panel: Optical Monitoring (OM) light curve.
-Bottom panel: Background-subtracted X-ray light curve using the flux in the entire $0-10\,$keV range  (PN, MOS1 and MOS2 combined).
+         Top panel: Optical Monitor light curve.
+Bottom panel: Background-subtracted X-ray light curve using the flux in the entire $0.2-10\,$keV range  (PN, MOS1 and MOS2 combined).
 The grey shaded portion defines the flare-only subset of the observations (see Section~\ref{sec:methods:epic} and Table~\ref{tab:specfit}).
@@
 \subsubsection{EPIC Spectra}
-From the EPIC observations, we used the standard spectral extraction procedure for point sources, as specified in the XMM SAS handbook.
+From the EPIC observations, we used the standard spectral extraction procedure for point sources, as specified in the XMM\textcolor{red}{-Newton} SAS handbook.
 We extracted spectra using \texttt{evselect} to select events in the circular event and background regions, \texttt{epatplot} to check for pile-up of multiple photons in one exposure, \texttt{backscale} to calculate the area of the source region, and \texttt{rmfgen} and \texttt{arfgen} to convolve the observed energies with the instrument response, and produce the spectra.
 Finally, we use \texttt{grppha} to rebin the data in the spectra to obtain at least 15 counts per bin.
-To prduce spectra for the quiescent and flaring portions of the observations separately, we used \texttt{tabgtigen} to select the time intervals to pass to \texttt{evselect}.
+To produce spectra for the quiescent and flaring portions of the observations separately, we used \texttt{tabgtigen} to select the time intervals to pass to \texttt{evselect}.
 
@@
 Since 2018, \textit{TESS} has been supplying publicly available red-optical high-precision time series photometry in an ongoing all-sky survey.
-Each uninterrupted observing Sector provides an approximately 27 day long light curve in a broad $6,000$-$10,000$\AA band pass.
+Each uninterrupted observing Sector provides an approximately 27 day long light curve in a broad $6000$-$10000\,$\AA band pass.
 
-We used 5 Sectors of optical monitoring of TIC 277 in 2-minute cadence.
+We used \textcolor{red}{all} 5 Sectors of optical monitoring of TIC 277 in 2-minute cadence.
 Sector 12 was observed in May 2019, Sectors 37 and 39 in April and June 2021, and Sectors 64 and 65 in April and May 2023~(Fig.~\ref{fig:tess_lcs}).
@@
 In this work, we analyse the coronal properties and flaring behavior of the late M dwarf TIC~277.
-From the X-ray observations with XMM-Newton, we obtain a soft X-ray spectrum, which we model with two thermal emitter components~(Section~\ref{sec:methods:epic}).
+From the X-ray observations with \textit{XMM-Newton}, we obtain a soft X-ray spectrum, which we model with two thermal emitter components~(Section~\ref{sec:methods:epic}).
 In TESS' optical time series photometry, we remove the variability introduced by the star's rotation, and find and characterize the flares in the so de-trended light curves~(Section~\ref{sec:methods:tess}).
-XMM-Newton captured a flare simultaneously in both the PN and MOS instruments in X-ray, and marginally the OM instrument in the optical.
+\textit{XMM-Newton} captured a flare simultaneously in both the PN and MOS instruments in X-ray, and marginally the OM instrument in the optical.
 We derive its soft X-ray energy using the luminosity derived from X-ray spectrum, and its bolometric energy from the OM data.
@@
 \label{sec:methods:epic}
-The stellar corona can be described as an optically thin thermal plasma that can consist of one or multiple temperature components that represent different regions.
-We use XSPEC version 12.12.1~\citep{arnaud1996xspec} to fit a two-temperature (2T) additive APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/0$=0.6$ ratio, which is more typical of M dwarfs~\citep{wood2018chandra}.
-We analyzed both the full data set, and the quiescent and flaring parts separately.
+The stellar corona can be described as an optically thin thermal plasma \textcolor{red}{that consists of multiple temperature components that represent different regions.
+The number of components one can identify in an X-ray spectrum depends on the brightness of the source.} We use XSPEC version 12.12.1~\citep{arnaud1996xspec} to fit a two-temperature (2T) additive \textcolor{red}{V}APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/0$=0.6$ ratio, which is more typical of M dwarfs~\citep{wood2018chandra}.
+We analyzed both the full data set, and the quiescent and flaring parts separately~\textcolor{red}{(see Fig.~\ref{fig:lightcurves})}.
 Neither subset could be adequately fit with a single temperature component (1T), and a 3T model did not improve the fit compared to the 2T-model.
@@
 \label{sec:methods:tess}
-We used AltaiPony~\citep{ilin2021altaipony} to de-trend the light curves using the de-trending function \texttt{custom\_detrending}\footnote{see also the documentation online at \url{https://altaipony.readthedocs.io/en/latest/tutorials/detrend.html}, accessed on July 3, 2023}, detailed in~\citet{ilin2022searching}.
+We used AltaiPony~\citep{ilin2021altaipony} to de-trend \textcolor{red}{the Pre-search Data Conditioning Simple Aperture Photometry (PDCSAP)} light curves using the de-trending function \texttt{custom\_detrending}\footnote{see also the documentation online at \url{https://altaipony.readthedocs.io/en/latest/tutorials/detrend.html}, accessed on July 3, 2023}, detailed in~\citet{ilin2022searching}.
 The procedure begins with a 3rd order spline fitting of coarse trends, followed by iterative sine fitting to remove rotational modulation, and two Savitzky-Golay filters~\citep{savitzky1964smoothing} with a 6h and 3h window, respectively to remove aperiodic short term trends, while masking potential flare candidates in each step.
@@
 First, all outliers above $>2.5\sigma$ are masked.
-Second, the six-fold of the number of data points masked in the first step are also masked.
-Third, three adjacent data points before and after the so expanded mask are added to the mask.
+\textcolor{red}{Second, to mask the flare decay, we append the six-fold number of data points masked in the first step to the mask.} Third, three adjacent data points before and after the so expanded mask are added to the mask.
 
-In the de-trended light curves, we searched for flare candidates as at least 3 consecutive $>3\sigma$ outliers above the noise level, filling in the noise level inside the previously masked areas with the mean of the noise levels adjacent to the mask.
-To the series of outliers, we kept adding data points to the flare candidate until a data point falls below $2\sigma$.
+In the de-trended light curves, we searched for flare candidates as at least 3 consecutive $>3\sigma$ outliers above the noise level\textcolor{red}{, defined as the rolling standard deviation of the de-trended light curve flux with a 2 h window}, filling in the noise level inside the previously masked areas with the mean of the noise levels adjacent to the mask.
+To the series of outliers, we kept \textcolor{red}{appending} data points to the flare candidate until a data point falls below $2\sigma$.
 
@@
 \label{sec:methods:flareenergies}
-For the OM flare, we median flux of the lightcurve as the quiescent flux $F_0$, against which we measured the equivalent duration $ED$ of the flare, i.e., the flare flux $F_{\rm flare}$, divided by $F_0$, integrated over the flare duration~\citep{gershberg1972results}:
+For the OM flare, we \textcolor{red}{used the} median flux of the lightcurve as the quiescent flux $F_0$, against which we measured the equivalent duration ($ED$) of the flare, i.e., the flare flux $F_{\rm flare}$, divided by $F_0$, integrated over the flare duration~\citep{gershberg1972results}:
 
@@
 
-We then converted $ED$ to flare energy using the bolometric flare energy following the procedure in~\citet{shibayama2013superflares}.
+We then converted \textcolor{red}{the} $ED$ to flare energy using the bolometric flare energy following the procedure in~\citet{shibayama2013superflares}.
 We used the throughput curve for the white light filter scaled to unity at the peak of transmission as an optimistic response curve $R_{\lambda, \rm OM}$, given that the degradation of this filter's response is poorly constrained\footnote{\textit{XMM-Newton} User's Handbook 3.5.3.1, \url{https://xmm-tools.cosmos.esa.int/external/xmm_user_support/documentation/uhb/omfilters.html}, accessed Aug 8, 2023}.
@@
 
-Analogously to Eq.~\ref{eq:ratio} for the OM flare, we used the \textit{TESS} response curve to obtain a flux ratio $f_{\rm TESS}$ of $6.3\cdot10^{-3}$ in the redder optical filter of \textit{TESS}, and convert $ED$ to bolometric flare energy using Eq.~\ref{eq:eflare}.
+Analogously to Eq.~\ref{eq:ratio} for the OM flare, we used the \textit{TESS} response curve to obtain a flux ratio $f_{\rm TESS}$ of $6.3\cdot10^{-3}$ in the optical filter of \textit{TESS}, and convert $ED$ to bolometric flare energy using Eq.~\ref{eq:eflare}.
 Besides the $ED$, \texttt{AltaiPony} also yields start and end time, defined as the difference between the first and last data point above the flare detection criterion defined in Section~\ref{sec:obs:tess}; and relative amplitude $a$, for each flare.
@@
 For the X-ray flare energy, we multiply flare duration $\Delta t$ by the flare luminosity in the EPIC data.
-The flare luminosity is equal to the difference between the quiescent and flaring X-ray luminosities in the 0.2-2 keV band, i.e.:
+The flare luminosity is equal to the difference between the quiescent and flaring X-ray luminosities in the $0.2-2.0\,$keV band, i.e.:
 
@@
 We derive its coronal temperature and luminosity from the EPIC spectrum~(Section~\ref{sec:res:XrayTL}).
-We also measure the flare energies in X-ray from the EPIC instruments, and optical from the OM, and compare them to the flare frequency distributions obtained from the 18 detected flares in \textit{TESS}~(Section~\ref{sec:res:flares}).
+We also measure the flare energies in X-ray from the EPIC instruments, and in optical from the OM, and compare them to the flare frequency distributions obtained from the 18 detected flares in \textit{TESS}~(Section~\ref{sec:res:flares}).
 
@@
 From the \textit{XMM-Newton} EPIC spectra, we could measure the coronal properties of TIC~277 in both its quiescent, and flaring states~(Table~\ref{tab:specfit}).
-The resulting fit~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about \Tcool and \Thot, respectively.
+The resulting fit for the full data set~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about \Tcool and \Thot\hspace{-1.cm}, respectively.
 Assuming these temperatures as ground truth, we also fitted 2T APEC models to the flaring and quiescent portions of the observations separately.
@@
 However, the emission measure weighted mean temperature $T_{\rm mean}$ changed from \Tqmean in the quiescent spectrum to \Tfmean in the flaring spectrum, since the hot component became more much dominant during the flare.
-From the quiescent PN spectra, we then also calculated soft X-ray flux of $F_X=$\FX, and a luminosity $L_X=$\LXquiet in the $0.2$-$2\,$keV band.
+From the quiescent PN spectra, we then also calculated soft X-ray flux of $F_X=$\FX, and a luminosity $L_X=$\LXquiet in the $0.2$-$2.0\,$keV band.
 %This suggests that the flaring corona emits the hotter component, while the cooler component corresponds to the quiescent corona.
@@
 This combination could lead to an unusual manifestation of the stellar dynamo that produces high latitude flares that other dynamo states cannot generate~\citep{weber2016modeling, weber2023understanding}.
-TIC 277 is underluminous, or supersaturated, in X-ray by about an order of magnitude compared to other activity saturated M dwarfs, which, in combination with the detection of a high latitude flare, may be an indication of polar updraft migration that is suspected to occur in rapid rotators~(Section~\ref{sec:discussion:xraylum}).
+TIC 277 is \textcolor{red}{fainter in X-ray by about an order of magnitude than M dwarfs on the canonical saturated branch}, which, in combination with the detection of a high latitude flare, may be an indication of polar updraft migration that is suspected to occur in rapid rotators~(Section~\ref{sec:discussion:xraylum}).
 
@@
 
-\begin{figure}
-    \script{FIGURE_LXLBol_rot.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/lx_lbol.png}
-        \caption{$L_X/L_{\rm bol}$ over rotation period of stars of different spectral types.
+% \begin{figure}
+%     \script{FIGURE_LXLBol_rot.py}
+%     \begin{centering}
+%         \includegraphics[width=\linewidth]{figures/lx_lbol.png}
+%         \caption{$L_X/L_{\rm bol}$ over rotation period of stars of different spectral types.
 For partially convective dwarfs, and fully convective mid-M dwarfs, we show the results from~\citet{wright2011stellaractivityrotation, wright2016solartype}, color-coded by $V-K$ magnitude.
@@
 The few late M dwarfs with known rotation period and X-ray luminosity show a tentative decline in $L_X/L_{\rm bol}$ relative to earlier fully convective dwarfs, which might be due to polar updraft migration~(see Section \ref{sec:discussion:xraylum}).
-        }
-        \label{fig:lx_lbol}
-    \end{centering}
-\end{figure}
+%         }
+%         \label{fig:lx_lbol}
+%     \end{centering}
+% \end{figure}
 
-TIC~277's coronal luminosity relative to bolometric luminosity of $L_X/L_{\rm bol}=$\LXLbol in quiescence is about an order of magnitude lower than the $L_X/L_{\rm bol}\sim 10^{-3}$ average in the saturated regime of partly and fully convective M dwarfs~\citep[Fig.~\ref{fig:lx_lbol},][]{wright2011stellaractivityrotation,wright2016solartype}.
-In the Hyades and Praesepe, \citet{nunez2022factory} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend in $L_X/L_{\rm bol}$ with decreasing rotation period for M3.5-M7 dwarfs.
+TIC~277's coronal luminosity relative to bolometric luminosity of $L_X/L_{\rm bol}=$\LXLbol in quiescence is about an order of magnitude lower than \textcolor{red}{the canonical} $L_X/L_{\rm bol}\sim 10^{-3}$ average in the saturated regime of partly and fully convective M dwarfs\citep{wright2011stellaractivityrotation,wright2016solartype,wright2018stellar}.% In the Hyades and Praesepe, \citet{nunez2022factory} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend in $L_X/L_{\rm bol}$ with decreasing rotation period for M3.5-M7 dwarfs.
 This tentatively points toward a supersaturation of the corona at fast rotation speeds, as has been suggested before for M~dwarfs~\citep{jeffries2011investigating}, but also earlier type stars~\citep{prosser1996rosat, argiroffi2016supersaturation} and even red giants~\citep{dixon2020rotationally}.
@@
 \citet{stelzer2022first} investigated late M and L dwarfs with eROSITA to show that $L_X/L_{\rm bol}$ increases towards later spectral types~\citep{magaudda2022firsta}, but begins to decrease again from M7 downward.
-In their compilation together with previous work from \citet{stelzer2012ultracool, cook2014trends, deluca2020extras, williams2014trends}, TIC~277 appears at the lower boundary of the $L_X/L_{\rm bol}$ distribution for M7 dwarfs.
+In their compilation, together with previous work from \citet{stelzer2012ultracool, cook2014trends, deluca2020extras, williams2014trends, berger2008simultaneous}, TIC~277 appears at the lower end\textcolor{red}{, but consistent with} the $L_X/L_{\rm bol}$ distribution for M7 dwarfs.
 However, the rotation periods are unknown for most of these late M dwarfs.
 Sometimes projected rotation periods are known~\citep{cook2014trends}, but true periods are only detected in a handful of these stars, which we list in Table~\ref{tab:latmlxlbol}.
-TIC~277 appears at the lower end of this small sample.
+TIC~277 appears \textcolor{red}{consistent with} this small sample.
 
@@
 
-If TIC~277 is indeed systematically underluminous, these results could mean that its dynamo produces a weaker magnetic field than that of other saturated M dwarfs.
-However, since the flaring activity and coronal temperature of TIC~277 are in line with saturated activity, this conjecture lacks evidence from these other activity indicators.
-An alternative explanation, that would at the same time explain the occurrence of a high latitude flare in \textit{TESS}, is polar updraft migration~\citep{stepien2001rosat}.
-Polar updraft migration implies that at high rotation rates, flux emergence becomes more efficient near the poles than at the equator~\citep{weber2016modeling}, draining the equatorial regions of magnetic flux required to produce a corona there.
+\textcolor{red}{In late M dwarfs like TIC~277, the decline in X-ray activity is expected as a result of decreasing efficiency in coronal heating in the transition region between a stellar and a brown dwarf atmosphere~(e.g.,~\citealt{williams2014trends}).
+However, the flaring activity and coronal temperature of TIC~277 are in line with saturated activity, implying that the efficiency of flare production has not diminished.} An alternative explanation, that would at the same time explain the occurrence of a high latitude flare in \textit{TESS}, is polar updraft migration~\citep{stepien2001rosat}.
+Polar updraft migration implies that at high rotation rates, flux emergence becomes more efficient near the poles than at the equator~\citep{weber2016modeling}\textcolor{red}{, producing flares there, but} draining the equatorial regions of magnetic flux required to produce a corona \textcolor{red}{at low latitudes}.
 As a result, the X-ray luminosity diminishes.
 Since TIC~277 is seen nearly equator-on~(Table~\ref{tab:starparams}), polar updraft could explain its fainter corona.
-Considering the high-latitude flare, viewing geometry, star-like flaring behavior and coronal temperatures together, we favor this scenario over a brown dwarf transition explanation, that would invoke a diminishing corona due to the overall cooler, more neutral atmosphere.
+Considering the high-latitude flare, viewing geometry, star-like flaring behavior and \textcolor{red}{star-like} coronal temperatures together, we \textcolor{red}{suggest this scenario as an alternative to the brown dwarf transition explanation}, that would invoke a diminishing corona due to the overall cooler, more neutral atmosphere.
 
@@
 
-%Compared to the X-ray luminosities of M7 dwarfs among the late M dwarfs observed by eROSITA in \citet{stelzer2022first}, TIC 277 appears mildly underluminous.
+%Compared to the X-ray luminosities of M7 dwarfs among the late M dwarfs observed by eROSITA in \citet{stelzer2022first}, TIC 277 appears mildly .
 This could be an effect of its fast rotation, and consequent supersaturation as in \citet{magaudda2022first}, but we lack the rotation periods to confirm this.

### 2023-12-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 The resulting fit for the full data set~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about \Tcool and \Thot\hspace{-1.cm}, respectively.
-Assuming these temperatures as ground truth, we also fitted 2T APEC models to the flaring and quiescent portions of the observations separately.
+Assuming these temperatures \textcolor{red}{for the prior distribution}, we also fitted 2T APEC models to the flaring and quiescent portions of the observations separately.
 The coronal temperatures did not update significantly using either subset.
@@
 
-We note that not using the full set of observations as ground truth resulted in an unconstrained hot coronal component during the flare.
+We note that not using the full set of observations as the prior resulted in an unconstrained hot coronal component during the flare.
 This can be interpreted as due to a lack of X-ray flux above $5\,$keV.

### 2023-12-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \label{sec:methods:flareenergies}
+\textcolor{red}{The OM data contain one marginal flare that we interpret as associated with the large X-ray flare shortly after~(inset in Fig.~\ref{fig:lightcurves}).
+In the TESS light curves, we found a total of 18 flares.
+To assess the flaring activity of TIC~277, we computed the bolometric energies for both OM and TESS flares.}
+
 For the OM flare, we \textcolor{red}{used the} median flux of the lightcurve as the quiescent flux $F_0$, against which we measured the equivalent duration ($ED$) of the flare, i.e., the flare flux $F_{\rm flare}$, divided by $F_0$, integrated over the flare duration~\citep{gershberg1972results}:
@@
 We derive its coronal temperature and luminosity from the EPIC spectrum~(Section~\ref{sec:res:XrayTL}).
-We also measure the flare energies in X-ray from the EPIC instruments, and in optical from the OM, and compare them to the flare frequency distributions obtained from the 18 detected flares in \textit{TESS}~(Section~\ref{sec:res:flares}).
+We also measure the flare energies in X-rays from the EPIC instruments, and in optical from the OM, and compare them to the flare frequency distributions obtained from the 18 detected flares in \textit{TESS}~(Section~\ref{sec:res:flares}).
 
@@
 The resulting fit for the full data set~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about \Tcool and \Thot\hspace{-1.cm}, respectively.
-Assuming these temperatures \textcolor{red}{for the prior distribution}, we also fitted 2T APEC models to the flaring and quiescent portions of the observations separately.
+Assuming these temperatures \textcolor{red}{for the prior distribution}, we also fitted 2T VAPEC models to the flaring and quiescent portions of the observations separately.
 The coronal temperatures did not update significantly using either subset.
@@
         \caption{
-         Best-fit joint two-temperature APEC spectral fit with Fe/O$=0.6$ for the full data set.
-The top panel shows the spectra taken with the different EPIC instruments.
-The bottom panel shows the residuals to the fit.
+         \textcolor{red}{XMM-Newton EPIC spectra.
+The top panel shows the spectra taken with the different EPIC instruments together with the best-fit two-temperature VAPEC model with Fe/O$=0.6$ for the full data set.} The bottom panel shows the residuals to the fit.
         }

### 2023-12-19 src/scripts/VALUES.py
formatting values

@@ -73,7 +73,7 @@ if __name__ == "__main__":
 
     # get Lx and make latex string with a single uncertainty
     Lx, Lxerr = row.Lx_erg_s, row.Lx_erg_s_err
-    Lx_str = fr"${Lx/1e26:.2f} \pm {Lxerr/1e26:.2f}" + r"\times 10^{26}\,\rm{erg s}^{-1}$"
+    Lx_str = fr"$({Lx/1e26:.2f} \pm {Lxerr/1e26:.2f})" + r"\times 10^{26}\,\rm{erg} \rm{s}^{-1}$"
 
     print(Lx_str)
 
@@ -86,7 +86,7 @@ if __name__ == "__main__":
 
     # get Lx and make latex string with a single uncertainty
     Lxq, Lxerrq = row.Lx_erg_s, row.Lx_erg_s_err
-    Lx_strq = fr"${Lx/1e27:.2f} \pm {Lxerr/1e27:.2f}" + r"\times 10^{27}\,\rm{erg s}^{-1}$"
+    Lx_strq = fr"$({Lx/1e27:.2f} \pm {Lxerr/1e27:.2f})" + r"\times 10^{27}\,\rm{erg s}^{-1}$"
 
     print(Lx_strq)
 
@@ -96,7 +96,7 @@ if __name__ == "__main__":
     # Fx -----------------------------------------------------------------------
 
     flux, fluxerr = row.flux_erg_s_cm2, row.flux_erg_s_cm2_err
-    flux_str = fr"${flux*1e14:.1f} \pm {fluxerr*1e14:.1f}" + r"\times 10^{-14} \,\rm{erg}\, \rm{cm}^{-2} \rm{s}^{-1}$"
+    flux_str = fr"$({flux*1e14:.1f} \pm {fluxerr*1e14:.1f})" + r"\times 10^{-14} \,\rm{erg}\, \rm{cm}^{-2} \rm{s}^{-1}$"
 
     print(flux_str)
 
@@ -109,7 +109,7 @@ if __name__ == "__main__":
     row = ilin2021[ilin2021.TIC == 277539431].iloc[0]
     Lbol, eLbol = row.Lbol_erg_s, row.eLbol_erg_s
 
-    Lbolstr = fr"${Lbol/1e30:.1f} \pm {eLbol/1e30:.1f}" + r" \times 10^{30}\,\rm{erg s}^{-1}$"
[…]

### 2023-12-19 src/scripts/VALUES.py
more formatting

@@ -73,7 +73,7 @@ if __name__ == "__main__":
 
     # get Lx and make latex string with a single uncertainty
     Lx, Lxerr = row.Lx_erg_s, row.Lx_erg_s_err
-    Lx_str = fr"$({Lx/1e26:.2f} \pm {Lxerr/1e26:.2f})" + r"\times 10^{26}\,\rm{erg} \rm{s}^{-1}$"
+    Lx_str = fr"$({Lx/1e26:.2f} \pm {Lxerr/1e26:.2f})" + r"\times 10^{26}\,\rm{erg}\,\rm{s}^{-1}$"
 
     print(Lx_str)
 
@@ -86,7 +86,7 @@ if __name__ == "__main__":
 
     # get Lx and make latex string with a single uncertainty
     Lxq, Lxerrq = row.Lx_erg_s, row.Lx_erg_s_err
-    Lx_strq = fr"$({Lx/1e27:.2f} \pm {Lxerr/1e27:.2f})" + r"\times 10^{27}\,\rm{erg s}^{-1}$"
+    Lx_strq = fr"$({Lx/1e27:.2f} \pm {Lxerr/1e27:.2f})" + r"\times 10^{27}\,\rm{erg}\,\rm{s}^{-1}$"
 
     print(Lx_strq)
 
@@ -109,7 +109,7 @@ if __name__ == "__main__":
     row = ilin2021[ilin2021.TIC == 277539431].iloc[0]
     Lbol, eLbol = row.Lbol_erg_s, row.eLbol_erg_s
 
-    Lbolstr = fr"$({Lbol/1e30:.1f} \pm {eLbol/1e30:.1f})" + r" \times 10^{30}\,\rm{erg s}^{-1}$"
+    Lbolstr = fr"$({Lbol/1e30:.1f} \pm {eLbol/1e30:.1f})" + r" \times 10^{30}\,\rm{erg}\,\rm{s}^{-1}$"
 
     with open(paths.output / "Lbol.tex", "w") as f:
         f.write(f"{Lbolstr}")

### 2023-12-19 src/scripts/VALUES.py
formatting energies

@@ -178,13 +178,13 @@ if __name__ == "__main__":
     E_om, eE_om = om.E_erg, om.eE_erg
 
     e, ee = E_epic / 1e30, eE_epic /1e30
-    epicstr = fr"${e:.1f}\pm{ee:.1f}" + r"\times 10^{30}\,$erg"
+    epicstr = fr"$({e:.1f}\pm{ee:.1f})" + r"\times 10^{30}\,$erg"
 
     with open(paths.output / "epic_flare.tex", "w") as f:
         f.write(epicstr)
 
     e, ee = E_om / 1e30, eE_om /1e30
-    omstr = fr"${e:.1f}\pm{ee:.1f}" + r"\times 10^{30}\,$erg"
+    omstr = fr"$({e:.1f}\pm{ee:.1f})" + r"\times 10^{30}\,$erg"
 
     with open(paths.output / "om_flare.tex", "w") as f:
         f.write(omstr)

### 2023-12-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 This can be interpreted as due to a lack of X-ray flux above $5\,$keV.
-Hot flares with temperatures of $>50\,$MK have been measured on DS~Tuc~A~\citep{pillitteri2011xray}.
+%Hot flares with temperatures of $>50\,$MK have been measured on DS~Tuc~A~\citep{pillitteri2011xray}.
 On the Sun, the highest flare associated coronal temperatures are around $35\,$MK~\citep{kay2003soft}.
@@
 \end{equation}
-following the procedure in~\citep{ilin2021flares}, implemented as \texttt{FFD.fit\_powerlaw} using the \textit{mcmc} argument in \texttt{AltaiPony}~\citep{ilin2021altaipony}, which is using the posterior distribution derived in \citet{wheatland2004bayesian}, and sampling the posterior distribution with the \texttt{emcee} package~\citep{foreman-mackey2013emcee}.
+following the procedure in~\citet{ilin2021flares}, implemented as \texttt{FFD.fit\_powerlaw} using the \textit{mcmc} argument in \texttt{AltaiPony}~\citep{ilin2021altaipony}, which is using the posterior distribution derived in \citet{wheatland2004bayesian}, and sampling the posterior distribution with the \texttt{emcee} package~\citep{foreman-mackey2013emcee}.
 
 The FFD power law fit converged after 13,500 steps on a slope $\alpha =$\ffdalpha\unskip.
-The frequency $R_{31.5}$ of flares per day above $\log_{10} E_{\rm flare} = 31.5\,\mathrm{erg}$ is about $\log_{10}R_{31.5}=$\ffdr\unskip.
-The slope is typical of other flaring stars, regardless of spectral type and rotation period~(see, e.g., Fig.~13 in \citealt{ilin2021flares}).
+The frequency $R_{31.5}$\textcolor{red}{, that is the frequency} of flares per day above $\log_{10} E_{\rm flare}\,[\mathrm{erg}] = 31.5$ is $\log_{10}R_{31.5}=$\ffdr\unskip.
+The slope is typical of other flaring stars, regardless of spectral type and rotation period~(see, e.g., Fig.~13 \textcolor{red}{of} \citealt{ilin2021flares}).
 $R_{31.5}$ is typical of flaring M dwarfs, in particular typical of late M dwarfs in the saturated activity regime, where $R_{31.5}$ becomes independent of rotation period~\citep{medina2020flare,murray2022study}.

### 2023-12-19 src/scripts/FIGURE_data_resid.py
use 1-sigma error bars for the data points

@@ -53,11 +53,11 @@ def plot_data_resid_3(file):
 
         # flux with errors
         ax[0].errorbar(data["E [keV]"], data["flux [counts/s/keV]"],
-                    yerr=data["e_flux"], fmt=m, markersize=6, alpha=1, c=c)
+                    yerr=data["e_flux"]/1.644854, fmt=m, markersize=6, alpha=1, c=c)
 
         # residuals
         ax[1].errorbar(resid["E [keV]"], resid["flux [counts/s/keV]"],
-                    yerr=resid["e_flux"], fmt=m, markersize=6, c=c)
+                    yerr=resid["e_flux"]/1.644854, fmt=m, markersize=6, c=c)
         
         
       # zero line for residuals
@@ -137,11 +137,11 @@ def plot_data_resid_1(file):
 
     # flux with errors
     ax[0].errorbar(data["E [keV]"], data["flux [counts/s/keV]"],
-                yerr=data["e_flux"], fmt=".", markersize=13, alpha=1, c="olive")
+                yerr=data["e_flux"] / 1.644854, fmt=".", markersize=13, alpha=1, c="olive")
 
     # residuals
     ax[1].errorbar(resid["E [keV]"], resid["flux [counts/s/keV]"],
-                yerr=resid["e_flux"], fmt=".", markersize=13, c="olive")
+                yerr=resid["e_flux"]/1.644854, fmt=".", markersize=13, c="olive")
 
     # zero line for residuals
     ax[1].axhline(0,c="k")

### 2023-12-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \affiliation{INAF – Osservatorio Astronomico di Palermo, Piazza del Parlamento 1, 90134 Palermo, Italy}
+
+\author[]{Desmond Dsouza}
+\affiliation{Leibniz Insitute for Astrophysics Potsdam (AIP), An der Sternwarte 16, 14482 Potsdam, Germany}
+\affiliation{Institute for Physics and Astronomy, University of Potsdam, Karl-Liebknecht-Strasse 24/25, 14476 Potsdam, Germany}
 
@@
 
-For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of \eepic in the $0.2-2\,$keV band, using $L_{X,\rm quiescent}=$\LXquiet\unskip, $L_{X,\rm flaring}=$\LXflaring\unskip, and $\Delta t =$\xraydur\unskip (see Fig.~\ref{fig:lightcurves}) in Eq.~\ref{eq:xrayflare}.
+For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of \eepic in the $0.2-2\,$keV band \textcolor{red}{by inserting} $L_{X,\rm quiescent}=$\LXquiet\unskip, $L_{X,\rm flaring}=$\LXflaring\unskip, and $\Delta t =$\xraydur\unskip\,(see Fig.~\ref{fig:lightcurves}) in Eq.~\ref{eq:xrayflare}.
 The energy of the corresponding OM flare is \eom\unskip.
 We use the 36ks observing baseline of \textit{XMM-Newton} to calculate a flare rate in OM.
-Fig.~\ref{fig:ffd} illustrates that it is roughly consistent, if slightly higher, than the extension of \textit{TESS}' flare frequency distributions to lower energies.
-About one flare per day of the order of the observed one or above can therefore be expected in future X-ray observations of TIC~277.
-However, this assumes that all optical flares above this energy in the optical data correspond to detected X-ray flares in the EPIC instruments, which is not always the case~\citep{paudel2021simultaneous}.
-\citet{guarcello2019simultaneous, kuznetsov2021stellar} find moderate correlation between flare energies in Kepler/K2 data and \textit{XMM-Newton}, but the majority of their sampled flare energies are two orders of magnitude above the flare in this work, both in X-ray and optical.
+Fig.~\ref{fig:ffd} illustrates that it is roughly consistent, if slightly higher, than the extension of \textit{TESS}' \textcolor{red}{FFD} to lower energies.
+About one flare per day \textcolor{red}{with energies} of the order of the observed one or above can therefore be expected in future X-ray observations of TIC~277.
+\textcolor{red}{However, usually, not all optical flares have X-ray counterparts}.~\citet{paudel2021simultaneous}.
+\citet{guarcello2019simultaneous, kuznetsov2021stellar} find moderate correlation between flare energies in I data and \textit{XMM-Newton}, but the majority of their sampled flare energies are two orders of magnitude \textcolor{red}{higher than of} the flare in this work, both in X-ray and optical.
 \citet{kuznetsov2021stellar} compare primary Kepler mission observations with \textit{XMM-Newton} to find a similar result.
-We note that the X-ray flare light curve follows the typical fast-rise exponential decay shape, and is therefore unlikely to be eclipsed by the star~\citep{johnstone2012soft}.
+We note that the X-ray flare light curve follows the typical fast-rise exponential decay shape, and is therefore unlikely to be eclipsed by the star.%~\citep{johnstone2012soft}.
 
@@
 The rate of OM flares (gray square) is calculated from the observing baseline of \textit{XMM-Newton}.
+\textcolor{red}{The flare energies are corrected for the pass bands of \textit{TESS} and OM, i.e., given as bolometric flare energies~(see Section~\ref{sec:methods:flareenergies}).}
         }
@@
 
-Bearing in mind the relatively low two-temperature resolution of TIC~277's X-ray spectrum and the dearth of X-ray spectra for stars at the bottom of the main sequence, it shows a coronal temperature make-up common for mid-to-late M dwarf in the saturated activity regime that it will possibly keep for gigayears until it spins down to very low rotation rates~\citep{medina2022galactic, engle2023living}.
+Bearing in mind the relatively low two-temperature resolution of TIC~277's X-ray spectrum and the dearth of X-ray spectra for stars at the bottom of the main sequence, it shows a coronal temperature make-up common for mid-to-late M dwarfa in the saturated activity regime\textcolor{red}{, which} it will possibly keep for gigayears until it spins down to very low rotation rates~\citep{medina2022galactic, engle2023living}.
 
@@
 Sometimes projected rotation periods are known~\citep{cook2014trends}, but true periods are only detected in a handful of these stars, which we list in Table~\ref{tab:latmlxlbol}.
-TIC~277 appears \textcolor{red}{consistent with} this small sample.
+TIC~277 appears \textcolor{red}{at the lower bound of} this small sample.
+\textcolor{red}{In Magaudda et al.
+(in prep.), rapidly rotating late M dwarfs observed with eROSITA appear with a range of luminosities centered around $L_X/L_{\rm bol}\approx 10^{-4}$.}
 
@@
         NLTT 33370 AB [3] & M7 & 0.16 & $-3.5$ & $0.2-2$ \\
-        LP 412-31  [4] & M8 & 0.61 & $-3.1$& \textcolor{red}{soft X-ray band?}\\\hline
+        LP 412-31  [4] & M8 & 0.61 & $-3.1$& \textcolor{green}{soft X-ray band?}\\\hline
     \end{tabular}
@@
 \label{sec:summary}
-We investigated whether TIC 277, a rapidly rotating M7 dwarf with a flare localized at $81^{\circ}$ latitude, shows unusual coronal or flaring properties that could explain the occurrence of this flare so close to the rotational pole.
+We investigated whether TIC 277, a rapidly rotating M7 dwarf with a flare localized at $81^{\circ}$ latitude, shows coronal or flaring properties that could explain the occurrence of this flare so close to the rotational pole.
 We obtained 36 ks of EPIC and OM observations with \textit{XMM-Newton}, and studied the five Sectors of red-optical 27-day light curves provided by \textit{TESS}.
-We found a mean quiescent coronal temperature of \Tqmean, and a flaring rate $f(> \log_{10} E_{\rm flare, bol} = 31.5\,\rm{erg})\approx$\ffdr with energy distribution with a slope of \ffdalpha, all typical of fully convective M dwarfs in the saturated regime.
+We found a mean quiescent coronal temperature of \Tqmean, and a flaring rate $f(> \log_{10} E_{\rm flare, bol} = 31.5\,\rm{erg})\approx$\ffdr with an energy distribution with a slope of \ffdalpha, all typical of M dwarfs in the saturated regime.
 We also detected a simultaneous X-ray and optical flare with an energy of \eepic in the $0.2-2\,$keV band, and bolometric energy of \eom from the OM observations.
@@
 
-An indication of an unusual corona stems from its X-ray luminosity relative to bolometric luminosity $L_X/L_{\rm bol}=$\LXLbol, which is about an order of magnitude lower than is typical for saturated activity M dwarfs.
+An indication of an unusual corona stems from its X-ray luminosity relative to bolometric luminosity $L_X/L_{\rm bol}=$\LXLbol.
+\textcolor{red}{It is an order of magnitude lower than the canonical value for saturated activity M dwarfs, but representative of late M dwarfs of spectral type M7 and later}.
 TIC 277 is an extreme rapid rotator with a rotation period of only $4.56\,$h.
 The detected high-latitude flare in \textit{TESS}~\citep{ilin2021giant} may hence be a product of high-latitude flux emergence driven by its fast rotation~\citep{weber2016modeling,weber2017suppression}.
-This effect is closely related to polar updraft migration, which suppresses coronal emission in equatorial regions~\citep{stepien2001rosat}.
-However, less than ten flare latitudes on stars other than the Sun are known to date~\citep{wolter2008doppler, ilin2021giant, johnson2021simultaneous}.
-Localization of more flares on the surfaces of a broader range of stars will allow us to resolve if polar updraft indeed occurs in these stars.
+\textcolor{red}{The low X-ray luminosity and the high-latitude flare taken together with the star's fast rotation, saturated flaring activity, and equator-on viewing angle can be explained by polar updraft migration, which suppresses coronal emission in equatorial regions, and confines the corona and flaring regions to near-polar regions~\citep{stepien2001rosat}.
+}
+
+\textcolor{red}{Overall, }less than ten flare latitudes on stars other than the Sun are known to date~\citep{wolter2008doppler, ilin2021giant, johnson2021simultaneous}.
+Localization of more flares on the surfaces of a broader range of stars will allow us to resolve if polar updraft migration indeed occurs in these stars.
 

### 2023-12-20 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\author[]{Desmond Dsouza}
+\author[0009-0008-2283-2840]{Desmond Dsouza}
 \affiliation{Leibniz Insitute for Astrophysics Potsdam (AIP), An der Sternwarte 16, 14482 Potsdam, Germany}
@@
 This is in stark contrast to solar flares that occur much closer to the equator, typically below $30^{\circ}$.
-The circumstances that allow flares at high latitudes to occur are poorly understood.
-We studied five Sectors of \textit{TESS} monitoring, and obtained 36 ks of \textit{XMM-Newton} observations, to find that TIC~277539431's corona does not differ significantly from other low mass stars on the \textcolor{red}{canonical saturated activity branch with respect to coronal temperatures and flaring activity, but shows lower luminosity in soft X-ray emission by about an order of magnitude, consistent with other late M dwarfs.} Combining the lack of X-ray flux, the high latitude flare, the star's viewing geometry, and the otherwise typical stellar corona, we suggest that the migration of flux emergence to the poles in rapid rotators can explain the observations of TIC~277539431.
+The mechanisms that allow flares at high latitudes to occur are poorly understood.
+We studied five Sectors of \textit{TESS} monitoring, and obtained 36 ks of \textit{XMM-Newton} observations, to find that TIC~277539431's corona does not differ significantly from other low mass stars on the \textcolor{red}{canonical saturated activity branch with respect to coronal temperatures and flaring activity, but shows lower luminosity in soft X-ray emission by about an order of magnitude, consistent with other late M dwarfs.
+The lack of X-ray flux, the high latitude flare, the star's viewing geometry, and the otherwise typical stellar corona taken together can be explained by} the migration of flux emergence to the poles in rapid rotators like TIC~277539431 \textcolor{red}{that drain the star's equatorial regions of magnetic flux, but preserve its ability to produce powerful flares}.
 
-    %Additionally, XMM-Newton captured a simultaneous X-ray and optical flare on TIC 277 with a bolometric energy of \eom and a soft X-ray energy of \eepic in the 0.2-2 keV band.
-In combination with the TESS data, we infer that TIC 277 produces flares at or above this energy approximately once a day.
-    %Our results indicate that TIC 277 is faint in X-ray copX-ray and flaring activity is representative of other rapidly rotating late M dwarfs, but that further measurements of localized flares are needed to understand the conditions under which high latitude flares in these stars can occur.
+    %Additionally, XMM-Newton captured a simultaneous X-ray and optical flare on TIC~277 with a bolometric energy of \eom and a soft X-ray energy of \eepic in the 0.2-2 keV band.
+In combination with the TESS data, we infer that TIC~277 produces flares at or above this energy approximately once a day.
+    %Our results indicate that TIC~277 is faint in X-ray copX-ray and flaring activity is representative of other rapidly rotating late M dwarfs, but that further measurements of localized flares are needed to understand the conditions under which high latitude flares in these stars can occur.
 
@@
 However, to be habitable in an Earth-like manner, merely the possibility of surface water in its liquid form does not suffice.
-Space weather of the host, that is, the high energy radiation and particles that impact the planet's atmosphere, affect its hospitality for life, for M dwarfs in particular~\citep{airapetian2020impact}.
+Space weather of the host, that is, the high energy radiation and particles that impact the planet's atmosphere, affect its hospitality for life, particularly for M dwarfs~\citep{airapetian2020impact}.
 If the energetic (photon or particle) flux is too high, the atmosphere may blow off~\citep[e.g., ][]{lammer2003atmospheric, garraffo2017threatening, ketzer2022influence} and water may be lost from the surface~\citep{doamaral2022contribution}.
-If not energetic enough, life may not emerge in the first place~\citep{rimmer2018origin}.
+If too low, life may not emerge in the first place~\citep{rimmer2018origin}.
 
-In M dwarfs, stellar activity evolution on the main sequence unfolds much more slowly than for \textcolor{red}{FGK stars}.
-While for the latter, magnetic activity, and with it the flares, winds, coronal mass ejections and energetic particle eruptions that make up their space weather, decline rapidly over a few hundred megayears, a fully convective M dwarf stays highly active for gigayears~\citep{magaudda2020relation, johnstone2021active, medina2022galactic}.
+In M dwarfs, stellar activity evolution on the main sequence unfolds much more slowly than \textcolor{red}{in FGK stars}.
+For the latter, magnetic activity, and with it the flares, winds, coronal mass ejections and energetic particle eruptions that make up their space weather, decline rapidly over a few hundred megayears.
+A fully convective M dwarf stays highly active for gigayears~\citep{magaudda2020relation, johnstone2021active, medina2022galactic}.
 Moreover, the zone where liquid water can reside is much closer to these faint stars.
@@
 The transition from star to brown dwarf is characterized by increasing rotation speed and decreasing coronal emission.
-\textcolor{red}{At the same time, the observed surge} surge of radio emission indicates a transition from stellar corona to planet-like magnetosphere at the bottom of the main sequence~\citep{zarka1998auroral,pineda2017panchromatic}.
+\textcolor{red}{At the same time, the observed surge} surge of radio emission at the bottom of the main sequence indicates a transition from stellar corona to planet-like magnetosphere~\citep{zarka1998auroral,pineda2017panchromatic}.
 Yet, brown dwarfs are regularly detected with energetic flares regardless of their expected low X-ray luminosity~\citep[e.g., ][]{hambaryan2004new, stelzer2006simultaneous, gizis2013kepler, paudel2020k2, schmidt2019largest}.
@@
 
-\citet{ilin2021giant} searched fully convective dwarfs that straddle the transition from star to brown dwarf with \textit{TESS}.
-They found four flares on four rapidly rotating ($P_{\rm rot}<10\,$h) M5-M7 dwarfs that lasted for multiple rotation periods of each star.
+In \citet{ilin2021giant}, we searched fully convective dwarfs that straddle the transition from star to brown dwarf with \textit{TESS}~\citep{ricker2015transiting}.
+We found four flares on four rapidly rotating ($P_{\rm rot}<10\,$h) M5-M7 dwarfs that lasted for multiple rotation periods of each star.
 The flares were observed to rotate in and out of view.
@@
 Our results indicate a preference for flares at high latitudes in these stars, as a chance finding \textcolor{red}{of these latitudes among latitudinally equidistributed flares was } unlikely ($\sim 0.1\%$).
-If flares and associated particle eruptions indeed have a latitude preference, the space weather of a planet in an aligned orbit would be less severe than previously thought.
+If flares and associated particle eruptions indeed had a latitude preference, the space weather of a planet in an aligned orbit would be less severe than previously thought.
 %Are high flaring latitudes  an effect of the dynamo at high rotation speeds?
@@
 
-\textcolor{red}{Among the studied objects, }TIC 277539431, TIC 277 for short, showed the highest latitude flare known to date at $\sim81^{\circ}$~\citep[][Table \ref{tab:starparams}]{ilin2021giant}.
+\textcolor{red}{Among the studied objects, }TIC~277539431, TIC~277 for short, showed the highest latitude flare known to date at $\sim81^{\circ}$~\citep[][Table \ref{tab:starparams}]{ilin2021giant}.
 It is an M7 dwarf with an extremely short rotation period of $4.56\,$h.
@@
 
-In this work, we follow TIC 277 with \textit{XMM-Newton} for 36 ks and use optical monitoring from \textit{TESS}~(Section~\ref{sec:obs}) to measure its coronal and flaring properties, respectively~(Section~\ref{sec:methods}).
-We contrast our results~(Section~\ref{sec:results}) with the literature to investigate if TIC~277 behaves like a low mass star or brown dwarf, investigate scenarios that are consistent with high latitude flare occurrence ~(Section~\ref{sec:discussion}), and \textcolor{red}{summarize our results} in Section~\ref{sec:summary}.
+In this work, we follow TIC~277 with \textit{XMM-Newton} for 36 ks and use optical monitoring from \textit{TESS}~(Section~\ref{sec:obs}) to measure its coronal and flaring properties, respectively~(Section~\ref{sec:methods}).
+We contrast our results~(Section~\ref{sec:results}) with the literature to investigate if TIC~277 behaves like a low mass star or brown dwarf, \textcolor{red}{consider scenarios that can explain the combined observations}~(Section~\ref{sec:discussion}), and \textcolor{red}{summarize our results} in Section~\ref{sec:summary}.
 
@@
 \centering
-    \caption{Stellar properties of TIC 277539431}
+    \caption{Stellar properties of TIC~277539431}
     \begin{tabular}{ll}\hline 
@@
 \textit{XMM-Newton} is an X-ray telescope, which \textcolor{red}{was} launched into orbit in 1999, equipped with six science instruments that operate simultaneously, among them the European Photon Imaging Camera with three cameras, PN, MOS1 and MOS2; and the Optical Monitor (OM) that provides near-UV and optical photometry, among other capabilities.
-We used \textit{XMM-Newton} to observe TIC 277 for 36 ks, i.e.
+We used \textit{XMM-Newton} to observe TIC~277 for 36 ks, i.e.
 two full rotation periods of the star, on August 5, 2022 (Proposal 090120; PI: E.
@@
 We extracted the events time series from EPIC using XMM SAS version 20\footnote{"Users Guide to the \textit{XMM-Newton} Science Analysis System", Issue 18.0, 2023 (ESA: \textit{XMM-Newton} SOC)} using the \texttt{evselect} task in the full $0.2-10\,$keV band.
-We selected a circular source region with 20 arcsec radius on all detectors, and a circular source-free background region 120 and 90 arcsec in MOS and PN, respectively.
+We selected a circular source region with 20 arcsec radius on all detectors, and a circular source-free background region with 120 and 90 arcsec radius in MOS and PN, respectively.
 We used the \texttt{epiclccorr} task in XMM SAS to correct for energy and time dependent loss of events\footnote{\url{https://heasarc.gsfc.nasa.gov/docs/xmm/sas/help/epiclccorr/node3.html}, accessed on July 3, 2023}, and extracted a soft X-ray time series with a time binning of $200\,$s.
 
-Simultaneously, we used the Optical Monitor (OM) aboard \textit{XMM-Newton}, to monitor TIC 277 in its white light filter at $10$\,s cadence.
+Simultaneously, we used the Optical Monitor (OM) aboard \textit{XMM-Newton}, to monitor TIC~277 in its white light filter at $10$\,s cadence.
 The OM white light filter is centered on $4060\,$\AA, with an equivalent width of $3470\,$\AA.
@@
 \subsubsection{EPIC Spectra}
-From the EPIC observations, we used the standard spectral extraction procedure for point sources, as specified in the XMM\textcolor{red}{-Newton} SAS handbook.
+From the EPIC observations, we used the standard spectral extraction procedure for point sources, as specified in the \textit{XMM-Newton} SAS handbook.
 We extracted spectra using \texttt{evselect} to select events in the circular event and background regions, \texttt{epatplot} to check for pile-up of multiple photons in one exposure, \texttt{backscale} to calculate the area of the source region, and \texttt{rmfgen} and \texttt{arfgen} to convolve the observed energies with the instrument response, and produce the spectra.
@@
 Since 2018, \textit{TESS} has been supplying publicly available red-optical high-precision time series photometry in an ongoing all-sky survey.
-Each uninterrupted observing Sector provides an approximately 27 day long light curve in a broad $6000$-$10000\,$\AA band pass.
+Each uninterrupted observing Sector provides an approximately 27 d long light curve in a broad $6000$-$10000\,$\AA\,band pass.
 
-We used \textcolor{red}{all} 5 Sectors of optical monitoring of TIC 277 in 2-minute cadence.
+We used \textcolor{red}{all} 5 Sectors of optical monitoring of TIC~277 in 2-minute cadence.
 Sector 12 was observed in May 2019, Sectors 37 and 39 in April and June 2021, and Sectors 64 and 65 in April and May 2023~(Fig.~\ref{fig:tess_lcs}).
@@
 From the X-ray observations with \textit{XMM-Newton}, we obtain a soft X-ray spectrum, which we model with two thermal emitter components~(Section~\ref{sec:methods:epic}).
-In TESS' optical time series photometry, we remove the variability introduced by the star's rotation, and find and characterize the flares in the so de-trended light curves~(Section~\ref{sec:methods:tess}).
+In \textit{TESS}' optical time series photometry, we remove the variability introduced by the star's rotation, and find and characterize the flares in the so de-trended light curves~(Section~\ref{sec:methods:tess}).
 \textit{XMM-Newton} captured a flare simultaneously in both the PN and MOS instruments in X-ray, and marginally the OM instrument in the optical.
-We derive its soft X-ray energy using the luminosity derived from X-ray spectrum, and its bolometric energy from the OM data.
-For the following comparison between optical and X-ray flaring activity, we also calculate the bolometric energies of the flares detected in TESS~(Section~\ref{sec:methods:flareenergies}).
+We derive its soft X-ray energy using the luminosity derived from the X-ray spectrum, and its bolometric energy from the OM data.
+For the following comparison between optical and X-ray flaring activity, we also calculate the bolometric energies of the flares detected in \textit{TESS}~(Section~\ref{sec:methods:flareenergies}).
 
@@
 The stellar corona can be described as an optically thin thermal plasma \textcolor{red}{that consists of multiple temperature components that represent different regions.
-The number of components one can identify in an X-ray spectrum depends on the brightness of the source.} We use XSPEC version 12.12.1~\citep{arnaud1996xspec} to fit a two-temperature (2T) additive \textcolor{red}{V}APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/0$=0.6$ ratio, which is more typical of M dwarfs~\citep{wood2018chandra}.
+The number of components one can identify in an X-ray spectrum depends on the brightness of the source.} We use XSPEC version 12.12.1~\citep{arnaud1996xspec} to fit a two-temperature (2T) additive \textcolor{red}{V}APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/O$=0.6$ ratio, which is more typical of M dwarfs~\citep{wood2018chandra}.
 We analyzed both the full data set, and the quiescent and flaring parts separately~\textcolor{red}{(see Fig.~\ref{fig:lightcurves})}.
@@
 
-\subsection{Lightcurve de-trending and flare finding: TESS}
+\subsection{Light curve de-trending and flare finding: TESS}
 \label{sec:methods:tess}
@@
 Black and red dots show the light curve with and without rotational variability and trends, respectively.
-The de-trended light curve is offset by $0.2$.
+The de-trended light curve is offset by $0.2$ \textcolor{red}{for visibility}.
 The large, rotationally modulated, flare, localized at about $81^{\circ}$ latitude by~\citet{ilin2021giant}, appears in the second half of the top panel~(Sector 12).}
@@
 In the TESS light curves, we found a total of 18 flares.
-To assess the flaring activity of TIC~277, we computed the bolometric energies for both OM and TESS flares.}
+To assess the flaring activity of TIC~277, we computed the bolometric energies for both the OM and the \textit{TESS} flares.}
 
-For the OM flare, we \textcolor{red}{used the} median flux of the lightcurve as the quiescent flux $F_0$, against which we measured the equivalent duration ($ED$) of the flare, i.e., the flare flux $F_{\rm flare}$, divided by $F_0$, integrated over the flare duration~\citep{gershberg1972results}:
+For the OM flare, we \textcolor{red}{used the} median flux of the light curve as the quiescent flux $F_0$, against which we measured the equivalent duration ($ED$) of the flare, i.e., the flare flux $F_{\rm flare}$, divided by $F_0$, integrated over the flare duration~\citep{gershberg1972results}:
 
@@
 We used the throughput curve for the white light filter scaled to unity at the peak of transmission as an optimistic response curve $R_{\lambda, \rm OM}$, given that the degradation of this filter's response is poorly constrained\footnote{\textit{XMM-Newton} User's Handbook 3.5.3.1, \url{https://xmm-tools.cosmos.esa.int/external/xmm_user_support/documentation/uhb/omfilters.html}, accessed Aug 8, 2023}.
-From that, we extracted a ratio $f_{\rm OM}$ of stellar to flare flux of $1.4\cdot 10^{-4}$ for a $10,000\,$K blackbody flare\footnote{A flare temperature of $T_{\rm flare}=10,000\,$K is a typical approximation for energetic M dwarf flares~\citep{kowalski2013timeresolved, howard2020evryflarea}.}:
+From that, we extracted a ratio $f_{\rm OM}$ of stellar to flare flux of $1.4\cdot 10^{-4}$ for a $10000\,$K blackbody flare\footnote{A flare temperature of $T_{\rm flare}=10000\,$K is a typical approximation for energetic M dwarf flares~\citep{kowalski2013timeresolved, howard2020evryflarea}.}:
 
@@
 Analogously to Eq.~\ref{eq:ratio} for the OM flare, we used the \textit{TESS} response curve to obtain a flux ratio $f_{\rm TESS}$ of $6.3\cdot10^{-3}$ in the optical filter of \textit{TESS}, and convert $ED$ to bolometric flare energy using Eq.~\ref{eq:eflare}.
-Besides the $ED$, \texttt{AltaiPony} also yields start and end time, defined as the difference between the first and last data point above the flare detection criterion defined in Section~\ref{sec:obs:tess}; and relative amplitude $a$, for each flare.
+Besides the $ED$, \texttt{AltaiPony} also yields start and end time, defined as the first and last data point above the flare detection criterion defined in Section~\ref{sec:obs:tess}; and relative amplitude $a$, for each flare.
 
-For the X-ray flare energy, we multiply flare duration $\Delta t$ by the flare luminosity in the EPIC data.
+For the X-ray flare energy, we multiply flare duration $\Delta t$, that is, the difference between start and end time, by the flare luminosity in the EPIC data.
 The flare luminosity is equal to the difference between the quiescent and flaring X-ray luminosities in the $0.2-2.0\,$keV band, i.e.:
@@
 \label{sec:results}
-With the \textit{XMM-Newton} and \textit{TESS} observations, we could constrain the coronal and flaring properties of TIC 277.
+With the \textit{XMM-Newton} and \textit{TESS} observations, we could constrain the coronal and flaring properties of TIC~277.
 We derive its coronal temperature and luminosity from the EPIC spectrum~(Section~\ref{sec:res:XrayTL}).
-We also measure the flare energies in X-rays from the EPIC instruments, and in optical from the OM, and compare them to the flare frequency distributions obtained from the 18 detected flares in \textit{TESS}~(Section~\ref{sec:res:flares}).
+We also measure the flare energies in X-rays from the EPIC instruments, and in optical from the OM, and compare them to the flare frequency distribution obtained from the 18 detected flares in \textit{TESS}~(Section~\ref{sec:res:flares}).
 
@@
 From the \textit{XMM-Newton} EPIC spectra, we could measure the coronal properties of TIC~277 in both its quiescent, and flaring states~(Table~\ref{tab:specfit}).
-The resulting fit for the full data set~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about \Tcool and \Thot\hspace{-1.cm}, respectively.
+The resulting fit for the full data set~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about \Tcool and \Thot\hspace{-.1cm}, respectively.
 Assuming these temperatures \textcolor{red}{for the prior distribution}, we also fitted 2T VAPEC models to the flaring and quiescent portions of the observations separately.
-The coronal temperatures did not update significantly using either subset.
+The coronal temperatures did not change significantly using either subset.
 However, the emission measure weighted mean temperature $T_{\rm mean}$ changed from \Tqmean in the quiescent spectrum to \Tfmean in the flaring spectrum, since the hot component became more much dominant during the flare.
-From the quiescent PN spectra, we then also calculated soft X-ray flux of $F_X=$\FX, and a luminosity $L_X=$\LXquiet in the $0.2$-$2.0\,$keV band.
+From the quiescent PN spectra, we then also calculated soft X-ray flux of $F_X=$\FX, and a luminosity $L_X=$\LXquiet in the $0.2-2.0\,$keV band.
 %This suggests that the flaring corona emits the hotter component, while the cooler component corresponds to the quiescent corona.
-In this interpretation, the presence of a hot component in the quiescent spectrum indicates that TIC 277 is flaring below the detection threshold, consistent with an extrapolation of the flare frequency distribution to lower temperatures~(see the following Section, and Fig.~\ref{fig:ffd}).
+In this interpretation, the presence of a hot component in the quiescent spectrum indicates that TIC~277 is flaring below the detection threshold, consistent with an extrapolation of the flare frequency distribution to lower temperatures~(see the following Section, and Fig.~\ref{fig:ffd}).
 
@@
     \caption{Flares detected with \textit{TESS}.
-$t_s$ is the starting time of the flare, $a$ is its relative amplitude, and $E_{\rm bol}$ is the bolometric energy assuming a 10,000 K blackbody emission from the flare.}
+$t_s$ is the starting time of the flare, $a$ is its relative amplitude, and $E_{\rm bol}$ is the bolometric energy assuming a $10000\,$K blackbody emission from the flare.}
     \input{output/tess_flares.tex}
@@
 The slope is typical of other flaring stars, regardless of spectral type and rotation period~(see, e.g., Fig.~13 \textcolor{red}{of} \citealt{ilin2021flares}).
-$R_{31.5}$ is typical of flaring M dwarfs, in particular typical of late M dwarfs in the saturated activity regime, where $R_{31.5}$ becomes independent of rotation period~\citep{medina2020flare,murray2022study}.
+$R_{31.5}$ is typical of flaring M dwarfs, \textcolor{red}{including} late M dwarfs in the saturated activity regime, where $R_{31.5}$ becomes independent of rotation period~\citep{medina2020flare,murray2022study}.
 
-For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of \eepic in the $0.2-2\,$keV band \textcolor{red}{by inserting} $L_{X,\rm quiescent}=$\LXquiet\unskip, $L_{X,\rm flaring}=$\LXflaring\unskip, and $\Delta t =$\xraydur\unskip\,(see Fig.~\ref{fig:lightcurves}) in Eq.~\ref{eq:xrayflare}.
+For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of \eepic in the $0.2-2.0\,$keV band \textcolor{red}{by inserting} $L_{X,\rm quiescent}=$\LXquiet\unskip, $L_{X,\rm flaring}=$\LXflaring\unskip, and $\Delta t =$\xraydur\unskip\,(see Fig.~\ref{fig:lightcurves}) in Eq.~\ref{eq:xrayflare}.
 The energy of the corresponding OM flare is \eom\unskip.
-We use the 36ks observing baseline of \textit{XMM-Newton} to calculate a flare rate in OM.
+We use the $36\,$ks observing baseline of \textit{XMM-Newton} to calculate a flare rate in OM.
 Fig.~\ref{fig:ffd} illustrates that it is roughly consistent, if slightly higher, than the extension of \textit{TESS}' \textcolor{red}{FFD} to lower energies.
 About one flare per day \textcolor{red}{with energies} of the order of the observed one or above can therefore be expected in future X-ray observations of TIC~277.
-\textcolor{red}{However, usually, not all optical flares have X-ray counterparts}.~\citet{paudel2021simultaneous}.
-\citet{guarcello2019simultaneous, kuznetsov2021stellar} find moderate correlation between flare energies in I data and \textit{XMM-Newton}, but the majority of their sampled flare energies are two orders of magnitude \textcolor{red}{higher than of} the flare in this work, both in X-ray and optical.
+\textcolor{red}{However, usually, not all optical flares have X-ray counterparts~\citep{paudel2021simultaneous}, and those that do show very variable ratios of optical vs.
+X-ray energy~\citep{guarcello2019simultaneous, kuznetsov2021stellar}.} We note that the X-ray flare light curve follows the typical fast-rise exponential decay shape, and is therefore unlikely to be eclipsed by the star \textcolor{red}{despite its fast rotation}.
+% find moderate correlation between flare energies in I data and \textit{XMM-Newton}, but the majority of their sampled flare energies are two orders of magnitude \textcolor{red}{higher than of} the flare in this work, both in X-ray and optical.
 \citet{kuznetsov2021stellar} compare primary Kepler mission observations with \textit{XMM-Newton} to find a similar result.
-We note that the X-ray flare light curve follows the typical fast-rise exponential decay shape, and is therefore unlikely to be eclipsed by the star.%~\citep{johnstone2012soft}.
+%~\citep{johnstone2012soft}.
 
@@
 Its coronal properties are therefore not known a priori.
-Our findings indicate that TIC 277's coronal temperature~(Section~\ref{sec:discussion:xraytemp}) and flaring activity~(Section~\ref{sec:discussion:flares}) are typical of saturated fully convective M dwarf stars.
-However, TIC 277 is rotating rapidly, even for the typically fast rotating late M dwarfs~\citep{medina2022galactic}, which is more typical of brown dwarfs.
-This combination could lead to an unusual manifestation of the stellar dynamo that produces high latitude flares that other dynamo states cannot generate~\citep{weber2016modeling, weber2023understanding}.
-TIC 277 is \textcolor{red}{fainter in X-ray by about an order of magnitude than M dwarfs on the canonical saturated branch}, which, in combination with the detection of a high latitude flare, may be an indication of polar updraft migration that is suspected to occur in rapid rotators~(Section~\ref{sec:discussion:xraylum}).
+Our findings indicate that TIC~277's coronal temperature~(Section~\ref{sec:discussion:xraytemp}) and flaring activity~(Section~\ref{sec:discussion:flares}) are typical of saturated fully convective M dwarf stars.
+However, TIC~277 is rotating rapidly, even for the typically fast rotating late M dwarfs~\citep{medina2022galactic}, which is more indicative of brown dwarfs.
+This combination could lead to an unusual manifestation of the stellar dynamo that produces high latitude flares~\citep{weber2016modeling, weber2023understanding}.
+TIC~277 is \textcolor{red}{fainter in X-ray by about an order of magnitude than M dwarfs on the canonical saturated branch}, which, in combination with the detection of a high latitude flare, may be an indication of polar updraft migration that is suspected to occur in rapid rotators~(Section~\ref{sec:discussion:xraylum}).
 
@@
 The resolution of different temperature components depends on the apparent brightness and integration time of the spectrum.
-Nearby stars like Proxima Cen can be modeled with a range of temperatures~\citep[e.g.,][]{gudel2004flares, drake2020pointing}, while one can often only capture few dominant temperatures in fainter objects.
+Nearby stars like Proxima Cen can be modeled with a range of temperatures~\citep[e.g.,][]{gudel2004flares, drake2020pointing}, while one can often only capture a few dominant temperatures in fainter objects.
 
-We find that the quiescent corona of TIC 277 is best described by two components, similar to many other active M dwarfs, in line with the standard picture for stellar flares~\citep{wargelin2008xray, robrade2010quiescent, behr2023muscles, magaudda2022firsta}.
+We find that the quiescent corona of TIC~277 is best described by two components, similar to many other active M dwarfs, in line with the standard picture for stellar flares~\citep{wargelin2008xray, robrade2010quiescent, behr2023muscles, magaudda2022firsta}.
 Both the cooler and the hotter component represent thermal radiation from hot plasma that evaporates from the chromosphere into the corona.
 The hotter component is driven by particles precipitating downward during flares that then heat the chromosphere to high temperatures, and cause additional evaporation into the coronal flaring loops~\citep{benz2016flare}.
-In our osbervation, this interpretation is supported by the observed delay between the optical and the X-ray flare~\citep[Fig.~\ref{fig:lightcurves}, ][]{hawley2003multiwavelength}.
-The optical flare originates from the lower chromosphere, which is heated by the precipitating particles.
+In our osbervations, this interpretation is supported by the observed delay between the optical and the X-ray flare~\citep[Fig.~\ref{fig:lightcurves}, ][]{hawley2003multiwavelength}: The optical flare originates from the lower chromosphere, which is heated by the precipitating particles.
 The heating then causes evaporation of particles into the flare loop.
@@
 
-The range of 6.5-11 MK for the emission measure weighted temperature of TIC 277 places it squarely in the saturated regime with other mid M dwarfs down to spectral type M6~\citep{wright2018stellar, magaudda2020relation, stelzer2022first,robrade2005xray,raassen2003xray,paudel2021simultaneous, foster2020corona}.
+The range of $6.5-11\,$MK for the emission measure weighted temperature of TIC~277 places it squarely in the saturated regime with other mid M dwarfs down to spectral type M6~\citep{wright2018stellar, magaudda2020relation, stelzer2022first,robrade2005xray,raassen2003xray,paudel2021simultaneous, foster2020corona}.
 Flares on the slowly rotating Proxima Cen (M5.5, $P_{\rm rot} = 83\,$d,~\citealt{anglada-escude2016terrestrial}), appear with a range of temperatures at 2-3 MK and 10-20 MK~\citep{gudel2004flares, fuhrmeister2011multiwavelength, fuhrmeister2022high, howard2022mouse} placing it in the transition region between saturated and unsaturated regimes.
@@
 Their intrinsic faintness and gradual decline in coronal emission toward the brown dwarf regime make them hard to detect~\citep{berger2010simultaneous, cook2014trends, stelzer2022first}.
-One example is NLTT 33370 AB, an M7 binary with a rotation period of about $3.8\,$h (close to TIC 277's $4.56\,$h, and of the same spectral type) is combined of a 3.1 MK and 14 MK component~\citep{williams2015simultaneous}.
-Another is TRAPPIST-1~(M8, $P_{\rm rot} = 3.3\,$d~\citep{luger2017sevenplanet}).
+One example is NLTT 33370 AB.
+This M7 binary with a rotation period of about $3.8\,$h (close to TIC~277's $4.56\,$h, and of the same spectral type) is combined of a 3.1 MK and 14 MK component~\citep{williams2015simultaneous}.
+Another is TRAPPIST-1~(M8, $P_{\rm rot} = 3.3\,$d~\citealt{luger2017sevenplanet}).
 Its corona appears cooler, with a 1.74 MK and a 9.6 MK component~\citep{wheatley2017strong}, and an emission measure weighted temperature of 5.36 MK~\citep{brown2023coronal}.
 
-Bearing in mind the relatively low two-temperature resolution of TIC~277's X-ray spectrum and the dearth of X-ray spectra for stars at the bottom of the main sequence, it shows a coronal temperature make-up common for mid-to-late M dwarfa in the saturated activity regime\textcolor{red}{, which} it will possibly keep for gigayears until it spins down to very low rotation rates~\citep{medina2022galactic, engle2023living}.
+Bearing in mind the relatively low two-temperature resolution of TIC~277's X-ray spectrum and the dearth of X-ray spectra for stars at the bottom of the main sequence, it shows a coronal temperature make-up common for mid-to-late M dwarfs in the saturated activity regime\textcolor{red}{, which} it will possibly keep for gigayears until it spins down to very low rotation rates~\citep{medina2022galactic, engle2023living}.
 
@@
 
-TIC~277's coronal luminosity relative to bolometric luminosity of $L_X/L_{\rm bol}=$\LXLbol in quiescence is about an order of magnitude lower than \textcolor{red}{the canonical} $L_X/L_{\rm bol}\sim 10^{-3}$ average in the saturated regime of partly and fully convective M dwarfs\citep{wright2011stellaractivityrotation,wright2016solartype,wright2018stellar}.% In the Hyades and Praesepe, \citet{nunez2022factory} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend in $L_X/L_{\rm bol}$ with decreasing rotation period for M3.5-M7 dwarfs.
+TIC~277's coronal luminosity relative to bolometric luminosity of $L_X/L_{\rm bol}=$\LXLbol in quiescence is about an order of magnitude lower than \textcolor{red}{the canonical} $L_X/L_{\rm bol}\sim 10^{-3}$ average in the saturated regime of partly and fully convective M dwarfs~\citep{wright2011stellaractivityrotation,wright2016solartype,wright2018stellar}.% In the Hyades and Praesepe, \citet{nunez2022factory} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend in $L_X/L_{\rm bol}$ with decreasing rotation period for M3.5-M7 dwarfs.
 This tentatively points toward a supersaturation of the corona at fast rotation speeds, as has been suggested before for M~dwarfs~\citep{jeffries2011investigating}, but also earlier type stars~\citep{prosser1996rosat, argiroffi2016supersaturation} and even red giants~\citep{dixon2020rotationally}.
@@
 
-At later spectral type, a decline in coronal activity is expected as the magnetosphere transitions from a stellar to a planetary one~\citep{pineda2017panchromatic}.
+At later spectral types, a decline in coronal activity is expected as the magnetosphere transitions from a stellar to a planetary one~\citep{pineda2017panchromatic}.
 \citet{stelzer2022first} investigated late M and L dwarfs with eROSITA to show that $L_X/L_{\rm bol}$ increases towards later spectral types~\citep{magaudda2022firsta}, but begins to decrease again from M7 downward.
@@
         ID & SpT &$P_{\rm rot}$ [d] & $\log_{10} \frac{L_X}{L_{\rm bol}} $ & energy range [keV]\\\hline
-        TIC 277 (this work) & M7 & 0.19 & $-4$ & $0.2-2$ \\
+        TIC~277 (this work) & M7 & 0.19 & $-4$ & $0.2-2$ \\
         TRAPPIST-1 [1] & M8 & 3.3 & $-4$ & $0.3-10$\\
@@
 
-\textcolor{red}{In late M dwarfs like TIC~277, the decline in X-ray activity is expected as a result of decreasing efficiency in coronal heating in the transition region between a stellar and a brown dwarf atmosphere~(e.g.,~\citealt{williams2014trends}).
+\textcolor{red}{In late M dwarfs like TIC~277, the decline in X-ray activity is expected as a result of decreasing efficiency of coronal heating~(e.g.,~\citealt{williams2014trends}).
 However, the flaring activity and coronal temperature of TIC~277 are in line with saturated activity, implying that the efficiency of flare production has not diminished.} An alternative explanation, that would at the same time explain the occurrence of a high latitude flare in \textit{TESS}, is polar updraft migration~\citep{stepien2001rosat}.
@@
 As a result, the X-ray luminosity diminishes.
-Since TIC~277 is seen nearly equator-on~(Table~\ref{tab:starparams}), polar updraft could explain its fainter corona.
+Since TIC~277 is seen nearly equator-on~(Table~\ref{tab:starparams}), polar updraft migration could explain its fainter corona.
 Considering the high-latitude flare, viewing geometry, star-like flaring behavior and \textcolor{red}{star-like} coronal temperatures together, we \textcolor{red}{suggest this scenario as an alternative to the brown dwarf transition explanation}, that would invoke a diminishing corona due to the overall cooler, more neutral atmosphere.
@@
 
-%Compared to the X-ray luminosities of M7 dwarfs among the late M dwarfs observed by eROSITA in \citet{stelzer2022first}, TIC 277 appears mildly .
+%Compared to the X-ray luminosities of M7 dwarfs among the late M dwarfs observed by eROSITA in \citet{stelzer2022first}, TIC~277 appears mildly .
 This could be an effect of its fast rotation, and consequent supersaturation as in \citet{magaudda2022first}, but we lack the rotation periods to confirm this.
@@
 \label{sec:summary}
-We investigated whether TIC 277, a rapidly rotating M7 dwarf with a flare localized at $81^{\circ}$ latitude, shows coronal or flaring properties that could explain the occurrence of this flare so close to the rotational pole.
+We investigated whether TIC~277, a rapidly rotating M7 dwarf with a flare localized at $81^{\circ}$ latitude, shows coronal or flaring properties that could explain the occurrence of this flare so close to the rotational pole.
 We obtained 36 ks of EPIC and OM observations with \textit{XMM-Newton}, and studied the five Sectors of red-optical 27-day light curves provided by \textit{TESS}.
@@
 
-An indication of an unusual corona stems from its X-ray luminosity relative to bolometric luminosity $L_X/L_{\rm bol}=$\LXLbol.
+An indication of an unusual corona stems from its X-ray luminosity relative to bolometric luminosity $L_X/L_{\rm bol}=$\LXLbol\hspace{-0.1cm}.
 \textcolor{red}{It is an order of magnitude lower than the canonical value for saturated activity M dwarfs, but representative of late M dwarfs of spectral type M7 and later}.
-TIC 277 is an extreme rapid rotator with a rotation period of only $4.56\,$h.
+TIC~277 is an extremely rapid rotator with a rotation period of only $4.56\,$h.
 The detected high-latitude flare in \textit{TESS}~\citep{ilin2021giant} may hence be a product of high-latitude flux emergence driven by its fast rotation~\citep{weber2016modeling,weber2017suppression}.
-\textcolor{red}{The low X-ray luminosity and the high-latitude flare taken together with the star's fast rotation, saturated flaring activity, and equator-on viewing angle can be explained by polar updraft migration, which suppresses coronal emission in equatorial regions, and confines the corona and flaring regions to near-polar regions~\citep{stepien2001rosat}.
+\textcolor{red}{The low X-ray luminosity and the high-latitude flare taken together with the star's fast rotation, saturated flaring activity, and equator-on viewing angle can be explained by polar updraft migration, which suppresses coronal emission in equatorial regions, and confines the corona and flaring regions to near-polar latitudes~\citep{stepien2001rosat}.
 }
@@
  E.I.
-acknowledges funding from the Deutsche Luft- und Raumfahrtgesellschaft (FKZ 50 OR 2209, project \textit{X-ray Loops}).
+\textcolor{red}{and D.D.} acknowledge funding from the Deutsche Luft- und Raumfahrtgesellschaft (FKZ 50 OR 2209, project \textit{X-ray Loops}).
 K.P.

### 2024-02-27 src/scripts/FIGURE_tess_lcs.py
added vertical lines to flares

@@ -17,6 +17,7 @@ import matplotlib.pyplot as plt
 import paths
 
 import numpy as np
+import pandas as pd
 
 
 if __name__ == "__main__":
@@ -31,6 +32,12 @@ if __name__ == "__main__":
     # detrended flux
     fig, axes = plt.subplots(5, 1, figsize=(13, 15))
 
+    # read in the flare table
+    df = pd.read_csv(paths.data / "tess_flares.csv")
+
+    # select the columns we want, sort by time
+    sel = df[["tstart", "ampl_rec", 'ed_rec', 'ed_rec_err', 'Sector']].sort_values("tstart")
+
     # loop over axes, light curves and sectors
     for ax, lcr, sector in zip(axes, lcrs, sectors):
 
@@ -39,6 +46,13 @@ if __name__ == "__main__":
                 ms=1.5)
         ax.plot(lcr['TIME'], lcr['DETRENDED_FLUX']/np.nanmedian(lcr["DETRENDED_FLUX"]) +0.2, 
                 'r.', ms=1.5,)
+        
+        # plot the flares as vertical lines at the bottom of the panel
+
+        flare_times = sel[sel["Sector"] == sector]["tstart"].values
+
+        for t in flare_times:
+                ax.axvline(t, color="blue", lw=0.5, alpha=0.5)
         
         # layout
         ax.set_ylabel("normalized flux", fontsize=13)

### 2024-02-27 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \affiliation{Leibniz Insitute for Astrophysics Potsdam (AIP), An der Sternwarte 16, 14482 Potsdam, Germany}
+\affiliation{ASTRON, Netherlands Institute for Radio Astronomy, Oude Hoogeveensedijk 4, Dwingeloo, 7991 PD, The Netherlands}
 
@@
 \affiliation{Institut für Astronomie \& Astrophysik, Eberhard-Karls-Universität Tübingen, Sand 1, 72076 Tübingen, Germany}
-\affiliation{INAF – Osservatorio Astronomico di Palermo, Piazza del Parlamento 1, 90134 Palermo, Italy}
+%\affiliation{INAF – Osservatorio Astronomico di Palermo, Piazza del Parlamento 1, 90134 Palermo, Italy}
 
@@
 The mechanisms that allow flares at high latitudes to occur are poorly understood.
-We studied five Sectors of \textit{TESS} monitoring, and obtained 36 ks of \textit{XMM-Newton} observations, to find that TIC~277539431's corona does not differ significantly from other low mass stars on the \textcolor{red}{canonical saturated activity branch with respect to coronal temperatures and flaring activity, but shows lower luminosity in soft X-ray emission by about an order of magnitude, consistent with other late M dwarfs.
-The lack of X-ray flux, the high latitude flare, the star's viewing geometry, and the otherwise typical stellar corona taken together can be explained by} the migration of flux emergence to the poles in rapid rotators like TIC~277539431 \textcolor{red}{that drain the star's equatorial regions of magnetic flux, but preserve its ability to produce powerful flares}.
+We studied five Sectors of \textit{TESS} monitoring, and obtained 36 ks of \textit{XMM-Newton} observations, to find that TIC~277539431's corona does not differ significantly from other low mass stars on the canonical saturated activity branch with respect to coronal temperatures and flaring activity, but shows lower luminosity in soft X-ray emission by about an order of magnitude, consistent with other late M dwarfs.
+The lack of X-ray flux, the high latitude flare, the star's viewing geometry, and the otherwise typical stellar corona taken together can be explained by the migration of flux emergence to the poles in rapid rotators like TIC~277539431 that drain the star's equatorial regions of magnetic flux, but preserve its ability to produce powerful flares.
 
@@
 Space weather of the host, that is, the high energy radiation and particles that impact the planet's atmosphere, affect its hospitality for life, particularly for M dwarfs~\citep{airapetian2020impact}.
-If the energetic (photon or particle) flux is too high, the atmosphere may blow off~\citep[e.g., ][]{lammer2003atmospheric, garraffo2017threatening, ketzer2022influence} and water may be lost from the surface~\citep{doamaral2022contribution}.
+If the energetic photon or particle flux is too high, the atmosphere may blow off~\citep[e.g., ][]{lammer2003atmospheric, garraffo2017threatening, ketzer2022influence} and water may be lost from the surface~\citep{doamaral2022contribution}.
 If too low, life may not emerge in the first place~\citep{rimmer2018origin}.
 
-In M dwarfs, stellar activity evolution on the main sequence unfolds much more slowly than \textcolor{red}{in FGK stars}.
+In M dwarfs, stellar activity evolution on the main sequence unfolds much more slowly than in FGK stars.
 For the latter, magnetic activity, and with it the flares, winds, coronal mass ejections and energetic particle eruptions that make up their space weather, decline rapidly over a few hundred megayears.
@@
 The transition from star to brown dwarf is characterized by increasing rotation speed and decreasing coronal emission.
-\textcolor{red}{At the same time, the observed surge} surge of radio emission at the bottom of the main sequence indicates a transition from stellar corona to planet-like magnetosphere~\citep{zarka1998auroral,pineda2017panchromatic}.
-Yet, brown dwarfs are regularly detected with energetic flares regardless of their expected low X-ray luminosity~\citep[e.g., ][]{hambaryan2004new, stelzer2006simultaneous, gizis2013kepler, paudel2020k2, schmidt2019largest}.
+At the same time, the observed surge of radio emission at the bottom of the main sequence indicates a transition from stellar corona to planet-like magnetosphere~\citep{zarka1998auroral,pineda2017panchromatic}.
+Yet, brown dwarfs are regularly detected with energetic flares regardless of their expected low X-ray luminosity~\citep[e.g., ][]{ stelzer2006simultaneous, gizis2013kepler, paudel2020k2, schmidt2019largest, audard2007chandra, deluca2020extras}.
 How these flares are produced in the apparent absence of a solar-type corona is unclear~\citep{mullan2018frequencies}.%.
@@
 
-In \citet{ilin2021giant}, we searched fully convective dwarfs that straddle the transition from star to brown dwarf with \textit{TESS}~\citep{ricker2015transiting}.
+In \citet{ilin2021giant}, we searched fully convective late M dwarfs that mark the transition from star to brown dwarf with \textit{TESS}~\citep{ricker2015transiting}.
 We found four flares on four rapidly rotating ($P_{\rm rot}<10\,$h) M5-M7 dwarfs that lasted for multiple rotation periods of each star.
@@
 These flares were found significantly closer to the pole than to the equator, in contrast to the Sun, where flares are usually found below $30^{\circ}$ latitude.
-Our results indicate a preference for flares at high latitudes in these stars, as a chance finding \textcolor{red}{of these latitudes among latitudinally equidistributed flares was } unlikely ($\sim 0.1\%$).
+Our results indicate a preference for flares at high latitudes in these stars, as a chance finding of these latitudes among latitudinally equidistributed flares was unlikely ($\sim 0.1\%$).
 If flares and associated particle eruptions indeed had a latitude preference, the space weather of a planet in an aligned orbit would be less severe than previously thought.
@@
 
-\textcolor{red}{Among the studied objects, }TIC~277539431, TIC~277 for short, showed the highest latitude flare known to date at $\sim81^{\circ}$~\citep[][Table \ref{tab:starparams}]{ilin2021giant}.
+Among the studied objects, TIC~277539431, TIC~277 for short, showed the highest latitude flare known to date at $\sim81^{\circ}$~\citep[][Table \ref{tab:starparams}]{ilin2021giant}.
 It is an M7 dwarf with an extremely short rotation period of $4.56\,$h.
-While the fast rotation and late spectral type could be indicative of \textcolor{red}{a diminishing corona, the detection of flares suggests otherwise.} %The star's age is unknown.
+While the fast rotation and late spectral type could be indicative of a diminishing corona, the detection of flares suggests otherwise.
+%The star's age is unknown.
 \citet{schneider2018hazmat} classify it as a field star, but its FUV and NUV fluxes are on the high end for field stars and would, if not as high, be compatible with the NUV/FUV emission from younger open cluster members.
@@
 
-In this work, we follow TIC~277 with \textit{XMM-Newton} for 36 ks and use optical monitoring from \textit{TESS}~(Section~\ref{sec:obs}) to measure its coronal and flaring properties, respectively~(Section~\ref{sec:methods}).
-We contrast our results~(Section~\ref{sec:results}) with the literature to investigate if TIC~277 behaves like a low mass star or brown dwarf, \textcolor{red}{consider scenarios that can explain the combined observations}~(Section~\ref{sec:discussion}), and \textcolor{red}{summarize our results} in Section~\ref{sec:summary}.
+In this work, we follow TIC~277 with \textit{XMM-Newton} for $36\,$ks and use optical monitoring from \textit{TESS}~(Section~\ref{sec:obs}) to measure its coronal and flaring properties, respectively~(Section~\ref{sec:methods}).
+We contrast our results~(Section~\ref{sec:results}) with the literature to investigate if TIC~277 behaves like a low mass star or brown dwarf, consider scenarios that can explain the combined observations~(Section~\ref{sec:discussion}), and summarize our results in Section~\ref{sec:summary}.
 
@@
     % [4] \citet{collaboration2016gaia,gaiacollaboration2023gaia}, 
-    [3] \citet{pecaut2013intrinsic}
+    [3] \citet{pecaut2013intrinsic},
     [*] see Section~\ref{sec:methods:flareenergies}
@@
 \label{sec:obs:xmm}
-\textit{XMM-Newton} is an X-ray telescope, which \textcolor{red}{was} launched into orbit in 1999, equipped with six science instruments that operate simultaneously, among them the European Photon Imaging Camera with three cameras, PN, MOS1 and MOS2; and the Optical Monitor (OM) that provides near-UV and optical photometry, among other capabilities.
-We used \textit{XMM-Newton} to observe TIC~277 for 36 ks, i.e.
+\textit{XMM-Newton} is an X-ray telescope that was launched into orbit in 1999.
+It was equipped with six science instruments that operate simultaneously, among them the European Photon Imaging Camera with three cameras, PN, MOS1 and MOS2; and the Optical Monitor (OM) that provides near-UV and optical photometry, among other capabilities.
+We used \textit{XMM-Newton} to observe TIC~277 for $36\,$ks, i.e.
 two full rotation periods of the star, on August 5, 2022 (Proposal 090120; PI: E.
@@
 
-Simultaneously, we used the Optical Monitor (OM) aboard \textit{XMM-Newton}, to monitor TIC~277 in its white light filter at $10$\,s cadence.
+Simultaneously, we used the OM aboard \textit{XMM-Newton} to monitor TIC~277 in its white light filter at $10$\,s cadence.
 The OM white light filter is centered on $4060\,$\AA, with an equivalent width of $3470\,$\AA.
-\textcolor{red}{In each light curve, OM observed uninterruptedly for $4390\,$s, followed by a short gap of $318\,$s before the next one.
+In each light curve, OM observed uninterruptedly for $4390\,$s, followed by a short gap of $318\,$s before the next one.
 We used the \texttt{omfchain} in XMM SAS to extract the 8 individual light curves.
-}
 
@@
         \caption{
-         Top panel: Optical Monitor light curve.
-Bottom panel: Background-subtracted X-ray light curve using the flux in the entire $0.2-10\,$keV range  (PN, MOS1 and MOS2 combined).
+         Top panel: \textit{XMM-Newton} Optical Monitor light curve.
+Bottom panel: Background-subtracted \textit{XMM-Newton} X-ray light curve using the flux in the entire $0.2-10\,$keV range  (PN, MOS1 and MOS2 combined).
 The grey shaded portion defines the flare-only subset of the observations (see Section~\ref{sec:methods:epic} and Table~\ref{tab:specfit}).
@@
 We extracted spectra using \texttt{evselect} to select events in the circular event and background regions, \texttt{epatplot} to check for pile-up of multiple photons in one exposure, \texttt{backscale} to calculate the area of the source region, and \texttt{rmfgen} and \texttt{arfgen} to convolve the observed energies with the instrument response, and produce the spectra.
-Finally, we use \texttt{grppha} to rebin the data in the spectra to obtain at least 15 counts per bin.
+Finally, we used \texttt{grppha} to rebin the data in the spectra to obtain at least 15 counts per bin.
 To produce spectra for the quiescent and flaring portions of the observations separately, we used \texttt{tabgtigen} to select the time intervals to pass to \texttt{evselect}.
@@
 
-We used \textcolor{red}{all} 5 Sectors of optical monitoring of TIC~277 in 2-minute cadence.
+We used all 5 Sectors of optical monitoring of TIC~277 at a $2\,$min cadence.
 Sector 12 was observed in May 2019, Sectors 37 and 39 in April and June 2021, and Sectors 64 and 65 in April and May 2023~(Fig.~\ref{fig:tess_lcs}).
@@
 \label{sec:methods:epic}
-The stellar corona can be described as an optically thin thermal plasma \textcolor{red}{that consists of multiple temperature components that represent different regions.
-The number of components one can identify in an X-ray spectrum depends on the brightness of the source.} We use XSPEC version 12.12.1~\citep{arnaud1996xspec} to fit a two-temperature (2T) additive \textcolor{red}{V}APEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/O$=0.6$ ratio, which is more typical of M dwarfs~\citep{wood2018chandra}.
-We analyzed both the full data set, and the quiescent and flaring parts separately~\textcolor{red}{(see Fig.~\ref{fig:lightcurves})}.
+The stellar corona can be described as an optically thin thermal plasma that consists of multiple temperature components that represent different regions.
+The number of components one can identify in an X-ray spectrum depends on the brightness of the source.
+We use XSPEC version 12.12.1~\citep{arnaud1996xspec} to fit a two-temperature (2T) additive VAPEC~\citep{smith2001collisional,foster2012updated} model to the joint PN and MOS spectra, using solar abundances from~\citet{grevesse1998standard}, but adjusting to an Fe/O$=0.6$ ratio, which is more typical of M dwarfs~\citep{wood2018chandra}.
+We analyzed both the full data set, and the quiescent and flaring parts separately~(see Fig.~\ref{fig:lightcurves}).
 Neither subset could be adequately fit with a single temperature component (1T), and a 3T model did not improve the fit compared to the 2T-model.
@@
 \label{sec:methods:tess}
-We used AltaiPony~\citep{ilin2021altaipony} to de-trend \textcolor{red}{the Pre-search Data Conditioning Simple Aperture Photometry (PDCSAP)} light curves using the de-trending function \texttt{custom\_detrending}\footnote{see also the documentation online at \url{https://altaipony.readthedocs.io/en/latest/tutorials/detrend.html}, accessed on July 3, 2023}, detailed in~\citet{ilin2022searching}.
-The procedure begins with a 3rd order spline fitting of coarse trends, followed by iterative sine fitting to remove rotational modulation, and two Savitzky-Golay filters~\citep{savitzky1964smoothing} with a 6h and 3h window, respectively to remove aperiodic short term trends, while masking potential flare candidates in each step.
+We used AltaiPony~\citep{ilin2021altaipony} to de-trend the Pre-search Data Conditioning Simple Aperture Photometry (PDCSAP) light curves (Fig.~\ref{fig:tess_lcs}) using the de-trending function \texttt{custom\_detrending}\footnote{see also the documentation online at \url{https://altaipony.readthedocs.io/en/latest/tutorials/detrend.html}, accessed on July 3, 2023}, detailed in~\citet{ilin2022searching}.
+The procedure begins with a 3rd order spline fitting of coarse trends, followed by iterative sine fitting to remove rotational modulation, and two Savitzky-Golay filters~\citep{savitzky1964smoothing} with a $6\,$h and $3\,$h window, respectively to remove aperiodic short term trends, while masking potential flare candidates in each step.
 The masking takes three steps each time.
 First, all outliers above $>2.5\sigma$ are masked.
-\textcolor{red}{Second, to mask the flare decay, we append the six-fold number of data points masked in the first step to the mask.} Third, three adjacent data points before and after the so expanded mask are added to the mask.
+Second, to mask the flare decay, we append the six-fold number of data points masked in the first step to the mask.
+Third, three adjacent data points before and after the so expanded mask are added to the mask.
 
-In the de-trended light curves, we searched for flare candidates as at least 3 consecutive $>3\sigma$ outliers above the noise level\textcolor{red}{, defined as the rolling standard deviation of the de-trended light curve flux with a 2 h window}, filling in the noise level inside the previously masked areas with the mean of the noise levels adjacent to the mask.
-To the series of outliers, we kept \textcolor{red}{appending} data points to the flare candidate until a data point falls below $2\sigma$.
+In the de-trended light curves, we searched for flare candidates as at least 3 consecutive $>3\sigma$ outliers above the noise level.
+Here the noise level is defined as the rolling standard deviation of the de-trended light curve flux within a $2$\,h window, filling in the noise level inside the previously masked areas with the mean of the noise levels adjacent to the mask.
+To the series of outliers, we kept appending data points to the flare candidate until a data point falls below $2\sigma$.
 
@@
 Black and red dots show the light curve with and without rotational variability and trends, respectively.
-The de-trended light curve is offset by $0.2$ \textcolor{red}{for visibility}.
-The large, rotationally modulated, flare, localized at about $81^{\circ}$ latitude by~\citet{ilin2021giant}, appears in the second half of the top panel~(Sector 12).}
+The de-trended light curve is offset by $0.2$ for better visibility.
+The large, rotationally modulated, flare, localized at about $81^{\circ}$ latitude presented by~\citet{ilin2021giant}, appears in the second half of the top panel~(Sector 12).}
         \label{fig:tess_lcs}
@@
 \label{sec:methods:flareenergies}
-\textcolor{red}{The OM data contain one marginal flare that we interpret as associated with the large X-ray flare shortly after~(inset in Fig.~\ref{fig:lightcurves}).
+The OM data contain one marginal flare that we interpret as associated with the large X-ray flare shortly after~(inset in Fig.~\ref{fig:lightcurves}).
 In the TESS light curves, we found a total of 18 flares.
-To assess the flaring activity of TIC~277, we computed the bolometric energies for both the OM and the \textit{TESS} flares.}
+To assess the flaring activity of TIC~277, we computed the bolometric energies for both the OM and the \textit{TESS} flares.
 
-For the OM flare, we \textcolor{red}{used the} median flux of the light curve as the quiescent flux $F_0$, against which we measured the equivalent duration ($ED$) of the flare, i.e., the flare flux $F_{\rm flare}$, divided by $F_0$, integrated over the flare duration~\citep{gershberg1972results}:
+For the OM flare, we used the median flux of the light curve as the quiescent flux $F_0$, against which we measured the equivalent duration ($ED$) of the flare, i.e., the flare flux $F_{\rm flare}$, divided by $F_0$, integrated over the flare duration~\citep{gershberg1972results}:
 
@@
 
-We then converted \textcolor{red}{the} $ED$ to flare energy using the bolometric flare energy following the procedure in~\citet{shibayama2013superflares}.
+We then converted the $ED$ to flare energy using the bolometric flare energy following the procedure in~\citet{shibayama2013superflares}.
 We used the throughput curve for the white light filter scaled to unity at the peak of transmission as an optimistic response curve $R_{\lambda, \rm OM}$, given that the degradation of this filter's response is poorly constrained\footnote{\textit{XMM-Newton} User's Handbook 3.5.3.1, \url{https://xmm-tools.cosmos.esa.int/external/xmm_user_support/documentation/uhb/omfilters.html}, accessed Aug 8, 2023}.
@@
 \begin{equation}
-    E_{\rm flare} = ED \cdot \pi R^2 \cdot \sigma_{kB} T_{\rm eff}^4 \cdot f_{\rm OM}
+    E_{\rm flare} = ED \cdot \pi R^2 \cdot k_{\text{B}} T_{\rm eff}^4 \cdot f_{\rm OM},
     \label{eq:eflare}
@@
 
-We note that due to the optimistic response curve assumption, the resulting $E_{\rm flare}$ is likely higher in reality.
+where $k_{\text{B}}$ is the Boltzmann constant. We note that due to the optimistic response curve assumption, the resulting $E_{\rm flare}$ is likely higher in reality.
 
@@
 \label{sec:results}
-With the \textit{XMM-Newton} and \textit{TESS} observations, we could constrain the coronal and flaring properties of TIC~277.
+In this work, we use the \textit{XMM-Newton} and \textit{TESS} observations to constrain the coronal and flaring properties of TIC~277.
 We derive its coronal temperature and luminosity from the EPIC spectrum~(Section~\ref{sec:res:XrayTL}).
@@
 The resulting fit for the full data set~(Fig.~\ref{fig:spec_joint_all}) consists of a cooler and a hotter component of about \Tcool and \Thot\hspace{-.1cm}, respectively.
-Assuming these temperatures \textcolor{red}{for the prior distribution}, we also fitted 2T VAPEC models to the flaring and quiescent portions of the observations separately.
+Assuming these temperatures for the prior distribution, we also fitted 2T VAPEC models to the flaring and quiescent portions of the observations separately.
 The coronal temperatures did not change significantly using either subset.
-However, the emission measure weighted mean temperature $T_{\rm mean}$ changed from \Tqmean in the quiescent spectrum to \Tfmean in the flaring spectrum, since the hot component became more much dominant during the flare.
-From the quiescent PN spectra, we then also calculated soft X-ray flux of $F_X=$\FX, and a luminosity $L_X=$\LXquiet in the $0.2-2.0\,$keV band.
+However, the emission measure weighted mean temperature $T_{\rm mean}$ changed from \Tqmean in the quiescent spectrum to \Tfmean in the flaring spectrum, when the hot component became much more dominant during the flare.
+From the quiescent PN spectra, we then also calculated the X-ray flux of $F_X=$\FX, and luminosity $L_X=$\LXquiet in the soft $0.2-2.0\,$keV band.
 %This suggests that the flaring corona emits the hotter component, while the cooler component corresponds to the quiescent corona.
@@
 
-We note that not using the full set of observations as the prior resulted in an unconstrained hot coronal component during the flare.
-This can be interpreted as due to a lack of X-ray flux above $5\,$keV.
+We note that not using the full set of observations as the prior resulted in an unconstrained hot coronal component during the flare, which could be due to a lack of X-ray flux above $5\,$keV.
 %Hot flares with temperatures of $>50\,$MK have been measured on DS~Tuc~A~\citep{pillitteri2011xray}.
@@
         \caption{
-         \textcolor{red}{XMM-Newton EPIC spectra.
-The top panel shows the spectra taken with the different EPIC instruments together with the best-fit two-temperature VAPEC model with Fe/O$=0.6$ for the full data set.} The bottom panel shows the residuals to the fit.
+         Time-averaged \textit{XMM-Newton} EPIC spectra.
+The top panel shows the spectra taken with the different EPIC instruments together with the best-fit two-temperature VAPEC model with Fe/O$=0.6$ for the full data set.
+The bottom panel shows the residuals to the fit.
         }
@@
 The FFD power law fit converged after 13,500 steps on a slope $\alpha =$\ffdalpha\unskip.
-The frequency $R_{31.5}$\textcolor{red}{, that is the frequency} of flares per day above $\log_{10} E_{\rm flare}\,[\mathrm{erg}] = 31.5$ is $\log_{10}R_{31.5}=$\ffdr\unskip.
-The slope is typical of other flaring stars, regardless of spectral type and rotation period~(see, e.g., Fig.~13 \textcolor{red}{of} \citealt{ilin2021flares}).
-$R_{31.5}$ is typical of flaring M dwarfs, \textcolor{red}{including} late M dwarfs in the saturated activity regime, where $R_{31.5}$ becomes independent of rotation period~\citep{medina2020flare,murray2022study}.
+The frequency $R_{31.5}$, that is the frequency of flares per day above $\log_{10} E_{\rm flare}\,[\mathrm{erg}] = 31.5$ is $\log_{10}R_{31.5}=$\ffdr\unskip.
+The slope is typical of other flaring stars, regardless of spectral type and rotation period~(see, e.g., Fig.~13 of \citealt{ilin2021flares}).
+$R_{31.5}$ is typical of flaring M dwarfs, including late M dwarfs in the saturated activity regime, where $R_{31.5}$ becomes independent of rotation period~\citep{medina2020flare,murray2022study}.
 
-For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of \eepic in the $0.2-2.0\,$keV band \textcolor{red}{by inserting} $L_{X,\rm quiescent}=$\LXquiet\unskip, $L_{X,\rm flaring}=$\LXflaring\unskip, and $\Delta t =$\xraydur\unskip\,(see Fig.~\ref{fig:lightcurves}) in Eq.~\ref{eq:xrayflare}.
-The energy of the corresponding OM flare is \eom\unskip.
+For the X-ray flare in Fig.~\ref{fig:lightcurves}, we obtain a total energy of \eepic in the $0.2-2.0\,$keV band by inserting the quiescent and flaring luminosities (Table~\ref{tab:specfit}), and $\Delta t =$\xraydur\unskip\,(see Fig.~\ref{fig:lightcurves}) into Eq.~\ref{eq:xrayflare}.
+With Eq.~\ref{eq:ratio}, we find an energy of \eom\unskip\, for the corresponding OM flare.
 We use the $36\,$ks observing baseline of \textit{XMM-Newton} to calculate a flare rate in OM.
-Fig.~\ref{fig:ffd} illustrates that it is roughly consistent, if slightly higher, than the extension of \textit{TESS}' \textcolor{red}{FFD} to lower energies.
-About one flare per day \textcolor{red}{with energies} of the order of the observed one or above can therefore be expected in future X-ray observations of TIC~277.
-\textcolor{red}{However, usually, not all optical flares have X-ray counterparts~\citep{paudel2021simultaneous}, and those that do show very variable ratios of optical vs.
-X-ray energy~\citep{guarcello2019simultaneous, kuznetsov2021stellar}.} We note that the X-ray flare light curve follows the typical fast-rise exponential decay shape, and is therefore unlikely to be eclipsed by the star \textcolor{red}{despite its fast rotation}.
-% find moderate correlation between flare energies in I data and \textit{XMM-Newton}, but the majority of their sampled flare energies are two orders of magnitude \textcolor{red}{higher than of} the flare in this work, both in X-ray and optical.
+Fig.~\ref{fig:ffd} illustrates that it is roughly consistent, if slightly higher, than the extension of \textit{TESS}' FFD to lower energies.
+About one flare per day with energies of the order of the observed one or above can therefore be expected in future X-ray observations of TIC~277.
+However, usually, not all optical flares have X-ray counterparts~\citep{paudel2021simultaneous}, and those that do show very variable ratios of optical vs.
+X-ray energy~\citep{guarcello2019simultaneous, kuznetsov2021stellar}.
+We note that the X-ray flare light curve follows the typical fast-rise exponential decay shape, and is therefore unlikely to be eclipsed by the star despite its fast rotation.
+% find moderate correlation between flare energies in I data and \textit{XMM-Newton}, but the majority of their sampled flare energies are two orders of magnitude higher than of the flare in this work, both in X-ray and optical.
 \citet{kuznetsov2021stellar} compare primary Kepler mission observations with \textit{XMM-Newton} to find a similar result.
@@
 The rate of OM flares (gray square) is calculated from the observing baseline of \textit{XMM-Newton}.
-\textcolor{red}{The flare energies are corrected for the pass bands of \textit{TESS} and OM, i.e., given as bolometric flare energies~(see Section~\ref{sec:methods:flareenergies}).}
+The flare energies are corrected for the pass bands of \textit{TESS} and OM, i.e., given as bolometric flare energies~(see Section~\ref{sec:methods:flareenergies}).
         }
@@
 TIC~277's spectral type and rotation places it in the middle of the transition from star to brown dwarf.
-Its coronal properties are therefore not known a priori.
 Our findings indicate that TIC~277's coronal temperature~(Section~\ref{sec:discussion:xraytemp}) and flaring activity~(Section~\ref{sec:discussion:flares}) are typical of saturated fully convective M dwarf stars.
-However, TIC~277 is rotating rapidly, even for the typically fast rotating late M dwarfs~\citep{medina2022galactic}, which is more indicative of brown dwarfs.
+However, TIC~277 is rotating rapidly, even for the typically fast rotating late M dwarfs~\citep{medina2022galactic}.
 This combination could lead to an unusual manifestation of the stellar dynamo that produces high latitude flares~\citep{weber2016modeling, weber2023understanding}.
-TIC~277 is \textcolor{red}{fainter in X-ray by about an order of magnitude than M dwarfs on the canonical saturated branch}, which, in combination with the detection of a high latitude flare, may be an indication of polar updraft migration that is suspected to occur in rapid rotators~(Section~\ref{sec:discussion:xraylum}).
+TIC~277 is fainter in X-ray by about an order of magnitude than earlier-type M dwarfs on the canonical saturated branch~\citep{magaudda2022firsta}, which, in combination with the detection of a high latitude flare, may be an indication of polar updraft migration that is suspected to occur in rapid rotators~(Section~\ref{sec:discussion:xraylum}).
 
@@
 \label{sec:discussion:xraytemp}
+
 Low mass stars' coronae are not uniform.
-Magnetic structures create a range of temperatures, starting at about 1 MK by definition and rising up to 30 MK in flares, in the heterogeneous Solar corona~\citep{vaiana1978recent}.
+Magnetic structures create a range of temperatures, starting at about 1 MK, and rising up to 30 MK in flares, in the heterogeneous Solar corona~\citep{vaiana1978recent}.
 Stellar point source X-ray observations are a superposition of various temperatures and densities.
@@
 
-We find that the quiescent corona of TIC~277 is best described by two components, similar to many other active M dwarfs, in line with the standard picture for stellar flares~\citep{wargelin2008xray, robrade2010quiescent, behr2023muscles, magaudda2022firsta}.
+We find that the \textit{XMM-Newton} observations of the quiescent corona of TIC~277 are best described by two temperature components, similar to many other active M dwarfs.
+Assuming that TIC~277 produces flares below the detection threshold of \textit{XMM-Newton} and thereby contribute to the temperature makeup of the quiescent corona, the presence of a second, hot component is consistent with the standard picture for stellar flares~\citep{wargelin2008xray, robrade2010quiescent, behr2023muscles, magaudda2022firsta}.
 Both the cooler and the hotter component represent thermal radiation from hot plasma that evaporates from the chromosphere into the corona.
-The hotter component is driven by particles precipitating downward during flares that then heat the chromosphere to high temperatures, and cause additional evaporation into the coronal flaring loops~\citep{benz2016flare}.
-In our osbervations, this interpretation is supported by the observed delay between the optical and the X-ray flare~\citep[Fig.~\ref{fig:lightcurves}, ][]{hawley2003multiwavelength}: The optical flare originates from the lower chromosphere, which is heated by the precipitating particles.
+The hotter component is driven by particles precipitating downward during flares that then heat the chromosphere to high temperatures, and cause evaporation into the coronal flaring loops, in addition to quiescent heating through other sources under debate, such as Alfv\'en waves or nanoflares~\citep{benz2016flare}.
+In our observations, this interpretation is supported by the observed delay between the optical and the X-ray flare~\citep[Fig.~\ref{fig:lightcurves}, ][]{stelzer2022great}: The optical flare originates from the lower chromosphere, which is heated by the precipitating particles.
 The heating then causes evaporation of particles into the flare loop.
-In the cooler component, the source of the more quiescent coronal heating is an open question, even for the Sun, where the respective contributions of reconnection and magnetohydrodynamical waves are under investigation~\citep{vandoorsselaere2020coronal}.
+%In the cooler component, the source of the more quiescent coronal heating is an open question, even for the Sun, where the respective contributions of reconnection and magnetohydrodynamical waves are under investigation~\citep{vandoorsselaere2020coronal}.
 %A third, very hot, possibly $>40\,$MK component could only be poorly constrained due to a lack of sensitivity to flux above $5\,$keV, and its short lifetime during the peak of the flare.
@@
 The range of $6.5-11\,$MK for the emission measure weighted temperature of TIC~277 places it squarely in the saturated regime with other mid M dwarfs down to spectral type M6~\citep{wright2018stellar, magaudda2020relation, stelzer2022first,robrade2005xray,raassen2003xray,paudel2021simultaneous, foster2020corona}.
-Flares on the slowly rotating Proxima Cen (M5.5, $P_{\rm rot} = 83\,$d,~\citealt{anglada-escude2016terrestrial}), appear with a range of temperatures at 2-3 MK and 10-20 MK~\citep{gudel2004flares, fuhrmeister2011multiwavelength, fuhrmeister2022high, howard2022mouse} placing it in the transition region between saturated and unsaturated regimes.
-Only at rotation periods beyond 100 days, flaring activity declines to yield coronal temperatures $<2\,$MK in M3-M6 dwarfs~\citep{wright2018stellar, foster2020corona}.
+Flares on the slowly rotating Proxima Cen (M5.5, $P_{\rm rot} = 83\,$d,~\citealt{anglada-escude2016terrestrial}), appear with a range of temperatures at 2-3 MK and 10-20 MK~\citep{gudel2004flares, fuhrmeister2011multiwavelength, fuhrmeister2022high, howard2022mouse}.
+Its X-ray luminosity~\citep{haisch1990rotational} and flaring activity~\citep{howard2018first, medina2020flare} place it in the transition region between saturated and unsaturated regimes.
+Only at rotation periods beyond 100 days, stellar activity declines to yield coronal temperatures $<2\,$MK in M3-M6 dwarfs~\citep{wright2018stellar, foster2020corona}.
 
@@
 One example is NLTT 33370 AB.
-This M7 binary with a rotation period of about $3.8\,$h (close to TIC~277's $4.56\,$h, and of the same spectral type) is combined of a 3.1 MK and 14 MK component~\citep{williams2015simultaneous}.
-Another is TRAPPIST-1~(M8, $P_{\rm rot} = 3.3\,$d~\citealt{luger2017sevenplanet}).
+This M7 binary with a rotation period of about $3.8\,$h (close to TIC~277's $4.56\,$h, and of the same spectral type) is composed of a 3.1 MK and 14 MK component~\citep{williams2015simultaneous}.
+Another is TRAPPIST-1~(M8, \mbox{$P_{\rm rot} = 3.3\,$d}~\citealt{luger2017sevenplanet}).
 Its corona appears cooler, with a 1.74 MK and a 9.6 MK component~\citep{wheatley2017strong}, and an emission measure weighted temperature of 5.36 MK~\citep{brown2023coronal}.
 
-Bearing in mind the relatively low two-temperature resolution of TIC~277's X-ray spectrum and the dearth of X-ray spectra for stars at the bottom of the main sequence, it shows a coronal temperature make-up common for mid-to-late M dwarfs in the saturated activity regime\textcolor{red}{, which} it will possibly keep for gigayears until it spins down to very low rotation rates~\citep{medina2022galactic, engle2023living}.
+Bearing in mind the relatively low two-temperature resolution of TIC~277's X-ray spectrum and the dearth of X-ray spectra for stars at the bottom of the main sequence, it shows a coronal temperature make-up common for mid-to-late M dwarfs in the saturated activity regime, which it will possibly keep for gigayears until it spins down to very low rotation rates~\citep{medina2022galactic, engle2023living}.
 
@@
 
-TIC~277's rapid rotation and UV flux is indicative of a younger age than TRAPPIST-1, although it has not been attributed to any star cluster so far~\citep{schneider2018hazmat}.
+TIC~277's rapid rotation, and UV flux hint at a younger age than TRAPPIST-1, although it has not been attributed to any moving group so far~\citep{schneider2018hazmat}, and the number of late M dwarfs with independently measured age and rotation is too small to infer age based on rotation alone~\citep{engle2023living}.
 Interpreting the upcoming observations of TRAPPIST-1 e, one of the rocky planets in the habitable zone of the system, with the James Webb Space Telescope will most likely involve its activity history.
@@
 
-TIC~277's coronal luminosity relative to bolometric luminosity of $L_X/L_{\rm bol}=$\LXLbol in quiescence is about an order of magnitude lower than \textcolor{red}{the canonical} $L_X/L_{\rm bol}\sim 10^{-3}$ average in the saturated regime of partly and fully convective M dwarfs~\citep{wright2011stellaractivityrotation,wright2016solartype,wright2018stellar}.% In the Hyades and Praesepe, \citet{nunez2022factory} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend in $L_X/L_{\rm bol}$ with decreasing rotation period for M3.5-M7 dwarfs.
+TIC~277's coronal luminosity relative to bolometric luminosity of $L_X/L_{\rm bol}=$\LXLbol in quiescence is about an order of magnitude lower than the canonical $L_X/L_{\rm bol}\sim 10^{-3}$ average in the saturated regime of partly and fully convective M dwarfs~\citep{wright2011stellaractivityrotation,wright2016solartype,wright2018stellar}.% In the Hyades and Praesepe, \citet{nunez2022factory} cover both fully convective and rapidly rotating M dwarfs, and find a downward trend in $L_X/L_{\rm bol}$ with decreasing rotation period for M3.5-M7 dwarfs.
 This tentatively points toward a supersaturation of the corona at fast rotation speeds, as has been suggested before for M~dwarfs~\citep{jeffries2011investigating}, but also earlier type stars~\citep{prosser1996rosat, argiroffi2016supersaturation} and even red giants~\citep{dixon2020rotationally}.
@@
 \citet{stelzer2022first} investigated late M and L dwarfs with eROSITA to show that $L_X/L_{\rm bol}$ increases towards later spectral types~\citep{magaudda2022firsta}, but begins to decrease again from M7 downward.
-In their compilation, together with previous work from \citet{stelzer2012ultracool, cook2014trends, deluca2020extras, williams2014trends, berger2008simultaneous}, TIC~277 appears at the lower end\textcolor{red}{, but consistent with} the $L_X/L_{\rm bol}$ distribution for M7 dwarfs.
+In their compilation, together with previous work from \citet{stelzer2012ultracool, cook2014trends, deluca2020extras, williams2014trends, berger2008simultaneous}, TIC~277 appears at the lower end, but consistent with the $L_X/L_{\rm bol}$ distribution for M7 dwarfs.
 However, the rotation periods are unknown for most of these late M dwarfs.
-Sometimes projected rotation periods are known~\citep{cook2014trends}, but true periods are only detected in a handful of these stars, which we list in Table~\ref{tab:latmlxlbol}.
-TIC~277 appears \textcolor{red}{at the lower bound of} this small sample.
-\textcolor{red}{In Magaudda et al.
-(in prep.), rapidly rotating late M dwarfs observed with eROSITA appear with a range of luminosities centered around $L_X/L_{\rm bol}\approx 10^{-4}$.}
+Sometimes projected rotation periods are known~\citep{cook2014trends}, but photometric periods have been published for a handful of these stars, which we list in Table~\ref{tab:latmlxlbol}.
+TIC~277 appears at the lower bound of this small sample, which may be a selection effect.
+In Magaudda et al.
+(in prep.), rapidly rotating late M dwarfs appear with a broad range of X-ray luminosities centered around $L_X/L_{\rm bol}\approx 10^{-4}$.
 
@@
     \centering
-    \caption{Late M dwarfs with known rotation period and X-ray luminosity.}
+    \caption{Late M dwarfs with published rotation period and X-ray luminosity.}
     \begin{tabular}{lcccc}
@@
         NLTT 33370 AB [3] & M7 & 0.16 & $-3.5$ & $0.2-2$ \\
-        LP 412-31  [4] & M8 & 0.61 & $-3.1$& \textcolor{green}{soft X-ray band?}\\\hline
+        LP 412-31  [4] & M8 & 0.61 & $-3.1$& $0.3-8$\\\hline
     \end{tabular}
@@
 
-\textcolor{red}{In late M dwarfs like TIC~277, the decline in X-ray activity is expected as a result of decreasing efficiency of coronal heating~(e.g.,~\citealt{williams2014trends}).
-However, the flaring activity and coronal temperature of TIC~277 are in line with saturated activity, implying that the efficiency of flare production has not diminished.} An alternative explanation, that would at the same time explain the occurrence of a high latitude flare in \textit{TESS}, is polar updraft migration~\citep{stepien2001rosat}.
-Polar updraft migration implies that at high rotation rates, flux emergence becomes more efficient near the poles than at the equator~\citep{weber2016modeling}\textcolor{red}{, producing flares there, but} draining the equatorial regions of magnetic flux required to produce a corona \textcolor{red}{at low latitudes}.
+In late M dwarfs like TIC~277, the decline in X-ray activity is expected as a result of decreasing efficiency of coronal heating~(e.g.,~\citealt{mohanty2002activity, williams2014trends}).
+However, the flaring activity and coronal temperature of TIC~277 are in line with saturated activity, suggesting that the efficiency of flare production has not diminished.
+An alternative explanation, that would at the same time explain the occurrence of a high latitude flare in \textit{TESS}, is polar updraft migration~\citep{stepien2001rosat}.
+Polar updraft migration implies that at high rotation rates, flux emergence becomes more efficient near the poles than at the equator~\citep{weber2016modeling} producing flares there, but draining the equatorial regions of magnetic flux required to produce a corona at low latitudes.
 As a result, the X-ray luminosity diminishes.
 Since TIC~277 is seen nearly equator-on~(Table~\ref{tab:starparams}), polar updraft migration could explain its fainter corona.
-Considering the high-latitude flare, viewing geometry, star-like flaring behavior and \textcolor{red}{star-like} coronal temperatures together, we \textcolor{red}{suggest this scenario as an alternative to the brown dwarf transition explanation}, that would invoke a diminishing corona due to the overall cooler, more neutral atmosphere.
+Considering the high-latitude flare, viewing geometry, star-like flaring behavior and star-like coronal temperatures together, we suggest this scenario as an alternative to the brown dwarf transition explanation, that would invoke a diminishing corona due to the overall cooler, more neutral atmosphere.
 
@@
 \label{sec:summary}
-We investigated whether TIC~277, a rapidly rotating M7 dwarf with a flare localized at $81^{\circ}$ latitude, shows coronal or flaring properties that could explain the occurrence of this flare so close to the rotational pole.
+We investigated whether TIC~277, a rapidly rotating M7 dwarf that exhibited a flare localized at $81^{\circ}$ latitude, shows coronal or flaring properties that could explain the occurrence of this flare so close to the rotational pole.
 We obtained 36 ks of EPIC and OM observations with \textit{XMM-Newton}, and studied the five Sectors of red-optical 27-day light curves provided by \textit{TESS}.
@@
 An indication of an unusual corona stems from its X-ray luminosity relative to bolometric luminosity $L_X/L_{\rm bol}=$\LXLbol\hspace{-0.1cm}.
-\textcolor{red}{It is an order of magnitude lower than the canonical value for saturated activity M dwarfs, but representative of late M dwarfs of spectral type M7 and later}.
+It is an order of magnitude lower than the canonical value for saturated activity M dwarfs, but representative of late M dwarfs of spectral type M7 and later.
 TIC~277 is an extremely rapid rotator with a rotation period of only $4.56\,$h.
 The detected high-latitude flare in \textit{TESS}~\citep{ilin2021giant} may hence be a product of high-latitude flux emergence driven by its fast rotation~\citep{weber2016modeling,weber2017suppression}.
-\textcolor{red}{The low X-ray luminosity and the high-latitude flare taken together with the star's fast rotation, saturated flaring activity, and equator-on viewing angle can be explained by polar updraft migration, which suppresses coronal emission in equatorial regions, and confines the corona and flaring regions to near-polar latitudes~\citep{stepien2001rosat}.
-}
+The low X-ray luminosity and the high-latitude flare taken together with the star's fast rotation, saturated flaring activity, and equator-on viewing angle could be explained by polar updraft migration, which suppresses coronal emission in equatorial regions, and confines the corona and flaring regions to near-polar latitudes~\citep{stepien2001rosat}.
 
-\textcolor{red}{Overall, }less than ten flare latitudes on stars other than the Sun are known to date~\citep{wolter2008doppler, ilin2021giant, johnson2021simultaneous}.
+Overall, less than ten flare latitudes on stars other than the Sun are known to date~\citep{wolter2008doppler, ilin2021giant, johnson2021simultaneous}.
 Localization of more flares on the surfaces of a broader range of stars will allow us to resolve if polar updraft migration indeed occurs in these stars.
@@
  E.I.
-\textcolor{red}{and D.D.} acknowledge funding from the Deutsche Luft- und Raumfahrtgesellschaft (FKZ 50 OR 2209, project \textit{X-ray Loops}).
+and D.D.
+acknowledge funding from the Deutsche Luft- und Raumfahrtgesellschaft (FKZ 50 OR 2209, project \textit{X-ray Loops}).
 K.P.
